### 2020-08-21 Chapter_0/04_Manuscript/99_PermeabilityModelAICs.tex
first commit

% latex table generated in R 3.6.3 by xtable 1.8-4 package
% Thu Jun  4 15:54:16 2020
\begin{tabular}{lrrrr}
  \toprule
Covariates & AIC & DeltaAIC & Weight & LogLik \\ 
  \midrule
cos(ta) + log(sl) + W + T + DTW + HB5 + S & 90091.40 & 0.00 & 1.00 & -45030.70 \\ 
  cos(ta) + log(sl) + W + T + DTW + HB5 + S + DTR & 90093.51 & 2.11 & 0.00 & -45029.75 \\ 
  cos(ta) + log(sl) + W + T + DTW + HB5 & 90094.79 & 3.38 & 0.00 & -45034.39 \\ 
  cos(ta) + log(sl) + W + T + DTW + HB5 + S + P & 90095.05 & 3.65 & 0.00 & -45030.52 \\ 
  cos(ta) + log(sl) + W + T + DTW + S & 90096.95 & 5.55 & 0.00 & -45035.47 \\ 
  cos(ta) + log(sl) + W + T + DTW + HB5 + DTR & 90097.02 & 5.62 & 0.00 & -45033.51 \\ 
  cos(ta) + log(sl) + W + T + DTW + HB5 + S + DTR + P & 90097.38 & 5.98 & 0.00 & -45029.69 \\ 
  cos(ta) + log(sl) + W + T + DTW + HB5 + S + RC & 90097.53 & 6.13 & 0.00 & -45029.76 \\ 
  cos(ta) + log(sl) + W + T + DTW + HB5 + P & 90098.46 & 7.06 & 0.00 & -45034.23 \\ 
  cos(ta) + log(sl) + W + T + DTW + HB5 + S + DTR + RC & 90099.62 & 8.22 & 0.00 & -45028.81 \\ 
  cos(ta) + log(sl) + W + T + DTW & 90099.97 & 8.57 & 0.00 & -45038.99 \\ 
  cos(ta) + log(sl) + W + T + DTW + HB5 + RC & 90100.68 & 9.28 & 0.00 & -45033.34 \\ 
  cos(ta) + log(sl) + W + T + DTW + DTR & 90101.68 & 10.28 & 0.00 & -45037.84 \\ 
  cos(ta) + log(sl) + W + T + DTW + P & 90103.07 & 11.67 & 0.00 & -45038.54 \\ 
  cos(ta) + log(sl) + W + T + DTW + HB5 + S + DTR + P + RC & 90103.48 & 12.08 & 0.00 & -45028.74 \\ 
  cos(ta) + log(sl) + W + T + DTW + RC & 90104.72 & 13.32 & 0.00 & -45037.36 \\ 
  cos(ta) + log(sl) + W + T + HB5 & 90123.71 & 32.31 & 0.00 & -45050.85 \\ 
  cos(ta) + log(sl) + W + T & 90128.17 & 36.77 & 0.00 & -45055.09 \\ 
  cos(ta) + log(sl) + W + T + S & 90129.36 & 37.96 & 0.00 & -45053.68 \\ 
  cos(ta) + log(sl) + W + T + DTR & 90130.11 & 38.71 & 0.00 & -45054.05 \\ 
  cos(ta) + log(sl) + W + T + P & 90131.75 & 40.35 & 0.00 & -45054.88 \\ 
  cos(ta) + log(sl) + W + T + RC & 90133.53 & 42.13 & 0.00 & -45053.77 \\ 
  cos(ta) + log(sl) + W + DTW & 90138.92 & 47.52 & 0.00 & -45060.46 \\ 
  cos(ta) + log(sl) + W + HB5 & 90149.30 & 57.90 & 0.00 & -45065.65 \\ 
  cos(ta) + log(sl) + W + S & 90151.14 & 59.74 & 0.00 & -45066.57 \\ 
  cos(ta) + log(sl) + W & 90153.70 & 62.30 & 0.00 & -45069.85 \\ 
  cos(ta) + log(sl) + W + DTR & 90156.32 & 64.92 & 0.00 & -45069.16 \\ 
  cos(ta) + log(sl) + W + P & 90157.23 & 65.83 & 0.00 & -45069.62 \\ 
  cos(ta) + log(sl) + W + RC & 90159.32 & 67.91 & 0.00 & -45068.66 \\ 
  cos(ta) + log(sl) + S & 90161.07 & 69.67 & 0.00 & -45073.54 \\ 
  cos(ta) + log(sl) + DTW & 90262.03 & 170.63 & 0.00 & -45124.01 \\ 
  cos(ta) + log(sl) + T & 90301.56 & 210.16 & 0.00 & -45143.78 \\ 
  cos(ta) + log(sl) + HB5 & 90307.55 & 216.15 & 0.00 & -45146.78 \\ 
  cos(ta) + log(sl) + DTR & 90316.37 & 224.97 & 0.00 & -45151.19 \\ 
  cos(ta) + log(sl) + P & 90316.62 & 225.22 & 0.00 & -45151.31 \\ 
  cos(ta) + log(sl) + RC & 90318.46 & 227.06 & 0.00 & -45150.23 \\ 
   \bottomrule
\end{tabular}