### 2020-08-21 Chapter_0/04_Manuscript/Manuscript_Backup_2020-01-12.tex

\documentclass[abstract=on,10pt,a4paper,bibliography=totocnumbered]{scrartcl}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{colortbl}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage{epstopdf}
\usepackage[onehalfspacing]{setspace}
\usepackage[paper=a4paper,left=30mm,right=30mm,top=20mm,bottom=25mm]{geometry}
\usepackage[round]{natbib}
\usepackage[hidelinks]{hyperref}
\usepackage[nameinlink]{cleveref}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage[labelfont=bf]{caption}
\usepackage{footnote}
\usepackage{tabularx}
\usepackage{gensymb}
\usepackage{todonotes}
\usepackage{tikz}
\usepackage{subcaption}
\usepackage{varwidth}
% \usepackage[nomarkers,figuresonly]{endfloat}
\usepackage{csvsimple}
\usepackage{arydshln}
\usepackage{titlesec}

% Define a new command to create a number inside a circle (for plots)
\newcommand*\circled[1]{\tikz[baseline=(char.base)]{
            \node[shape=circle,draw,inner sep=1pt] (char) {#1};}}

% Creating some TikZ styles
\tikzset{
  nonterminal/.style = {rectangle
    , minimum size = 6mm
    , very thick
    , draw = black!
  }
}

% Increase the space between two rows in tables
\renewcommand{\arraystretch}{1.2}

% Changing the style of captions in figures etc.
\captionsetup{labelfont=bf, format=plain, font=small}

% Adjust the distance between title and paragraph
\RedeclareSectionCommand[
  beforeskip=-.5\baselineskip,
  afterskip=.5\baselineskip]{subsection}
\RedeclareSectionCommand[
  beforeskip=-.25\baselineskip,
  afterskip=.25\baselineskip]{subsubsection}

\doublespacing
\begin{document}
\pagenumbering{gobble}
\bibliographystyle{apalike}

\begin{titlepage}

% Define a new command to produce a horizontal line
\newcommand{\HRule}{\rule{\linewidth}{0.5mm}}

% Center everything on the page
\center

%------------------------------------------------------------------------------
%	HEADING SECTIONS
%------------------------------------------------------------------------------

% Name of your university/college
\textsc{\LARGE  }\\[1.5cm]

% Major heading
\textsc{\Large Department of Evolutionary Biology and Environmental
Studies}\\[0.5cm]

% Minor heading
\textsc{\large Master's Thesis}\\[0.5cm]

%------------------------------------------------------------------------------
%	TITLE SECTION
%------------------------------------------------------------------------------
\doublespacing
\HRule \\[0.4cm]

% Title of the document
{\textsc{ \LARGE Ranging Wide: Movement Corridors of Dispersing African Wild
Dogs (\textit{Lycaon pictus}) in the Kavango-Zambezi Transfrontier Conservation
Area}}\\[0.4cm]

\HRule \\[1.5cm]
\singlespacing

%------------------------------------------------------------------------------
%	AUTHOR SECTION
%------------------------------------------------------------------------------

\begin{minipage}{0.4\textwidth}
\begin{flushleft} \large

% Main author
\emph{Author:}\\
David D. Hofmann\\
david.hofmann2@uzh.ch\\
12-728-846

\end{flushleft}
\end{minipage}
~
\begin{minipage}{0.4\textwidth}
\begin{flushright} \large

% Supervisors
\emph{Supervisors:} \\
Dominik Behr\\
Ph.D. Gabriele Cozzi\\
Prof. Arpat Ozgul

\end{flushright}
\end{minipage}\\[2cm]

%------------------------------------------------------------------------------
%	DATE SECTION
%------------------------------------------------------------------------------

% Today's date
{\large \today}\\[5cm]

%------------------------------------------------------------------------------
%	LOGO SECTION
%------------------------------------------------------------------------------

% Logo of the university
\includegraphics{UZHVector.eps}\\[1cm]

%------------------------------------------------------------------------------

% Fill the rest of the page with whitespace
\vfill

\end{titlepage}

\begin{abstract}
The African wild dog \textit{(Lycaon pictus)} is an endangered species that
currently only survives in a few scattered subpopulations in Sub-Saharan Africa.
Long-term viability of the species thus requires the identification and
preservation of key dispersal corridors that connect such subpopulations. We
collected GPS data of 16 dispersing coalitions originating from a free-ranging
population in northern Botswana and used integrated step selection analysis to
create a permeability surface spanning over the entire Kavango-Zambezi
Transfrontier Conservation Area (KAZA-TFCA), the largest transboundary
conservation area on earth. Based on our permeability surface, we calculated
least-cost paths and least-cost corridors between selected locations within the
KAZA-TFCA. Our results show that dispersal corridors typically run through open
shrubs/grassland and in the vicinity of water. Corridors rarely cross
waterbodies, areas with high human presence, or areas densely vegetated by
trees. Several dispersal corridors meet in the Okavango-Linyanty ecosystem,
which highlights the importance of northern Botswana for the conservation of
this charismatic species.
\end{abstract}
\newpage

\onehalfspacing
\tableofcontents
\doublespacing
\newpage

\pagenumbering{arabic}

\section{Introduction}
% Importance of Connectivity
Successful conservation of fragmented and isolated subpopulations requires
maintenance and enhancement of connectivity between habitats occupied by these
subpopulations \citep{Hanski.1998}. In this context, connectivity describes the
degree to which the landscape matrix facilitates or impedes movement of
individuals and thereby genes of the focal species between habitat patches
\citep{Taylor.1993, Clobert.2012}. Thus, sufficient connectivity facilitates
dispersal, which in turn promotes genetic exchange, reinforcement and rescuing
of small populations, and colonization or recolonization of unoccupied habitats
\citep{Brown.1977, Hanski.1998, MacArthur.2001, Frankham.2002, Leigh.2012}. All
these processes crucially contribute to long-term population viability, which
emphasizes the importance of dispersal and connectivity. However, increasing
anthropogenic pressure and ongoing habitat loss and fragmentation reduce
connectivity on a worldwide scale \citep{Fahrig.2003, Barnosky.2012}. As a
result, management actions for the conservation of endangered species are
increasingly aimed towards preserving and restoring connectivity through the
identification and preservation of key dispersal corridors that link isolated
habitats \citep{Fahrig.2003, Heller.2009}. An unprecedented initiative to
restore connectivity on a large scale is represented by the creation of the
Kavango-Zambezi Transfrontier Conservation Area (KAZA-TFCA), a protected area
which traverses five African countries and spans over 520,000
km\textsuperscript{2}. While such initiatives are expected to bring benefits to
wide-ranging species such as elephants (\textit{Loxodonta africana}), African
wild dogs (\textit{Lycaon pictus}), and lions (\textit{Panthera leo}), the
extent to which they are effective for targeted species and whether boundaries
reflect natural dispersal corridors is rarely assessed.

% Connectivity Assessment
In recent years, there has been a growing body of literature that uses animal
movement data to identify dispersal corridors and to assess connectivity on
large scales \citep{Chetkiewicz.2006}. Identification of potential dispersal
corridors between suitable habitat patches traditionally relies on the
estimation of a permeability surface (elsewhere referred to as resistance
surface, when inverted), which renders the ease or willingness at which the
focal species traverses a specific landscape. Latest empirical methods to
predict landscape permeability based on recorded GPS relocation data include
\textit{resource selection functions} (RSF, \cite{Boyce.2002}), \textit{step
selection functions} (SSF, \cite{Fortin.2005}), and \textit{path selection
functions} (PSF, \cite{Cushman.2010}). All methods have in common that they
infer habitat preferences of the focal species by comparing covariates (e.g.
environmental and anthropogenic) at visited locations to the same covariates at
randomly selected locations that are deemed available but presumably unused by
the animal. Based on inferred habitat preferences the methods further assume
that habitats preferred by the focal species provide high permeability, whereas
avoided habitats provide low permeability. Regardless of the method chosen (i.e.
RSF, SSF, or PSF), all approaches require adequate GPS relocation data that is
representative for the process being studied. More specifically, it has been
shown that GPS relocation data collected on dispersing individuals is better
suited for detecting dispersal corridors than GPS relocation data collected on
resident individuals \citep{Elliot.2014}. Few studies have, however, partitioned
their GPS relocation data according to behavioral mode such as \textit{resident}
vs. \textit{dispersing} or \textit{active} vs. \textit{inactive} periods of
movement \citep{Wilson.2012, Vasudev.2015}. As a consequence, most permeability
surfaces upon which dispersal corridors are identified are created using GPS
relocation data that was collected on resident individuals. This introduces
severe biases and substantially reduces the power to reveal meaningful
corridors, for dispersing individuals have different needs and motivations
compared to resident individuals \citep{Killeen.2014, Elliot.2014, Cozzi.2020}.
For instance, \cite{Elliot.2014} demonstrated marked differences in predicted
permeability surfaces for dispersing lions in the KAZA-TFCA depending on whether
permeability was assessed using data collected during residence or dispersal.
\cite{Jackson.2016} made similar observations for African wild dogs and revealed
higher permeability for dispersers than for residents. Moreover,
\cite{Abrahms.2017} found a significantly greater overlap between predicted and
observed corridors for three dispersing wild dogs when predictions were based on
information collected only during wild dog activity, rather than across the
entire day.

% African wild dogs
A highly endangered flagship species for conservation efforts across the
KAZA-TFCA, but for which systematic GPS relocation data on dispersing
individuals is largely unavailable, is the African wild dog
\citep{VanDerMeer.2016}. While once present across the entire Sub-Saharan
continent, the species has been widely extirpated through human persecution,
habitat destruction, and disease outbreaks \citep{Woodroffe.2012}. As a result,
the African wild dog has become one of Africa's most endangered large carnivores
and currently only survives in small, spatially scattered subpopulations
\citep{Woodroffe.2012}. Within these subpopulations, wild dogs form cooperative
breeding groups of up to thirty individuals in which a single dominant pair
monopolizes the majority of reproduction \citep{Frame.1979, Fuller.1992,
Creel.2002}, while subordinates of both sexes help to feed newborn pups
\citep{Malcolm.1982}. After reaching 15 months of age, male and female offspring
disperse from their natal pack in same-sex coalitions and search for unrelated
mates and a suitable territory to settle (\citeauthor{McNutt.1996}, 1996;
\citeauthor{Behr.2019}, unpublished). During such dispersal events, wild dogs
are able to cover several hundred kilometers, sometimes crossing risky areas
that are strongly influenced by humans \citep{DaviesMostert.2012, Masenga.2016,
Cozzi.2020}. Despite the importance of dispersal for the remaining wild dog
subpopulations, little is known about the corridors of dispersing individuals
and connectivity between subpopulations in the KAZA-TFCA in general.

% Covariate Motivation
Wild dog movements and spatial behavior are known to be influenced by a suite of
environmental and anthropogenic factors. For instance, \cite{Abrahms.2017} found
that wild dog movements during residence mainly depended on land cover,
protection status and anthropogenic features. Similar observations on resident
packs were made by \cite{Pretorius.2019}, who's results indicate that pack
presence was contingent on factors such as agricultural activity, road presence,
and protection status of the area. With respect to dispersing individuals
\cite{Whittington.2011} demonstrated that elevation, land cover, road density,
and human density best explained the distribution of recorded sightings in South
Africa. Also based on data from dispersing wild dogs, \cite{Jackson.2016} showed
that land cover, human density and road density significantly influenced the
probability of occurrence of dispersing wild dogs. Finally, \cite{Masenga.2016}
reported a low willingness of dispersers to cross human-dominated landscapes.
Large scale spatial information on the above mentioned factors can be obtained
either through freely available online services or through remote sensed
satellite imagery.

% Our Study
We used GPS relocation data collected on 16 wild dogs in as many dispersing
coalitions from a free ranging population in northern Botswana to estimate the
degree of selection or avoidance for environmental and anthropogenic landscape
features. We then used the obtained information in combination with relevant
spatial covariates to create a permeability surface for the entire KAZA-TFCA.
Lastly, we used the permeability surface as input for the computation of
least-cost paths and least-cost corridors between selected locations. This
allowed us to highlight key dispersal corridors within and outside the
boundaries of KAZA-TFCA and concluded with the identification of a ``dispersal
hub'' and of dispersal corridors that run outside formally protected areas.

\newpage
\section{Methods}
\subsection{Study Area}
% Both Study Areas
We differentiated between a \textit{core study area}, where all dispersers
originated from, and an \textit{extended study area}, which was represented by a
bounding box encompassing the KAZA-TFCA \citep{Cozzi.2020} and for which we
calculated connectivity.

% Core Study Area
The \textit{core study area} was located in the eastern section of the Okavango
Delta in northern Botswana (centered at -19\degree 25'30''S, 23\degree 37'30''E;
elevation ca. 950 m, orange rectangle in \Cref{StudyArea}b) and comprised parts
of the Moremi Game Reserve and adjacent wildlife management areas. The Okavango
Delta is one of the main strongholds of African wild dogs in southern Africa and
likely acts as a source for the recolonization of surrounding habitats
\citep{Woodroffe.2012, Cozzi.2013}. It is the earth's largest inland delta and
is characterized by diverse habitats, ranging from permanent river systems,
mopane woodland, and acacia forests, to open grasslands and seasonal swamps
\citep{Broekhuis.2013}. Rainfall is seasonal and lasts from November to March,
whereas the remainder of the year remains dry \citep{McNutt.1996}. The extent of
the flood greatly changes within and between years depending on the amount of
rain that descends from the catchment areas in Angola and reaches the distal
ends of the delta between July and August (see \Cref{FloodPulse} in
\Cref{Appendix:FloodmappingAlgorithm} for an illustration of a typical flood
pulse). The flood thereby drastically affects surrounding landscapes, so that
during maximum extent (ca. 12'000 km\textsuperscript{2}) the delta becomes a
patchy conglomerate of swamps, open water, and islands, whereas it shrinks
substantially (ca. 5'000 km\textsuperscript{2}) during dry periods
\citep{Gumbricht.2004, Wolski.2017}. Most of the Okavango Delta's unique basin
is under protection, such that human influence in our core study area was mainly
restricted to the surroundings of Maun and the smaller villages of Khwai,
Mababe, and Sankuyo.

% Large Study Area
The \textit{extended study area} (centered at -17\degree 13'9''S, 23\degree
56'4''E; elevation ca. 1,030 m, red rectangle in \Cref{StudyArea}a) was confined
by a bounding box encompassing the entire KAZA-TFCA. The KAZA-TFCA lies in the
basins of the Okavango and Zambezi rivers and embraces parts of Angola,
Botswana, Namibia, Zimbabwe, and Zambia. With a total area of over 520,000
km\textsuperscript{2} it contains several already-existing national parks (NPs)
and other protected areas. It constitutes the earth's largest transboundary
conservation area, which resulted in a rectangular study area of over 1.3 Mio
km\textsuperscript{2}. Similar to the core study area, the extended study area
hosts various landscapes, including savanna, grassland, and dry or moist
woodland habitats. Despite 36 NPs and other protected areas, there is
considerable human influence in some regions of the KAZA-TFCA originating from
farms, high human density, and road traffic.

\begin{figure}[h]
  \begin{center}
    \begin{tikzpicture}
        \node[anchor=south west,inner sep=0] (image) at (0,0,0) {
        \includegraphics[width=\textwidth]{99_StudyArea.pdf}
        };
        \begin{scope}[x={(image.south east)},y={(image.north west)}]
            % % next four lines will help you to locate the point needed by forming a grid.
            % % comment these four lines in the final picture.
            % \draw[help lines,xstep=.1,ystep=.1] (0,0) grid (1,1);
            % \draw[help lines,xstep=.05,ystep=.05] (0,0) grid (1,1);
            % \foreach \x in {0,1,...,9} { \node [anchor=north] at (\x/10,0) {0.\x}; }
            % \foreach \y in {0,1,...,9} { \node [anchor=east] at (0,\y/10) {0.\y};}
            % % upto here
            \draw[black] (0.202, 0.255) -- (0.36, 0.955);
            \draw[black] (0.202, 0.205) -- (0.36, 0.070);
        \end{scope}
    \end{tikzpicture}
    \caption{Overview of our study area. (a) The red rectangle depicts our
    \textit{extended study area}, which was confined by the extent of the
    KAZA-TFCA. Purple areas indicate remaining wild dog populations according to
    the IUCN \citep{Woodroffe.2012}. (b) The orange rectangle illustrates our
    \textit{core study area}, from which all dispersers departed. Yellow lines
    depict all dispersal trajectories that were recorded between the years 2011
    and 2019. Since game reserves in Botswana virtually serve the same purpose as
    National Parks, we will use the terms interchangeably for this region.}
    \label{StudyArea}
  \end{center}
\end{figure}

\subsection{GPS Relocation Data}
% Collection
We collected GPS relocation data of 16 wild dog coalitions (i.e. 7 female and 9
male coalitions) that dispersed from the core study area between 2011 and 2019.
Potential dispersing individuals were immobilized according to protocols
described in \cite{Osofsky.1996} and fitted with GPS/Satellite radio collars
(\textit{Vertex Lite; Vectronic Aerospace GmbH, Berlin}) while they were still
with their natal pack. All required procedures were undertaken and supervised by
a Botswana-registered wildlife veterinarian. Immobilized individuals were
monitored and all were observed to successfully re-join their packs within one
hour after immobilization. GPS collars were programmed to record GPS relocation
data every 4 hours during dispersal and to regularly transmit them via Iridium
satellite system to a base station. The system thus allowed remote-tracking of
collared individuals, even where field conditions were prohibitive or dispersal
trajectories crossed international borders.

% Separation of Dispersers/Residents
Because we were mainly interested in recording dispersal trajectories, we
discarded any GPS data collected while individuals were still with their natal
packs \citep{Cozzi.2020}. Dispersal was deemed to have started when a coalition
had left its natal home range, and continued until it exhibited clear signs of
resident behavior. Some coalitions undertook exploratory movements prior to
proper dispersal. Such movements have been observed in other species and are
reported to be very similar to proper dispersal \citep{Killeen.2014}. We
therefore classified exploratory movements as dispersal too.

% Some Summary Statistics
In total, we collected 4'691 GPS relocations during dispersal (see
\Cref{Trajectories}). In our analysis, we did not differentiate between male and
female dispersing coalitions, for previous research found little differences
between sexes during dispersal \citep{Woodroffe.2019, Cozzi.2020}. On average,
dispersal coalitions included three individuals (min = 1, max = 7), dispersed
for 41 days (min = 7, max = 133), and covered a mean cumulative distance of 370
km (min = 105, max = 610).

\begin{figure}[h]
  \begin{center}
    \includegraphics[width = \textwidth]{99_Trajectories.pdf}
    \caption{Illustration of the dispersal trajectories that we recorded. Each
    color represents a different dispersal coalition. All coalitions departed
    from our core study area, which is indicated by the white square. The
    coalition dispersing towards the far east of the map covered over 360 km in
    under 10 days.}
    \label{Trajectories}
  \end{center}
\end{figure}

\subsection{Spatial Covariates}
% General stuff
We used a set of geo-referenced covariates to investigate habitat preferences of
dispersing wild dogs (see \Cref{Covariates}). Covariates fell into one of the
three categories \textit{land cover}, \textit{protection status}, and
\textit{human influence}. Note that we did not include elevation as covariate
due to the absence of noteworthy elevational gradients in our study area. For
each covariate we obtained spatial raster layers from freely available online
services or from remote sensed satellite imagery. To ensure a consistent
resolution (i.e. cell-size or grain) across covariates, all layers were
coarsened or interpolated to match a resolution of 250m x 250m. Processing and
manipulation of data, as well as all spatial and statistical analyses were
performed in R, version 3.6.1 \citep{R.2019}.

\begin{figure}[h]
  \begin{center}
    \includegraphics[width = \textwidth]{99_Covariates.pdf}
    \caption{Overview of covariates that we used in our models. Although all
    covariates were prepared for the extended study area, we present only an
    extract to maximize visibility. The white square in each plot depicts our
    core study area. (a) Averaged layer of all binary maps of water and dryland
    that we acquired using our floodmapping algorithm. (b) Percentage cover of
    trees. (c) Percentage cover of shrubs/grassland. Note that anything not
    covered by trees or shrubs/grassland was deemed to be bareland. (d)
    Protection status according to the Peace Parks Foundation. (e) Human
    influence proxy, including human density, farms, and roads. (f) Distance to
    nearest road (white lines depict actual roads).}
    \label{Covariates}
  \end{center}
\end{figure}

\newpage
\subsubsection{Land Cover}
% Overview
Land cover was represented by four layers: a binary layer distinguishing water
and dryland, and three continuous layers showing the percentage cover by
different vegetation types, together adding up to 100\%.

% Water Layers
Water included open waters, rivers, wetlands, and swamps, whereas anything else
was presumed to be dryland. Because the flood extent of the Okavango Delta is
highly variable within and between years, we created weekly (8 days)
``floodmaps'', following the algorithm developed in \cite{Wolski.2017}. The
algorithm uses thresholding of SWIR reflectances of MODIS Terra satellite
imagery (MCD43A4, \cite{Schaaf.2015}) to separate water and dryland. As we
rewrote and slightly amended the original Python code in R, we describe the
underlying principles and indicate the discrepancies to the original algorithm
in \Cref{Appendix:FloodmappingAlgorithm}. While water in the Okavango Delta was
mapped dynamically, we used a static watermap for its surroundings. This static
map was based on Globeland's land cover dataset \citep{Chen.2015}, from which we
only retained the categories \textit{wetland} and \textit{waterbodies} and
collectively reclassified them to \textit{water}. Globeland had an original
resolution of 30m x 30m, so we coarsened the layer to 250m x 250m using the mode
of each 250m x 250m cell. Lastly, we improved river representation using the
MERIT Hydro dataset \citep{Yamazaki.2019}, from which we added all rivers with a
width of over 10m to our watermaps.

% Vegetation Layers
The three vegetation layers that we used in addition to the water/dryland maps
were derived from the MODIS Terra Vegetation Continuous Fields dataset (MOD44B,
\cite{Dimiceli.2015}). Combined, the three layers added up to 100\% and depicted
the percentage cover of tree-vegetation (henceforth trees), non-tree-vegetation
(henceforth shrubs/grassland), and non-vegetated (henceforth bareland). To
ensure that areas covered by the flood would get 0\% vegetation cover (i.e.
100\% bareland cover), we used our floodmap that aligned with the creation date
of the layers (30 March 2017) and manually defined pixels covered by the flood
as being 100\% bareland. The MODIS vegetation layers had a resolution of 250m x
250m, so we did not need to coarsen or interpolate them.

\subsubsection{Protection Status}
To assess whether dispersers preferred protected over unprotected (pastoral)
regions, we derived a binary layer separating the two categories. Corresponding
data on protection status was downloaded in shapefile format from the Peace
Parks Foundation \citep{PeaceParks.2019}. We simplified original categories to a
binary classification of protected and pastoral areas (see
\Cref{Reclassification} in \Cref{Appendix:Reclassification}). Protected areas
mainly included forest reserves, game reserves, wildlife management areas, and
NPs, whereas everything not covered by such categories was considered to be
pastoral area. Finally, we rasterized the two categories onto a binary raster (1
= protected, 0 = pastoral) with a resolution of 250m x 250m.

\subsubsection{Human Influence}
We defined human influence using a proxy that incorporated anthropogenic
pressures arising from (1) human density, (2) farming activities, and (3) roads.
Previous research revealed that veterinary fences have no influence on wild dog
movements \citep{Cozzi.2013b}, so we did not include them in our analysis.

% Human Density
(1) We obtained spatial human density estimates through Facebook's 30m x 30m
high resolution population density dataset \citep{Facebook.2019}. The dataset
contained some human density that arose from lodges and other touristic
activities that we did not want to depict as human pressures, so we manually
removed such regions from the data. The resulting layer was coarsened to 250m x
250m by summing up human densities within 250m x 250m cells.

% Farming
(2) Information on farming activities was sourced from the Globeland
\citep{Chen.2015} and Cropland \citep{Xiong.2017} land cover datasets, from
which we only retained pixels that were classified as either \textit{cultivated
land} or \textit{croplands}. Both layers had a resolution of 30m x 30m and were
coarsened to 250m x 250m by summing up values within 250m x 250m cells. The
layers were then merged into a single binary layer depicting presence (= 1) or
absence (= 0) of farms within 250m x 250m cells.

% Roads
(3) Geo-referenced data on roads was downloaded in shapefile format from Open
Street Map \citep{OpenStreetMap.2019}. We omitted minor roads and only retained
main tar roads and trunks belonging to the biggest four categories (see
\Cref{RoadsDescription} in \Cref{Appendix:RoadCategories}). Finally, we
rasterized roads and trunks onto a binary raster (1 = roads, 0 = no roads) with
250m x 250m resolution.

To calculate a single proxy of human influence, we added up the layers
describing \textit{human density} (continuous), \textit{farming} (binary), and
\textit{roads} (binary). To reduce the influence of outliers, totaled values
were limited to a maximum of 50, which visually resulted in a good balance
between high and low anthropogenic influence and was therefore considered
appropriate for our analysis. To render the fact that humans influence their
surroundings beyond their presence, we followed \cite{Elliot.2014} and applied
to each pixel a 5 km spatial buffer within which we summed up and log-transformed
human-influence values.

\subsection{Habitat Selection Model}
\label{Modeling}
% Explanation of iSSF and Extraction
We used an integrated step selection function approach (iSSF, \cite{Avgar.2016})
to investigate dispersers' selection or avoidance for the above mentioned
spatial covariates. In the iSSF framework, covariates along realized steps are
contrasted with covariates along alternative random steps that the animal could
have taken, but decided not to. A step is here defined as the connecting line
between two consecutive GPS relocations \citep{Turchin.1998}. In contrast to
regular SSFs, iSSFs require to include movement metrics as covariates in the
corresponding conditional logistic regression model. Their inclusion in turn
allows simultaneous inference on habitat and movement preferences, as well as to
reduce potential biases in estimated habitat preferences \citep{Forester.2009,
Warton.2013, Avgar.2016}.

In order to conduct an iSSF analysis, we followed recommendations described in
Appendix S1 in \cite{Avgar.2016}. We prepared our GPS relocation data for
iSSF-analysis using the R-package \textit{amt} \citep{Amt.2019} and coerced
relocations recorded during dispersal to steps that were regularly spaced four
hours apart (allowing for a minor mismatch of up to 15 minutes). Steps that were
separated by more than four hours (e.g. due to GPS failure) were omitted from
further analysis. Each remaining step was paired with 24 random steps, generated
by sampling turning angles from a uniform distribution U(\(-\pi, \pi\)) and step
lengths from a gamma distribution that was fitted using realized step lengths
(shape = 0.3677, scale = 6,302)). Together, a realized and its 24 associated
random steps formed a stratum and received a unique identifier.

We extracted spatial covariates along realized and random steps as listed in
\Cref{ExtractedCovars}. For continuous covariates we calculated the average
value along the step, for binary covariates the percentage cover along the step.
Since our floodmaps were dynamic, we extracted values from the floodmap that was
closest in date to the actual step. We further derived a binary variable
indicating whether or not a step crossed a road and we identified the average
distance of each step to the nearest source of water or to the nearest road. We
square-rooted the values for \textit{DistanceToRoads} and
\textit{DistanceToWater} to render a decreasing marginal impact of distance. We
standardized extracted covariates to a mean of zero and a standard deviation of
one using a z-score transformation. Furthermore, we screened covariates for
correlation using Pearson's Correlation Coefficient. None of the covariates were
overly correlated (\(|r| > 0.6\), \cite{Latham.2011}) and we retained all of
them for modeling. To run a proper iSSF analysis, we also included two movement
metrics, namely the cosine of the turning angle (\(cos(ta)\)) and the logarithm
of the step length (\(log(sl)\)), as covariates in our models
\citep{Avgar.2016}. The movement metric \(cos(ta\)) served to describe the
directionality of a step, as it transformed the circular measure of (\(-\pi,
+\pi\)) into a linear measure (-1, 1), where positive values indicated forward
movements, whereas negative values indicated backwards movements
\citep{Turchin.1998}. The movement metric \(log(sl\)), on the other hand, was
included as an indicator of the preferred step length. Since steps were
regularly spaced by four hours, \(log(sl\)) can also be interpreted as movement
rate.

\begin{table}[h]
  \begin{center}
    \caption{Overview of spatial covariates extracted along steps (see
    \Cref{Sources} in \Cref{Appendix:Sources} for an overview of the sources and
    resampling methods). *The covariates \textit{Water} and \textit{Dryland}
    added up to 100\%, which is why only \textit{Water} was included as
    explanatory variable in our models. The same applied for the group
    \textit{Shrubs/Grassland}, \textit{Trees}, and \textit{Bareland}, where we
    omitted \textit{Bareland} for modeling. Finally, from the group
    \textit{Protected} and \textit{Pastoral}, we only included
    \textit{Protected} in our models.}
    \label{ExtractedCovars}
    \resizebox{\textwidth}{!} {
      \begin{tabular}{lllll}
      \hline
      \textbf{Category} &
        \textbf{Covariate} &
          \textbf{Description} &
            \textbf{Values} \\
              % \textbf{Source(s)} \\
      \midrule
      \multirow{5}{*}{Land Cover}
        & Water
          & Percentage cover of water along step
            & 0-100\%\\
              % & \cite{Chen.2015}, \cite{Schaaf.2015}, \cite{Yamazaki.2019} \\
        & Dryland*
          & Percentage cover of dryland along step
            & 0-100\%\\
              % & Globeland, \cite{Schaaf.2015}, \cite{Yamazaki.2019} \\
        & DistanceToWater
          & Average distance of step to nearest water source
            & \(\geq\) 0m\\
              % & Globeland, \cite{Schaaf.2015}, \cite{Yamazaki.2019} \\
        & Shrubs/Grassland
          & Average non-tree vegetation along step
            & 0-100\%\\
              % & \cite{Dimiceli.2015} \\
        & Trees
          & Average tree-vegetation along step
            & 0-100\%\\
              % & \cite{Dimiceli.2015} \\
        & Bareland*
          & Average non-vegetated area along step
            & 0-100\%\\
              % & \cite{Dimiceli.2015} \\
      \hdashline
      \multirow{2}{*}{Protection Status}
        & Protected
          & Percentage cover of protected area along step
            & 0-100\%\\
              % & \cite{PeaceParks.2019} \\
        & Pastoral*
          & Percentage cover of unprotected area along step
            & 0-100\%\\
              % & \cite{PeaceParks.2019} \\
      \hdashline
      \multirow{3}{*}{Human Influence}
        & HumansBuffer5000
          & Average buffered (5 km) human influence along step
            & \(\geq\) 0\\
              % & \cite{Facebook.2019}, \cite{OpenStreetMap.2019}, \cite{Chen.2015}, \cite{Xiong.2017}\\
        & DistanceToRoads
          & Average distance of step to nearest road
            & \(\geq\) 0m\\
              % & \cite{Facebook.2019}, \cite{OpenStreetMap.2019}, \cite{Chen.2015}, \cite{Xiong.2017}\\
        & RoadCrossing
          & Binary value indicating whether a step crossed a road
            & 0, 1\\
              % & \cite{Facebook.2019}, \cite{OpenStreetMap.2019}, \cite{Chen.2015}, \cite{Xiong.2017}\\
      \hline
      \end{tabular}
    }
  \end{center}
\end{table}

% Permeability Model
\noindent We used the iSSF framework and parameterized a habitat selection model
that further served to predict landscape permeability. This selection model
operated under the assumption that dispersing wild dogs assigned a selection
score \(w(x)\) of the following exponential form to each realized and random
step:

\begin{equation}
\label{EQ2}
  w(x) = exp(\beta_1 x_1 + \beta_2 x_2 + ... + \beta_n x_n)
\end{equation}

\noindent That is, the selection score \(w(x)\) of a step depended on its
associated covariates (\(x_1, x_2, ..., x_n\)), as well as on the animal's
preferences for these covariates (\(\beta_1, \beta_2, ..., \beta_n\)). The
probability that a step \(i\) was realized \(P(Y_{i} = 1\)) was then contingent
on the step's selection score, as well as on the selection scores of all
alternative steps in the stratum:

\begin{equation}
\label{EQ3}
  P(Y_{i} = 1 | Y_{1} + Y_{2} + ... + Y_{i} = 1) =
  \frac{w(x_{i})}{w(x_{1}) + w(x_{2}) + ... + w(x_{i})}
\end{equation}

% Conditional Logistic Regression Model
\noindent Habitat and movement preferences of interest, i.e. the \(\beta\)'s,
were estimated by comparing realized (scored 1) and random (scored 0) steps in a
conditional logistic regression model \citep{Fortin.2005}. In this model,
positive \(\beta\)-coefficients indicate selection of a covariate, negative
\(\beta\)-coefficients avoidance of a covariate. Selected covariates therefore
contribute positively to expected landscape permeability, whereas avoided
covariates contribute negatively to landscape permeability.

% Mixed Effects Conditional Logistic Regression Model
To deal with multiple individuals, we applied mixed effects conditional logistic
regression analysis following \cite{Muff.2019}. Their approach allowed to
include individuals for whom some covariates did not vary, as well as to model
random intercepts and slopes for each covariate. We implemented their method
using the R-package \textit{glmmTMB} \citep{Mollie.2017}, where we fixed the
variance of the random intercept to an artificially high value of
10\textsuperscript{12} (for an explanation, see \cite{Muff.2019}) and used
animal ID to model random intercepts and slopes.

% AIC Model Selection
We defined the movement metrics \(cos(ta)\) and \(log(sl)\) as core covariates
and ran forward model selection based on Akaike's Information Criterion (AIC,
\cite{Burnham.2002}) for all other spatial covariates. That is, we started with
a basic model and iteratively increased model complexity upon reaching a
full model. We then ranked models according to AIC, assessed relative model
weights, and identified the most parsimonious model. Due to convergence issues
we were unable to model interactions between spatial covariates.

% Validation
To validate the predictive power of the most parsimonious selection model, we
ran k-fold cross-validation for case-control studies as described in
\cite{Fortin.2009}. Using 80\% of randomly selected strata, we parameterized a
selection model and predicted selection scores \(w(x)\) for all steps in the
remaining 20\% of strata. According to predicted selection scores we assigned
ranks 1-25 within each stratum, with rank 1 indicating the highest selection
score. We identified the realized step's rank in each stratum and tallied rank
frequencies for realized steps over all strata. Finally, we carried out a
Spearman-rank correlation analysis between ranks and associated frequencies and
we recorded the correlation coefficient (\(r_{s, realized}\)). We repeated this
procedure 100 times with replacement and computed the mean correlation
coefficient (\(\bar{r}_{s, realized}\)), as well as its 95\% confidence
interval. For comparison, we also repeated the same procedure 100 times assuming
completely randomized preferences. We implemented randomized preferences by
omitting the realized step from each stratum and identifying the rank of a
randomly chosen random step within each stratum (now only ranks 1-24). Again, we
calculated Spearman's rank correlation coefficient (\(r_{s, random}\)), its mean
across repetitions (\(\bar{r}_{s, random}\)), and its 95\% confidence interval.
Ultimately, the validation proofed a significant prediction in case the
confidence intervals of \(\bar{r}_{s, realized}\) and \(\bar{r}_{s, random}\)
did not overlap.

\subsection{Permeability Surface}
We used the most parsimonious selection model and applied \Cref{EQ2} to predict
a permeability surface over the entire extent of the KAZA-TFCA. In other words,
we calculated the selection score \(w(x)\) for each 250m x 250m pixel covering
our extended study area given the pixel's spatial covariates. Note that the
movement metrics, i.e. \(cos(ta)\) and \(log(sl)\), did not enter the prediction
of landscape permeability as their inclusion in our habitat selection model
solely served to reduce the bias in estimated habitat preferences. By definition
of permeability, high selection scores resulted in high permeability, whereas
low selection scores resulted in low permeability. To reduce the influence of
outliers in predicted scores, we followed \cite{Squires.2013} and curtailed
predicted scores between the 1\textsuperscript{st} and 99\textsuperscript{th}
percentile of their original values. To predict permeability in the first place,
we needed to collapse all dynamic watermaps into a single layer, which we
achieved by identifying areas that were inundated in at least a quarter of all
our floodmaps. In order to be able to compare permeability across different
regions, we rescaled the permeability surface to a range between 0 and 1 using
\(f(x) = \frac{x - min(x)}{max(x) - min(x)}\). The most impermeable landscape
therefore received a value of 0, whereas the most permeable landscape received a
value of 1.

\subsection{Least-Cost Paths \& Least-Cost Corridors}
To identify movement corridors within the extent of KAZA-TFCA, we calculated
least-cost paths (LCPs) and least-cost corridors (LCCs) between selected points
located in protected areas of our extended study area \citep{Adriaensen.2003,
Sawyer.2011}. Selected points were generated by overlaying the study area with a
regular grid of points, equally spaced 100 km apart and located within protected
areas that contiguously covered at least 700 km\textsuperscript{2}. The
threshold of 700 km\textsuperscript{2} was based on home-range requirements of
African wild dogs reported in \cite{Pomilia.2015} and served to eliminate
locations that were assumed unsuitable for supporting viable wild dog
populations. Finally, we defined centroids as source points for those protected
areas larger than 700 km\textsuperscript{2} that did not get any source points
from the regular grid. In total, we generated 68 points, which resulted in 2,278
unique pairwise combinations and therefore 2,278 unique LCPs and LCCs.

LCP analysis between source points was implemented using the R-package
\textit{gdistance} \citep{vanEtten.2018}. The package translated the (unscaled)
permeability surface into a network of nodes and used graph theory, in
particular Dijkstra's algorithm \citep{Dijkstra.1959}, to find shortest
effective distances between source points based on costs of moving from cell to
cell. The package thus required to convert the permeability surface to a
transition layer, where permeability values were translated into transition
probabilities for moving from one cell to another. In our case, the transition
probability of moving between two adjacent cells depended on their averaged
permeability. We allowed individuals to move from each cell to the cell's 8
surrounding neighbors (i.e. Moores neighborhood) and applied a geographic
correction to account for the fact that diagonal neighbors were more remote than
orthogonal neighbors. Because African wild dogs have been observed to cover
massive dispersal distances \citep{DaviesMostert.2012, Masenga.2016,
Cozzi.2020}, we did not limit LCPs to a maximal effective cost. After
computation, we tallied overlapping LCPs to identify high-frequency routes.

LCPs are often criticized because they identify single-pixel-wide paths and thus
obliterate potential alternative routes with only marginally higher costs
\citep{Pinto.2009}. Moreover, they implicitly assume that dispersers have
complete knowledge of the landscape and associated movement costs and therefore
always move along the least-costly route \citep{Carroll.2012}. Yet, individuals
rarely use a single optimum path, so that LCPs are often of limited biological
relevance \citep{Pinto.2009, Pullinger.2010}. To address these shortcomings, we
also calculated LCCs \citep{Pinto.2009, Sawyer.2011}, again using the R-package
\textit{gdistance}. In contrast to LCPs, LCCs overcome the
single-pixel-width-issue and reveal additional, slightly suboptimal dispersal
routes. To identify LCCs, we first computed for each source point a cumulative
cost map (see \Cref{LCCExample}), which indicated the total minimal costs
required to get from the selected point to any other location on the map. A LCC
between two points was then obtained by adding up their cumulative cost maps and
masking out all cell-values exceeding the lowest cell-value by more than 5\%. We
repeated this procedure for each possible pairwise combination of source points
and thereby identified LCCs between all 68 selected points. The resulting
corridor-maps were normalized to a range from zero to one using \(f(x) = \frac{x
- min(x)}{max(x) - min(x)}\) and totaled up into a single connectivity map.

% \begin{figure}[hbtp]
%   \begin{center}
%     \begin{tikzpicture}
%         \node[anchor=south west,inner sep=0] (image) at (0,0,0) {
%           \begin{minipage}{0.54\textwidth}
%             \begin{center}
%               \includegraphics[width = 0.5\textwidth]
%                 {99_LeastCostPaths(Example1).pdf}\\
%               \includegraphics[width = 1.0\textwidth]
%                 {99_LeastCostPaths(Example2).pdf}\\
%             \end{center}
%           \end{minipage}
%         };
%         \begin{scope}[x={(image.south east)},y={(image.north west)}]
%            %  % next four lines will help you to locate the point needed by
%            %  % forming a grid. comment these four lines in the final picture.↓
%            % \draw[help lines,xstep=.1,ystep=.1] (0,0) grid (1,1);
%            % \draw[help lines,xstep=.05,ystep=.05] (0,0) grid (1,1);
%            % \foreach \x in {0,1,...,9} { \node [anchor=north] at (\x/10,0) {0.\x}; }
%            % \foreach \y in {0,1,...,9} { \node [anchor=east] at (0,\y/10) {0.\y};}
%            %  % upto here
%             % \draw[black, line width = 0.5pt] (0.389, 0.627) -- (0.080, 0.477);
%             % \draw[black, line width = 0.5pt] (0.582, 0.627) -- (0.980, 0.477);
%             % \draw[white, line width = 1pt] (0.675, 0.290) -- (0.720, 0.265);
%             % \draw[white, line width = 1pt] (0.840, 0.255) -- (0.800, 0.260);
%             % \draw[white, line width = 1pt] (0.807, 0.205) -- (0.752, 0.205);
%         \end{scope}
%     \end{tikzpicture}
%     \caption{Figures illustrating the process of identifying a least-cost
%     corridor between the source points A and B. (a) Example of a permeability
%     surface, which serves to determine the costs of movement. (b1) Cumulative
%     cost map for point A, depicting the total minimal costs necessary to get
%     from point A to every other location. (b2) Cumulative cost map for point B,
%     depicting the total minimal costs necessary to get from point B to every
%     other location. (c) Summed cost maps of points A and B. (d) Masked out
%     corridor containing pixels that do not exceed the cheapest pixel by more
%     than 5\%.}
%     \label{LCCExample}
%   \end{center}
% \end{figure}

\begin{figure}[hbtp]
  \begin{center}
  \begin{minipage}{.25\linewidth}
    \begin{subfigure}[t]{.9\linewidth}
        \includegraphics[width=\textwidth]{99_LeastCostPaths(Example1).pdf}
    \end{subfigure}
  \end{minipage}
  \begin{minipage}{.25\linewidth}
    \begin{subfigure}[t]{.9\linewidth}
        \includegraphics[width=\textwidth]{99_LeastCostPaths(Example2).pdf}
    \end{subfigure}
  \end{minipage}
  \begin{minipage}{.25\linewidth}
    \begin{subfigure}[t]{.9\linewidth}
        \includegraphics[width=\textwidth]{99_LeastCostPaths(Example3).pdf}
    \end{subfigure}
  \end{minipage}
  \caption{Figures illustrating the process of identifying a least-cost corridor
  between the source points A and B following \citep{Pinto.2009}. (a) Example of
  a permeability surface, which serves to determine the costs of movement. (b1)
  Cumulative cost map for point A, depicting the total minimal costs necessary
  to get from point A to every other location. (b2) Cumulative cost map for
  point B, depicting the total minimal costs necessary to get from point B to
  every other location. (c) Summed cost maps of points A and B. (d) Masked out
  corridor containing pixels that do not exceed the cheapest pixel by more than
  5\%.}
  \label{LCCExample}
  \end{center}
\end{figure}

\newpage
\section{Results}
\subsection{Habitat Selection Model}
Our most parsimonious habitat selection model (\(\Delta AIC > 2\) than any
alternative model; see \Cref{ModelAICs} in \Cref{AICs}) retained the covariates
\textit{Water}, \textit{DistanceToWater}, \textit{Trees},
\textit{Shrubs/Grassland}, \textit{HumansBuff5000}, and the fixed covariates
\(cos(ta)\) and \(log(sl)\) (\Cref{PermeabilityResults}a). Parameter estimates
showed that dispersing wild dogs moved in a directional and fast manner, as
indicated by a positive selection for small turning angles, i.e. high
\(cos(ta)\) (\(\beta\) = 0.14; 95\% CI = 0.07 to 0.21,
\Cref{PermeabilityResults}a) and longer steps, i.e. high \(log(sl)\) (\(\beta\)
= 0.05, 95\% CI = 0.02 to 0.09, \Cref{PermeabilityResults}a). Given the chance,
dispersers avoided moving through water (\(\beta\) = -0.5, 95\% CI -0.67 to
-0.32, \Cref{PermeabilityResults}a) but selected for locations in its vicinity,
although the effect was not significant (\(\beta\) = -0.24, 95\% CI = -0.67 to
0.19, \Cref{PermeabilityResults}a). Dispersers avoided areas that were densely
covered by trees (\(\beta\) = -0.18, CI = -0.32 to -0.04,
\Cref{PermeabilityResults}a) and preferred areas covered by shrubs/grassland
(\(\beta\) = 0.40, 95\% CI = 0.24 to 0.56, \Cref{PermeabilityResults}a).
Finally, dispersers avoided areas that were influenced by human presence and
activities (\(\beta\) = -0.47, 95\% CI = -0.87 to -0.07,
\Cref{PermeabilityResults}a).

\begin{figure}[h]
  \begin{center}
    \includegraphics[width = \textwidth]{99_PermeabilityResults.pdf}
    \caption{(a) Estimated selection coefficients from the most parsimonious
    permeability model. Negative coefficients indicate avoidance of a covariate,
    positive coefficients selection of a covariate. Whiskers delineate the
    95\%-CIs for estimated parameters. The intercept is meaningless in the iSSF
    framework and was thus omitted. (b) Results from the k-fold cross validation
    for case control studies. The left graph shows rank frequencies of
    \textit{realized} steps according to predictions, whereas the right graph
    shows rank frequencies of \textit{randomly selected} steps according to
    predictions. \(\bar{r}_s\) indicates the mean correlation coefficient
    resulting from 100 repetitions of the k-fold cross validation. The blue
    smoothing line was fitted using a LOESS regression and solely serves to aid
    the eye in detecting the trends. Correlation coefficients suggest that our
    prediction is significant and robust, evidenced by the fact that the
    confidence intervals of \(\bar{r}_{s, realized}\) and \(\bar{r}_{s,
    random}\) do not overlap and that there is strong and significant
    correlation between ranks and associated frequency for realized steps.}
    \label{PermeabilityResults}
  \end{center}
\end{figure}

\newpage
\noindent Results from the k-fold cross-validation suggested that our prediction
was significant and robust, as highlighted by the fact that the 95\%-CIs
intervals of \(\bar{r}_{s, realized}\) (-0.45, CI = -0.48 to -0.42) and
\(\bar{r}_{s, random}\) (0.04, CI = 0.01 to 0.08) did not overlap
(\Cref{PermeabilityResults}b). Likewise, the significant correlation between
ranks and corresponding frequencies for realized steps suggested a good fit
between predictions and observations.

\subsection{Permeability Surface}
Our prediction revealed substantial differences in predicted permeability across
the entire study area (\Cref{PermeabilityMap}). Comparisons of median (Mdn)
permeability values (\Cref{PermeabilityComp}) showed that permeability inside
the boundaries of the KAZA-TFCA (Mdn = 0.21) was three times as high as
permeability outside it (Mdn = 0.07). A comparison across countries
(\Cref{PermeabilityComp}) showed that Angola (Mdn = 0.29) and Botswana (Mdn =
0.25) were characterized by comparably highly permeable landscapes, whereas
Zimbabwe (Mdn = 0.05) and Zambia (Mdn = 0.04) were relatively impermeable.
Namibia (Mdn = 0.16) ranged somewhere in between the two extremes. Across all
countries, protected areas provided higher permeability (Mdn = 0.21) than
human-dominated landscapes (Mdn = 0.07). A visual assessment of our permeability
surface revealed that northern Botswana had the highest density of permeable
landscapes (\Cref{PermeabilityMap}, bright yellow colors), despite the waters
of the Okavango Delta and Linyanti Swamp themselves provided highly impermeable
landscapes (\Cref{PermeabilityMap}, dark blue colors).

% \newgeometry{left=30mm,right=30mm,top=0mm,bottom=20mm}
% \begin{figure}[hbtp]
%   \begin{center}
%     \begin{tikzpicture}
%         \node[anchor=south west,inner sep=0] (image) at (0,0,0) {
%           \begin{minipage}{0.95\textwidth}
%             \includegraphics[width = \textwidth]{99_PermeabilityMap.pdf}
%             \includegraphics[width = \textwidth]{99_PermeabilityMap2.pdf}
%           \end{minipage}
%         };
%         \begin{scope}[x={(image.south east)},y={(image.north west)}]
%            %  % next four lines will help you to locate the point needed by
%            %  % forming a grid. comment these four lines in the final picture.↓
%            % \draw[help lines,xstep=.1,ystep=.1] (0,0) grid (1,1);
%            % \draw[help lines,xstep=.05,ystep=.05] (0,0) grid (1,1);
%            % \foreach \x in {0,1,...,9} { \node [anchor=north] at (\x/10,0) {0.\x}; }
%            % \foreach \y in {0,1,...,9} { \node [anchor=east] at (0,\y/10) {0.\y};}
%            %  % upto here
%             \draw[black, line width = 1pt] (0.389, 0.627) -- (0.080, 0.477);
%             \draw[black, line width = 1pt] (0.582, 0.627) -- (0.980, 0.477);
%             \draw[white, line width = 1pt] (0.675, 0.290) -- (0.720, 0.265);
%             \draw[white, line width = 1pt] (0.840, 0.255) -- (0.800, 0.260);
%             \draw[white, line width = 1pt] (0.807, 0.205) -- (0.752, 0.205);
%         \end{scope}
%     \end{tikzpicture}
%     \caption{(a) Predicted permeability surface for the extent of the KAZA-TFCA.
%     Permability was predicted by calculating selection scores \(w(x)\) for each
%     pixel based on the pixel's underlying covariates. Areas that dispersers find
%     easy to traverse are depicted in bright colors. Bold white lines delineate
%     the borders of the KAZA-TFCA, whereas dashed white lines show country
%     borders. (b) Detailed view of the Okavango Delta and Linyanti Swamp region,
%     which appears to provide the most suitable landscape for dispersers. The
%     white square highlights our core study area.}
%     \label{PermeabilityMap}
%   \end{center}
% \end{figure}
% \restoregeometry

% \newgeometry{left=30mm,right=30mm,top=0mm,bottom=20mm}
\begin{figure}[hbtp]
  \begin{center}
    \begin{tikzpicture}
        \node[anchor=south west,inner sep=0] (image) at (0,0,0) {
          \begin{minipage}{0.7\textwidth}
            \includegraphics[width = \textwidth]{99_PermeabilityMap.pdf}
          \end{minipage}
        };
        \begin{scope}[x={(image.south east)},y={(image.north west)}]
           %  % next four lines will help you to locate the point needed by
           %  % forming a grid. comment these four lines in the final picture.↓
           % \draw[help lines,xstep=.1,ystep=.1] (0,0) grid (1,1);
           % \draw[help lines,xstep=.05,ystep=.05] (0,0) grid (1,1);
           % \foreach \x in {0,1,...,9} { \node [anchor=north] at (\x/10,0) {0.\x}; }
           % \foreach \y in {0,1,...,9} { \node [anchor=east] at (0,\y/10) {0.\y};}
           %  % upto here
          % \node[text=white, circle, scale=1, font=\bfseries, inner sep = 0.5pt
          %   , draw=white] (No1) at (0.440, 0.350){1};
          % \node[text=white, circle, scale=1, font=\bfseries, inner sep = 0.5pt
          %   , draw=white] (No2) at (0.520, 0.438){2};
        \end{scope}
    \end{tikzpicture}
    \caption{Predicted permeability surface for the extent of the KAZA-TFCA.
    Permability was predicted by calculating selection scores \(w(x)\) for each
    pixel based on the pixel's underlying covariates. Areas that dispersers find
    easy to traverse are depicted in bright colors. Bold white lines delineate
    the borders of the KAZA-TFCA, whereas dashed white lines show country
    borders.}
    \label{PermeabilityMap}
  \end{center}
\end{figure}
% \restoregeometry

% \begin{table}[h]
%   \caption{Comparison of median permeability across countries, separated
%   according to areas within and outside the KAZA-TFCA, as well as within and
%   outside protected areas. High values indicate high permeability, whereas low
%   values correspond to low permeability.}
%   \label{PermeabilityComp}
%   \begin{center}
%     \resizebox{\textwidth}{!} {
%       \begin{tabular}{llllllllr}
%       & & \multicolumn{2}{c}{KAZA-TFCA} && \multicolumn{2}{c}{Protection Status} \\
%       \cline{3-4} \cline{6-7}
%       Country & & Inside & Outside & & Protected & Pastoral & & Overall \\
%       \hline
%       Angola & & 1.67 & 0.59 & & 1.67 & 0.60 & & 1.03 \\
%       Namibia & & 0.91 & 0.49 & & 0.98 & 0.44 & & 0.59 \\
%       Botswana & & 1.17 & 0.69 & & 1.37 & 0.62 & & 0.91 \\
%       Zimbabwe & & 0.25 & 0.18 & & 0.32 & 0.17 & & 0.19 \\
%       Zambia & & 0.26 & 0.12 & & 0.23 & 0.13 & & 0.15 \\
%       \hline
%       Overall & & 0.75 & 0.26 & & 0.74 & 0.27 & & 0.38 \\
%       \end{tabular}
%     }
%   \end{center}
% \end{table}

\begin{table}[h]
  \caption{Comparison of median permeability across countries, separated
  according to areas within and outside the KAZA-TFCA, as well as within and
  outside protected areas. High values indicate high permeability, whereas low
  values correspond to low permeability.}
  \label{PermeabilityComp}
  \begin{center}
    \resizebox{\textwidth}{!} {
      \begin{tabular}{llllllllr}
      & & \multicolumn{2}{c}{KAZA-TFCA} && \multicolumn{2}{c}{Protection Status} \\
      \cline{3-4} \cline{6-7}
      Country & & Inside & Outside & & Protected & Pastoral & & Overall \\
      \hline
      Angola & & 0.46 & 0.18 & & 0.46 & 0.18 & & 0.29 \\
      Namibia & & 0.26 & 0.13 & & 0.27 & 0.12 & & 0.16 \\
      Botswana & & 0.33 & 0.19 & & 0.38 & 0.17 & & 0.25 \\
      Zimbabwe & & 0.06 & 0.05 & & 0.09 & 0.04 & & 0.05 \\
      Zambia & & 0.07 & 0.03 & & 0.06 & 0.03 & & 0.04 \\
      \hline
      Overall & & 0.21 & 0.07 & & 0.21 & 0.07 & & 0.11 \\
      \end{tabular}
    }
  \end{center}
\end{table}

\subsection{Least-Cost Paths \& Least-Cost Corridors}
Because LCPs identify the cheapest set of pixels between two selected points,
they typically run along the center of LCCs and traverse the same regions. The
two methods thus essentially return very similar information, but we show both
for the sake of clarity and comparison and because LCPs help more precisely
identifying discrete corridors. Our analysis revealed three major corridors
(\Cref{LeastCost}): one runs SE-NW and connects the Okavango-Linyanti ecosystem
in Botswana with Luengue-Luiana NP in Angola. One runs W-E between Chobe NP in
Botswana and Zimbabwe’s Hwange NP. One runs NE-SW, completely across unprotected
areas, and connects Kafue NP in Zambia with more central regions of the
KAZA-TFCA. With 637 LCPs passing through it, the Okavango-Linyanti region is the
highest frequented corridor. Several minor routes branch off these three major
corridors; this includes a southward connection between the Okavango-Linyanti
and the Central Kalahari Game Reserve, a southwesterly corridor connecting
Luengue-Luiana NP with Namibia’s Kaudom NP, and a northeasterly extension of the
Hwange corridor into Zimbabwe’s Matusadona NP. Our model did not detect any
significant direct corridors between Zimbabwe and Zambia and only very limited
W-E direct connection between the Okavango region and Namibia’s Khaudum NP. We
also found little to no direct connectivity between peripheral points; that is,
most paths and corridors connecting two adjacent peripheral points run through
more central regions before heading towards their destination at the periphery.
With the exception of the Central Kalahari Game Reserve corridor, our model did
not detect any significant connectivity outside the boundaries of KAZA-TFCA.

\newgeometry{left=30mm,right=30mm,top=0mm,bottom=20mm}
\begin{figure}[hbtp]
  \begin{center}
    \begin{tikzpicture}
        \node[anchor=south west,inner sep=0] (image) at (0,0,0) {
          \begin{minipage}{0.95\textwidth}
            \includegraphics[width = \textwidth]{99_LeastCostPaths.pdf}
            \includegraphics[width = \textwidth]{99_LeastCostCorrs.pdf}
          \end{minipage}
        };
        \begin{scope}[x={(image.south east)},y={(image.north west)}]
          % % next four lines will help you to locate the point needed by
          % % forming a grid. comment these four lines in the final picture.
          % \draw[help lines,xstep=.1,ystep=.1] (0,0) grid (1,1);
          % \draw[help lines,xstep=.05,ystep=.05] (0,0) grid (1,1);
          % \foreach \x in {0,1,...,9} {\node [anchor=north] at (\x/10,0) {0.\x};}
          % \foreach \y in {0,1,...,9} {\node [anchor=east] at (0,\y/10) {0.\y};}
          % % upto here
          % \node[text=black, circle, scale=1, font=\bfseries, inner sep = 0.5pt
          %   , draw=black] (No1) at (0.400, 0.700){1};
          % \node[text=black, circle, scale=1, font=\bfseries, inner sep = 0.5pt
          %   , draw=black] (No2) at (0.370, 0.760){2};
          % \node[text=black, circle, scale=1, font=\bfseries, inner sep = 0.5pt
          %   , draw=black] (No3) at (0.585, 0.780){3};
          % \node[text=black, circle, scale=1, font=\bfseries, inner sep = 0.5pt
          %   , draw=black] (No4) at (0.647, 0.697){4};
          % \node[text=black, circle, scale=1, font=\bfseries, inner sep = 0.5pt
          %   , draw=black] (No5) at (0.910, 0.770){5};
          % \node[text=black, circle, scale=1, font=\bfseries, inner sep = 0.5pt
          %   , draw=black] (No6) at (0.580, 0.658){6};
          % \node[text=black, circle, scale=1, font=\bfseries, inner sep = 0.5pt
          %   , draw=black] (No7) at (0.410, 0.640){7};
          % \node[text=black, circle, scale=1, font=\bfseries, inner sep = 0.5pt
          %   , draw=black] (No8) at (0.380, 0.665){8};
          % \draw[black, line width = 1pt, densely dotted] (No1) -- (0.500, 0.710);
          % \draw[black, line width = 1pt, densely dotted] (No2) -- (0.438, 0.760);
          % \draw[black, line width = 1pt, densely dotted] (No3) -- (0.565, 0.805);
          % \draw[black, line width = 1pt, densely dotted] (No4) -- (0.665, 0.685);
          % \draw[black, line width = 1pt, densely dotted] (No5) -- (0.885, 0.765);
          % \draw[black, line width = 1pt, densely dotted] (No6) -- (0.540, 0.650);
          % \draw[black, line width = 1pt, densely dotted] (No7) -- (0.450, 0.645);
          % \draw[black, line width = 1pt, densely dotted] (No8) -- (0.345, 0.690);
        \end{scope}
    \end{tikzpicture}
    \caption{(a) Source points (black dots) and corresponding least-cost paths
    leaving from protected areas (light grey). Note that only contiguous
    protected areas covering more than 700 km\textsuperscript{2} are depicted.
    Continuous thin black lines indicate the borders of the KAZA-TFCA, whereas
    dashed black lines delineate country-borders. (b) Least-cost corridors
    between the same source points as illustrated in subfigure (a). For ease of
    spatial reference we also labeled national parks that are mentioned in the
    main text (dark-grey).}
    \label{LeastCost}
  \end{center}
\end{figure}
\restoregeometry

\newpage
\section{Discussion}
% Recap of what we did
We combined GPS relocation data of dispersing African wild dog coalitions with
spatial environmental and anthropogenic covariates to gain information on
habitat selection or avoidance during dispersal. We then used this knowledge to
create a permeability surface and assess connectivity by means of identifying
dispersal corridors across the world’s largest transboundary conservation area,
the Kavango-Zambezi Transfrontier Conservation Area.

% Finding Regarding Human Influence
Our results indicate that dispersers avoided anthropogenic influences,
highlighting that human presence or activities constitute some of the main
obstacles to connectivity. While \cite{Masenga.2016} found similar avoidance of
human-dominated landscapes by dispersers in East Africa,
\cite{DaviesMostert.2012} reported a high willingness of dispersers to cross
such landscapes in South Africa. Our findings support the notion that dispersers
forgo human-dominated landscapes and we conjecture that differences to
observations by Davies-Mostert and colleagues are be due to the unavailability
of alternative ``natural'' routes in South Africa. That is, dispersers in South
Africa may be forced to cross human-dominated landscapes despite a strong
aversion against human presence.

% Finding Regarding Water
Water was identified as being another major obstacle to dispersal. In earlier
studies, water has been shown to be almost impermeable for resident packs and it
is assumed that residents may in fact use it as easy-to-defend territory
boundaries \citep{Cozzi.2013}. Although dispersers have different needs and
motivations than residents, they have also been shown to be partially hindered
by larger water bodies \citep{Abrahms.2017, Cozzi.2020}. Our findings thus align
with previous research and stress the importance to accurately and dynamically
represent water bodies, notably for such pulsing ecosystems as represented by
the Okavango Delta.

% Finding Regarding Distance To Water
While we found that dispersers avoided moving \textit{through} water, we
revealed a preference for moving \textit{close to} water. We hypothesize that
high prey-availability (e.g. impala) in proximity to water explains this result.
It is well documented that water is a major driver of wildlife and that it
determines the distribution of many herbivores, especially in dry biomes
\citep{Western.1975}. For the Okavango Delta in particular, it has been shown
that prey abundance strongly correlates with the extent of the flood
\citep{Bonyongo.2005}. An increasing concentration of ungulates around remaining
water-sources during the dry season \citep{Ogutu.2014} therefore might attract
dispersing wild dogs. Following the same logic, however, apex predators (e.g.
lions, spotted hyenas) and resident wild dogs may also prefer proximity to water
\citep{Valeix.2009} and may occasionally force dispersing wild dogs to move
distant to water \citep{Ndaimani.2016} to avoid intraguild competition
\citep{Creel.1996, Mills.1997}. We could not control for the presence or absence
of apex predators or conspecifics during dispersal as we lacked corresponding
data, but these considerations might explain why we did not find a statistically
clear effect of distance to water.

% Finding Regarding Different Permeabilities
According to our prediction of landscape permeability, permeability of northern
Botswana and southern Angola are relatively high. High permeability in these
countries is mainly caused by a combination of low human presence and
activities, low tree cover, high shrubs/grassland cover and close distance to
water. Although swamps, wetlands, and permanent water provide little
permeability themselves, their surroundings act as strong attractants to
dispersers. The low permeability that characterizes Zambia and Zimbabwe, on the
other hand, is mainly caused by severe anthropogenic presence and other human
activities. Despite the fact that the KAZA-TFCA covers many high permeability
hot-spots, several permeable regions remain uncovered by the KAZA-TFCA, offering
exciting opportunities for future expansions of already existing protected
areas.

% Finding Regarding Corridor Overlap with KAZA
Our results moreover suggest that all major dispersal corridors of African wild
dogs run within the boundaries of the KAZA-TFCA, demonstrating the potential
value of such an initiative. Within the KAZA-TFCA, however, some key
connectivity areas are not formally protected and their protection status may be
subject to future changes under the influences of an expanding human population,
farming, and cropping industry.

% Finding Regarding Dispersal Hub
Importantly, we identified the Okavango-Linyanti region as potential dispersal
hub through which dispersers gain access to more peripheral regions of the
KAZA-TFCA. While this might partly be owed to the area's central position within
the KAZA-TFCA, it is likely also due to its environmental characteristics,
namely due to human activities and the presence of impermeable water bodies such
as the Okavango Delta and the Linyanti Swamp, which potentially funnel dispersal
movements. The region is also instrumental for connecting subpopulations
remaining in Zambia’s Kafue and Zimbabwe’s Hwange NPs. These subpopulations are
separated by the Zambezi River, forcing dispersers to detour via northern
Botswana to reach more peripheral areas. The key role of the Okavango-Linyanti
region for the overall connectivity of the KATA-TFCA thus calls for actions to
secure its protection status in the future. While the region is currently a
Wildlife Management Area devoted to photographic tourism, it has neither the
status of a National Park nor that of a Game Reserve and so potentially faces
growing pressure towards exploitation of its natural resources.

% Finding Regarding Non-Protected Corridor Between Zambia and Botswana
A similar case of non-formally protected but key dispersal landscape is
represented by the area south of Kafue NP in Zambia, for which a disruption of
its main and narrow dispersal corridor would result in a considerable isolation
of its subpopulations, for no other corridor links Kafue NP to the rest of the
KAZA-TFCA.

% Finding Regarding
Our results also highlighted a potential southwards corridor between the
Okavango-Linyanti ecosystem and the Central Kalahari Game Reserve. Some areas
through which the corridor runs are neither part of KAZA-TFCA nor of any
protected status. Human presence and activities along the national road that
longitudinally traverses the corridor as well as the related increased
mortality caused by direct persecution and collisions \citep{Cozzi.2020} may
limit realized connectivity. In fact, despite having dispersed south of the
Okavango Delta, none of the monitored dispersing coalitions crossed the national
road and reached the Central Kalahari Game Reserve \citep{Cozzi.2020}.
\citep{Elliot.2014} identified a similar corridor for dispersing lions, which
shows that protecting the area would benefit lions and wild dogs alike.

Although overall connectivity gradually decreased towards the borders of the
study area, this might partly be a result of boundary-effects \citep{Koen.2010}.

In general, our network of dispersing corridors shows high similarities to
predicted movement corridors of dispersing lions \citep{Elliot.2014}, which not
only reinforces confidence in our own predictions but also suggests potential
synergies between conservation of the two species.

% The goal of our study was to (1) assess landscape permeability for dispersing
% wild dogs and to (2) subsequently use this knowledge to examine connectivity
% across the greater extent of the KAZA-TFCA. To achieve these goals, we
% parameterized a permeability model in which we used GPS data of 16 dispersing
% wild dog coalitions to quantify their habitat preferences with respect to land
% cover, protection status, and human influence. While we found significant
% impacts of different land covers and human influence, the influence of
% protection status was not statistically clear. We further applied the
% parameterized permeability model and predicted a permeability surface for the
% entire KAZA-TFCA, which depicted the ease at which dispersers crossed a specific
% landscape. Finally, based on this permeability map, we computed LCPs and LCCs
% between protectorates and thereby highlighted dispersal corridors of African
% wild dogs.
%
% \subsection{Habitat Selection Model}
% % Permeability Model: Land Cover, Water
% Our most parsimonious permeability model revealed that, with respect to land
% cover, dispersers strictly avoided areas covered by water (including swamps,
% wetlands). This result corroborates previous research on wild dogs (e.g.
% \cite{Abrahms.2017}) and is rather unsurprising, as the species is not well
% adapted for moving through water. For residents, water is almost impermeable and
% frequently serves as natural boundary to their home-ranges \citep{Cozzi.2013}.
% It is conceivable that dispersers are more willing than residents to cross
% waterbodies \citep{Cozzi.2020}, but we did not include residents in our analyses
% and could not test for such differences. Regardless of that, water appears to
% pose a serious obstacle for dispersers and is therefore likely to reduce
% landscape permeability for dispersing wild dogs.
%
% % Permeability Model: Land Cover, DistanceToWater
% While dispersers avoided moving \textit{through} water, they preferred moving
% \textit{close to} water as indicated by avoidance of high distance to water.
% While its effect was insignificant, inclusion of the covariate
% \textit{DistanceToWater} strictly increased the AIC in all of our models and
% suggests that distance to water influenced permeability negatively. We
% hypothesize that high prey-availability (e.g. impala, kudu) in proximity to
% water explains this result. It is well documented that water is a major driver
% of wildlife and that it determines the distribution of many herbivores,
% especially in dry biomes \citep{Western.1975}. For the Okavango Delta in
% particular, it has been shown that prey abundance strongly correlates with the
% extent of the flood \citep{Bonyongo.2005} and that an increasing concentration
% of ungulates around remaining water-sources during the dry season might attract
% dispersing wild dogs \citep{Ogutu.2014}. Following the same logic, however, apex
% predators (e.g. lions, spotted hyenas) and resident wild dogs may also prefer
% closeness to water \citep{Valeix.2009}, which may occasionally force dispersers
% to move distant to water \citep{Ndaimani.2016} to avoid intraguild competition
% \citep{Creel.1996, Mills.1997}. We could not control for the presence or absence
% of apex predators or conspecifics because we lacked corresponding data, but
% these considerations might explain why we did not find a statistically clear
% effect of distance to water.
%
% % Permeability Model: Land Cover, Trees & Shrubs
% In line with existing literature on moving wild dogs \citep{Abrahms.2017}, we
% found that dispersers avoided habitats with high percentages of tree cover,
% whereas they preferred areas with high percentages of shrubs/grassland cover.
% Comparable preferences were reported for residents, which typically reside in
% savanna-like habitats but rarely inhabit areas with dense tree cover
% \citep{Estes.2012}. This may be explained by wild dogs' hunting behavior, where
% open habitats are favored because they offer increased visibility and thus
% facilitate hunting \citep{Estes.1967}. Additionally, we found a strong negative
% correlation between percentage cover by shrubs/grassland and percentage cover by
% bareland (although we did not include bareland in our models). The observed
% preference for shrubs/grassland could therefore also indicate an aversion to
% bareland. Again, we hypothesize that this aversion is driven by low
% prey-availability in non-vegetated habitats \citep{Pettorelli.2011}.
% Unfortunately, we lacked detailed vegetation data and could not distinguish
% vegetation types further.
%
% % Permeability Model: Human Influence
% Finally, considering human influence, we found that dispersers strongly avoided
% human pressures arising from human density, farming, and roads. Previous
% research on dispersing wild dogs was inconclusive about the impact of human
% presence on permeability, as some researchers reported a high willingness of
% dispersers to cross human-dominated landscapes \citep{DaviesMostert.2012},
% whereas others found that humans are avoided almost entirely during dispersal
% \citep{Masenga.2016}. Our results support the notion that dispersers strongly
% avoid human-dominated areas, but we hypothesize that this finding depends on the
% availability of alternative routes. The dipsersers monitored by
% \cite{DaviesMostert.2012} originated from habitats almost exclusively surrounded
% by humans, whereas our dipsersers departed from regions with various
% possibilities to move through natural habitats. A lack of alternative routes
% combined with strong inbreeding avoidance intrinsic in African wild dogs
% \citep{McNutt.1996} might have forced the dispersers reported in
% \cite{DaviesMostert.2012} to cross human-dominated landscapes despite a strong
% aversion against humans. Although we found that our dispersers strictly avoided
% humans, \citeauthor{Cozzi.2020} (in review) reports that human persecution
% remains the main cause of mortality for dispersers, suggesting that avoidance is
% not strong enough to prevent human-induced fatalities. We argue that human
% influence remains a source of reduced landscape permeability and that the
% proceeding transformation of natural landscapes into human altered habitats
% continues to undermine wild dogs' ability to dipserse. Since our human influence
% proxy did not include minor gravel roads, our finding that humans are avoided
% does not contradict the observation that wild dogs preferably move on gravel
% roads during activity \citep{Abrahms.2016}. With four-hourly GPS relocations we
% felt that the temporal resolution was insufficient to depict such fine scale
% effects \citep{Thurfjell.2014}.
%
% \subsection{Permeability Surface}
% % Permeability Surface
% Our prediction of landscape permeability revealed vast regional differences.
% Across the entire study area, northern Botswana stood out as it comprises the
% highest density of highly permeable landscapes. The hospitality of this region
% is caused by a combination of low human influence, low tree cover, high
% shrubs/grassland cover and close distance to water. Thus, although swamps,
% wetlands, and water (e.g. Okavango Delta, Linyanti Swamp) in northern Botswana
% provide little permeability, their surroundings act as strong attractants to
% dispersers. It is therefore not very surprising that the region serves as source
% for the recolonization of surrounding habitats \citep{Cozzi.2013}. Moreover, we
% found much higher permeability inside the KAZA-TFCA than outside its borders,
% which suggests a good fit between the planned protectorates and areas that are
% suitable for dispersal. We made similar observations for protection status,
% where predicted permeabilities inside protectorates are much higher than in
% pastoral regions. Yet, several highly permeable regions remain uncovered by the
% KAZA-TFCA or other protected areas, proposing exciting opportunities for future
% expansions of already existing protectorates. Across countries, we revealed that
% Zambia and Zimbabwe offer the least permeable landscapes. Visual inspection of
% our covariate-layers disclosed that this is mainly owed to severe anthropogenic
% influence in Zambia and Zimbabwe. In contrast, Botswana and Angola host vast
% uninhabited areas, resulting in more permeable landscapes for dispersers.
% However, all of our predictions were based on observations from dispersing wild
% dogs departing from northern Botswana, which might have introduced a bias in
% estimated habitat preferences and resulting predictions.
%
% \subsection{Least-Cost Paths \& Least-Cost Corridors}
% % Least Cost Paths & Least Cost Corridors
% Our LCP- and LCC-analyses disclosed several major movement corridors within the
% extended study area and suggested a decent congruence between predicted
% corridors and the layout of the KAZA-TFCA. That is, only few corridors run
% outside the planned boundaries of the KAZA-TFCA and several currently
% unprotected corridors will become protected with realization of the KAZA-TFCA.
% For example, the high-frequency corridor running from southern Zambia into the
% Kafue NP (\Cref{LeastCost}a \circled{3}) currently contains extended unprotected
% sections that will be transformed into protectorates under the KAZA-TFCA. We
% thus expect that dispersal between such regions will be vastly facilitated
% thanks to the KAZA-TFCA. Our network of LCPs and LCCs furthermore highlighted
% the Chobe-Moremi system in northern Botswana, which is traversed by several
% LCPs and LCCs and seems to hold the potential to serve as central hub through
% which dispersers gain access to more remote regions of the KAZA-TFCA. Still, our
% predictions also indicated that movements between wild dog subpopulations may
% often involve extended detours. This is especially the case for regions in the
% northern sectors of the KAZA-TFCA. For instance, only few LCPs connect the
% Zimbabwean and Zambian sectors, such that moving directly from the Matusadona NP
% to the Kafue NP seems almost impossible and likely requires to make a detour
% through the Chobe-Moremi region. Similarly, few LCPs directly link the Mavinga
% NP in Angola and the Kafue NP in Zambia. Moreover, we discovered potential
% dispersal routes between the Chobe-Moremi ecosystem and the Central Kalahari
% NP that are not yet proclaimed protected areas. Nonetheless, a protected link
% between the Central Kalahari NP and the Chobe-Moremi ecosystem might grant
% southern populations invaluable access to the Chobe-Moremi-Hub and
% subsequently to the entire KAZA-TFCA. In fact, \cite{Elliot.2014} identified a
% similar corridor for dispersing lions, proposing that protecting the area would
% benefit lions and wild dogs alike. The absence of direct LCPs and LCCs between
% some regions of the KAZA-TFCA is likely owed to an impermeable landscape between
% those regions and demonstrates that protection of such areas in their current
% state would be futile. That is, impermeable landscapes first need to be
% transformed into more permeable habitats before protection pays off. Although,
% we noted that connectivity gradually decreases towards the borders of our study
% area, this might partly be owed to boundary-effects \citep{Koen.2010}.
% Interestingly, our network of LCPs and LCCs shows surprising similarities to
% predicted movement corridors of dispersing lions \citep{Elliot.2014}.
% Remarkably, corridors that we identified as high-frequency corridors are also
% visible in the network presented by \cite{Elliot.2014}. This not only reinforces
% confidence in our own predictions but it also suggests potential synergies
% between conservation of the two species.

% Future Research: Genetic Resistance
While we assessed landscape permeability purely based on GPS relocation
data, future studies could challenge and test our predictions using genetic data
\citep{Spear.2010}. Similar comparisons have been conducted for other species
and revealed a surprisingly high overlap between predicted landscape
permeability based on genetic and movement data \citep{Cushman.2010}. Genetic
data would also deepen our understanding of the degree to which subpopulations
admix and of the overall genetic diversity found in different subpopulations
\citep{Girman.2001}. Based on our connectivity network, we would expect to find
a low genetic diversity for populations that live in insufficiently connected
habitats, whereas populations residing at dispersal hotspots, such as the
Okavango-Linyanti ecosystem, should exhibit a relatively diverse gene-pool.

% Future Research: Better Covariates
Future research should also strive to improve representation of geo-referenced
covariates. Because we predicted connectivity for such an extensive area, we
were strongly limited in available and reliable land-cover data. As a result,
our representation of vegetation remained rather basic with only three
vegetation types (tree-cover, shrubs/grassland, bareland). Although there are a
few large-scale, high-resolution land cover datasets, such as Globeland's 30m
\citep{Chen.2015} and the European Space Agency's and 20m \citep{ESA.2019} land
cover datasets, we found their classifications often erroneous for our study
area. This is probably owed to the KAZA-TFCA's unique landscape and
corresponding difficulties in remote sensing satellite imagery using traditional
indices \citep{Wolski.2017}. Nevertheless, land cover is a very powerful
predictor of animal movement \citep{Thurfjell.2014}, which stresses the need for
more reliable land-cover data.

% Future Research: Competitor Presence
In a similar manner, we would desire better knowledge of the whereabouts of
other species. In particular, we would desire location data of apex predators,
potential prey, and conspecifics. It is well documented that resident wild dogs
frequently compete with other carnivores \citep{Creel.1998} and we would expect
to find competition between dispersing wild dogs and apex predators too.
Furthermore, knowledge about the spatial distribution of prey could verify our
hypothesis that dispersers move along water due prey-availability. Lastly, we
could not control for the likelihood of encountering conspecifics during
dispersal . Yet, for dispersing meerkats, a species with surprisingly similar
social structure to wild dogs, it has been shown that the social landscape, i.e.
the presence or absence of conspecifics, strongly shapes behavior during
dispersal \citep{Cozzi.2018}. More specifically, it has been shown that
dispersers select for high density of conspecifics while still inside their
home-range, but that they avoid high densities of conspecifics once they leave
their natal home-range (we replicated this analysis for our core study area; see
\Cref{Appendix:SocialLandscape}). Spatially explicit knowledge about the
distribution of foreign wild dogs is thus desirable and would certainly deepen
our understanding of dispersal behavior. However, constructing meaningful and up
to date social landscapes is extremely challenging and requires vast amounts of
movement data, especially for such a wide-ranging species as the African wild
dog \citep{Pomilia.2015}.

% Future Research: Global Resistance Mapping
Besides urging for more and better data, we also want to emphasize that our
analytical approach is neither limited to African wild dogs, nor to our study
area. All covariates used throughout this study are available on a global scale
and they are likely to be important determinants of landscape permeability for
other species. Our framework could thus easily be applied to assess connectivity
for a wide range of species, provided suitable relocation data of the species
exists. Expanding our analyses to other species would surely yield interesting
insights on the congruence of movement corridors of different species and
thereby highlight areas that are used by multiple species for dispersal.

% Future Research: Simulating Dispersal
To this end, we identified movement corridors between pre-defined start and
end-points, which assumes that dispersers have a specific end-point in mind.
However, pre-defining end-points might not even be necessary, as one could
invert the iSSF and use it as a mechanistic movement model to simulate dispersal
without known end-point. That is, one could define a set of source points at
locations where wild-dogs are assumed or known to reside and simulate dispersers
according to habitat-preferences that were derived using the iSSF framework.
End-points would not be restricted to a certain location and movement corridors
would emerge naturally as the result of a myriad of simulated dispersal events
(for an example of simulated dispersal events see
\Cref{Appendix:DispersalSimulation}).

% Conclusion
Our results demonstrate that indeed the majority of dispersal corridors for
African wild dogs are covered by the boundaries of the KAZA-TFCA, suggesting
that the KAZA-TFCA will severely contribute to the long-term viability of this
species. Nevertheless, some unprotected dispersal routes do exist outside the
KAZA-TFCA and suggest opportunities for further expansions in the future.
Moreover, our connectivity network of least-cost paths and least-cost corridors
revealed especially the Okavango-Linyanti region as central dispersal hub
through which dispersers gain access to more remote regions of the KAZA-TFCA. We
therefore stress the importance of northern Botswana for the conservation of
this charismatic species. Finally, our investigations showed that human pressure
remains a severe impediment to dispersal and strictly reduces landscape
connectivity. Successful conservation of the African wild dog is thus contingent
on the willingness of local authorities, policy makers, and land managers to
provide vast areas that are liberated from human strains. Ultimately, our work
contributes to the growing field of connectivity literature and has important
implications for the management and conservation of the endangered African wild
dog.

\newpage
\section{Acknowledgements}
All of this work would not have been possible without the aid of several helpful
people. I would first like to thank Prof. Arpat Ozgul, the leader and mastermind
of our research group, whose expertise in statistics, biological processes and
scientific writing was invaluable for my work. Moreover, I want to thank
Gabriele Cozzi for his excellent guidance, all the interesting discussions, his
constructive criticism, valuable revisions in the manuscript, and of course his
constant supply of freshly brewed coffee. Particular gratitude goes to Dominik
Behr, who supervised my work on a daily basis and greatly contributed to all
aspects of the thesis. I genuinely enjoyed the frequent meetings and brainstorms
and I really feel privileged to have had the opportunity of working with such a
great scientist. I am also greatly indebted to Prof. John Fieberg, who visited
our department this summer and invested a great amount of time to consult and
improve all statistical aspects of the thesis, as well as to deepen my personal
understanding of movement ecology. Furthermore, I thank the Okavango Research
Institute, in particular Piotr Wolski, for sharing all of their Python scripts,
floodmaps and knowledge on developing dynamic floodmaps. Also, I want to thank
Stefan Sommer for sharing and teaching his hard earned knowledge on scientific
writing. Additionally, I want to thank all of my colleague master students who
have been wonderful companions and made the time at university a blast.
Likewise, I thank my parents for supporting all of my undertakings and for
providing first class shelter and food. Last, but most definitely not least, I
also greatly thank my girlfriend, who was always there for me and provided
welcome distractions for my mind.

\newpage
\begingroup
\singlespacing
\bibliography{Literatur}
\endgroup

\newpage
\appendix
\section{Appendix}
\subsection{Floodmapping Algorithm}
\label{Appendix:FloodmappingAlgorithm}
To implement the floodmapping algorithm, we defined two sets of polygons located
in the region of the Okavango Delta (see \Cref{Floodmapping}). The first set
consisted of areas known to be permanent dryland, whereas the second set
consisted of permanent waters. Since we were unable to retrieve the original
polygons used in \cite{Wolski.2017}, we geo-referenced one of their
illustrations and traced the polygons there. After recreation of the polygons,
we used the R-package \textit{getSpatialData} \citep{Schwalb.2018} to download
and pre-process all relatively cloud-free MODIS Terra images (MCD43A4,
\cite{Schaaf.2015}) available for the period of our dispersal events. Assessment
of cloud cover was based on visual inspection of MODIS images on ORI's website
\citep{ORI.2019}. The MCD43A4 MODIS product is particularly useful for dynamic
floodmapping because it is updated with new composite images every 8th day
\citep{Wolski.2017}. After download, we classified each MODIS image into a
binary map of water (flood) and dryland using a threshold that was identified as
follows. First, we extracted all reflectance values of MODIS Terra Band 7 within
the water- and dryland-polygons. Second, we computed histograms of
water-reflectances and dryland-reflectances and empirically verified that
reflectances of the two groups were sufficiently distinct. More specifically, we
checked if superimposing the two histograms resulted in a bimodal plot. This was
said to be achieved if the 99\textsuperscript{th} percentile of
water-reflectances did not severely exceed the 1\textsuperscript{st} percentile
of dryland-reflectances (\(p_{0.99, water} - \frac{10}{255} < p_{0.01,
dryland}\)). Third, if peaks were sufficiently different, we calculated a
threshold (\(t\)) using \Cref{EQ1}:

\begin{equation}
\label{EQ1}
t = \widetilde{p}_{water} + 0.3 * (\widetilde{p}_{dryland} -
\widetilde{p}_{water})
\end{equation}

\noindent where \(\widetilde{p}_{water}\) and \(\widetilde{p}_{dryland}\) were
the median reflectances of water and dryland, respectively. We then classified
all pixels of MODIS Terra Band 7 with a value greater than \(t\) as dryland and
all pixels with a value smaller than \(t\) as water.

\begin{figure}[h]
  \begin{center}
    \includegraphics[width = \textwidth]{99_Floodmapping.pdf}
    \caption{Figures describing the floodmapping algorithm. (a) The colored
    polygons indicate permanent waters and permanent dryland. Below these
    polygons we extracted reflectance values of MODIS Terra Band 7 and used
    their repsective medians to calculate a classification threshold. (b)
    Example of a classified MODIS Terra Band 7 image after application of the
    threshold.}
    \label{Floodmapping}
  \end{center}
\end{figure}

% Dynamic Wetmask
Importantly, bimodality was not always achieved and in some cases no floodmap
could be calculated. In fact, it appears that non-bimodality caused the ORI
algorithm to fail since the end of 2018, such that no floodmaps have been
generated since then (ORI, personal comm.). We hypothesized that this was caused
by the application of static water-polygons that did not cover permanent waters
correctly anymore. We therefore revised the algorithm and allowed for a more
dynamic polygonization of water. That is, for each MODIS image we calculated new
water-polygons comprising areas that were covered by the flood in 99\% of the
floodmaps from the previous five years. All of the necessary floodmaps from
previous years were kindly provided to us by ORI. Using this slightly amended
approach, we were able to address some of the bimodality issues and to classify
several additional floodmaps for the period of our study. Because MODIS Terra
Band 7 had a resolution of 500m x 500m, we interpolated all maps to 250m x 250m.

% Validation
To validate and compare the performance of our own algorithm to the original
ORI-algorithm, we randomly sampled 48 dates for which ORI prepared classified
images. To make sure that months were equally represented in the sampled dates,
we employed stratified sampling based on months (regardless of the year) and
randomly sampled four maps for each month. For the sampled dates we downloaded
and classified MODIS Terra Band 7 images and compared our classified images to
those provided by ORI (\Cref{FloodmappingValidation}). For each pair of maps we
created a difference map indicating false positives and false negatives and
computed the relative number of wrongly classified pixels. We achieved an
overall accuracy of 97\%, which presumably is an underestimate of the true
performance, as we introduced some errors when resampling the ORI-maps to our
reference raster.

\begin{figure}[ht]
  \begin{center}
    \includegraphics[width = \textwidth]{99_FloodmappingValidation.pdf}
    \caption{Validation procedure of our floodmapping algorithm. (a) Classified
    image that was provided to us by ORI. (b) Image for the same date but now
    classified using our own algorithm. (c) Difference image indicating false
    positives and false negatives in our own classification.}
    \label{FloodmappingValidation}
  \end{center}
\end{figure}

% Static Watermap
\noindent While water in the Okavango Delta was mapped dynamically, we used a
static watermap for its surroundings. This static map was based on Globeland's
land cover dataset \citep{Chen.2015}, which contains ten land cover categories.
From the original layer we only retained the categories \textit{wetland} and
\textit{waterbodies} and collectively reclassified them to \textit{water}.
Globeland had an original resolution of 30m x 30m, so we coarsened the layer to
250m x 250m using the mode of each 250m x 250m cell. Lastly, we improved river
representation using the MERIT Hydro dataset \citep{Yamazaki.2019}, from which
we added all rivers with a width of over 10m to our watermaps. The entire
process resulted in several hundred watermaps, of which 72 represented the
closest watermap to one of our GPS relocations.

\begin{figure}[h]
  \begin{center}
    \includegraphics[width = \textwidth]{99_FloodPulse.pdf}
    \caption{Sequence of floodmaps showing a typical flood pulse throughout the
    year. The flood arrives from the north-western corner (so called
    ``pan-handle'') of the Okavango Delta and slowly descents through the delta
    in south-eastern direction, where it nourishes several tributaries. The
    extent of the flood peaks around August or September, after which it slowly
    retracts. Between December and March the reflectance properties of water and
    dryland change, which is why often no accurate floodmaps can be obtained for
    these months using remote sensing techniques.}
    \label{FloodPulse}
  \end{center}
\end{figure}

\newpage
\subsection{Reclassification of Protected Areas}
\label{Appendix:Reclassification}
\begin{table}[h]
  \caption{Reclassification of protected areas. Data on protection status was
  obtained from the \cite{PeaceParks.2019}. We simplified original categories to
  a binary classification of protected and unprotected (pastoral) areas. Any
  area that was not represented by one of the categories listed in
  \Cref{Reclassification} was deemed to be pastoral landscape.}
  \label{Reclassification}
  \begin{center}
    \resizebox{\textwidth}{!}{
      \begin{tabular}{llp{11cm}}
      \toprule
      \textbf{Original Category} &
        \textbf{Reclassification} &
          \textbf{Comment} \\
      \midrule
      Communal Conservancy &
        Protected &
          No comment\\
      Community Forest &
        Protected &
          No comment\\
      Conservancy &
        Protected &
          No comment\\
      Forest Reserve &
        Protected &
          No comment\\
      Game Management Area &
        Protected &
          No comment\\
      Game Park &
        Protected &
          No comment\\
      Game Ranch &
        Protected &
          No comment\\
      Game Reserve &
        Protected &
          No comment\\
      Hunting Reserve &
        Protected &
          No comment\\
      National Park &
        Protected &
          No comment\\
      National Park (not yet proclaimed) &
        \textit{Removed} &
          Since not yet proclaimed, we did not include this area as protected\\
      Overlap (GMA/FR) &
        Protected &
          Mixture between game management area and forest reserve\\
      Proposed &
        Protected &
          No comment\\
      Protected Area &
        Protected &
          No comment\\
      Recreation Park &
        \textit{Removed} &
          Marine reserve\\
      Safari Area &
        Protected &
          No comment\\
      Sanctuary &
        Protected &
          No comment\\
      State Forest &
        Protected &
          No comment\\
      Wildlife Management Area &
        Protected &
          No comment\\
      World Heritage Site &
        Protected &
          No comment\\
      Freehold Conservancy &
        Protected &
          No comment\\
      \bottomrule
      \end{tabular}
    }
  \end{center}
\end{table}

\newpage
\subsection{Road Categories}
\label{Appendix:RoadCategories}
\begin{table}[h]
  \caption{Description of road types, as sourced from Open Street Map's mapping
  guide (\url{https://wiki.openstreetmap.org/wiki/Key:highway}). Roads types
  that were considered for the purpose of this study are shaded in light gray.}
  \label{RoadsDescription}
  \begin{center}
    \resizebox{\textwidth}{!}{
      \begin{tabular}{llp{13cm}}
      \toprule
      \textbf{Group} &
        \textbf{Subgroup} &
          \textbf{Description} \\
      \midrule
      \rowcolor[gray]{0.80}
      Roads &
        motorway &
          A restricted access major divided highway, normally with 2 or more
          running lanes plus emergency hard shoulder. Equivalent to the
          Freeway, Autobahn, etc. \\
      \rowcolor[gray]{0.80}
      Roads &
        trunk &
          The most important roads in a country's system that aren't
          motorways. Need not necessarily be a divided highway. \\
      \rowcolor[gray]{0.80}
      Roads &
        primary &
          The next most important roads in a country's system. Often link
          larger towns. \\
      \rowcolor[gray]{0.80}
      Roads &
        secondary &
          The next most important roads in a country's system. Often link
          towns. \\
      Roads &
        tertiary &
          The next most important roads in a country's system. Often link
          smaller towns and villages \\
      Roads &
        unclassified &
          The least important thorough roads in a country's system, i.e. minor
          roads of a lower classification than tertiary, but which serve a
          purpose other than access to properties. Often link villages and
          hamlets. \\
      Roads &
        residential &
          Roads which serve as an access to housing, without function of
          connecting settlements. Often lined with housing. \\
      Roads &
        service &
          For access roads to, or within an industrial estate, camp site,
          business park, car park etc. \\
      \rowcolor[gray]{0.80}
      Link roads &
        motorway\_link &
          The link roads (sliproads/ramps) leading to/from a motorway from/to
          a motorway or lower class highway. Normally with the same motorway
          restrictions. \\
      \rowcolor[gray]{0.80}
      Link roads &
        trunk\_link &
          The link roads (sliproads/ramps) leading to/from a trunk road
          from/to a trunk road or lower class highway. \\
      \rowcolor[gray]{0.80}
      Link roads &
        primary\_link &
          The link roads (sliproads/ramps) leading to/from a primary road
          from/to a primary road or lower class highway. \\
      \rowcolor[gray]{0.80}
      Link roads &
        secondary\_link &
          The link roads (sliproads/ramps) leading to/from a secondary road
          from/to a secondary road or lower class highway. \\
      Link roads &
        tertiary\_link &
          The link roads (sliproads/ramps) leading to/from a tertiary road
          from/to a tertiary road or lower class highway. \\
      Special road types &
        living\_street &
          For living streets, which are residential streets where pedestrians
          have legal priority over cars, speeds are kept very low and where
          children are allowed to play on the street. \\
      Special road types &
        pedestrian &
          For roads used mainly/exclusively for pedestrians in shopping and
          some residential areas which may allow access by motorised vehicles
          only for very limited periods of the day. \\
      Special road types &
        track &
          Roads for mostly agricultural or forestry uses. \\
      Special road types &
        bus\_guideway &
          A busway where the vehicle is guided by the way (though not a railway)
          and is not suitable for other traffic. \\
      Special road types &
        escape &
          For runaway truck ramps, runaway truck lanes, emergency escape
          ramps, or truck arrester beds. It enables vehicles with braking
          failure to safely stop. \\
      Special road types &
        raceway &
          A course or track for racing \\
      Special road types &
        road &
          A road/way/street/motorway/etc. of unknown type. It can stand for
          anything ranging from a footpath to a motorway. \\
      \bottomrule
      \end{tabular}
    }
  \end{center}
\end{table}

\newpage
\subsection{Covariate Sources}
\label{Appendix:Sources}
\begin{table}[h]
  \begin{center}
    \caption{Sources from which we obtained our set of geo-referenced
    covariates, as well as an indicator of their original resolution and the
    method used for adjusting the resolution to 250m x 250m.}
    \label{Sources}
    \resizebox{\textwidth}{!} {
      \begin{tabular}{llp{4cm}p{4cm}p{7cm}}
      \hline
      \textbf{Category} &
        \textbf{Covariate} &
          \textbf{Resolution} &
            \textbf{Resampling} &
              \textbf{Source(s)} \\
      \midrule
      Land Cover
        & Water
          & 30m x 30m (static), 500m x 500m (dynamic)
            & Mode of each 250x x 250m cell (static), disaggregation to 250m x 250m (dynamic)
              & \cite{Chen.2015}, \cite{Schaaf.2015}, \cite{Yamazaki.2019} \\
      Land Cover
        & Dryland
          & 30m x 30m (static), 500m x 500m (dynamic)
            & Mode of each 250x x 250m cell (static), disaggregation to 250m x 250m (dynamic)
              & \cite{Chen.2015}, \cite{Schaaf.2015}, \cite{Yamazaki.2019} \\
      Land Cover
        & DistanceToWater
          & 30m x 30m (static), 500m x 500m (dynamic)
            & Mode of each 250x x 250m cell (static), disaggregation to 250m x 250m (dynamic)
              & \cite{Chen.2015}, \cite{Schaaf.2015}, \cite{Yamazaki.2019} \\
      Land Cover
        & Shrubs/Grassland
          & 250m x 250m
            & NA
              & \cite{Dimiceli.2015} \\
      Land Cover
        & Trees
          & 250m x 250m
            & NA
              & \cite{Dimiceli.2015} \\
      Land Cover
        & Bareland
          & 250m x 250m
            & NA
              & \cite{Dimiceli.2015} \\
      \hdashline
      Protection Status
        & Protected
          & NA
            & NA
              & \cite{PeaceParks.2019} \\
      Protection Status
        & Pastoral
          & NA
            & NA
              & \cite{PeaceParks.2019} \\
      \hdashline
      Human Influence
        & HumansBuffer5000
          & 30m x 30m
            & Sum of each 250m x 250m cell
              & \cite{Facebook.2019}, \cite{OpenStreetMap.2019}, \cite{Chen.2015}, \cite{Xiong.2017}\\
      Human Influence
        & DistanceToRoads
          & NA
            & NA
              & \cite{Facebook.2019}, \cite{OpenStreetMap.2019}, \cite{Chen.2015}, \cite{Xiong.2017}\\
      Human Influence
        & RoadCrossing
          & NA
            & NA
              & \cite{Facebook.2019}, \cite{OpenStreetMap.2019}, \cite{Chen.2015}, \cite{Xiong.2017}\\
      \hline
      \end{tabular}
    }
  \end{center}
\end{table}

\newpage
\subsection{Model Selection Results}
\label{AICs}
\begin{table}[h]
  \caption{Results from the forward model selection procedure based on AIC
  \citep{Burnham.2002} for our permeability model. Weights were calculated
  using: \(\text{AIC-Weight}_i = \frac{e^{-0.5} \Delta \text{AIC}_i} {\sum_{i =
  1}^n e^{-0.5} \Delta \text{AIC}_i} \). The most parsimonious model (shaded in
  light gray) outperformed all other models (\(\Delta AIC > \) 2) and received a
  weight of one. The model in the last row did not converge and no AIC value
  could be obtained. (*) represents the fixed covariates \(cos(ta)\) and
  \(log(sl)\).}
  \label{ModelAICs}
  \begin{center}
    \resizebox{\textwidth}{!}{
      \begin{tabular}{lllll}
      \toprule
      \textbf{Covariates} &
        \textbf{AIC} &
          \textbf{\(\Delta\)AIC} &
            \textbf{AIC-Weight} &
              \textbf{LogLik} \\
      \midrule
      \rowcolor[gray]{0.80}
      (*), Water, DistanceToWater, Shrubs/Grassland, HumansBuff5000, Trees &
        89841.85 &
          0.00 &
            1.00 &
              -44905.93 \\
      (*), Water, DistanceToWater, Shrubs/Grassland, HumansBuff5000, Trees, DistanceToRoads &
        89844.46 &
          2.61 &
            0.00 &
              -44905.23 \\
      (*), Water, DistanceToWater, Shrubs/Grassland, HumansBuff5000, Trees, Protected &
        89845.78 &
          3.93 &
            0.00 &
              -44905.89 \\
      (*), Water, DistanceToWater, Shrubs/Grassland, HumansBuff5000 &
        89847.96 &
          6.11 &
            0.00 &
              -44910.98 \\
      (*), Water, DistanceToWater, Shrubs/Grassland, HumansBuff5000, Trees,
      RoadCrossing &
        89848.24 &
          6.38 &
            0.00 &
              -44905.12 \\
      (*), Water, DistanceToWater, Shrubs/Grassland, HumansBuff5000, Trees,
      DistanceToRoads, Protected &
        89848.45 &
          6.60 &
            0.00 &
              -44905.22 \\
      (*), Water, DistanceToWater, Shrubs/Grassland, Trees
        & 89850.51 &
          8.66 &
            0.00 &
              -44912.26 \\
      (*), Water, DistanceToWater, Shrubs/Grassland, HumansBuff5000,
      DistanceToRoads &
        89850.81 &
          8.96 &
            0.00 &
              -44910.40 \\
      (*), Water, DistanceToWater, Shrubs/Grassland, HumansBuff5000, Trees, DistanceToRoads, RoadCrossing &
        89850.87 &
          9.02 &
            0.00 &
              -44904.44 \\
      (*), Water, DistanceToWater, Shrubs/Grassland, HumansBuff5000, Protected &
        89851.91 &
          10.06 &
            0.00 &
              -44910.96 \\
      (*), Water, DistanceToWater, Shrubs/Grassland, HumansBuff5000, RoadCrossing &
        89854.42 &
          12.57 &
            0.00 &
              -44910.21 \\
      (*), Water, DistanceToWater, Shrubs/Grassland &
        89856.89 &
          15.03 &
            0.00 &
              -44917.44 \\
      (*), Water, DistanceToWater, Shrubs/Grassland, DistanceToRoads &
        89858.78 &
          16.92 &
            0.00 &
              -44916.39 \\
      (*), Water, DistanceToWater, Shrubs/Grassland, Protected &
        89860.44 &
          18.59 &
            0.00 &
              -44917.22 \\
      (*), Water, DistanceToWater, Shrubs/Grassland, RoadCrossing &
        89862.66 &
          20.81 &
            0.00 &
              -44916.33 \\
      (*), Water, DistanceToWater, HumansBuff5000 &
        89870.14 &
          28.29 &
            0.00 &
              -44924.07 \\
      (*), Water, DistanceToWater &
        89877.72 &
          35.87 &
            0.00 &
              -44929.86 \\
      (*), Water, DistanceToWater, DistanceToRoads &
        89880.07 &
          38.22 &
            0.00 &
              -44929.04 \\
      (*), Water, DistanceToWater, Protected &
        89881.39 &
          39.54 &
            0.00 &
              -44929.70 \\
      (*), Water, DistanceToWater, Trees &
        89881.57 &
          39.72 &
            0.00 &
              -44929.79 \\
      (*), Water, DistanceToWater, RoadCrossing &
        89883.48 &
          41.63 &
            0.00 &
              -44928.74 \\
      (*), Water, HumansBuff5000 &
        89887.27 &
          45.42 &
            0.00 &
              -44934.64 \\
      (*), Water &
        89894.64 &
          52.79 &
            0.00 &
              -44940.32 \\
      (*), Water, DistanceToRoads &
        89897.24 &
          55.39 &
            0.00 &
              -44939.62 \\
      (*), Water, Protected &
        89898.47 &
          56.62 &
            0.00 &
              -44940.24 \\
      (*), Water, Trees &
        89898.53 &
          56.67 &
            0.00 &
              -44940.26 \\
      (*), Water, RoadCrossing &
        89900.73 &
          58.88 &
            0.00 &
              -44939.37 \\
      (*), Shrubs/Grassland &
        89999.47 &
          157.62 &
            0.00 &
              -44992.73 \\
      (*), DistanceToWater &
        90001.72 &
          159.87 &
            0.00 &
              -44993.86 \\
      (*), HumansBuff5000 &
        90056.07 &
          214.22 &
            0.00 &
              -45021.04 \\
      (*), Trees &
        90059.55 &
          217.69 &
            0.00 &
              -45022.77 \\
      (*), DistanceToRoads &
        90067.62 &
          225.76 &
            0.00 &
              -45026.81 \\
      (*), Protected &
        90067.85 &
          226.00 &
            0.00 &
              -45026.93 \\
      (*), RoadCrossing &
        90069.86 &
          228.00 &
            0.00 &
              -45025.93 \\
      (*), Water, Shrubs/Grassland &
        \(NA\) &
          \(NA\) &
            0.00 &
              \(NA\) \\
      \bottomrule
      \end{tabular}
    }
  \end{center}
\end{table}

\newpage
\subsection{Further Investigations}
We have been working on further investigations that did not fit the main text
and were thus moved to the Appendix of this thesis. \Cref{SchematicOverview}
gives an overview of how these additional investigations integrate into our main
work. Our main work is described in the blue boxes, which resulted in the
identification of LCPs and LCCs. However, as we have emphasized in the
discussion, the applicability of the iSSF framework expands beyond the
estimation of movement and habitat preferences and it could, in fact, be used to
simulate animal movement. We thus engaged in fitting a movement model that was
specifically intended to simulate dispersal events (red boxes). Using this
model, we simulated 1000 dispersal events across the surroundings of our core
study area and qualitatively assessed the outcome. Corresponding methods and
results are reported in \Cref{Appendix:DispersalSimulation}. Furthermore, we
motivated that data on the social landscape of dispersers would deepen our
understanding of how dispersers are influenced by the presence or absence of
conspecifics, apex predators, and prey. Although we lacked proper GPS relocation
data of apex predators and prey, we collected a considerable amount of movement
data from resident wild dogs residing in our core study area. These data cover
the periods of almost all of our dispersal events and thus hold the potential
for modeling a social landscape for our dispersers. Hence, we modeled such a
landscape for the extent of our core study area and used it as additional
covariate to parameterize a social model (yellow box). Corresponding methods and
results are reported in \Cref{Appendix:SocialLandscape}.

\begin{figure}[h]
  \begin{center}
    \begin{tikzpicture}[node distance=5mm and 5mm]
      \node (mod2) [nonterminal
        , minimum width = 30mm
        , fill          = red!20
        ] {Movement Model};
      \node (mod1) [nonterminal
        , above         = of mod2
        , fill          = blue!20
        , minimum width = 30mm
        ] {Permeability Model};
      \node (mod3) [nonterminal
        , below         = of mod2
        , fill          = yellow!20
        , minimum width = 30mm
        ] {Social Model};
      \node (dat1) [nonterminal
        , above left    = of mod2
        , fill          = blue!20
        , yshift        = -5mm
        , minimum width = 30mm
        ] {GPS Data};
      \node (dat2) [nonterminal
        , below left    = of mod2
        , fill          = blue!20
        , yshift        = +5mm
        , minimum width = 30mm
        ] {Spatial Covariates};
      \node (for1) [nonterminal
        , right         = of mod1
        , fill          = blue!20
        , minimum width = 35mm
        ] {Permeability Map};
      \node (forfor1) [nonterminal
        , right         = of for1
        , fill          = blue!20
        , minimum width = 35mm
        , yshift        = +5mm
        ] {Least-Cost Paths};
      \node (forfor2) [nonterminal
        , right         = of for1
        , fill          = blue!20
        , minimum width = 35mm
        , yshift        = -5mm
        ] {Least-Cost Corridors};
      \node (for2) [nonterminal
        , right         = of mod2
        , fill          = red!20
        , minimum width = 35mm
        ] {Dispersal Simulation};
      \node (box) [draw
        , dotted
        , very thick
        , minimum width   = 35mm
        , minimum height  = 35mm
        ] {};
      \node [above = of box, yshift = -5mm] {iSSF};
      \path (dat1) edge[->] (box);
      \path (dat2) edge[->] (box);
      \path (mod1) edge[->] (for1);
      \path (mod2) edge[->] (for2);
      \path (for1) edge[->] (forfor1);
      \path (for1) edge[->] (forfor2);
    \end{tikzpicture}
    \caption{Schematic overview of our methodological approaches. The blue boxes
    highlight our main line of work which resulted in the identification of
    least-cost paths and least-cost corridors. The red boxes show how we
    additionally used the iSSF framework to specifically analyze movement
    behavior and subsequently to simulate dispersal events. Finally, the yellow
    box indicates our attempt to use the iSSF framework to model and analyze the
    influence of the social landscape on dispersal behavior.}
    \label{SchematicOverview}
  \end{center}
\end{figure}

\newpage
\subsection{Movement Model \& Dispersal Simulation}
\label{Appendix:DispersalSimulation}
\subsubsection{Methods}
To gain more detailed insights into the movement behavior of dispersing wild
dogs and to simulate dispersal events, we fitted a more complex movement model.
The movement model still contained the most parsimonious permeability model, but
in addition included interaction terms between movement metrics (i.e.
\(cos(ta)\), \(log(sl\))) and environmental covariates (e.g. water). For
instance, for the environmental covariate \textit{Water}, we created the
interaction terms \textit{cos(ta):Water} and \textit{log(sl):Water}. These
interactions served to learn how dispersers moved differently, depending on
environmental covariates (for comparable work see \cite{Prokopenko.2017}). We
applied the same modeling framework as explained in \Cref{Modeling} and ran
forward model selection based on AIC to identify the most parsimonious movement
model. That is, we departed from the permeability model and iteratively added
interactions with movement metrics. Note that we applied forward model selection
only on the newly identified interactions, but not for the terms that were
already in the permeability model. Since we attempted to keep our simulation
simple, we did not average models with positive AIC-weights but simply
identified the most parsimonious movement model.

Using the most parsimonious movement model, we simulated 1000 dispersal events
departing from our core study area. The simulation basically resembled an
inverted iSSF function and was set up as follows. First, we sampled a random
starting location within our core study area. Second, we assigned a random
orientation in order to be able to calculate turning angles. Third, we created
an array of 25 random steps (i.e. one stratum) following the same procedure as
described in \Cref{Modeling}. Fourth, we extracted environmental covariates
along each of the generated random steps and calculated the metrics listed in
\Cref{ExtractedCovars}. Fifth, we applied \Cref{EQ2} using parameter estimates
from the movement model and calculated the selection score \(w(x)\) of each of
the 25 random steps. Sixth, we identified and realized the step with the highest
selection score and determined the new location. We then repeated steps three to
six until 200 steps were realized. To deal with dispersers that approached a map
boundary, our simulation included the options to either stop the simulation and
skip to the next disperser, or to remove any random step ranging outside the
study area, thereby forcing the simulated disperser to stay within the
boundaries. None of the simulated dispersers approached a map boundary within
200 steps and the decision did not influence our findings. Although we ran the
simulation only for dispersers departing from our core study area, one could
easily expand the simulation to include multiple sources and thereby check how
well these sources are connected.

\newpage
\subsubsection{Results}
Along with the original covariates contained in the permeability model, our most
parsimonious movement model (see \Cref{ModelAICs2} and
\Cref{MovementModelResults}) retained three additional interactions (see red box
in \Cref{MovementModel}a). Since the sign and direction of the effects contained
in the permeability model did not markedly change, we only report on the newly
added interactions between environmental covariates and movement metrics. For an
easier interpretation of how these interactions influenced habitat selection and
movement patterns during dispersal, we prepared surface plots as depicted in
\Cref{MovementModel}b. These surface plots revealed that dispersers preferred
long steps (i.e. a high movement rate) when water cover was low, whereas they
preferred short steps (i.e. a low movement rate) when water cover was high (see
\Cref{MovementModel}b1). Furthermore, dispersers were indifferent between high
and low turning angles in proximity to water, but clearly preferred low turning
angles (i.e. a high \(cos(ta)\)) when distance to water was high (see
\Cref{MovementModel}b2). Finally, dispersers preferably realized steps with low
turning angles when human influence was low, whereas they preferably realized
low turning angles when human influence was high (see \Cref{MovementModel}b3).

\begin{figure}[hbtp]
  \begin{center}
    \begin{tikzpicture}
        \node[anchor=south west,inner sep=0] (image) at (0,0,0) {
          \begin{minipage}{0.95\textwidth}
            \begin{center}
              \includegraphics[width = 0.80\textwidth]
                {99_MovementModel.pdf}\\
              \includegraphics[width = 1\textwidth]
                {99_MovementModel(Interactions).pdf}
            \end{center}
          \end{minipage}
        };
        \begin{scope}[x={(image.south east)},y={(image.north west)}]
           %  % next four lines will help you to locate the point needed by
           %  % forming a grid. comment these four lines in the final picture.↓
           % \draw[help lines,xstep=.1,ystep=.1] (0,0) grid (1,1);
           % \draw[help lines,xstep=.05,ystep=.05] (0,0) grid (1,1);
           % \foreach \x in {0,1,...,9} { \node [anchor=north] at (\x/10,0) {0.\x}; }
           % \foreach \y in {0,1,...,9} { \node [anchor=east] at (0,\y/10) {0.\y};}
           %  % upto here
            \draw [red, dashed] (0.13, 0.50) rectangle (0.405, 0.645);
        \end{scope}
    \end{tikzpicture}
    \caption{(a) Results from the most parsimonious movement model. Negative
    coefficients indicate avoidance of a covariate, positive coefficients
    selection of a covariate. Whiskers delineate the 95\%-CIs for estimated
    parameters. The intercept is meaningless in the iSSF framework and was thus
    omitted. (b) Corresponding surface plots for two-way interactions. Light
    areas depict a combination of covariates that were preferred by dispersers,
    whereas darker areas correspond to avoided combinations. The absolute values
    along the axes have no meaning as they were normalized to 0 (low) and 1
    (high).}
    \label{MovementModel}
  \end{center}
\end{figure}

\newpage
\noindent \Cref{Simulations}a shows the outcome of the dispersal simulations.
For better visibility of high-frequency routes, we rasterized all simulated
trajectories and computed a heatmap depicting areas that were crossed by many
simulated trajectories. Unsurprisingly, we found that the dispersal trajectories
clearly reflected preferences that we identified earlier. For instance, most
trajectories were quite directional, avoided crossing a waterbody, ran along
waterbodies and avoided human-dominated areas. More interestingly, however, the
resulting heatmap closely resembled our permeability surface (see
\Cref{Simulations}b). That is, areas providing high permeability were typically
crossed by more simulated trajectories. Yet, the heatmap also showed that some
regions are almost impossible to reach, even if they provide highly permeable
landscapes. For example, although the western section of the Okavango Delta
holds highly permeable landscapes, no simulated trajectory reached these areas.
In contrast, many trajectories ventured towards east, some of them even crossing
the border to Zimbabwe. The reason that no trajectory reached the eastern part
of the Delta is likely owed to severe human influence and an impermeable flood
that together posed an insurmountable barrier and prevented dispersers from
traversing the landscape.

\begin{figure}[h]
  \begin{center}
    \includegraphics[width = 0.83\textwidth]{99_Simulations.pdf}
    \caption{(a) Heatmap of 1000 simulated dispersal events departing from the
    core study area (white square) in northern Botswana (countries with dashed
    white lines). (b) Permeability map for the same extent, showing that
    simulated dispersal trajectories typically cover highly permeable
    landscapes.}
    \label{Simulations}
  \end{center}
\end{figure}

\newpage
\begin{table}[h]
  \caption{Results from the forward model selection procedure based on AIC
  \citep{Burnham.2002} for our movement model. Weights were calculated using:
  \(\text{AIC-Weight}_i = \frac{e^{-0.5} \Delta \text{AIC}_i} {\sum_{i = 1}^n
  e^{-0.5} \Delta \text{AIC}_i} \). (*) represents all covariates from the most
  parsimonious permeability model, i.e. \(cos(ta)\), \(log(sl\)),
  \textit{Water}, \textit{DistanceToWater}, \textit{Shrubs/Grassland},
  \textit{HumansBuff5000}, and \textit{Trees}. To keep the table short we only
  included models with positive AIC weights. We abstained from averaging models
  with AIC-weight \(>\) 0 because we attempted to keep our simulation simple.}
  \label{ModelAICs2}
  \begin{center}
    \resizebox{\textwidth}{!}{
      \begin{tabular}{llllll}
      \toprule
      \textbf{Covariates} &
        \textbf{AIC} &
          \textbf{\(\Delta\)AIC} &
            \textbf{AIC-Weight} &
              \textbf{LogLik} \\
      \midrule
      \rowcolor[gray]{0.80}
        (*),
        log(sl):Water, cos(ta):DistanceToWater, cos(ta):HumansBuff5000 &
          89814.19 &
            0.00 &
              0.13 &
                -44889.10 \\
        (*), log(sl):Water, cos(ta):DistanceToWater, cos(ta):Trees &
          89814.37 &
            0.17 & 0.12 &
              -44889.18 \\
        (*), log(sl):Water, cos(ta):DistanceToWater &
          89814.46 &
            0.27 &
              0.11 &
                -44890.23 \\
        (*), log(sl):Water, cos(ta):DistanceToWater, cos(ta):HumansBuff5000, cos(ta):Trees &
          89814.87 &
            0.68 &
              0.09 &
                -44888.44 \\
        (*), log(sl):Water, cos(ta):DistanceToWater, cos(ta):HumansBuff5000, cos(ta):Shrubs &
          89815.32 &
            1.13 &
              0.07 &
                -44888.66 \\
        (*), log(sl):Water, cos(ta):DistanceToWater, cos(ta):Shrubs &
          89815.34 &
            1.15 &
              0.07 &
                -44889.67 \\
        (*), log(sl):Water, cos(ta):DistanceToWater, cos(ta):HumansBuff5000, log(sl):Shrubs &
          89815.40 &
            1.20 &
              0.07 &
                -44888.70 \\
        (*), log(sl):Water, cos(ta):DistanceToWater, log(sl):Shrubs &
          89815.67 &
            1.48 &
              0.06 &
                -44889.84 \\
        (*), log(sl):Water, cos(ta):DistanceToWater, cos(ta):HumansBuff5000, log(sl):HumansBuff5000 &
          89815.94 &
            1.75 &
              0.05 &
                -44888.97 \\
        (*), log(sl):Water, cos(ta):DistanceToWater, cos(ta):HumansBuff5000, cos(ta):Trees, log(sl):Shrubs &
          89816.08 &
            1.89 &
              0.05 &
                -44888.04 \\
        (*), log(sl):Water, cos(ta):DistanceToWater, cos(ta):HumansBuff5000, log(sl):DistanceToWater &
          89816.14 &
            1.95 &
              0.05 &
                -44889.07 \\
        (*), log(sl):Water, cos(ta):DistanceToWater, cos(ta):HumansBuff5000, cos(ta):Water &
          89816.16 &
            1.97 &
              0.05 &
                -44889.08 \\
        (*), log(sl):Water, cos(ta):DistanceToWater, log(sl):HumansBuff5000 &
          89816.18 &
            1.99 &
              0.05 &
                -44890.09 \\
        (*), log(sl):Water, cos(ta):DistanceToWater, cos(ta):HumansBuff5000, log(sl):Trees &
          89816.19 &
            2.00 &
              0.05 &
                -44889.10 \\
      \bottomrule
      \end{tabular}
    }
  \end{center}
\end{table}

\begin{table}[h]
  \caption{Selection coefficients as derived from the most parsimonious movement
  model. The movement model was a modified permeability model to which we added
  additional interaction terms between movement metrics and environmental
  covariates (e.g. \(Water:cost(ta\_)\)) in order to learn about how movement
  behavior changes depending on the environmental covariates.}
  \label{MovementModelResults}
  \begin{center}
    \resizebox{0.5\textwidth}{!}{
      \begin{tabular}{llllll}
      \toprule
      Covariate & Coefficient & SE & p-value \\
      \midrule
      cos(ta) & 0.149 & 0.045 & 0.001 \\
      log(sl) & 0.053 & 0.019 & 0.005 \\
      Water & -0.265 & 0.090 & 0.003 \\
      DistanceToWater & -0.272 & 0.226 & 0.230 \\
      Shrubs & 0.391 & 0.082 & 0.000 \\
      HumansBuff5000 & -0.466 & 0.203 & 0.022 \\
      Trees & -0.177 & 0.072 & 0.013 \\
      log(sl):Water & -0.040 & 0.008 & 0.000 \\
      cos(ta):DistanceToWater & 0.115 & 0.038 & 0.003 \\
      cos(ta):HumansBuff5000 & -0.047 & 0.031 & 0.134 \\
       \bottomrule
      \end{tabular}
    }
  \end{center}
\end{table}

\newpage
\subsection{Social Model}
\label{Appendix:SocialLandscape}
\subsubsection{Methods}
We used GPS relocation data of resident wild dogs and calculated a
geo-referenced stack of layers to which we refer as the \textit{social
landscape}. The social landscape was a collection of raster-layers, each
depicting the probability of encountering a resident group in a given pixel and
month. The social landscape was updated every month, using a moving window that
took into account six months of relocation data. For instance, the social
landscape dated July 1\textsuperscript{st} consisted of GPS relocation data
collected between January 1\textsuperscript{st} and June 30\textsuperscript{th}.
Because we collected relocation data of residents only inside our core study
area, we modeled the social landscape only for this region. GPS relocations were
collected at a 24-hourly interval during residence, so we kept one GPS
relocation per pack and day. To compute monthly updated social landscapes, we
calculated each resident pack's utilization distribution (UD) and corresponding
95\% home range (HR) using the kernelUD function from the R-package
\textit{adehabitatHR} \citep{Calenge.2019}. The package required to set a
smoothing parameter \(h\), which determined the sharpness of the resulting UD
and 95\%-HR. To come up with a meaningful smoothing parameter \(h\), we followed
the procedure described in \cite{Cozzi.2018}. That is, we first applied the
reference value \( h_{ref} \), which the package automatically determined. We
then obtained the corresponding UD and 95\%-HR and assessed whether the 95\%-HR
comprised of a single polygon or whether it split up into multiple polygons. In
case the 95\%-HR segregated into multiple polygons, we assumed that the
smoothing parameter was set to strictly. On the other hand, if the 95\%-HR
comprised of a single polygon, we assumed that we could further tighten the
smoothing parameter. We thus iteratively adjusted the smoothing parameter until
a value right before the 95\%-HR split up into multiple polygons. However, we
maximally varied \( h \) within 30\% and 120\% of the original \( h_{ref} \) to
avoid extreme over- or under-smoothing. Using the resulting \(h\) we computed
the final UD and corresponding 95\%-HR. Each pack's UD of a given month was
converted to a spatial probability density using \(f(x) =
\frac{x_i}{\sum_1^n{x_i}}\), so that pixel-values summed up to one. We then used
\Cref{EQ4} to merge all UDs of the same month and to calculate the probability
of not encountering any wild dogs in a given pixel and month.

\begin{equation}
\label{EQ4}
x_i = 1 - \prod_{j = 1}^{n} (1 - x_{ij})
\end{equation}

\noindent where \(x_i =\) is the probability of encountering another pack in
pixel \(i\) and \(x_{ij} =\) the probability of encountering pack \(j\) in pixel
\(i\). Finally, we inverted this probability using \((1-x_i)\) to get the
probability of encountering at least one resident pack in a given pixel and
month. Note that we kept track of each packs' 95\%-HR throughout all months,
which allowed us to identify through which HRs dispersers moved during
dispersal. An example of the final layers is given in
\Cref{SocialLandscapeExample}.

\begin{figure}[h]
  \begin{center}
    \includegraphics[width = 0.95\textwidth]{99_SocialLandscapeExample.pdf}
    \caption{(a) Example of a layer depicting the probability of encountering a
    resident pack in July 2017. The layer was derived from utility distributions
    of all packs residing in the core study area. (b) Example of a layer
    depicting all 95\% home ranges of the packs residing in the core study area.
    This layer served to identify whether a disperser still moved through his
    natal pack's home range, or whether it moved through a foreign packs home
    range. The layers in (a) and (b) were updated each month to take into
    account only the latest GPS relocation data of resident individuals.}
    \label{SocialLandscapeExample}
  \end{center}
\end{figure}

\noindent To fit a social model using the newly created social landscape, we
reran our iSSF described in \Cref{Modeling}. However, this time we discarded any
stratum that did not fully lie within the core study area. The social model that
we fitted was again a modification of the most parsimonious permeability model.
Despite the covariates contained in the permeability model, we added the social
covariates listed in \Cref{SocialCovars}. Because we hypothesized that the
preference for encountering other wild dogs was different within and outside the
disperser's natal home range, we included the interaction terms
\textit{ProbEncounter:OwnHomerange} and \textit{ProbEncounter:ForeignHomeranges}
in the social model. Since we did not anticipate to use the model for
predictions, we did not engage in model selection.

\begin{table}[h]
  \begin{center}
    \caption{Overview of extracted social covariates. Values were extracted from
    the layer that was closest in date to the actual step.}
    \label{SocialCovars}
    \resizebox{\textwidth}{!} {
      \begin{tabular}{lllll}
      \hline
      \textbf{Category} &
        \textbf{Covariate} &
          \textbf{Description} &
            \textbf{Values} \\
              % \textbf{Source(s)} \\
      \midrule
      Social Features
        & Own Homerange
          & Percentage cover of natal pack’s 95\% home-range along step
            & 0-100\% \\
      Social Features
        & ForeignHomeranges
          & Number of foreign packs’ 95\% home-ranges along step
            & \(\geq\) 0 \\
      Social Features
        & ProbEncounter
          & Average probability of encountering conspecifics along step
            & 0-100\% \\
      \hline
      \end{tabular}
    }
  \end{center}
\end{table}

\newpage
\subsubsection{Results}
Results from our social model are depicted in \Cref{SocialModel}a (see also
\Cref{SocialModelResults}). Again, parameter estimates for the covariates
contained in the permeability model did not severely change and we will only
report on the newly added social covariates (see red box in
\Cref{SocialModel}a). Visual inspection of the corresponding surface plot (see
\Cref{SocialModel}b1) revealed that dispersers tried to maximize the probability
of encounter when moving outside their natal pack's HR. Yet, this preference was
even stronger when dispersers moved inside their natal pack's HR. A second
surface plot (see \Cref{SocialModel}b2) showed that dispersers maximized the
probability of encounter when the number of foreign HRs was low, but that they
minimized the probability of encounter when the number of foreign HRs was high.
In summary, dispersers appeared to be drawn towards other wild dogs, but only if
the number of overlapping foreign HRs was low.

\begin{figure}[hbtp]
  \begin{center}
    \begin{tikzpicture}
        \node[anchor=south west,inner sep=0] (image) at (0,0,0) {
          \begin{minipage}{0.95\textwidth}
            \begin{center}
              \includegraphics[width = 0.75\textwidth]
                {99_SocialModel.pdf}\\
              \includegraphics[width = 0.8\textwidth]
                {99_SocialModel(Interactions).pdf}
            \end{center}
          \end{minipage}
        };
        \begin{scope}[x={(image.south east)},y={(image.north west)}]
           %  % next four lines will help you to locate the point needed by
           %  % forming a grid. comment these four lines in the final picture.↓
           % \draw[help lines,xstep=.1,ystep=.1] (0,0) grid (1,1);
           % \draw[help lines,xstep=.05,ystep=.05] (0,0) grid (1,1);
           % \foreach \x in {0,1,...,9} { \node [anchor=north] at (\x/10,0) {0.\x}; }
           % \foreach \y in {0,1,...,9} { \node [anchor=east] at (0,\y/10) {0.\y};}
           %  % upto here
            \draw [red, dashed] (0.15, 0.60) rectangle (0.43, 0.75);
        \end{scope}
    \end{tikzpicture}
    \caption{(a) Estimated parameters from the social model. Negative
    coefficients indicate avoidance of a covariate, positive coefficients
    selection of a covariate. Whiskers delineate the 95\%-CIs for estimated
    parameters. The intercept is meaningless in the iSSF framework and was thus
    omitted. (b) Corresponding surface plots for two-way interactions (red box).
    Light areas in the surface plots depict preferred combinations of the
    covariates, whereas darker areas correspond to avoided combinations. The
    absolute values along the axes of the surface plots have no meaning as they
    were scaled. Rather, low values along the axis correspond to low values of
    the respective covariate. The same is true for the scalebar.}
    \label{SocialModel}
  \end{center}
\end{figure}

\newpage
\begin{table}[h]
  \caption{Selection coefficients as derived from the social model. The social
  model was a modified permeability model to which we added additional social
  covariates in order to learn about how the social landscape infleunces
  dispersal behavior.}
  \label{SocialModelResults}
  \begin{center}
    \resizebox{0.6\textwidth}{!}{
      \begin{tabular}{llllll}
        \toprule
        Covariate & Coefficient & SE & p-value \\
        \midrule
        cos(ta) & 0.148 & 0.066 & 0.025 \\
        log(sl) & 0.055 & 0.019 & 0.003 \\
        Water & -0.472 & 0.113 & 0.000 \\
        DistanceToWater & -0.193 & 0.411 & 0.639 \\
        Shrubs & 0.470 & 0.110 & 0.000 \\
        HumansBuff5000 & -1.013 & 0.295 & 0.001 \\
        Trees & -0.465 & 0.193 & 0.016 \\
        ProbEncounter & 0.724 & 0.268 & 0.007 \\
        OwnHomerange & 0.056 & 0.095 & 0.557 \\
        ForeignHomeranges & -0.099 & 0.132 & 0.452 \\
        ProbEncounter:OwnHomerange & 0.189 & 0.179 & 0.291 \\
        ProbEncounter:ForeignHomeranges & -0.504 & 0.281 & 0.072 \\
        \bottomrule
      \end{tabular}
    }
  \end{center}
\end{table}

\newpage
\subsection{Statutory Declaration}
I assure that this thesis is a result of my personal work and that no other than
the indicated aids have been used for its completion. Furthermore, I assure that
all quotations and statements that have been inferred literally or in a general
manner from published or unpublished writings are marked as such. Beyond this, I
assure that the work has not been used, neither completely nor in parts, to pass
any previous examination.\\
\vspace{15mm}\\
Zürich, \today

\vspace{5mm}
\includegraphics[width = 0.2\textwidth]{Signature.png}

\noindent David Hofmann

\end{document}