### 2020-08-21 Chapter_0/04_Manuscript/Manuscript_Backup_2020-06-05.tex

\documentclass[abstract=on,10pt,a4paper,bibliography=totocnumbered]{article}
\usepackage[paper=a4paper,left=35mm,right=35mm,top=25mm,bottom=30mm]{geometry}
\usepackage[doublespacing]{setspace}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[round]{natbib}
\usepackage{amsmath}
\usepackage{colortbl}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{gensymb}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage[hidelinks]{hyperref}
\usepackage[nameinlink]{cleveref}
% \usepackage{lineno}
\usepackage{multirow}
\usepackage{arydshln}
\usepackage[nomarkers, nolists]{endfloat}

%------------------------------------------------------------------------------
%	Some Styling
%------------------------------------------------------------------------------
% Creating some TikZ styles
\tikzset{
  nonterminal/.style = {rectangle
    , minimum size = 6mm
    , very thick
    , draw = black!
  }
}

% Changing the style of captions in figures etc.
\captionsetup{labelfont=bf, format=plain, font=small}

%------------------------------------------------------------------------------
%	Titlepage: Header
%------------------------------------------------------------------------------
\title{Bound within Boundaries: How Well Do Boundaries of Protected Areas Match
Movement Corridors of Their Most Mobile Protected Species?}

% List of Authors
\author{
  David Hofmann\textsuperscript{1\S*} \and
  Dominik Behr\textsuperscript{1,2*} \and
  John W. McNutt\textsuperscript{2} \and
  Arpat Ozgul\textsuperscript{1} \and
  Gabriele Cozzi\textsuperscript{1,2}
}

% Reduce spacing between authors
\makeatletter
\def\and{%
  \end{tabular}%
  \hskip -0.5em \@plus.17fil\relax
  \begin{tabular}[t]{c}}
\makeatother

% Current Date
\date{\today}

% And here the masterpiece begins
\begin{document}

% Change page numbering
\pagenumbering{gobble}

% Required to be able to cite
\bibliographystyle{apalike}

% Create Titlepage
\maketitle

%------------------------------------------------------------------------------
%	Titlepage: Additional Info
%------------------------------------------------------------------------------
\begin{flushleft}

\vspace{0.5cm}

\textsuperscript{1} Department of Evolutionary Biology and Environmental
Studies, University of Zurich, Winterthurerstarsse 190, 8057 Zurich,
Switzerland.

\textsuperscript{2} Botswana Predator Conservation Trust, Private Bag 13, Maun,
Botswana.

\textsuperscript{\S} Corresponding author (david.hofmann2@uzh.ch)

\textsuperscript{*} Shared co-authorship

\vspace{4cm}

\textbf{Running Title:} Connectivity across a Transfrontier Conservation Area.

\vspace{0.5cm}

\textbf{Keywords:} Lycaon pictus, KAZA-TFCA, permeability surface, least-cost
corridors, integrated step selection function, landscape connectivity

\end{flushleft}

%------------------------------------------------------------------------------
%	Abstract
%------------------------------------------------------------------------------
\newpage
\begin{abstract}
Protecting large portions of natural land connecting natural reserves is an
increasingly used strategy to maintain and restore connectivity among wildlife
subpopulations. One such initiative area is the Kavango-Zambezi Transfrontier
Conservation Area (KAZA-TFCA) in Africa, the world's largest transboundary
conservation area on earth, spanning five countries and 520'000
km\textsuperscript{2}. While boundaries of such conservation areas are often
determined by expert opinion as well as socio-political needs and constraints,
the extent to which these areas match species' movement corridors is rarely
examined. This is mainly due to a lack of adequate empirical data, particularly
on dispersal. Between 2011 and 2019, we collected high-resolution GPS relocation
data on 16 dispersing African wild dog coalitions (\textit{Lycaon pictus}) to
(i) measure differences in landscape permeability within and outside the
boundaries of the KAZA-TFCA and (ii) to assess to which degree the area of the
KAZA-TFCA reflects natural movement corridors of this highly mobile and
endangered species. We show that permeability within the KAZA-TFCA boundaries
was more than double compared to areas outside it. Furthermore, we identify a
five-fold permeability difference among the five KAZA-TFCA countries. These
differences were mainly caused by different degrees of human activities,
hampering dispersal movements. We further show that all major movement corridors
of wild dogs run within the KAZA-TFCA boundaries, although some minor corridors
remain outside formally protected areas. Our work highlights the importance of
empirically assessing the adequacy of protected areas and the need for
international coordination that stem from different contributions to
connectivity across regions and countries.
\end{abstract}

%------------------------------------------------------------------------------
%	Main Text
%------------------------------------------------------------------------------
\newpage

% Change page numbering
\pagenumbering{arabic}

% Create linenumbers
% \linenumbers

\section{Introduction}
As a result of the ongoing human-caused habitat deterioration and fragmentation,
successful management and conservation of isolated subpopulations requires
maintenance and enhancement of connectivity between them \citep{Hanski.1998,
Fahrig.2003, Lindenmayer.2013}. Recently, efforts have been made to facilitate
animal movements and dispersal, and to secure connectivity by protecting
contiguous natural or semi-natural areas \citep{Heller.2009, Doerr.2011,
Rudnick.2012, Cozzi.2020}. These efforts have resulted in the creation of an
ever growing number of large and often transboundary protected areas
\citep{Wolmer.2003}. While boundaries of large scale conservation areas have
been drawn largely based on expert opinion and socio-political needs and
constraints, the usefulness of subjective boundary assessments have revealed
deficiencies in the past \citep{Pullinger.2010}. Due to the very nature of large
scale protected areas, which sometimes span several thousands of square
kilometers and traverse multiple countries, a thorough quantitative assessment
of their effectiveness and adequacy of their boundaries is challenging
\citep{Rudnick.2012}. This challenge is complicated by a lack of animal movement
data at the appropriate spatial scale and an insufficient understanding of the
ranging behavior, movement patterns, and dispersal corridors of the species for
which these protected areas have been created \citep{Vasudev.2015, Tshipa.2017,
Cozzi.2020}. Nevertheless, an accurate representation of movement corridors and
dispersal routes based on empirical global positioning system (GPS) data is
fundamental to objectively assess how well boundaries of protected areas reflect
natural animal movements \citep{Beier.2008, Doerr.2011, Rudnick.2012}.

In recent years, there has been a growing body of research that uses animal
relocation data to identify movement corridors and assess connectivity at large
scales \citep{Chetkiewicz.2006, Doerr.2011, Squires.2013, Elliot.2014,
Benz.2016, Osipova.2019}. Identification of potential corridors relies on the
estimation of permeability (respectively resistance) surfaces, which return the
ease or willingness at which the focal species traverses a specific landscape
\citep{Sawyer.2011}. Such surfaces are typically created on the basis of habitat
preferences, which can be estimated using a suite of selection functions
\citep{Boyce.2002, Fortin.2005, Cushman.2010}. In selection functions, habitat
preferences are inferred by comparing spatial covariates (e.g. environmental and
anthropogenic) at locations visited by the animal to the same spatial covariates
at randomly selected locations in the animal's availability domain
\citep{Zeller.2012, Thurfjell.2014}. Importantly, selection functions rely on
adequate landscape and GPS relocation data that are representative of the
process being studied \citep{Diniz.2020}. More specifically, GPS relocation data
collected on dispersing individuals has been shown to outperform GPS relocation
data collected on resident individuals when aiming at detecting large scale
dispersal corridors \citep{Elliot.2014, Diniz.2020}. However, few connectivity
studies have partitioned their GPS relocation data according to behavioral modes
(e.g. resident vs. dispersing) of the studied species \citep{Wilson.2012,
Vasudev.2015}. As such, most permeability surfaces upon which dispersal
corridors are identified are created using GPS relocation data that were
collected on resident individuals. This introduces severe biases and
substantially reduces the power to reveal meaningful movement corridors, for
dispersing individuals have different needs and drives compared to resident
individuals \citep{Killeen.2014, Elliot.2014, Cozzi.2020}. Consequently, these
biases have limited our ability to assess the adequacy of the boundaries of
protected areas in securing connectivity for their protected species.

One initiative that aims at restoring and enhancing connectivity across large
scales is the Kavango-Zambezi Transfrontier Conservation Area (KAZA-TFCA),
which, spanning over 520'000 km\textsuperscript{2} and five countries,
constitutes the world's largest transfrontier conservation area. While the
KAZA-TFCA was originally set to facilitate movements of African elephants
(\textit{Loxodonta africana}; \citealp{Tshipa.2017}), it is also key to the
conservation of other wide-ranging species such as African wild dogs
(\textit{Lycaon pictus}; \citealp{Woodroffe.2012, Cozzi.2020}), lions
(\textit{Panthera leo}; \citealp{Elliot.2014, Cushman.2018}), and cheetahs
(\textit{Acinonyx jubatus}; \citealp{Weise.2017}). To date, however, few studies
have attempted to assess the adequacy of the KAZA-TFCA's boundaries using
relevant GPS movement data at the appropriate spatial scale \citep{Elliot.2014,
Tshipa.2017}. Thus, how well the KAZA-TFCA's boundaries reflect natural movement
patterns and dispersal corridors of its protected species is virtually unknown.

Across the KAZA-TFCA, the African wild dog (\textit{Lycaon pictus}) represents a
highly endangered flagship species for conservation efforts, but for which
systematically collected GPS relocation data on dispersing individuals are
unavailable. Once widespread across the entire Sub-Saharan continent, wild
dogs have been widely extirpated through human persecution, habitat destruction,
and disease outbreaks \citep{Woodroffe.2012}. As a result, the species has
become one of Africa's most endangered large carnivores, and currently only
survives in small, spatially scattered subpopulations \citep{Woodroffe.2012}.
Within these subpopulations, wild dogs form cooperative breeding packs of up to
thirty individuals \citep{Frame.1979, Fuller.1992, Creel.2002}, whose social
structure is strongly shaped by the process of dispersal \citep{McNutt.1996,
Woodroffe.2019, Behr.2020}. Both males and females typically disperse from their
natal pack, either alone or in same-sex dispersing coalitions, and search for
unrelated mates and a suitable territory to settle \citep{McNutt.1996,
Cozzi.2020, Behr.2020}. During dispersal, wild dogs can cover several hundred
kilometers, sometimes crossing national boundaries and areas that are strongly
influenced by humans \citep{DaviesMostert.2012, Masenga.2016, Woodroffe.2019,
Cozzi.2020}. Despite the importance of dispersal for the long-term viability of
this species, little empirical information is available on habitat selection and
potential movement barriers of dispersing wild dogs. The few studies that have
collected GPS relocation data during dispersal have shown that dispersers
quickly move over large distances \citep{Woodroffe.2019}, generally avoid
human-dominated landscapes \citep{Masenga.2016, Oneill.2020, Cozzi.2020} and
areas densely forested \citep{Oneill.2020}, but prefer proximity to water and
dirt roads \citep{Oneill.2020}.

Here, we used GPS relocation data collected on 16 dispersing wild dogs in as
many dispersing coalitions from a free-ranging population in northern Botswana
to assess the adequacy of the boundaries of the KAZA-TFCA, the largest protected
area in the world. We estimated the degree of selection or avoidance for
environmental and anthropogenic landscape features. We then applied the obtained
habitat preferences to create a permeability surface spanning the entire
KAZA-TFCA and investigated how landscape permeability varies regionally and
internationally. Finally, we calculated least-cost paths and least-cost
corridors to identify major dispersal corridors, pinpoint dispersal hubs, and
verify whether the KAZA-TFCA successfully covers these features.

\section{Methods}
\subsection{Study Area}
The study area (centered at -17\degree 13'9''S, 23\degree 56'4''E; elevation ca.
1'030 m; red dotted rectangle in \Cref{StudyArea}a) was outlined by a
rectangular bounding box stretching over 1.3 Mio km\textsuperscript{2} and
encompassing the entire KAZA-TFCA (\Cref{StudyArea}b). The KAZA-TFCA lies in the
basins of the Okavango and Zambezi rivers and includes parts of Angola,
Botswana, Namibia, Zimbabwe, and Zambia. With a total area of over 520'000
km\textsuperscript{2} it covers several already-existing national parks and
other protected areas. It constitutes the earth's largest transboundary
conservation area and is characterized by diverse landscapes, including savanna,
grassland, and dry or moist woodland habitats. Rainfall in the study area is
seasonal and lasts from November to March, whereas the rest of the year remains
dry \citep{Mendelsohn.2010}. The KAZA-TFCA also encloses the Okavango Delta,
which is one of the main strongholds of African wild dogs in southern Africa and
likely acts as a source for the recolonization of surrounding habitats
\citep{Woodroffe.2012, Cozzi.2013}. The Okavango Delta represents a highly
dynamic hydrological flood-pulsing system \citep{McNutt.1996, Wolski.2017}. The
extent of the flood in the delta greatly changes within and between years
depending on the amount of rain that descends from the catchment areas in Angola
and reaches the distal ends of the delta between July and August (see Figure S5
for an illustration of a typical flood pulse). The flood drastically affects
surrounding landscapes, so that during maximum extent (ca. 12'000
km\textsuperscript{2}) the delta becomes a patchy conglomerate of swamps, open
water, and islands, whereas these structures run dry when the flood retracts to
its minimum extent (ca. 5'000 km\textsuperscript{2}; \citealp{Gumbricht.2004,
Wolski.2017, McCarthy.1998}). Despite 36 NPs and other protected areas, there is
considerable human influence in some regions of the KAZA-TFCA, mainly
originating from farms, high human density, and road traffic.

\begin{figure}[h]
  \begin{center}
    \begin{tikzpicture}
        \node[anchor=south west,inner sep=0] (image) at (0,0,0) {
        \includegraphics[width=\textwidth]{99_StudyArea.pdf}
        };
        \begin{scope}[x={(image.south east)},y={(image.north west)}]
            % % next four lines will help you to locate the point needed by forming a grid.
            % % comment these four lines in the final picture.
            % \draw[help lines,xstep=.1,ystep=.1] (0,0) grid (1,1);
            % \draw[help lines,xstep=.05,ystep=.05] (0,0) grid (1,1);
            % \foreach \x in {0,1,...,9} { \node [anchor=north] at (\x/10,0) {0.\x}; }
            % \foreach \y in {0,1,...,9} { \node [anchor=east] at (0,\y/10) {0.\y};}
            % % upto here
            \draw[black] (0.202, 0.255) -- (0.36, 0.955);
            \draw[black] (0.202, 0.205) -- (0.36, 0.070);
        \end{scope}
    \end{tikzpicture}
    \caption{Overview of our study area. (a) The red dotted rectangle depicts
    the study area, which was confined by a bounding box encompassing the entire
    KAZA-TFCA. Gray areas indicate remaining wild dog populations according to
    the IUCN \citep{Woodroffe.2012}. (b) The white rectangle illustrates the
    area within which dispersal coalitions were collared. Since Game Reserves in
    Botswana virtually serve the same purpose as National Parks, we use the
    terms interchangeably for this region.}
    \label{StudyArea}
  \end{center}
\end{figure}

\subsection{GPS Relocation Data}
We used a population of free-ranging African wild dogs inhabiting the Okavango
Delta in northern Botswana as a source population for dispersing individuals.
This population has been extensively studied since 1989 \citep{McNutt.1996,
McNutt.2008, Cozzi.2013, Cozzi.2020, Behr.2020}. Between 2011 and 2019, we
collected GPS relocation data on 16 coalitions of dispersing African wild dogs
(7 female and 9 male coalitions; see also \cite{Abrahms.2017} and
\cite{Cozzi.2020}). Candidate dispersing individuals were identified based on
criteria reported in \cite{Behr.2020}, immobilized according to protocols
described in \cite{Osofsky.1996}, and fitted with GPS/Satellite radio collars
(\textit{Vertex Lite; Vectronic Aerospace GmbH, Berlin}) while still in their
natal pack. All procedures were undertaken and supervised by a
Botswana-registered wildlife veterinarian. Immobilized individuals were
monitored and all were observed to successfully re-join their packs within one
hour after immobilization. During dispersal, GPS collars were programmed to
record GPS relocation data every 4 hours and to regularly transmit them via
Iridium satellite system to a base station. The system thus allowed
remote-tracking of collared individuals, even where field conditions were
prohibitive or dispersal trajectories crossed international borders.

Because we were interested in dispersal behavior, we discarded any GPS data
collected while individuals were still with their natal packs and also after
settlement in a new territory after dispersal \citep{Cozzi.2020}. We identified
the exact time of emigration and settlement based on direct field observations
and through visual inspection of the net squared displacement (NSD) metric. NSD
is a metric returning the euclidean distance of a relocation to a reference
point \citep{Borger.2012}, which in our case was the center of the dispersing
coalition's natal home range. Thus, dispersal was deemed to have started when a
coalition had left its natal home range and continued until the NSD metric
remained stationary, implying that the coalition had settled (Figure S1). Some
coalitions undertook exploratory movements prior to proper dispersal. Such
movements have been observed in other species and are reported to be very
similar to proper dispersal \citep{Killeen.2014}. Therefore, we also classified
exploratory movements as dispersal. In total, we collected 4'169 GPS relocations
during dispersal, i.e. between emigration and settlement (Figure S2 \& Table
S1), resulting in an average of 260 locations per dispersing coalition (range:
37 - 729). On average, coalitions dispersed for 48 days (min = 9, max = 137),
covered a mean euclidean distance of 54 km (min = 3, max = 263) and a mean
cumulative distance of 596 km (min = 130, max = 1'962). In our analysis, we did
not differentiate between male and female dispersing coalitions, for previous
research found little differences between sexes during dispersal
\citep{Woodroffe.2019, Cozzi.2020}.

\subsection{Spatial Covariates}
To investigate habitat preferences of dispersing wild dogs, we used a set of
geo-referenced covariates (\Cref{Covariates}) that we aggregated in the
categories \textit{land cover}, \textit{protection status}, and \textit{human
influence}. We did not include terrain features as covariates due to the absence
of noteworthy elevational gradients in our study area. For each covariate, we
prepared spatial raster layers from freely available online services or from
remotely sensed satellite imagery. To ensure a consistent resolution (i.e.
cell-size or grain) across covariates, we coarsened or interpolated all layers
to match a resolution of 250m x 250m. We performed processing and manipulation
of data as well as all spatial and statistical analyses using R, version 3.6.1
\citep{R.2019}.

\begin{figure}[h]
  \begin{center}
    \includegraphics[width = \textwidth]{99_Covariates.pdf}
    \caption{Overview of spatial covariates that we included in our models. We
    prepared all covariates for the entire study area but for better visibility
    we only plot them for the surroundings of the Okavango Delta. The white
    rectangle in each plot depicts the area within which dispersal coalitions
    were collared. (a) Averaged layer of all dynamic (binary) water maps. (b)
    Percentage cover of trees. (c) Percentage cover of shrubs/grassland.
    Anything that was not covered by trees or shrubs/grassland was deemed to be
    bare land. (d) Protection status of the area. (e) Human influence proxy
    composed of human density, farms, and roads. (f) Distance to nearest road
    (white lines depict actual roads).}
    \label{Covariates}
  \end{center}
\end{figure}

\subsubsection{Land Cover}
We divided land cover into water and dryland. Water included rivers, wetlands,
and swamps. Because the inundation extent of the flood in the Okavango Delta is
highly variable within and between years, we created dynamic ``flood maps'' that
were updated each 8\textsuperscript{th} day following a remote sensing algorithm
developed by \citep{Wolski.2017}. The algorithm uses thresholding of short
wavelength infrared reflectances of MODIS Terra satellite imagery (MCD43A4;
\citealp{Schaaf.2015}) to distinguish areas covered by water and dryland
(details in Appendix A.3). While we created dynamic flood maps for the Okavango
Delta, we assumed the extent of all other water bodies (e.g. Chobe river,
Zambezi river) to be static within and between years. This static representation
was based on Globeland's land cover dataset \citep{Chen.2015}, from which we
only retained the categories \textit{wetland} and \textit{water bodies} and
collectively reclassified them to \textit{water}. Globeland had an original
resolution of 30m x 30m, so we coarsened the layer to 250m x 250m using the mode
of each 250m x 250m cell. We further improved river representation by employing
the rasterized MERIT Hydro dataset \citep{Yamazaki.2019} from which we added all
rivers with a width of over 10m to our Globeland layer. We merged dynamic and
static water maps into a large rasterstack, covering the entire study area. We
also created a rasterstack rendering the covariate \textit{distance to water} by
calculating the eculidean distance of each raster cell in the study area to the
nearest source of water.

We further subdivided dryland into three layers as derived from the MODIS Terra
Vegetation Continuous Fields dataset (MOD44B; \citealp{Dimiceli.2015}). The
three layers depicted percentage cover of tree-vegetation (henceforth
\textit{trees}), non-tree-vegetation (henceforth \textit{shrubs/grassland}), and
non-vegetated (henceforth \textit{bare land}) and added up to 100\% of dryland
coverage. We used our flood map that aligned with the creation date of these
MODIS layers and defined anything covered by water as 0\% vegetated. The MODIS
vegetation layers had a resolution of 250m x 250m and no coarsening or
interpolation was required.

\subsubsection{Protection Status}
We created a binary layer separating protected from unprotected land. We
downloaded corresponding data on protection status from the Peace Parks
Foundation (\url{www.peaceparks.org}; \citealp{PeaceParks.2019}). Protected
areas included forest reserves, game reserves, wildlife management areas, and
national parks. We classified anything not covered by these categories as
unprotected (e.g. communal pastoral land, private land). We rasterized the two
categories to the binary raster \textit{protection status} (1 = protected, 0 =
pastoral) with a resolution of 250m x 250m.

\subsubsection{Human Influence}
We created a raster layer representing human influence by integrating
information on (1) human density, (2) farming, and (3) roads.

\begin{enumerate}[label = (\arabic*)]

  \item We obtained spatial human density estimates through a publicly available
  30m x 30m high-resolution population density dataset
  (\url{www.dataforgood.fb.com}; \citealp{Facebook.2019}). We coarsened the
  layer to 250m x 250m by summing up human density values within each 250m x
  250m cell.

  \item We sourced spatial information on farms from the Globeland
  \citep{Chen.2015} and Cropland \citep{Xiong.2017} land cover datasets from
  which we retained areas that were classified as either \textit{cultivated
  land} or \textit{croplands}. Any other land cover class was not pertinent to
  farming and therefore omitted. Because both layers had a resolution of 30m x
  30m we coarsened them to 250m x 250m by assigning a value of 1 to any 250m x
  250m cell that covered farmland and a value 0 otherwise. Thus, the final layer
  depicted presence (= 1) or absence (= 0) of farms within each 250m x 250m
  cell.

  \item We obtained geo-referenced data on roads from Open Street Map
  \citep{OpenStreetMap.2019}, downloaded through Geofabrik
  (\url{www.geofabrik.de}). We only retained main tarmac roads and omitted
  smaller roads (Table S2) as these are scarcely frequented and do not represent
  an obstacle to wild dog movements \citep{Abrahms.2016}. We rasterized main
  tarmac roads to the binary raster \textit{roads} (1 = roads, 0 = no roads)
  with 250m x 250m resolution. Finally, we created the covariate
  \textit{distance to roads} by calculating the eculidean distance of each
  raster cell in the study area to the nearest road.

\end{enumerate}

\noindent Because layers (1), (2), and (3) depicted features that are typically
spatially clustered and because not all dispersing coalitions moved within
meaningful distance to each of these features, we totaled the layers
\textit{human density}, \textit{farming}, and \textit{roads}. This resulted in a
single covariate called \textit{human influence}. Note that totalling the layers
implied that the three covariates received different weights (see Appendix A.6
for details). To reduce the influence of outliers, totaled values were limited
to a maximum of 50, which visually resulted in a good balance between high and
low anthropogenic influence and was therefore considered appropriate for our
analysis. To render the fact that humans influence their surroundings beyond
their presence, we followed \cite{Elliot.2014} and applied to each raster cell a
5 km focal buffer within which we summed up and log-transformed human-influence
values (Figure S6).

\subsection{Habitat Selection Model}
We used an integrated step selection function (iSSF; \citealp{Avgar.2016}) to
investigate dispersers' selection or avoidance for the above-mentioned spatial
covariates. In the iSSF framework, GPS relocations are coerced to  ``steps'',
where a step is defined as the connecting line between two consecutive GPS
relocations \citep{Turchin.1998}. Subsequently, habitat preferences are
estimated by contrasting covariates along realized steps to the same covariates
along alternative ``random''  steps \citep{Fortin.2005, Thurfjell.2014,
Avgar.2016}. Thus, we paired each realized step with 24 random steps. We
generated random steps by sampling turning angles from a uniform distribution
U(\(-\pi, +\pi\)) and step lengths from a gamma distribution that was fitted
using realized steps \citep{Avgar.2016}. Together, a realized step and its 24
associated random steps formed a stratum of 25 steps that received a unique
identifier. Along all steps we extracted the above mentioned covariates (see
also Table S3). We also derived a binary covariate indicating whether or not a
step crossed a road and we identified the average distance of each step to the
nearest source of water and the nearest road. We square-rooted the values for
\textit{distance to roads} and \textit{distance to water} to render a decreasing
marginal impact of distance. Additionally, we derived a binary indicator of
whether a step crossed a road or not. We standardized extracted values using a
z-score transformation, and checked for correlation using Pearson's Correlation
Coefficient \(r\). None of the covariates were overly correlated (\(|r| > 0.6\);
\citealp{Latham.2011}) and we retained all of them for modeling. Our habitat
selection was then based on the assumption that dispersing wild dogs assigned a
selection score \(w(x)\) of the following exponential form to each realized and
random step \citep{Fortin.2005, Thurfjell.2014, Avgar.2016}:

\begin{equation}
\label{EQ1}
  w(x) = exp(\beta_1 x_1 + \beta_2 x_2 + ... + \beta_n x_n)
\end{equation}

\noindent The selection score \(w(x)\) of a step depended on its associated
covariates (\(x_1, x_2, ..., x_n\)), as well as on the animal's preferences for
these covariates (\(\beta_1, \beta_2, ..., \beta_n\)). To estimate habitat
preferences (i.e. the \(\beta\)'s) of interest, we used mixed effects
conditional logistic regression analysis as suggested by \cite{Muff.2020}. We
implemented their method using the R-package \textit{glmmTMB}
\citep{Mollie.2017} and used dispersing coalition ID to model random intercepts
and slopes. We defined two movement metrics, namely the cosine of the turning
angle (\(cos(ta)\)) and the log of the step length (\(log(sl)\)), as core
covariates and ran stepwise forward model selection based on Akaike's
Information Criterion (AIC; \citealp{Burnham.2002}) for all other covariates.
The inclusion of movement metrics served to reduce biases in estimated habitat
preferences that may have arised due to movement preferences \citep{Avgar.2016}.

To validate the predictive power of the most parsimonious habitat selection
model, we ran k-fold cross-validation for case-control studies as described in
\cite{Fortin.2009}. This validation procedure compares rank frequencies of
realized steps under observed and random preferences and proves a significant
prediction in case the confidence intervals of the spearman-rank correlation
coefficients \(\bar{r}_{s, realized}\) and \(\bar{r}_{s, random}\) do not
overlap (details in Appendix A.8).

\subsection{Permeability Surface}
Using the most parsimonious habitat selection model, we predicted a permeability
surface spanning the entire extent of the KAZA-TFCA. That is, we applied
\Cref{EQ1} to our spatial covariates and calculated the selection score \(w(x)\)
for each raster cell in our study area. Because our flood maps were dynamic, we
collapsed all flood maps into a single static layer using areas that were
covered by water in at least a quarter of all our flood maps. We also calculated
a corresponding layer indicating the euclidean distance of each raster cell to
the nearest source of water on the static water map. To reduce the influence of
outliers in predicted permeability scores we followed \cite{Squires.2013} and
curtailed predicted scores between the 1\textsuperscript{st} and
99\textsuperscript{th} percentile of their original values. To compare
permeability across different regions, we rescaled the permeability surface to a
range between 0 and 1. The most impermeable landscape thus received a value of
0, the most permeable landscape a value of 1. Based on the permeability surface
we calculated median permeability values within and outside the KAZA-TFCA,
within and outside formally protected areas, and within each of the five
KAZA-TFCA countries.

\subsection{Least-Cost Paths and Corridors}
To identify movement corridors of dispersing wild dogs, we specified source
points and calculated factorial least-cost paths (LCPs) as well as factorial
least-cost corridors (LCCs) among all source points (Figure S.7;
\citealp{Adriaensen.2003, Sawyer.2011, Elliot.2014}. We generated source points
by overlaying the study area with a regular grid of points spaced at 100 km. We
only considered those points that fell within protected areas \(>\) 700
km\textsuperscript{2} as source points, which conforms with home-range
requirements of African wild dogs reported in \cite{Pomilia.2015}. Finally, we
defined centroids as source points for those protected areas \(>\) 700
km\textsuperscript{2} that were not assigned any source points from the regular
grid. In total, we generated 68 source points, which resulted in 2'278 unique
pairwise combinations and therefore 2'278 unique LCPs and LCCs.

We implemented factorial LCP analysis between source points using the R-package
\textit{gdistance} (Figure S.7; \citealp{vanEtten.2017}). The package translated
the (unscaled) permeability surface into a network of nodes to find shortest
effective distances between source points based on probabilities of moving from
cell to cell. In our case, the transition probability of moving between two
adjacent cells depended on their averaged permeability. We allowed individuals
to move from each cell to the cell's eight surrounding neighbors (i.e. Moores
neighborhood) and applied a geographic correction to account for the fact that
diagonal neighbors were more remote than orthogonal neighbors. Because African
wild dogs have been observed to cover large dispersal distances
\citep{DaviesMostert.2012, Masenga.2016, Cozzi.2020}, we did not limit LCPs to a
maximal effective cost. After computation, we tallied overlapping LCPs and
identified high-frequency routes.

We also calculated factorial LCCs \citep{Pinto.2009, Sawyer.2011, Elliot.2014},
again using the R-package \textit{gdistance} (Figure S.7;
\citealp{vanEtten.2017}). To identify LCCs, we first computed for each source
point a cumulative cost map, which indicated the total minimal costs required to
get from the source point to any other location in the study area. We then
obtained an LCC between two source points by adding up their cumulative cost
maps and masking out all cell-values exceeding the lowest cell-value by more
than 5\% \citep{Pinto.2009}. We repeated this procedure for each possible unique
pairwise combination of source points and thereby identified LCCs between all 68
selected source points. We normalized the resulting corridor-maps to range from
zero to one and tallied them into a single connectivity map.

\section{Results}
\subsection{Habitat Selection Model}
Our most parsimonious habitat selection model (\(\Delta AIC > 2\) than any
alternative model; Table S4) retained the covariates \textit{water},
\textit{distance to water}, \textit{trees}, \textit{shrubs/grassland}, and
\textit{human influence}, beside the fixed covariates \textit{cos(ta)} and
\textit{log(sl)} (\Cref{PermeabilityResults}a). Parameter estimates showed that
dispersing wild dogs moved in a directional and fast manner, as indicated by a
positive selection for small turning angles, i.e. high \(cos(ta)\) (\(\beta\) =
0.14; 95\% CI = 0.07 to 0.21, \Cref{PermeabilityResults}a) and longer steps,
i.e. high \(log(sl)\) (\(\beta\) = 0.06, 95\% CI = 0.02 to 0.09,
\Cref{PermeabilityResults}a). Dispersers avoided moving through water (\(\beta\)
= -0.52, 95\% CI -0.77 to -0.26, \Cref{PermeabilityResults}a) but selected for
locations in its vicinity, although the latter effect was not significant
(\(\beta\) = -0.32, 95\% CI = -0.72 to 0.08, \Cref{PermeabilityResults}a).
Dispersers avoided areas that were densely covered by trees (\(\beta\) = -0.31,
CI = -0.46 to -0.15, \Cref{PermeabilityResults}a) and preferred areas covered by
shrubs/grassland (\(\beta\) = 0.25, 95\% CI = 0.07 to 0.42,
\Cref{PermeabilityResults}a). Finally, dispersers avoided areas that were
influenced by human presence and activities (\(\beta\) = -0.41, 95\% CI = -0.78
to -0.05, \Cref{PermeabilityResults}a).

\begin{figure}[h]
  \begin{center}
    \includegraphics[width = \textwidth]{99_PermeabilityResults.pdf}
    \caption{(a) Estimated selection coefficients from the most parsimonious
    habitat selection model. Negative coefficients indicate avoidance of a
    covariate, positive coefficients selection of a covariate. Whiskers
    delineate the 95\%-CIs for estimated parameters. (b) Results from the k-fold
    cross validation for case-control studies. The left graph shows rank
    frequencies of \textit{realized} steps according to predictions, whereas the
    right graph shows rank frequencies of \textit{randomly selected} steps
    according to predictions. \(\bar{r}_s\) indicates the mean correlation
    coefficient resulting from 100 repetitions of the k-fold cross validation.
    The blue smoothing line was fitted using a locally weighted polynomial
    regression and serves to aid the eye in detecting the trends. Correlation
    coefficients suggest that our prediction was significant and robust,
    evidenced by the fact that the confidence intervals of \(\bar{r}_{s,
    realized}\) and \(\bar{r}_{s, random}\) did overlap and by the fact that
    there was strong and significant correlation between ranks and associated
    frequency for realized steps.}
    \label{PermeabilityResults}
  \end{center}
\end{figure}

Results from the k-fold cross-validation suggested that our prediction was
significant and robust, as highlighted by the fact that the 95\%-CIs intervals
of \(\bar{r}_{s, realized}\) and \(\bar{r}_{s, random}\) did not overlap
(\Cref{PermeabilityResults}b). Likewise, the significant correlation between
ranks and corresponding frequencies for realized steps suggested a good fit
between predictions and observations (\Cref{PermeabilityResults}b).

\subsection{Permeability Surface}
Our prediction of landscape permeability revealed substantial differences across
regions in the study area (\Cref{PermeabilityMap}). Comparisons of median
permeability values (\Cref{PermeabilityComp}) showed that permeability inside
KAZA-TFCA is more than two times as high as permeability outside it.
Permeability varies by country and reveals a five-fold permeability difference
among them. Angola and Botswana are characterized by comparably highly permeable
landscapes, Zimbabwe and Zambia are relatively impermeable, and Namibia ranges
in between the two extremes (\Cref{PermeabilityComp}). Visual inspection of our
covariate layers indicated that high permeability in Botswana and Angola is
mainly caused by a combination of low human presence and activities, low tree
cover, high shrubs/grassland cover, and a close distance to water. Although
swamps, wetlands, and permanent water themselves provide little permeability,
their surroundings act as strong attractants to dispersers. The low permeability
that characterizes Zambia and Zimbabwe, on the other hand, is mainly caused by
substantial human influences. Albeit the KAZA-TFCA covers a majority of
permeability hot-spots, several highly permeable regions remain uncovered by its
borders. Across all countries, protected areas provide roughly double the
permeability of human-dominated landscapes (\Cref{PermeabilityComp}).

\begin{figure}[hbtp]
  \begin{center}
    \begin{tikzpicture}
        \node[anchor=south west,inner sep=0] (image) at (0,0,0) {
          \begin{minipage}{\textwidth}
            \includegraphics[width = \textwidth]{99_PermeabilityMap.pdf}
          \end{minipage}
        };
        \begin{scope}[x={(image.south east)},y={(image.north west)}]
        \end{scope}
    \end{tikzpicture}
    \caption{Predicted permeability surface for the extent of the KAZA-TFCA.
    Permeability was predicted by calculating selection scores \(w(x) =
    exp(\beta_1 x_1 + \beta_2 x_2 + ... + \beta_n x_n)\) for each raster cell
    based on the raster cell's underlying covariates (\(x_i\)) and estimated
    habitat preferences (\(\beta_i\)). Areas that dispersers find easy to
    traverse are depicted in bright colors. Bold white lines delineate the
    borders of the KAZA-TFCA, whereas dashed white lines show country borders.}
    \label{PermeabilityMap}
  \end{center}
\end{figure}

\begin{table}[h]
  \begin{center}
  \caption{Comparison of median permeability (interquantile range in brackets)
  across countries, separated into areas within and outside the KAZA-TFCA, as
  well as within and outside formally protected areas. High values indicate high
  permeability, whereas low values correspond to low permeability.}
  \label{PermeabilityComp}
  \begin{tabular}{llllll}
    \toprule
    \multicolumn{1}{c}{} & \multicolumn{2}{c}{KAZA-TFCA} &
    \multicolumn{2}{c}{Protection Status} & \multicolumn{1}{c}{} \\
    \cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
    Country & Inside & Outside & Protected & Pastoral & Overall\\
    \midrule
    Angola & 0.36 (0.41) & 0.12 (0.32) & 0.36 (0.41) & 0.12 (0.33) & 0.20 (0.39)\\
    Botswana & 0.25 (0.30) & 0.15 (0.16) & 0.28 (0.35) & 0.15 (0.18) & 0.19 (0.25)\\
    Namibia & 0.22 (0.30) & 0.13 (0.18) & 0.24 (0.30) & 0.11 (0.15) & 0.16 (0.24)\\
    Zambia & 0.05 (0.09) & 0.03 (0.06) & 0.04 (0.10) & 0.03 (0.05) & 0.03 (0.07)\\
    Zimbabwe & 0.07 (0.16) & 0.06 (0.05) & 0.08 (0.17) & 0.05 (0.05) & 0.06 (0.07)\\
    \hline
    Overall & 0.16 (0.30) & 0.07 (0.15) & 0.15 (0.30) & 0.07 (0.15) & 0.10 (0.22)\\
    \bottomrule
  \end{tabular}
  \end{center}
\end{table}

\subsection{Least-Cost Paths \& Least-Cost Corridors}
Our least-cost analysis revealed three major movement corridors of which all
were well-within the KAZA-TFCA boundaries (\Cref{LeastCost}). Furthermore, we
found little to no direct connectivity between peripheral points; that is, most
paths and corridors connecting two adjacent peripheral points run through more
central regions before heading towards their destination at the periphery
(\Cref{LeastCost}). One major corridor runs SE-NW and connects the
Okavango-Linyanti ecosystem in Botswana with Luengue-Luiana NP in Angola. A
further corridor runs W-E between Chobe NP in Botswana and Zimbabwe's Hwange NP.
A third major corridor runs NE-SW, completely across unprotected areas, and
connects Kafue NP in Zambia with more central regions of the KAZA-TFCA. Several
minor corridors branch off from these three major corridors; these include a
southward connection between the Okavango-Linyanti and the Central Kalahari Game
Reserve, a southwesterly corridor connecting Luengue-Luiana NP with Namibia's
Kaudom NP, and a northeasterly extension of the Hwange corridor into Zimbabwe's
Matusadona NP. According to our predictions, the corridor towards Kafue NP and
the corridor passing the Okavango-Linyanti region are the highest frequented
routes within KAZA (maybe numbers?). Our model did not detect any significant
direct corridors between Zimbabwe and Zambia or Zambia and Angola, and only a
very limited W-E direct connection between the Okavango region and Namibia's
Khaudum NP. Except for the corridor into the Central Kalahari National Park, our
model did not detect any significant connectivity outside the boundaries of
KAZA-TFCA.

\begin{figure}[hbtp]
  \begin{center}
    \begin{minipage}{0.95\textwidth}
      \includegraphics[width = 0.95\textwidth]{99_LeastCostPaths.pdf}
      \includegraphics[width = 0.95\textwidth]{99_LeastCostCorrs.pdf}
    \end{minipage}
    \caption{(a) Source points (black dots) and corresponding least-cost paths
    leaving from protected areas (light grey). Note that only contiguous
    protected areas covering more than 700 km\textsuperscript{2} are depicted.
    Continuous thin black lines indicate the borders of the KAZA-TFCA, whereas
    dashed black lines delineate country-borders. (b) Least-cost corridors
    between the same source points as illustrated in subfigure (a). For ease of
    spatial reference, we also labeled some national parks (dark-grey).}
    \label{LeastCost}
  \end{center}
\end{figure}

\section{Discussion}
We used GPS relocation data collected on dispersing African wild dogs to
investigate whether their main movement corridors are contained within the
boundaries of the world's largest transboundary conservation area, namely the
KAZA-TFCA. Our analysis suggests that the KAZA-TFCA indeed encompasses all major
movement corridors of African wild dogs, demonstrating the potential value of
such an initiative. However, our approach is neither limited to the African wild
dog, nor to our study area. All covariates used throughout this study are
readily available on a global scale and many of them are likely to be important
determinants of movement behavior, landscape permeability, and connectivity for
other species \citep{Thurfjell.2014, Zeller.2012}. Interestingly, our predicted
network of least cost-paths and corridors for African wild dogs shows surprising
similarities to corridors of dispersing lions inhabiting the same ecosystem
\citep{Elliot.2014}. This not only reinforces confidence in our own predictions
but also suggests potential synergies for the conservation of these two, and
possibly more, species. Expanding this analytical framework to additional
species will yield important insights on the consistency of inter-specific
movement corridors, thus highlighting areas that are exceptionally valuable for
the conservation of several species.

Our approach of identifying movement corridors based on pre-defined start and
end points implicitly assumes that individuals know the end point of their
dispersal journey and that they have almost complete knowledge of associated
movement costs. Considering the fact that dispersers often move into unknown
territory \citep{Woodroffe.2019, Cozzi.2020}, this may not necessarily be the
case \citep{Abrahms.2017, Woodroffe.2019}. However, specification of pre-defined
end points might not be necessary, as it should be possible to simulate
dispersal from known source-points, yet without restricting the domain of
potential end points \citep{Signer.2017}. As a consequence, movement corridors
would emerge naturally as the result of a myriad of simulated dispersal events.
While this approach is conceptually straightforward, it requires a comprehensive
mechanistic understanding of dispersal movements, which is conditional on our
ability to collect additional dispersal data and adequately represent the
landscape through which individuals disperse. An approach that does not require
the creation of end points will also be instrumental in identifying corridors
outside formally protected areas. In contrast to classical corridor mapping,
individual-based movement simulations would further allow taking into account
movement preferences of dispersers as well as rendering the temporal aspect of
dispersal (for instance to examine how quickly dispersers are able to move
between locations). The latter consideration is particularly important in
ecosystems where connectivity may change in response to seasonal variations in
the landscape \citep{Osipova.2019}.

Our results emphasize that human presence and activities constitute some of the
main barriers to connectivity. Our result that human influence is detrimental to
dispersal conforms to findings in wild dog populations from eastern Africa
\citep{Masenga.2016, Oneill.2020} but conflicts with findings by
\cite{DaviesMostert.2012}, who reported a high willingness of dispersers to
cross human-dominated landscapes throughout South Africa. We believe that such
differences are due to the unavailability of alternative ``natural'' routes,
which may have forced dispersers to cross human dominated landscapes despite a
strong aversion to do so. This may hint at a considerable ability of dispersing
wild dogs to adapt to prevailing habitat condition and could imply a limited
model transferability between dispersersers from our study system to dispersers
found in other ecosystems, such as in South Africa \citep{Whittington.2011,
DaviesMostert.2012} or Kenya \citep{Woodroffe.2019, Oneill.2020}. However, we
believe that our study population adequately represents wild dog populations
residing within KAZA-TFCA where more natural dispersal routes are available.
Successful conservation of this species is thus contingent on policymakers' and
local authorities' ability and willingness to provide and conserve areas that
remain free from anthropogenic pressures. This is not only paramount in light of
providing connectivity and facilitating dispersal, but also in terms of reducing
human-caused mortality during dispersal. In fact, previous studies have shown
that human-caused mortality during dispersal, especially in proximity to
settlements and roads, represents a major threat to the dispersal ability of
this species \citep{Woodroffe.2019, Cozzi.2020}.

Despite anthropogenic pressures, we identified water as another major obstacle
to dispersal. This is consistent with earlier studies showing that water bodies
are almost impenetrable by resident packs \citep{Abrahms.2017, Cozzi.2020} and
only infrequently crossed by dispersing individuals \citep{Cozzi.2020}. An
accurate and dynamic representation of water is thus imperative and particularly
relevant in flood-pulsing ecosystems such as the Okavango Delta. Although
dispersers avoided moving through water, we revealed a preference for moving in
proximity to water. This preference is likely due to high prey-availability in
vicinity of water \citep{Western.1975, Bonyongo.2005}. However, by the same
reasoning apex predators (e.g. lions, spotted hyenas) and resident wild dogs may
prefer proximity to water \citep{Valeix.2010} and occasionally force dispersing
wild dogs to move into prey-poor areas that are more distant to water
\citep{Ndaimani.2016} to avoid aggressive intraguild interactions
\citep{Creel.1996, Mills.1997, Webster.2012}. We could not control for the
presence or absence of apex predators or conspecifics during dispersal as we
lacked corresponding data, but these considerations may explain why we did not
find a significant effect of \textit{distance to water}. Given the influence
that resident conspecifics, predators, competitors, and prey can have during
dispersal \citep{Cozzi.2018, Armansin.2019} future studies should strive to
collect and incorporate corresponding spatial information into analyses of
landscape connectivity to more precisely reconstruct dispersal corridors.

Locally, we identified the Okavango-Linyanti region as a potential dispersal hub
through which dispersersing wild dogs gain access to more peripheral regions of
the KAZA-TFCA. While this might partly be owed to the area's central position
within the KAZA-TFCA, it is likely also due to its unique environmental and
anthropogenic characteristics. It appears that the absence of human activities
and the presence of relatively impermeable water bodies, such as the Okavango
Delta and the Linyanti Swamp funnel dispersal movements, resulting in a highly
frequented corridor. The region is also instrumental for connecting wild dog
subpopulations in Zambia's Kafue and Zimbabwe's Hwange NPs. These subpopulations
are currently separated by the Zambezi River, forcing dispersers to detour via
northern Botswana to reach more peripheral areas. The key role of the
Okavango-Linyanti region for overall connectivity within the KATA-TFCA thus
calls for actions to secure its protection status in the future. While the
region is currently a Wildlife Management Area devoted to photographic tourism,
it has neither the status of a National Park nor that of a Game Reserve and so
potentially faces growing pressure towards exploitation of its natural resources
under the influences of an expanding human population, farming, and cropping
industry. A similar case of non-formally protected but key dispersal landscape
is represented by the area south of Kafue NP in Zambia, for which a disruption
of its main and narrow dispersal corridor would result in considerable isolation
of its subpopulations. We also revealed a potential southwards corridor between
the Okavango-Linyanti ecosystem and the Central Kalahari National Park.
\cite{Elliot.2014} identified a similar corridor for dispersing lions,
suggesting that upholding and protecting a link between those ecosystems is
pivotal. Some areas through which the corridor runs are neither part of the
KAZA-TFCA nor profit from any form of protected status. In fact, human presence
and activities along the national road that longitudinally traverses this
corridor may limit realized connectivity \citep{Cozzi.2020}. Finally, our
connectivity map suggested that connectivity within the extent of the KAZA-TFCA
gradually decreases towards the borders of our study area, yet this might partly
be a result of boundary effects \citep{Koen.2010}.

%------------------------------------------------------------------------------
%	Gabs wanted to delete this
%------------------------------------------------------------------------------
% Because we predicted permeability and connectivity for such an extensive area,
% we were strongly limited in available and reliable land cover data. As a
% result, our representation of vegetation remained rather basic with only three
% vegetation classes (tree-cover, shrubs/grassland, bare land). Although some
% large-scale, high-resolution land cover datasets exist, such as Globeland's
% 30m \citep{Chen.2015} and the European Space Agency's and 20m \citep{ESA.2019}
% land cover datasets, we found their classifications often erroneous for our
% study area. This is presumably owed to the KAZA-TFCA's unique landscape and
% corresponding difficulties in remote sensing satellite imagery using
% traditional indices \citep{Wolski.2017}.

\section{Conclusion}
Our work showed that the majority of dispersal corridors for African wild dogs
are covered by the boundaries of the KAZA-TFCA, suggesting that the KAZA-TFCA
will significantly contribute to the long-term viability of this species.
Moreover, our connectivity network highlights the Okavango-Linyanti region as a
central dispersal hub through which dispersers gain access to more remote
regions of the KAZA-TFCA. This fidning emphasizes the importance of northern
Botswana for the conservation of this charismatic species. Finally, our
investigations showed that human influence constitutes one of the main barriers
to dispersal and thus reduces landscape connectivity. Successful conservation of
the wild dogs is therefore contingent on the willingness of local authorities,
policymakers, and land managers to preserve areas that are largely liberated
from human strains. Ultimately, our work contributes to the growing field of
connectivity literature and has important implications for the management and
conservation of the highly endangered African wild dog.

\section{Acknowledgements}
We thank the Ministry of Environment and Tourism of Botswana for granting
permission to conduct this research. We thank C. Botes, I. Clavadetscher, and G.
Camenisch for assisting with wild dog immobilizations. We also thank B. Abrahms
for sharing her data of three dispersing wild dogs. Furthermore, we are indebted
to Prof. J. Fieberg, who consulted all statistical aspects of this work and P.
Wolski from the Okavango Research Institute who assisted us in generating
dynamic flood maps. This study was funded by Basler Stiftung für Biologische
Forschung, Claraz Foundation, Idea Wild, Jacot Foundation, National Geographic
Society, Parrotia Stiftung, Stiftung Temperatio, Wilderness Wildlife Trust
Foundation, Forschungkredit der Universität Zürich, and a Swiss National Science
Foundation Grant (31003A\_182286) to A. Ozgul.

\newpage
\begingroup
\singlespacing
\bibliography{Literatur}
\endgroup

\end{document}