### 2020-08-21 Chapter_1/04_Manuscript/99_MovementModelAICs.tex

% latex table generated in R 3.6.3 by xtable 1.8-4 package
% Thu Jul  9 09:17:26 2020
\begin{tabular}{lrrrr}
  \toprule
Covariates & AIC & DeltaAIC & Weight & LogLik \\ 
  \midrule
cos(ta) + log(sl) + W + T + DTW + HI + S + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) & 89303.73 & 0.00 & 0.19 & -44631.86 \\ 
  cos(ta) + log(sl) + W + T + DTW + HI + S + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) & 89304.49 & 0.76 & 0.13 & -44633.24 \\ 
  cos(ta) + log(sl) + W + T + DTW + HI + S + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + T:log(sl\_) & 89304.75 & 1.02 & 0.12 & -44633.38 \\ 
  cos(ta) + log(sl) + W + T + DTW + HI + S + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) + DTW:log(sl\_) & 89305.30 & 1.57 & 0.09 & -44631.65 \\ 
  cos(ta) + log(sl) + W + T + DTW + HI + S + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) + T:cos(ta\_) & 89305.37 & 1.64 & 0.09 & -44631.68 \\ 
  cos(ta) + log(sl) + W + T + DTW + HI + S + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) + HI:log(sl\_) & 89305.37 & 1.64 & 0.09 & -44631.69 \\ 
  cos(ta) + log(sl) + W + T + DTW + HI + S + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) & 89305.45 & 1.72 & 0.08 & -44634.73 \\ 
  cos(ta) + log(sl) + W + T + DTW + HI + S + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) + S:cos(ta\_) & 89305.72 & 1.99 & 0.07 & -44631.86 \\ 
  cos(ta) + log(sl) + W + T + DTW + HI + S + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) + S:log(sl\_) & 89305.72 & 2.00 & 0.07 & -44631.86 \\ 
  cos(ta) + log(sl) + W + T + DTW + HI + S + log(sl\_):MA + W:log(sl\_) + DTW:cos(ta\_) + HI:cos(ta\_) + T:log(sl\_) + W:cos(ta\_) & 89305.72 & 2.00 & 0.07 & -44631.86 \\ 
   \bottomrule
\end{tabular}