### 2020-09-02 Chapter_0/04_Manuscript/99_PermeabilityModelAICs.tex
Worked on the revision for BES paper

% latex table generated in R 3.6.3 by xtable 1.8-4 package
% Wed Sep  2 09:04:38 2020
\begin{tabular}{lrrrr}
  \toprule
Covariates & AIC & DeltaAIC & Weight & LogLik \\ 
  \midrule
cos(ta) + sl + log(sl) + W + T + DTW + HI + S & 90068.15 & 0.00 & 1.00 & -45017.08 \\ 
  cos(ta) + sl + log(sl) + W + T + DTW + HI & 90071.84 & 3.69 & 0.00 & -45020.92 \\ 
  cos(ta) + sl + log(sl) + W + T + DTW + HI + S + DTR & 90071.94 & 3.79 & 0.00 & -45016.97 \\ 
  cos(ta) + sl + log(sl) + W + T + DTW + HI + S + P & 90071.94 & 3.79 & 0.00 & -45016.97 \\ 
  cos(ta) + sl + log(sl) + W + T + DTW + HI + S + RC & 90073.46 & 5.30 & 0.00 & -45015.73 \\ 
  cos(ta) + sl + log(sl) + W + T + DTW + HI + DTR & 90075.66 & 7.50 & 0.00 & -45020.83 \\ 
  cos(ta) + sl + log(sl) + W + T + DTW + HI + P & 90075.66 & 7.51 & 0.00 & -45020.83 \\ 
  cos(ta) + sl + log(sl) + W + T + DTW + HI + S + DTR + P & 90075.79 & 7.64 & 0.00 & -45016.89 \\ 
  cos(ta) + sl + log(sl) + W + T + DTW + S & 90076.71 & 8.56 & 0.00 & -45023.36 \\ 
  cos(ta) + sl + log(sl) + W + T + DTW + HI + RC & 90076.84 & 8.69 & 0.00 & -45019.42 \\ 
  cos(ta) + sl + log(sl) + W + T + DTW + HI + S + DTR + RC & 90077.20 & 9.05 & 0.00 & -45015.60 \\ 
  cos(ta) + sl + log(sl) + W + T + DTW & 90080.08 & 11.92 & 0.00 & -45027.04 \\ 
  cos(ta) + sl + log(sl) + W + T + DTW + HI + S + DTR + P + RC & 90080.96 & 12.81 & 0.00 & -45015.48 \\ 
  cos(ta) + sl + log(sl) + W + T + DTW + DTR & 90082.95 & 14.79 & 0.00 & -45026.47 \\ 
  cos(ta) + sl + log(sl) + W + T + DTW + P & 90083.31 & 15.16 & 0.00 & -45026.66 \\ 
  cos(ta) + sl + log(sl) + W + T + HI & 90103.28 & 35.13 & 0.00 & -45038.64 \\ 
  cos(ta) + sl + log(sl) + W + T & 90109.40 & 41.25 & 0.00 & -45043.70 \\ 
  cos(ta) + sl + log(sl) + W + T + S & 90110.35 & 42.20 & 0.00 & -45042.17 \\ 
  cos(ta) + sl + log(sl) + W + T + DTR & 90112.55 & 44.40 & 0.00 & -45043.27 \\ 
  cos(ta) + sl + log(sl) + W + T + P & 90113.11 & 44.96 & 0.00 & -45043.56 \\ 
  cos(ta) + sl + log(sl) + W + T + RC & 90113.60 & 45.45 & 0.00 & -45041.80 \\ 
  cos(ta) + sl + log(sl) + W + DTW & 90118.55 & 50.40 & 0.00 & -45048.28 \\ 
  cos(ta) + sl + log(sl) + W + HI & 90128.70 & 60.54 & 0.00 & -45053.35 \\ 
  cos(ta) + sl + log(sl) + W + S & 90132.22 & 64.06 & 0.00 & -45055.11 \\ 
  cos(ta) + sl + log(sl) + W & 90134.85 & 66.69 & 0.00 & -45058.42 \\ 
  cos(ta) + sl + log(sl) + W + DTR & 90138.31 & 70.16 & 0.00 & -45058.16 \\ 
  cos(ta) + sl + log(sl) + W + P & 90138.50 & 70.35 & 0.00 & -45058.25 \\ 
  cos(ta) + sl + log(sl) + W + RC & 90139.30 & 71.15 & 0.00 & -45056.65 \\ 
  cos(ta) + sl + log(sl) + S & 90141.98 & 73.83 & 0.00 & -45061.99 \\ 
  cos(ta) + sl + log(sl) + DTW & 90225.64 & 157.49 & 0.00 & -45103.82 \\ 
  cos(ta) + sl + log(sl) + T & 90271.73 & 203.58 & 0.00 & -45126.86 \\ 
  cos(ta) + sl + log(sl) + HI & 90273.18 & 205.02 & 0.00 & -45127.59 \\ 
  cos(ta) + sl + log(sl) + P & 90285.24 & 217.08 & 0.00 & -45133.62 \\ 
  cos(ta) + sl + log(sl) + DTR & 90285.33 & 217.18 & 0.00 & -45133.67 \\ 
  cos(ta) + sl + log(sl) + RC &  &  & 0.00 &  \\ 
  cos(ta) + sl + log(sl) + W + T + DTW + RC &  &  & 0.00 &  \\ 
   \bottomrule
\end{tabular}