### 2020-09-10 Chapter_0/04_Manuscript/Manuscript.tex
Worked on revisions

\documentclass[abstract=on,10pt,a4paper,bibliography=totocnumbered]{article}
\usepackage[paper=a4paper,left=35mm,right=35mm,top=25mm,bottom=30mm]{geometry}
\usepackage[doublespacing]{setspace}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[round]{natbib}
\usepackage{amsmath}
\usepackage{colortbl}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{gensymb}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage[hidelinks]{hyperref}
\usepackage[nameinlink]{cleveref}
\usepackage{lineno}
\usepackage{multirow}
\usepackage{arydshln}
\usepackage[nomarkers, nolists]{endfloat}
\usepackage{todonotes}

%------------------------------------------------------------------------------
%	Some Styling
%------------------------------------------------------------------------------
% Creating some TikZ styles
\tikzset{
  nonterminal/.style = {rectangle
    , minimum size = 6mm
    , very thick
    , draw = black!
  }
}

% Changing the style of captions in figures etc.
\captionsetup{labelfont=bf, format=plain, font=small}

% Change how equations are referenced
\renewcommand{\theequation}{Equation \arabic{equation}}%

%------------------------------------------------------------------------------
%	Titlepage: Header
%------------------------------------------------------------------------------
\title{Bound within Boundaries: How Well Do Protected Areas Match Movement
Corridors of Their Most Mobile Protected Species?}

% List of Authors
\author{
  David D. Hofmann\textsuperscript{1,\S,*} \and
  Dominik M. Behr\textsuperscript{1,2,*} \and
  John W. McNutt\textsuperscript{2} \and
  Arpat Ozgul\textsuperscript{1} \and
  Gabriele Cozzi\textsuperscript{1,2}
}

% Reduce spacing between authors
\makeatletter
\def\and{%
  \end{tabular}%
  \hskip -0.5em \@plus.17fil\relax
  \begin{tabular}[t]{c}}
\makeatother

% Current Date
\date{\today}

% And here the masterpiece begins
\begin{document}

% Change page numbering
\pagenumbering{gobble}

% Required to be able to cite
\bibliographystyle{apalike}

% Create Titlepage
\maketitle

%------------------------------------------------------------------------------
%	Titlepage: Additional Info
%------------------------------------------------------------------------------
\begin{flushleft}

\vspace{0.5cm}

\textsuperscript{1} Department of Evolutionary Biology and Environmental
Studies, University of Zurich, Winterthurerstarsse 190, 8057 Zurich,
Switzerland.

\textsuperscript{2} Botswana Predator Conservation Trust, Private Bag 13, Maun,
Botswana.

\textsuperscript{\S} Corresponding author (david.hofmann2@uzh.ch)

\textsuperscript{*} Shared first authorship

\vspace{4cm}

\textbf{Running Title:} Connectivity across a Transfrontier Conservation Area.

\vspace{0.5cm}

\textbf{Keywords:} dispersal, habitat selection, integrated step selection
function, Kavango-Zambezi Transfrontier Conservation Area, landscape
connectivity, least-cost corridors, Lycaon pictus, permeability surface,
protected areas, wildlife management

\end{flushleft}

%------------------------------------------------------------------------------
%	Abstract
%------------------------------------------------------------------------------
\newpage
\begin{abstract}
\noindent 1. Conserving and managing large portions of land to connect wildlife
reserves is an increasingly used strategy to maintain and restore connectivity
among wildlife populations. Boundaries of such conservation areas are often
determined based on expert opinion and socio-political constraints, yet the
extent to which they match species' movement corridors is rarely examined. This
is mainly due to a lack of data, particularly on wide-ranging movement behavior
such as dispersal. Nevertheless, empirically assessing the adequacy of protected
areas is key for the implementation of targeted management actions and efficient
use of limited conservation funds.

\noindent 2. Between 2011 and 2019, we collected high-resolution GPS movement
data on 16 dispersing African wild dog (\textit{Lycaon pictus}) coalitions from
a free-ranging population in the Kavango-Zambezi Transfrontier Conservation Area
(KAZA-TFCA). Spanning five countries and 520'000 km\textsuperscript{2} the
KAZA-TFCA is the world's largest transboundary conservation area and a prime
example for international conservation efforts. We used integrated step
selection analysis to estimate habitat selection of dispersers and to create a
permeability surface for the entire KAZA-TFCA. We compared landscape
permeability across different regions within the KAZA-TFCA as well as outside
its boundaries. Lastly, we calculated least-cost paths and corridors to verify
that major movement routes were adequately encompassed within the KAZA-TFCA.

\noindent 3. Permeability within the boundaries of the KAZA-TFCA was more than
double compared to areas outside it. Furthermore, we observed a five-fold
permeability difference among the five KAZA-TFCA countries. We also showed
that major movement corridors of wild dogs run within the KAZA-TFCA, although
some minor routes remained outside formally protected areas.

\noindent 4. Differences in permeability were mainly related to different
degrees of human activities across regions, and to the presence or absence of
rivers, swamps or open water. The relation between permeability and other
landscape features was less evident.\todo{Editor indicated that the
relationships are correlative and not necessarily causal.}

\noindent 5. \textit{Synthesis and Applications}: In this study, we showed how
pertinent dispersal data of a highly mobile species can be used to empirically
evaluate the adequacy of already-existing or planned protected areas.
Furthermore, we observed regional differences in landscape permeability that
highlight the need for a coordinated effort towards maintaining or restoring
connectivity, especially where transboundary dispersal occurs.\todo{According to
suggestions of reviewer 2}
\end{abstract}

%------------------------------------------------------------------------------
%	Main Text
%------------------------------------------------------------------------------
\newpage

% Change page numbering
\pagenumbering{arabic}

% Create linenumbers
\linenumbers

\section{Introduction}
Connectivity among subpopulations is a crucial pre-requisite for many species to
thrive and persist \citep{Fahrig.2003}. Accordingly, preserving and protecting
movement corridors between wildlife reserves has become a task of utmost
\todo{Reviewer one pointed out an awkward use of the word utmost. I have changed
it according to his suggestion.} importance \citep{Doerr.2011, Rudnick.2012},
resulting in an ever-growing number of large and often transboundary protected
areas. While boundaries of such areas are often drawn according to expert
opinion and socio-political needs, subjective assessments have revealed
deficiencies in the past \citep{Clevenger.2002, Pullinger.2010}. Thus, an
empirical evaluation of the adequacy of already-existing or planned protected
areas using pertinent animal movement data is paramount for targeted use of
valuable and scarce conservation funds \citep{Pullinger.2010}.

In recent years, a growing body of research has used animal relocation data to
identify movement corridors and assess connectivity at large scales (e.g.
\citealp{Chetkiewicz.2006, Squires.2013, Elliot.2014}). The use of relocation
data has shifted the focus from a structural to a functional view on
connectivity, taking into account how species respond to environmental features
\citep{Taylor.2006, Fattebert.2015}.\todo{As per reviewer 2} Identification of
potential movement corridors typically relies on the estimation of permeability
surfaces, which return the ease or willingness at which the focal species
traverses a specific landscape \citep{Sawyer.2011}. Such surfaces are created
based on species' relative selection strengths \citep{Avgar.2017}\todo{reviewer
1}, which can be quantified using a suite of selection functions
\citep{Zeller.2012}. Specifically, selection strengths \todo{reviewer 1} are
estimated by comparing spatial covariates (e.g. environmental and anthropogenic)
at locations visited by the animal to the same spatial covariates at locations
avilable to the animal \todo{Reviewer 1 pointed out correctly that we should
talk about available locations, rather than randomly selected locations}
\citep{Zeller.2012}. Importantly, selection functions rely on adequate landscape
and relocation data that are representative of the process being studied
\citep{Diniz.2020}. Altough selection during residence and dispersal may
coincide \citep{Fattebert.2015},\todo{second reviewer pointed out that space use
can be alike} it appears that relocation data collected on dispersing
individuals outperforms data collected on resident individuals in the detection
of large-scale movement corridors \citep{Elliot.2014, Diniz.2020}. Nevertheless,
dispersal data is inherently difficult to collect and remains scarce in the
connectivity literature \citep{Vasudev.2015}. As such, most permeability
surfaces upon which movement corridors are identified are created using
relocation data collected on resident individuals. This has likely limited our
ability to meaningfully assess the effectiveness of protected areas in securing
functional connectivity for their protected species.\todo{Reviewer 2 desired to
change from: this introduces - to: this can introduce}

One initiative that aims at restoring and enhancing connectivity across large
scales is the Kavango-Zambezi Transfrontier Conservation Area (KAZA-TFCA), which
constitutes the world's largest transfrontier conservation area, spanning over
520'000 km\textsuperscript{2} and five countries (\url{www.kavangozambezi.org}).
While the KAZA-TFCA was originally set to facilitate movements of African
elephants (\textit{Loxodonta africana}; \citealp{Tshipa.2017}), it is also key
to the conservation of other wide-ranging species such as African wild dogs
(\textit{Lycaon pictus}; \citealp{Woodroffe.2012, Cozzi.2020}), lions
(\textit{Panthera leo}; \citealp{Elliot.2014, Cushman.2018}), and cheetahs
(\textit{Acinonyx jubatus}; \citealp{Weise.2017}). To date, however, few studies
have attempted to assess the adequacy of the KAZA-TFCA using relevant global
positioning system (GPS) relocation data of its protected species at large
spatial scales \citep{Elliot.2014, Tshipa.2017}.\todo{as per editor} Thus, how
well the boundaries of the KAZA-TFCA reflect natural movement patterns and
dispersal corridors of its most mobile protected species is virtually unknown.

Across the KAZA-TFCA, the African wild dog (\textit{Lycaon pictus}) represents a
highly mobile and endangered flagship species for conservation efforts. Once
widespread across the entire Sub-Saharan continent, wild dogs have been widely
extirpated through human persecution, habitat destruction, and disease outbreaks
\citep{Woodroffe.2012}. As a result, the species has become one of Africa's most
endangered large carnivores, and currently only survives in small, spatially
scattered subpopulations \citep{Woodroffe.2012}. Within these subpopulations,
wild dogs form cooperative breeding packs of up to thirty individuals
\citep{Creel.2002}, whose social structure is strongly governed by the process
of dispersal \citep{McNutt.1996, Behr.2020}. Both males and females disperse
from their natal pack, either alone or in same-sex dispersing coalitions, and
search for unrelated mates and a suitable territory to settle
\citep{McNutt.1996, Cozzi.2020, Behr.2020}. During dispersal, wild dogs can
cover several hundred kilometers and cross international borders
\citep{Masenga.2016, Woodroffe.2019, Cozzi.2020}. Despite the importance of
dispersal for the long-term viability of this species, little empirical
information is available on habitat selection and potential movement barriers
during dispersal. The few studies that have collected dispersal data have shown
that dispersers quickly move over large distances, avoid human-dominated
landscapes and areas densely covered by trees, but prefer proximity to water
\citep{Masenga.2016, Woodroffe.2019, Oneill.2020, Cozzi.2020}.

Here, we collected and analyzed GPS relocation data on 16 dispersing wild dogs
in as many dispersing coalitions from a free-ranging population in northern
Botswana to assess the adequacy of the KAZA-TFCA in securing connectivity. We
estimated relative selection strengths\todo{reviewer 1} towards environmental
and anthropogenic landscape features, and used the obtained coefficients to
predict a permeability surface spanning the entire KAZA-TFCA. We then
investigated how landscape permeability varied regionally and internationally,
and we compared permeability within and outside the KAZA-TFCA boundaries.
Finally, we calculated least-cost paths and corridors to identify major movement
routes and to verify that these were successfully covered by the KAZA-TFCA.

\section{Methods}
\subsection{Study Area}
The study area (centered at -17\degree 13'9''S, 23\degree 56'4''E;
\Cref{StudyArea}a) was outlined by a rectangular bounding box stretching over
1.3 Mio km\textsuperscript{2} and encompassing the entire KAZA-TFCA
(\Cref{StudyArea}b). The KAZA-TFCA lies in the basins of the Okavango and
Zambezi rivers and includes parts of Angola, Botswana, Namibia, Zimbabwe, and
Zambia. With a total area of over 520'000 km\textsuperscript{2} it constitutes
the earth's largest transboundary conservation area and is characterized by
diverse landscapes, including savanna, grassland, and dry or moist woodland
habitats. Rainfall in the study area is seasonal and lasts from November to
March. The KAZA-TFCA also comprises the Okavango Delta, which represents a
highly dynamic hydrological flood-pulsing system \citep{McNutt.1996,
Wolski.2017}. The extent of the flood in the delta greatly changes within and
between years, depending on the amount of rain that descends from the catchment
areas in Angola and reaches the distal ends of the delta between July and August
(Figure S4). The flood drastically affects surrounding landscapes, so that
during maximum extent (ca. 12'000 km\textsuperscript{2}) the delta becomes a
patchy conglomerate of swamps, open water, and islands, whereas these structures
run dry when the flood retracts to its minimum extent (ca. 5'000
km\textsuperscript{2}; \citealp{Wolski.2017}). Despite 36 national parks (NPs)
and other protected areas, there is considerable human influence in some regions
of the KAZA-TFCA, mainly originating from farms, human density, and road
traffic.

\begin{figure}[h]
  \begin{center}
    \begin{tikzpicture}
        \node[anchor=south west,inner sep=0] (image) at (0,0,0) {
        \includegraphics[width=\textwidth]{99_StudyArea.pdf}
        };
        \begin{scope}[x={(image.south east)},y={(image.north west)}]
            % % next four lines will help you to locate the point needed by forming a grid.
            % % comment these four lines in the final picture.
            % \draw[help lines,xstep=.1,ystep=.1] (0,0) grid (1,1);
            % \draw[help lines,xstep=.05,ystep=.05] (0,0) grid (1,1);
            % \foreach \x in {0,1,...,9} { \node [anchor=north] at (\x/10,0) {0.\x}; }
            % \foreach \y in {0,1,...,9} { \node [anchor=east] at (0,\y/10) {0.\y};}
            % % upto here
            \draw[black] (0.202, 0.255) -- (0.36, 0.955);
            \draw[black] (0.202, 0.205) -- (0.36, 0.070);
        \end{scope}
    \end{tikzpicture}
    \caption{Overview of our study area. (a) The red dotted rectangle depicts
    the study area, which was confined by a bounding box encompassing the entire
    KAZA-TFCA. Gray areas indicate remaining wild dog populations according to
    the IUCN \citep{Woodroffe.2012}. (b) The white rectangle illustrates the
    area within which dispersing coalitions were collared. Since Game Reserves
    in Botswana virtually serve the same purpose as National Parks, we use the
    terms interchangeably for this region.}
    \label{StudyArea}
  \end{center}
\end{figure}

\subsection{GPS Relocation Data}
We used a population of free-ranging African wild dogs inhabiting the Okavango
Delta in northern Botswana as a source population for dispersing individuals.
This population has been extensively studied since 1989 \citep{McNutt.1996,
Cozzi.2013, Cozzi.2020, Behr.2020}. Between 2011 and 2019, we systematically
collected GPS relocation data on 16 coalitions of dispersing African wild dogs
(7 female and 9 male coalitions). Candidate dispersing individuals were
identified based on age, number of same‐sex siblings, pack size, and presence of
unrelated individuals of the opposite sex in their pack \citep{McNutt.1996,
Behr.2020}\todo{The editor desired to include criteria}. Selected individuals
were immobilized according to protocols described in \cite{Osofsky.1996}, and
fitted with GPS/Satellite radio collars (\textit{Vertex Lite; Vectronic
Aerospace GmbH, Berlin, Germany}) while still with their natal pack. All
required procedures were undertaken and supervised by a Botswana-registered
wildlife veterinarian. Collars produced a tag weight of 300 gramms, accounting
for 1.2\% of a wild dog's average body weight. They were attached using leather
belts that included a drop off mechanism, triggered by a decomposing piece of
cloth.\todo{Editor desired more information on GPS collar} During dispersal, GPS
collars were programmed to record a GPS relocation every 4 hours and to
regularly transmit data via Iridium satellite system to a base station.

Because we were interested in dispersal behavior only, we discarded any GPS data
collected while individuals were still with their natal packs and after
settlement in a new territory \citep{Cozzi.2020}. We identified the exact time
of emigration and settlement based on direct field observations and through
visual inspection of the net squared displacement (NSD) metric. NSD quantifies
the squared \todo{Reviewer 1 pointed out correctly that we need to talk about
the \textit{squared} euclidean distance} Euclidean distance of a relocation to a
reference point \citep{Borger.2012}, which in our case was the center of the
dispersing coalition's natal home range. Thus, dispersal was deemed to have
started when a coalition had left its natal home range and continued until the
NSD metric remained stationary, indicating that the coalition had successfully
settled (Figure S1). In our analysis, we did not differentiate between male and
female dispersing coalitions, for previous research found little differences
between sexes during dispersal \citep{Woodroffe.2019, Cozzi.2020}.

\subsection{Spatial Covariates}
To investigate relative selection strengths \todo{As per reviewer 1} of
dispersing wild dogs, we used a set of geo-referenced covariates
(\Cref{Covariates}) that we aggregated in the categories \textit{land cover},
\textit{protection status}, and \textit{human influence}. Land cover comprised
the covariates water cover, distance to water, percentage cover by
shrubs/grassland, and percentage cover by trees. To capture the pulsing behavior
of the Okavango Delta, we frequently updated layers for water cover and
corresponding layers depicting distance to water. Protection status contained a
binary indicator of whether an area was protected or not. Human influence
included covariates rendering human density, the presence of roads, and the
distance to roads.\todo{Reviewer 2 found the description of the covariates
unclear} We prepared all covariates as spatial raster layers from freely
available online services and from remotely sensed satellite imagery. To ensure
a consistent resolution (i.e. cell-size or grain) across covariates, we
coarsened or interpolated all layers to a resolution of 250m x 250m. For further
details on the preparation of each covariate and their sources, see Appendix
A.3. We performed processing and manipulation of data as well as all spatial and
statistical analyses using R, version 3.6.1 \citep{R.2019}.

\begin{figure}[h]
  \begin{center}
    \includegraphics[width = \textwidth]{99_Covariates.pdf}
    \caption{Overview of spatial covariates that we included in our models. We
    prepared all covariates for the entire study area but for better visibility
    we only plot them for the surroundings of the Okavango Delta. The white
    rectangle in each plot depicts the area within which dispersing coalitions
    were collared. (a) Averaged layer of all dynamic (binary) water maps. (b)
    Percentage cover of trees. (c) Percentage cover of shrubs/grassland.
    Anything that was not covered by trees or shrubs/grassland was deemed to be
    bare land. (d) Protection status of the area. (e) Human influence proxy
    composed of human density, farms, and roads. (f) Distance to nearest road
    (white lines depict actual roads).}
    \label{Covariates}
  \end{center}
\end{figure}

\subsection{Habitat Selection Model}
We used an integrated step selection function (iSSF; \citealp{Avgar.2016}) to
investigate dispersers' relative selection strentghs towards \todo{Reviewer 1
suggested to use relative selection strength, rather than preference} the
above-mentioned spatial covariates. That is, we paired each realized step (i.e.
the connecting line between two consecutive GPS relocations;
\citealp{Turchin.1998}) with 24 random steps that were generated by sampling
turning angles from a uniform distribution U(\(-\pi, +\pi\)) and step lengths
from a gamma distribution fitted to realized steps \citep{Avgar.2016}. A
realized step and its 24 associated random steps formed a stratum and received a
unique identifier. Along each step, we extracted the above-mentioned covariates
(Table S3), standardized extracted values using a z-score transformation, and
checked for correlation using Pearson's Correlation Coefficient \(r\). None of
the covariates were overly correlated (\(|r| > 0.6\); \citealp{Latham.2011}) and
we retained all of them for modeling. Our habitat selection model then assumed
that dispersing wild dogs assigned a selection score \(w(x)\) of the following
exponential form to each realized and random step \citep{Fortin.2005}:

\begin{equation}
\label{EQ1}
  w(x) = exp(\beta_1 x_1 + \beta_2 x_2 + ... + \beta_n x_n)
\end{equation}

\noindent The selection score \(w(x)\) of a step depended on its associated
covariates (\(x_1, x_2, ..., x_n\)), as well as on the animal's relative
selection strengths \todo{As per reviewer 1} towards these covariates
(\(\beta_1, \beta_2, ..., \beta_n\)). To estimate relative selection strengths
\todo{As per reviewer 1} (i.e. the \(\beta\)'s) for each covariate, we used
mixed effects conditional logistic regression analysis as suggested by
\cite{Muff.2020}. We implemented their method using the R-package
\textit{glmmTMB} \citep{Mollie.2017} and used dispersing coalition ID to model
random slopes. We also modelled random intercepts with an arbitrary high
variance of 10\textsuperscript{6} to make use of the poisson trick (see
\citealp{Muff.2020}). We defined three movement metrics, namely the cosine of
the turning angle (\(cos(ta)\)), the step length (\(sl\)) and the natural
logarithm of the step length (\(log(sl)\)), as core covariates and ran stepwise
forward model selection based on Akaike's Information Criterion (AIC;
\citealp{Burnham.2002}) for all other covariates.\todo{Reviewer 1 asked for more
details in the random effects. I therefore decided to update this section too}
The inclusion of movement metrics served to reduce biases in estimated habitat
selection coefficients \todo{as per revierwer 1} that may have arisen due to
movement behavior \citep{Avgar.2016}. To validate the predictive power of the
most parsimonious habitat selection model, we ran k-fold cross-validation for
case-control studies as described in \cite{Fortin.2009} (details in Appendix
A.5).

\subsection{Permeability Surface}
Using the most parsimonious habitat selection model, we predicted a permeability
surface spanning the entire extent of the KAZA-TFCA. That is, we applied
\ref{EQ1} to our spatial covariates and calculated the selection score \(w(x)\)
for each raster cell. Because our representation of water was dynamic (to
properly render the pulsing behavior of the Okavango Delta) we collapsed all
dynamic water maps into a single static map using areas that were covered by
water in at least 10\% of the cases. Using the resulting static map we also
calculated a layer returning the distance to water. Because the extent of the
delta covers merely 6\% of the entire KAZA-TFCA, the use of static maps should
introduce little bias in predicted permeability. To reduce the influence of
outliers in predicted permeability scores, we followed \cite{Squires.2013} and
curtailed predicted scores between the 1\textsuperscript{st} and
99\textsuperscript{th} percentile of their original values. To compare
permeability across different regions, we normalized the permeability surface to
a range between 0 (most impermeable) and 1 (most permeable), and we determined
median permeability within and outside the KAZA-TFCA, within and outside
formally protected areas, and within each of the five KAZA-TFCA countries.

\subsection{Least-Cost Paths and Corridors}
To identify movement corridors of dispersing wild dogs, we specified source
points and calculated factorial least-cost paths (LCPs) as well as factorial
least-cost corridors (LCCs) among them \citep{Elliot.2014}. We generated source
points by overlaying the study area with a regular grid of points spaced at 100
km. We only considered those points that fell within protected areas \(>\) 700
km\textsuperscript{2}, which conforms with home-range requirements of African
wild dogs \citep{Pomilia.2015}. Finally, we defined centroids as source points
for those protected areas \(>\) 700 km\textsuperscript{2} that were not assigned
any source points from the regular grid. Because wild dogs residing outside of
protected areas are rare and likely unviable \citep{VanDerMeer.2014}, we
consider our selection of source points within protected areas to be
appropriate. In total, we generated 68 source points, which resulted in 2'278
unique pairwise combinations and therefore 2'278 unique LCPs and LCCs. We
computed factorial LCPs and LCCs between source points using the R-package
\textit{gdistance} (further details in Appendix A.7). After computation, we
tallied overlapping LCPs and LCCs, respectively, into single connectivity maps.

\section{Results}
\subsection{Dispersal Events}
In total, we collected 4'169 GPS relocations during dispersal (Figure S2 \&
Table S1), resulting in an average of 261 (\(SD\) = 207) locations per
dispersing coalition. Coalitions on average dispersed for 48 days (\(SD\) = 44),
covered a mean Euclidean distance of 54 km (\(SD\) = 71) and a cumulative
distance of 597 km (\(SD\) = 508). One female coalition dispersed far east into
the Hwange Hwange National Park and covered a cumulative distance of over 360 km
in under 9 days.\todo{Editor desired a quick summary of the dispersal data.}

\subsection{Habitat Selection Model}
Our most parsimonious habitat selection model (\(\Delta AIC > 2\) than any
alternative model; Table S4) retained the covariates \textit{water},
\textit{distance to water}, \textit{trees}, \textit{shrubs/grassland}, and
\textit{human influence}, beside the fixed covariates \textit{cos(ta)},
\textit{sl}, and \textit{log(sl)} (\Cref{PermeabilityResults}a). The positive
coefficients for \(sl\) (\(\beta\) = 0.03, 95\% CI = -0.09 to 0.15) and
\(log(sl)\) (\(\beta\) = 0.12, 95\% CI = 0.02 to 0.22) indicated that our
initial (selection-biased) gamma distribution produced step-lengths slightly
shorter than a selection-free gamma distribution would. Similarly, the positive
coefficient for \(cos(ta)\) (\(\beta\) = 0.10; 95\% CI = 0.05 to 0.15) suggested
that our uniform distribution for turning angles produced turning angles with
too little directionality. \todo{Reviewer 1 nicely explained that the
coefficients for log(sl) and cos(ta) imply that the initial distributions were
selection-biased} With respect to environmental covariates, dispersers avoided
moving through water (\(\beta\) = -0.53, 95\% CI -0.79 to -0.27) but selected
for locations in its vicinity, although the latter effect was not significant
(\(\beta\) = -0.33, 95\% CI = -0.73 to 0.08). Dispersers avoided areas that were
densely covered by trees (\(\beta\) = -0.31, CI = -0.47 to -0.15) and preferred
areas covered by shrubs/grassland (\(\beta\) = 0.25, 95\% CI = 0.07 to 0.42).
Finally, dispersers avoided areas that were influenced by humans (\(\beta\) =
-0.45, 95\% CI = -0.82 to -0.08). With the exepction of \textit{distance to
water} (\(SD_{RE}\) = 0.57), random effects revealed little variability between
different dispersal coalitions' selection strengths (i.e. all with \(SD_{RE}
\leq 0.22\), Figure S7).\todo{Reviewer 1 desired more information on the random
effects}

\begin{figure}[h]
  \begin{center}
    \includegraphics[width = \textwidth]{99_PermeabilityResults.pdf}
    \caption{(a) Estimated selection coefficients from the most parsimonious
    habitat selection model. Negative coefficients indicate avoidance of a
    covariate, positive coefficients selection of a covariate. Whiskers
    delineate the 95\%-CIs for estimated parameters. Significance codes: ** \(p
    < 0.05\), *** \(p < 0.01\).(b) Results from the k-fold cross validation for
    case-control studies. The left graph shows rank frequencies of
    \textit{realized} steps according to predictions, whereas the right graph
    shows rank frequencies of \textit{randomly selected} steps according to
    predictions. \(\bar{r}_s\) indicates the mean correlation coefficient
    resulting from 100 repetitions of the k-fold cross validation. The blue
    smoothing line was fitted using a locally weighted polynomial regression and
    serves to aid the eye in detecting the trends. Correlation coefficients
    suggest that our prediction was significant and robust, evidenced by the
    fact that the confidence intervals of \(\bar{r}_{s, realized}\) and
    \(\bar{r}_{s, random}\) did overlap and by the fact that there was strong
    and significant correlation between ranks and associated frequency for
    realized steps.}
    \label{PermeabilityResults}
  \end{center}
\end{figure}

Results from the k-fold cross-validation suggested that our prediction was
significant and robust, as highlighted by the fact that the 95\%-CIs intervals
of \(\bar{r}_{s, realized}\) and \(\bar{r}_{s, random}\) did not overlap
(\Cref{PermeabilityResults}b). Likewise, the significant correlation between
ranks and corresponding frequencies for realized steps suggested a good fit
between predictions and observations (\Cref{PermeabilityResults}b).

\subsection{Permeability Surface}
Our prediction of landscape permeability revealed substantial differences across
regions in the study area (\Cref{PermeabilityMap}). Comparisons of median
permeability values (\Cref{PermeabilityComp}) showed that permeability inside
the KAZA-TFCA was more than two times as high as permeability outside it.
Permeability varied by country, with a five-fold permeability difference among
them. Angola and Botswana were characterized by comparably highly permeable
landscapes, Zimbabwe and Zambia were relatively impermeable, and Namibia ranged
in between the two extremes (\Cref{PermeabilityComp}). Visual inspection of our
covariate layers indicated that high permeability in Angola and Botswana was
mainly related to a combination of low human influence, low tree cover, high
shrubs/grassland cover, and a close distance to water. Although swamps,
wetlands, and permanent water themselves provided little permeability, their
surroundings acted as strong attractants to dispersers. The low permeability
that characterized Zambia and Zimbabwe, on the other hand, was mainly caused by
substantial human influence. Albeit the KAZA-TFCA covered most permeability
hot-spots, several highly permeable regions remained uncovered by its borders.
Across all countries, protected areas provided roughly double the permeability
of unprotected landscapes (\Cref{PermeabilityComp}).

\begin{figure}[hbtp]
  \begin{center}
    \begin{tikzpicture}
        \node[anchor=south west,inner sep=0] (image) at (0,0,0) {
          \begin{minipage}{\textwidth}
            \includegraphics[width = \textwidth]{99_PermeabilityMap.pdf}
          \end{minipage}
        };
        \begin{scope}[x={(image.south east)},y={(image.north west)}]
        \end{scope}
    \end{tikzpicture}
    \caption{Predicted permeability surface for the extent of the KAZA-TFCA.
    Permeability was predicted by calculating selection scores \(w(x) =
    exp(\beta_1 x_1 + \beta_2 x_2 + ... + \beta_n x_n)\) for each raster cell
    based on the raster cell's underlying covariates (\(x_i\)) and estimated
    selection strength (\(\beta_i\)). Areas that dispersers find easy to
    traverse are depicted in bright colors. Bold white lines delineate the
    borders of the KAZA-TFCA, whereas dashed white lines show country borders.}
    \label{PermeabilityMap}
  \end{center}
\end{figure}

\begin{table}[h]
  \begin{center}
  \caption{Comparison of median permeability (interquantile range in brackets)
  across countries, separated into areas within and outside the KAZA-TFCA, as
  well as within and outside formally protected areas. High values indicate high
  permeability, whereas low values correspond to low permeability.}
  \label{PermeabilityComp}
  \begin{tabular}{llllll}
    \toprule
    \multicolumn{1}{c}{} & \multicolumn{2}{c}{KAZA-TFCA} &
    \multicolumn{2}{c}{Protection Status} & \multicolumn{1}{c}{} \\
    \cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
    Country & Inside & Outside & Protected & Pastoral & Overall\\
    \midrule
    Angola & 0.35 (0.41) & 0.12 (0.32) & 0.35 (0.41) & 0.12 (0.32) & 0.19 (0.38)\\
    Botswana & 0.24 (0.30) & 0.14 (0.16) & 0.27 (0.35) & 0.14 (0.18) & 0.18 (0.25)\\
    Namibia & 0.20 (0.30) & 0.12 (0.17) & 0.22 (0.30) & 0.10 (0.14) & 0.14 (0.24)\\
    Zambia & 0.05 (0.09) & 0.02 (0.05) & 0.04 (0.09) & 0.03 (0.05) & 0.03 (0.06)\\
    Zimbabwe & 0.06 (0.16) & 0.05 (0.04) & 0.07 (0.17) & 0.04 (0.04) & 0.05 (0.06)\\
    \hline
    Overall & 0.15 (0.29) & 0.06 (0.14) & 0.14 (0.30) & 0.06 (0.14) & 0.08 (0.21)\\
    \bottomrule
  \end{tabular}
  \end{center}
\end{table}

\subsection{Least-Cost Paths \& Least-Cost Corridors}
Our least-cost analysis revealed three major movement corridors of which all
were well-contained within the KAZA-TFCA boundaries (\Cref{LeastCost}). One
major corridor ran SE-NW and connected the Okavango-Linyanti ecosystem in
Botswana with Luengue-Luiana NP in Angola. A second corridor ran W-E between
Chobe NP in Botswana and Zimbabwe's Hwange NP. A third major corridor ran NE-SW,
completely across unprotected areas, and connected Kafue NP in Zambia with more
central regions of the KAZA-TFCA. Several minor corridors branched off from
these three major corridors; these included a southward connection between the
Okavango-Linyanti and the Central Kalahari Game Reserve, a southwesterly
corridor connecting Luengue-Luiana NP with Namibia's Khaudum NP, and a
northeasterly extension of the Hwange corridor into Zimbabwe's Matusadona NP.
According to our predictions, the landscapes in the Okavango-Linyanti region
were the highest frequented dispersal routes within the KAZA-TFCA
(\Cref{LeastCost}b). Our model did not detect any significant direct corridors
between Zimbabwe and Zambia or Zambia and Angola, and only a very limited W-E
direct connection between the Okavango region and Namibia's Khaudum NP. Except
for the corridor into the Central Kalahari National Park, our model did not
detect any significant connectivity outside the boundaries of the KAZA-TFCA.
Furthermore, we found little to no direct connectivity between peripheral
points; that is, most paths and corridors connecting two adjacent peripheral
points ran through more central regions before heading towards their destination
at the periphery (\Cref{LeastCost}).

\begin{figure}[hbtp]
  \begin{center}
    \begin{minipage}{0.95\textwidth}
      \includegraphics[width = 0.95\textwidth]{99_LeastCostPaths.pdf}
      \includegraphics[width = 0.95\textwidth]{99_LeastCostCorrs.pdf}
    \end{minipage}
    \caption{(a) Source points (black dots) and corresponding least-cost paths
    leaving from protected areas (dark grey borders). Note that only contiguous
    protected areas covering more than 700 km\textsuperscript{2} are depicted.
    Continuous thin black lines indicate the borders of the KAZA-TFCA, whereas
    dashed black lines delineate country-borders. (b) Least-cost corridors
    between the same source points as illustrated in subfigure (a). For ease of
    spatial reference, we also labeled some national parks (NPs, in dark-grey).}
    \label{LeastCost}
  \end{center}
\end{figure}

\section{Discussion}
We used GPS relocation data collected on dispersing African wild dogs to
investigate whether their main movement corridors are contained within the
boundaries of the world's largest transboundary conservation area, namely the
KAZA-TFCA. Our analysis suggests that the KAZA-TFCA indeed encompasses all major
corridors of African wild dogs, demonstrating the potential value of such an
initiative. We thus exemplified how pertinent dispersal data of a highly mobile
species can be used to assess the adequacy of already existing or planned
protected areas. Our approach is neither limited to the African wild dog, nor to
our study area and thus applicable to any study system. All covariates used
throughout this study are readily available on a global scale and many of them
are likely to be important determinants of movement behavior, landscape
permeability, and connectivity for other species \citep{Zeller.2012,
Thurfjell.2014}. Interestingly, our predicted network of least cost-paths and
corridors for African wild dogs shows surprising similarities to corridors of
dispersing lions inhabiting the same ecosystem \citep{Elliot.2014,
Cushman.2018}. This not only reinforces confidence in our own predictions but
also suggests potential synergies for the conservation of these two, and
possibly more, species. Expanding our analytical framework to additional species
will likely yield important insights on the consistency of inter-specific
movement corridors, thus highlighting areas that are exceptionally valuable for
the conservation of several species (e.g. \citealp{Brennan.2020}).

Our results emphasize that human influences constitute some of the main barriers
to connectivity among wild dog populations. According to our model, dispersers
avoided corssing human dominated landscapes whenever given the
choice.\todo{According to reviewer 1 we should point out that our model compares
available vs. unavailable} This conforms to findings on dispersing wild dogs
from eastern Africa \citep{Masenga.2016, Oneill.2020} but conflicts with
findings from South Africa by \cite{DaviesMostert.2012}, who reported a high
willingness of dispersers to cross human-dominated landscapes. We believe that
such differences are due to the unavailability of alternative routes through
natural landscapes, which may have forced dispersers in South Africa to cross
human dominated landscapes despite a strong aversion to do so. In this regard,
our representation of dispersal corridors and the resulting connectivity appear
conservative, as dispersers may be able to make the best out of a bad situation
and cross landscapes characterized by considerably unfavorable conditions
\citep{Palomares.2000, Elliot.2014}. Unfortunately, \cite{DaviesMostert.2012}
only recorded net dispersal distances, thereby precluding a more indepth
understanding of the impact of humans on dispersal behavior.\todo{As per
reviewer 2} Nevertheless, successful conservation of this species relies on
policymakers' and local authorities' willingness and ability to provide and
conserve natural areas that remain free from anthropogenic pressures. This is
not only paramount in light of increasing connectivity and facilitating
dispersal, but also in terms of reducing human-caused mortality during
dispersal. In fact, previous studies have shown that human-caused mortality
represents a major threat to wild dogs' ability to disperse
\citep{Woodroffe.2019, Cozzi.2020}.

Besides human influence, we identified water as additional obstacle to
dispersal. This corroborates earlier studies showing that water bodies are
almost impenetrable to resident packs \citep{Abrahms.2017} and only infrequently
crossed by dispersing individuals \citep{Cozzi.2020}. An accurate and dynamic
representation of water is thus imperative and particularly relevant in seasonal
or flood-pulsing ecosystems such as the Okavango Delta.

Although dispersers avoided moving through water, they selected locations in its
vicinity. This behavior may be caused by the occurence of prey close to water
\citep{Bonyongo.2005}. For the same reason, however, competitors such as lions,
spotted hyenas, and resident wild dogs may also use areas close to water
\citep{Valeix.2010}, thereby occasionally forcing dispersing wild dogs to move
into prey-poorer areas away from water \citep{Creel.2002, Mills.1997}. This
reasoning may also explain why the effect of distance to water was
insignificant.\todo{Reviewer 1 suggested to add this} Given the influence that
resident conspecifics, competitors, and prey can have on dispersers
\citep{Cozzi.2018, Armansin.2019} future studies should strive to collect and
incorporate intra- and interspecific relationships into analyses of landscape
connectivity.

Overall, our findings on habitat selection during dispersal coincide with
findings from dipsersing wild dogs in Kenya \citep{Oneill.2020}, suggesting that
there are strong commonalities dispersers from these very different ecosystems.
Thus, despite wild dogs' ability to cope with diverse habitats and adapt to
changing conditions \citep{Woodroffe.2011}, the fundamental factors included in
our study appear to influence wild dogs from other ecosystems alike.
Nevertheless, expanding our analysis to dispersers emigrating from other source
populations would invaluably contribute to our understanding of
dispersal.\todo{Editor desired info on applicability}

Locally, we identified the Okavango-Linyanti region as a potential dispersal hub
through which dispersing wild dogs gain access to more peripheral regions of the
KAZA-TFCA. It appears that the absence of human activities, the central position
within the KAZA-TFCA, and the presence of relatively impermeable water bodies
(e.g. Okavango Delta, Linyanti Swamp) funnel dispersal movements, resulting in a
highly frequented corridor. Furthermore, the lack of permeable areas between
peripheral source points often resulted in corridors stretching longer Euclidean
distances, in an arc-shaped route via a stretch of suitable habitat through the
Okavango-Linyanty ecosystem. This is an expected outcome in case structural and
functional connectivity coincide, i.e. when dispersers follow suitable habitats
to disperse \citep{Fattebert.2015, Hauenstein.2019}.\todo{As per reviewer 2} The
key role of the Okavango-Linyanti region for overall connectivity within the
KAZA-TFCA thus calls for actions to secure its protection status in the future.
While the region is currently a Wildlife Management Area, it has neither the
status of a National Park nor that of a Game Reserve. A similar case of
non-formally protected but key dispersal landscape is represented by the area
south of Kafue NP in Zambia, for which a disruption of its main and narrow
dispersal corridor would result in considerable isolation of its subpopulations.
We also revealed a potential southwards corridor between the Okavango-Linyanti
ecosystem and the Central Kalahari National Park. \cite{Elliot.2014} and
\cite{Cushman.2018} identified a similar corridor for dispersing lions,
suggesting that upholding and protecting a link between those ecosystems is
pivotal. Some areas through which the corridor runs are neither part of the
KAZA-TFCA nor profit from any form of protection status. Human presence and
activities along the national road that longitudinally traverses this corridor
may limit functional connectivity \citep{Cozzi.2020}.

Our approach of identifying movement corridors based on pre-defined start and
end points implicitly assumes that individuals know the end point of their
dispersal journey and that they have almost complete knowledge of associated
movement costs \citep{Panzacchi.2016}. Since dispersers often move into unknown
territory, this may not necessarily be the case \citep{Abrahms.2017,
Cozzi.2020}. However, specification of pre-defined end points might not be
necessary, as the parametrized iSSF model can be used as mechanistic movement
model to simulate dispersal events from known source points, yet without
restricting the domain of potential end points \citep{Signer.2017}.
Consequently, movement corridors would emerge more naturally as the result of a
myriad of simulated dispersal events \citep{Zeller.2020}. Individual based
simulations could also serve to generate permeability surfaces that are based on
steady state utilistation distributions \citep{Avgar.2016, Signer.2017}. Such
surfaces have been shown to reduce the risk of overstimating permeability,
particularly in areas that lie far from suitable habitats \cite{Signer.2017}.
While a simulation-based approach is conceptually straightforward, computational
requirements are tremendous, especially for such a large extent as constituted
by the KAZA-TFCA. \todo{Reviewer 1 pointed out that the permeability surface
likely overestimates permeability in areas far from suitable habitat.
Additionally, reviewer 2 mentioned that our dataset by far suffices for
simulations and we should not argue with sample size.}

Our work shows how dispersal data of a highly mobile species can be used to
\todo{to instead of t, pointed out by the editor and reviewer 1} identify
movement corridors and to assess the adequacy of protected areas. In our case,
the predicted movement corridors of African wild dogs were well contained within
the boundaries of the world's largest transboundary conservation area, namely
the KAZA-TFCA, suggesting that it will significantly contribute to the long-term
viability of this species. Moreover, our connectivity network allowed revealing
potential dispersal hubs through which dispersers gain access to more remote
regions of the study area. Finally, our investigations showed that human
influence constitutes one of the main barriers to dispersal and substantially
reduces landscape connectivity. Successful conservation of wide-ranging species,
such as exemplified by the African wild dog, will therefore be contingent on the
willingness of local authorities, policymakers, and land managers to preserve
areas that remain free from human strains. Ultimately, our work contributes to
the growing field of connectivity studies and provides and easily expandable
framework for assessing the adequacy of already-existing or planned protected
areas.

\section{Authors' Contributions}
D.D.H., D.M.B., A.O. and G.C. conceived the study and designed methodology;
D.M.B., G.C., and J.W.M. collected the data; D.D.H. and D.M.B. analysed the
data; G.C. and A.O. assisted with modelling; D.D.H., D.M.B., and G.C. wrote the
first draft of the manuscript and all authors contributed to the drafts at
several stages and gave final approval for publication.

\section{Data Availability}
GPS movement data of dispersing coalitions will be made available on dryad at
the time of publication. We will also provide example R-code on Github to
illustrate our least-cost modelling approach
(\url{https://github.com/DavidDHofmann}).\todo{As per editor}

\section{Acknowledgements}
We thank the Ministry of Environment and Tourism of Botswana for granting
permission to conduct this research. We thank C. Botes, I. Clavadetscher, and G.
Camenisch for assisting with wild dog immobilizations. We also thank B. Abrahms
for sharing her data of three dispersing wild dogs. Furthermore, we are indebted
to Prof. J. Fieberg, who consulted all statistical aspects of this work and P.
Wolski, from the Okavango Research Institute, who assisted us in generating
dynamic water maps. We also thank the anonymous reviewers for their valuable
comments and suggestions. This study was funded by Albert-Heim Stiftung, Basler
Stiftung für Biologische Forschung, Claraz Foundation, Idea Wild, Jacot
Foundation, National Geographic Society, Parrotia Stiftung, Stiftung Temperatio,
Wilderness Wildlife Trust Foundation, Forschungkredit der Universität Zürich,
and a Swiss National Science Foundation Grant (31003A\_182286) to A. Ozgul.

\newpage
\begingroup
\singlespacing
\bibliography{Literatur}
\endgroup

\end{document}