### 2021-01-04 Chapter_0/04_Manuscript/Manuscript.tex
Implemented minor revisions

\documentclass[abstract=on,10pt,a4paper,bibliography=totocnumbered]{article}
\usepackage[paper=a4paper,left=35mm,right=35mm,top=25mm,bottom=30mm]{geometry}
\usepackage[doublespacing]{setspace}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[round]{natbib}
\usepackage{amsmath}
\usepackage{colortbl}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{gensymb}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage[hidelinks]{hyperref}
\usepackage[nameinlink]{cleveref}
\usepackage{lineno}
\usepackage{multirow}
\usepackage{arydshln}
\usepackage[nomarkers, nolists]{endfloat}
\usepackage{todonotes}

%------------------------------------------------------------------------------
%	Some Styling
%------------------------------------------------------------------------------
% Creating some TikZ styles
\tikzset{
  nonterminal/.style = {rectangle
    , minimum size = 6mm
    , very thick
    , draw = black!
  }
}

% Changing the style of captions in figures etc.
\captionsetup{labelfont=bf, format=plain, font=small}

% Change how equations are referenced
\renewcommand{\theequation}{Equation \arabic{equation}}%

%------------------------------------------------------------------------------
%	Titlepage: Header
%------------------------------------------------------------------------------
\title{Bound within Boundaries: How Well Do Protected Areas Match Movement
Corridors of Their Most Mobile Protected Species?}

% List of Authors
\author{
  David D. Hofmann\textsuperscript{1,\S,*} \and
  Dominik M. Behr\textsuperscript{1,2,*} \and
  John W. McNutt\textsuperscript{2} \and
  Arpat Ozgul\textsuperscript{1} \and
  Gabriele Cozzi\textsuperscript{1,2}
}

% Reduce spacing between authors
\makeatletter
\def\and{%
  \end{tabular}%
  \hskip -0.5em \@plus.17fil\relax
  \begin{tabular}[t]{c}}
\makeatother

% Current Date
\date{\today}

% And here the masterpiece begins
\begin{document}

% Change page numbering
\pagenumbering{gobble}

% Required to be able to cite
\bibliographystyle{apalike}

% Create Titlepage
\maketitle

%------------------------------------------------------------------------------
%	Titlepage: Additional Info
%------------------------------------------------------------------------------
\begin{flushleft}

\vspace{0.5cm}

\textsuperscript{1} Department of Evolutionary Biology and Environmental
Studies, University of Zurich, Winterthurerstarsse 190, 8057 Zurich,
Switzerland.

\textsuperscript{2} Botswana Predator Conservation Trust, Private Bag 13, Maun,
Botswana.

\textsuperscript{\S} Corresponding author (david.hofmann2@uzh.ch)

\textsuperscript{*} Shared first authorship

\vspace{4cm}

\textbf{Running Title:} Connectivity across a Transfrontier Conservation Area.

\vspace{0.5cm}

\textbf{Keywords:} dispersal, habitat selection, integrated step selection
function, Kavango-Zambezi Transfrontier Conservation Area, landscape
connectivity, least-cost corridors, Lycaon pictus, permeability surface,
protected areas, wildlife management

\end{flushleft}

%------------------------------------------------------------------------------
%	Abstract
%------------------------------------------------------------------------------
\newpage
\begin{abstract}
\noindent 1. Conserving and managing large portions of land to connect wildlife
reserves is an increasingly used strategy to maintain and restore connectivity
among wildlife populations. Boundaries of such conservation areas are often
determined based on expert opinion and socio-political constraints, yet the
extent to which they match species' movement corridors is rarely examined. This
is mainly due to a lack of data, particularly on wide-ranging movement behavior
such as dispersal. Nevertheless, empirically assessing the adequacy of protected
areas is key for the implementation of targeted management actions and efficient
use of limited conservation funds.

\noindent 2. Between 2011 and 2019, we collected high-resolution GPS data on 16
dispersing African wild dog (\textit{Lycaon pictus}) coalitions from a
free-ranging population in the Kavango-Zambezi Transfrontier Conservation Area
(KAZA-TFCA). Spanning five countries and 520'000 km\textsuperscript{2} the
KAZA-TFCA is the world's largest transboundary conservation area and a prime
example for international conservation efforts. We used integrated step
selection analysis to estimate habitat selection of dispersers and to create a
permeability surface for the KAZA-TFCA. We compared landscape permeability
across different regions within the KAZA-TFCA as well as outside its boundaries.
Lastly, we calculated least-cost paths and corridors to verify that major
movement routes were adequately encompassed within the KAZA-TFCA.

\noindent 3. Permeability within the boundaries of the KAZA-TFCA was more than
double compared to areas outside it. Furthermore, we observed a five-fold
permeability difference among the five KAZA-TFCA countries. We also showed that
major movement corridors of wild dogs ran within the KAZA-TFCA, although some
minor routes remained formally unprotected.

\noindent 4. Differences in permeability were mainly related to different
degrees of human activities across regions, and to the presence or absence of
rivers, swamps and open water. The relationship between permeability and other
landscape features was less pronounced.

\noindent 5. \textit{Synthesis and Applications}: In this study, we showed how
pertinent dispersal data of a highly mobile species can be used to empirically
evaluate the adequacy of already-existing or planned protected areas.
Furthermore, we observed regional differences in landscape permeability that
highlight the need for a coordinated effort towards maintaining or restoring
connectivity, especially where transboundary dispersal occurs.
\end{abstract}

%------------------------------------------------------------------------------
%	Main Text
%------------------------------------------------------------------------------
\newpage

% Change page numbering
\pagenumbering{arabic}

% Create linenumbers
\linenumbers

\section{Introduction}
Connectivity among subpopulations is a crucial pre-requisite for many species to
thrive and persist \citep{Fahrig.2003}. Accordingly, preserving and protecting
movement corridors between wildlife reserves has become a task of utmost
importance \citep{Doerr.2011, Rudnick.2012}, resulting in an ever-growing number
of large and often transboundary protected areas. While boundaries of such areas
are often drawn according to expert opinion and socio-political needs,
subjective assessments have revealed deficiencies in the past
\citep{Clevenger.2002, Pullinger.2010}. Thus, an empirical evaluation of the
adequacy of already-existing or planned protected areas using pertinent animal
movement data is paramount for targeted use of valuable and scarce conservation
funds \citep{Pullinger.2010}.

In recent years, a growing body of research has used animal relocation data to
identify movement corridors and assess connectivity at large scales (e.g.
\citealp{Chetkiewicz.2006, Squires.2013, Elliot.2014}). Identification of
potential movement corridors typically relies on the estimation of permeability
surfaces, which return the ease or willingness at which the focal species
traverses a specific landscape \citep{Sawyer.2011}. Such surfaces are created
based on species' relative selection strengths \citep{Avgar.2017}, which can be
quantified using a suite of selection functions \citep{Zeller.2012}.
Specifically, selection strengths are estimated by comparing spatial covariates
(e.g. environmental and anthropogenic) at locations visited by the animal to the
same spatial covariates at locations available to the animal
\citep{Zeller.2012}. Importantly, selection functions require adequate landscape
and relocation data that are representative of the process being studied
\citep{Diniz.2020}. Altough selection during residence and dispersal may
coincide \citep{Fattebert.2015}, it appears that relocation data collected on
dispersing individuals outperforms data collected on resident individuals in the
detection of large-scale movement corridors \citep{Elliot.2014, Abrahms.2017,
Diniz.2020}. Nevertheless, dispersal data is inherently difficult to collect and
remains scarce in the connectivity literature \citep{Vasudev.2015}. As such,
most permeability surfaces upon which movement corridors are identified are
created using relocation data collected on resident individuals. This has likely
limited our ability to meaningfully assess the effectiveness of protected areas
in securing connectivity for their protected species.

One initiative that aims at restoring and enhancing connectivity across large
scales is the Kavango-Zambezi Transfrontier Conservation Area (KAZA-TFCA), which
constitutes the world's largest transfrontier conservation area, spanning over
520'000 km\textsuperscript{2} and five countries (\url{www.kavangozambezi.org}).
While the KAZA-TFCA was originally set to facilitate movements of African
elephants (\textit{Loxodonta africana}; \citealp{Tshipa.2017}), it is also key
to the conservation of other wide-ranging species such as African wild dogs
(\textit{Lycaon pictus}; \citealp{Woodroffe.2012, Cozzi.2020}), lions
(\textit{Panthera leo}; \citealp{Elliot.2014, Cushman.2018}), and cheetahs
(\textit{Acinonyx jubatus}; \citealp{Weise.2017}). To date, however, few studies
have attempted to assess the adequacy of the KAZA-TFCA using global positioning
system (GPS) relocation data of its protected species at large spatial scales
\citep{Elliot.2014, Tshipa.2017, Brennan.2020}. Thus, how well the boundaries of
the KAZA-TFCA reflect natural movement patterns and dispersal corridors of its
most mobile protected species is virtually unknown.

Across the KAZA-TFCA, the African wild dog (\textit{Lycaon pictus}) represents a
highly mobile and endangered flagship species for conservation efforts. Once
widespread across the entire Sub-Saharan continent, wild dogs have been widely
extirpated through human persecution, habitat destruction, and disease outbreaks
\citep{Woodroffe.2012}. For these reasons, viable populations mainly occur in
spatially scattered subpopulations within protected areas \citep{Woodroffe.1999,
Woodroffe.2012, VanDerMeer.2014}. Within these subpopulations, wild dogs form
cooperative breeding packs of up to thirty individuals \citep{Creel.2002}, whose
social structure is strongly governed by the process of dispersal
\citep{McNutt.1996, Behr.2020}. Both males and females disperse from their natal
pack, either alone or in same-sex dispersing coalitions, and search for
unrelated mates and a suitable territory to settle \citep{McNutt.1996,
Cozzi.2020, Behr.2020}. During dispersal, wild dogs can cover several hundred
kilometers and cross international borders \citep{Masenga.2016, Woodroffe.2019,
Cozzi.2020}. Despite the importance of dispersal for the long-term viability of
this species, little empirical information is available on habitat selection and
potential movement barriers during dispersal. The few studies that have
collected dispersal data have shown that dispersers quickly move over large
distances, avoid human-dominated landscapes and areas densely covered by trees,
but prefer proximity to water \citep{Masenga.2016, Woodroffe.2019, Oneill.2020,
Cozzi.2020}.

Here, we collected and analyzed GPS relocation data on 16 dispersing wild dogs
in as many dispersing coalitions from a free-ranging population in northern
Botswana to assess the adequacy of the KAZA-TFCA in securing connectivity. We
estimated relative selection strengths towards environmental and anthropogenic
landscape features, and used the obtained coefficients to predict a permeability
surface spanning the entire KAZA-TFCA. We then investigated how landscape
permeability varied regionally and internationally, and we compared permeability
within and outside the KAZA-TFCA boundaries. Finally, we calculated least-cost
paths and corridors to identify major movement routes and to verify that these
were successfully covered by the KAZA-TFCA.

\section{Methods}
\subsection{Study Area}
The study area (centered at -17\degree 13'9''S, 23\degree 56'4''E;
\Cref{StudyArea}a) was outlined by a rectangular bounding box stretching over
1.3 Mio km\textsuperscript{2} and encompassing the entire KAZA-TFCA
(\Cref{StudyArea}b). The KAZA-TFCA lies in the basins of the Okavango and
Zambezi rivers and includes parts of Angola, Botswana, Namibia, Zimbabwe, and
Zambia. With a total area of over 520'000 km\textsuperscript{2} it constitutes
the earth's largest transboundary conservation area and is characterized by
diverse landscapes, including savanna, grassland, and dry or moist woodland
habitats. Rainfall in the study area is seasonal and lasts from November to
March. The KAZA-TFCA also comprises the Okavango Delta, which represents a
highly dynamic hydrological flood-pulsing system \citep{McNutt.1996,
Wolski.2017}. The extent of the flood in the delta greatly changes within and
between years, depending on the amount of rain that descends from the catchment
areas in Angola and reaches the distal ends of the delta between July and August
(Figure S4). The flood drastically affects surrounding landscapes, so that
during maximum extent (ca. 12'000 km\textsuperscript{2}) the delta becomes a
patchy conglomerate of swamps, open water, and islands, whereas these structures
run dry when the flood retracts to its minimum extent (ca. 5'000
km\textsuperscript{2}; \citealp{Wolski.2017}). Despite 36 national parks (NPs)
and other protected areas, there is considerable human influence in some regions
of the KAZA-TFCA, mainly originating from farms, human density, and road
traffic.

\begin{figure}[h]
  \begin{center}
    \begin{tikzpicture}
        \node[anchor=south west,inner sep=0] (image) at (0,0,0) {
        \includegraphics[width=\textwidth]{99_StudyArea.pdf}
        };
        \begin{scope}[x={(image.south east)},y={(image.north west)}]
            % % next four lines will help you to locate the point needed by forming a grid.
            % % comment these four lines in the final picture.
            % \draw[help lines,xstep=.1,ystep=.1] (0,0) grid (1,1);
            % \draw[help lines,xstep=.05,ystep=.05] (0,0) grid (1,1);
            % \foreach \x in {0,1,...,9} { \node [anchor=north] at (\x/10,0) {0.\x}; }
            % \foreach \y in {0,1,...,9} { \node [anchor=east] at (0,\y/10) {0.\y};}
            % % upto here
            \draw[black] (0.202, 0.255) -- (0.36, 0.955);
            \draw[black] (0.202, 0.205) -- (0.36, 0.070);
        \end{scope}
    \end{tikzpicture}
    \caption{Overview of our study area. (a) The red dotted rectangle depicts
    the study area, which was confined by a bounding box encompassing the entire
    KAZA-TFCA. Gray areas indicate remaining wild dog populations according to
    the IUCN \citep{Woodroffe.2012}. (b) The white rectangle illustrates the
    area within which dispersing coalitions were collared. Since Game Reserves
    in Botswana virtually serve the same purpose as National Parks, we use the
    terms interchangeably for this region.}
    \label{StudyArea}
  \end{center}
\end{figure}

\subsection{GPS Relocation Data}
We used a population of free-ranging African wild dogs inhabiting the Okavango
Delta in northern Botswana as a source population for dispersing individuals.
This population has been extensively studied since 1989 \citep{McNutt.1996,
Cozzi.2013, Cozzi.2020, Behr.2020}. Between 2011 and 2019, we systematically
collected GPS relocation data on 16 coalitions of dispersing African wild dogs
(7 female and 9 male coalitions). Candidate dispersing individuals were
identified based on age, number of same‐sex siblings, pack size, and presence of
unrelated individuals of the opposite sex in their pack \citep{McNutt.1996,
Behr.2020}. Selected individuals were immobilized according to protocols
described in \cite{Osofsky.1996}, and fitted with GPS/Satellite radio collars
(\textit{Vertex Lite; Vectronic Aerospace GmbH, Berlin, Germany}) while still
with their natal pack. Collars weighed 330g, accounting for about 1.5\% of a
wild dog's body weight. A 5cm long decomposable cotton piece was added to the
collar belt to guarantee collar drop-off after about 12-18 months. All required
procedures were undertaken and supervised by a Botswana-registered wildlife
veterinarian. During dispersal, GPS collars were programmed to record a GPS
relocation every 4 hours and to regularly transmit data via iridium satellite
system to a base station.

Because we were interested in dispersal behavior only, we discarded any GPS data
collected while individuals were still with their natal packs and after
settlement in a new territory \citep{Cozzi.2020}. We identified the exact time
of emigration and settlement based on direct field observations and through
visual inspection of the net squared displacement (NSD) metric. NSD quantifies
the squared Euclidean distance of a relocation to a reference point
\citep{Borger.2012}, which in our case was the center of the dispersing
coalition's natal home range. Thus, dispersal was deemed to have started when a
coalition had left its natal home range and continued until the NSD metric
remained stationary, indicating that the coalition had successfully settled
(Figure S1). In our analysis, we did not differentiate between male and female
dispersing coalitions, for previous research found little differences between
sexes during dispersal \citep{Woodroffe.2019, Cozzi.2020}.

\subsection{Spatial Covariates}
To investigate relative selection strengths of dispersing wild dogs, we used a
set of geo-referenced covariates (\Cref{Covariates}) that we aggregated in the
categories \textit{land cover}, \textit{protection status}, and
\textit{anthropogenic}. \textit{Land cover} comprised the covariates water cover
(binary), distance to water (continuous), percentage cover by shrubs/grassland
(continuous), and percentage cover by trees (continuous). To capture the pulsing
behavior of the Okavango Delta, we classified satellite imagery and frequently
updated layers for water cover and corresponding layers depicting distance to
water. \textit{Protection status} contained a binary covariate, indicating
whether an area was protected or not. \textit{Anthropogenic} included covariates
rendering the presence of roads (binary), the distance to roads (continuous),
and a proxy for human influence (continuous) that took into account human
density, farming, and roads. We prepared all covariates as spatial raster layers
from freely available online services and from remotely sensed satellite
imagery. To ensure a consistent resolution (i.e. cell-size or grain) across
covariates, we coarsened or interpolated all layers to a resolution of 250m x
250m. For further details on the preparation and source of each covariate, see
Appendix A.3. We performed processing and manipulation of data as well as all
spatial and statistical analyses using R, version 3.6.1 \citep{R.2019}.

\begin{figure}[h]
  \begin{center}
    \includegraphics[width = \textwidth]{99_Covariates.pdf}
    \caption{Overview of spatial covariates that we included in our models. We
    prepared all covariates for the entire study area but for better visibility
    we only plot them for the surroundings of the Okavango Delta. The white
    rectangle in each plot depicts the area within which dispersing coalitions
    were collared. (a) Averaged layer of all dynamic (binary) water maps. (b)
    Percentage cover of trees. (c) Percentage cover of shrubs/grassland.
    Anything that was not covered by trees or shrubs/grassland was deemed to be
    bare land. (d) Protection status of the area. (e) Human influence proxy
    composed of human density, farms, and roads. (f) Distance to nearest road
    (white lines depict actual roads).}
    \label{Covariates}
  \end{center}
\end{figure}

\subsection{Habitat Selection Model}
We used an integrated step selection function (iSSF; \citealp{Avgar.2016}) to
investigate dispersers' relative selection strentghs towards the above-mentioned
spatial covariates. That is, we paired each realized step (i.e. the connecting
line between two consecutive GPS relocations; \citealp{Turchin.1998}) with 24
random steps that were generated by sampling turning angles from a uniform
distribution U(\(-\pi, +\pi\)) and step lengths from a gamma distribution fitted
to realized steps \citep{Avgar.2016}. A realized step and its 24 associated
random steps formed a stratum and received a unique identifier. Along each step,
we extracted the above-mentioned covariates (Table S3), standardized extracted
values using a z-score transformation, and checked for correlation using
Pearson's Correlation Coefficient \(r\). None of the covariates were overly
correlated (\(|r| > 0.6\); \citealp{Latham.2011}) and we retained all of them
for modeling. Our habitat selection model then assumed that dispersing wild dogs
assigned a selection score \(w(x)\) of the following exponential form to each
realized and random step \citep{Fortin.2005}:

\begin{equation}
\label{EQ1}
  w(x) = exp(\beta_1 x_1 + \beta_2 x_2 + ... + \beta_n x_n)
\end{equation}

\noindent The selection score \(w(x)\) of a step depended on its associated
covariates (\(x_1, x_2, ..., x_n\)), as well as on the animal's relative
selection strengths towards these covariates (\(\beta_1, \beta_2, ...,
\beta_n\)). To estimate relative selection strengths (i.e. the \(\beta\)'s) for
each covariate, we used mixed effects conditional logistic regression analysis
as suggested by \cite{Muff.2020}. We implemented their method using the
R-package \textit{glmmTMB} \citep{Mollie.2017} and used dispersing coalition ID
to model random slopes. We also modelled random intercepts with an arbitrary
high variance of 10\textsuperscript{6} to make use of the poisson trick (see
\citealp{Muff.2020}). We defined three movement metrics, namely the cosine of
the turning angle (\(cos(ta)\)), the step length (\(sl\)) and the logarithm of
the step length (\(log(sl)\)), as core covariates and ran stepwise forward model
selection based on Akaike's Information Criterion (AIC; \citealp{Burnham.2002})
for all other covariates. The inclusion of movement metrics served to reduce
biases in estimated habitat selection coefficients that may have arisen due to
movement behavior \citep{Avgar.2016}. To validate the predictive power of the
most parsimonious habitat selection model, we ran k-fold cross-validation for
case-control studies as described in \cite{Fortin.2009} (details in Appendix
A.5).

\subsection{Permeability Surface}
Using the most parsimonious habitat selection model, we predicted a permeability
surface spanning the entire extent of the KAZA-TFCA. That is, we applied
\ref{EQ1} to our spatial covariates and calculated the selection score \(w(x)\)
for each raster cell. Because our representation of water was dynamic, we
collapsed all dynamic water maps into a single map using areas that were covered
by water in at least 10\% of the cases. We used the resulting map to also
calculate a layer returning the distance to water. Because the delta only covers
5\% of the KAZA-TFCA, we considered the use of a single water map to be
appropriate. To reduce the influence of outliers in predicted permeability
scores, we followed \cite{Squires.2013} and curtailed predicted scores between
the 1\textsuperscript{st} and 99\textsuperscript{th} percentile of their
original values. To compare permeability across different regions, we normalized
the permeability surface to a range between 0 (most impermeable) and 1 (most
permeable), and we determined median permeability within and outside the
KAZA-TFCA, within and outside formally protected areas, and within each of the
five KAZA-TFCA countries.

\subsection{Least-Cost Paths and Corridors}
To identify potential movement corridors of dispersing wild dogs, we specified
source points and calculated factorial least-cost paths (LCPs) as well as
factorial least-cost corridors (LCCs) among them \citep{Elliot.2014}. To specify
source points, we applied the omnidirectional \textit{go-through} approach
proposed by \citep{Koen.2014} and placed 68 regularly spaced source points along
the map border. This approach has been shown to reduce potential biases arising
from the selection of inappropriate source locations \citep{Koen.2014,
Pitman.2017}. To assess the sensitivity of our results with respect to the
location of source points, we also applied an alternative method and placed 68
source points within protected areas that are large enough to sustain viable
wild dog populations (further details and corresponding results in Appendix
A.9). In either case, the 68 source points resulted in 2'278 unique pairwise
combinations and therefore 2'278 unique LCPs and LCCs. We computed factorial
LCPs and LCCs between source points using the R-package \textit{gdistance}
(details in Appendix A.6). After computation, we tallied overlapping
LCPs and LCCs, respectively, into single connectivity maps. Because LCPs return
discrete paths, whereas LCCs return continuous corridors, we present both
methods. R-codes showcasing the main steps for sections 2.5 to 2.6 can be
downloaded from GitHub
(\url{https://github.com/DavidDHofmann/LeastCostAnalysis}).

\section{Results}
\subsection{Dispersal Events}
In total, we collected 4'169 GPS relocations during dispersal (Figure S2 \&
Table S1), resulting in an average of 261 (\(SD\) = 207) locations per
dispersing coalition. Coalitions on average dispersed for 48 days (\(SD\) = 44),
covered a mean Euclidean distance of 54 km (\(SD\) = 71) and a cumulative
distance of 597 km (\(SD\) = 508).

\subsection{Habitat Selection Model}
Our most parsimonious habitat selection model (\(\Delta AIC > 2\) than any
alternative model; Table S4) retained the covariates \textit{water},
\textit{distance to water}, \textit{trees}, \textit{shrubs/grassland}, and
\textit{human influence}, beside the fixed covariates \textit{cos(ta)},
\textit{sl}, and \textit{log(sl)} (\Cref{PermeabilityResults}a). Dispersers
avoided moving through water (\(\beta\) = -0.53, 95\% \(CI\) -0.79 to -0.27) but
selected for locations in its vicinity, although the latter effect was not
significant (\(\beta\) = -0.33, 95\% \(CI\) = -0.73 to 0.08). Dispersers avoided
areas that were densely covered by trees (\(\beta\) = -0.31, \(CI\) = -0.47 to
-0.15) and preferred areas covered by shrubs/grassland (\(\beta\) = 0.25, 95\%
\(CI\) = 0.07 to 0.42). Finally, dispersers avoided areas that were influenced
by humans (\(\beta\) = -0.45, 95\% \(CI\) = -0.82 to -0.08). Except for
\textit{distance to water} (\(SD\)\textsubscript{RandomEffect} = 0.57), we
observed little variation between dispersal coalitions' relative selection
strengths (\(SD\)\textsubscript{RandomEffect} \(<\) 0.22 for all other
covariates, see also Figure S8).

\begin{figure}[h]
  \begin{center}
    \includegraphics[width = \textwidth]{99_PermeabilityResults.pdf}
    \caption{(a) Estimated selection coefficients from the most parsimonious
    habitat selection model. Negative coefficients indicate avoidance of a
    covariate, positive coefficients selection of a covariate. Whiskers
    delineate the 95\%-CIs for estimated parameters. Significance codes: ** \(p
    < 0.05\), *** \(p < 0.01\).(b) Results from the k-fold cross validation for
    case-control studies. The left graph shows rank frequencies of
    \textit{realized} steps according to predictions, whereas the right graph
    shows rank frequencies of \textit{randomly selected} steps according to
    predictions. \(\bar{r}_s\) indicates the mean correlation coefficient
    resulting from 100 repetitions of the k-fold cross validation. The blue
    smoothing line was fitted using a locally weighted polynomial regression and
    serves to aid the eye in detecting the trends. Correlation coefficients
    suggest that our prediction was significant and robust, evidenced by the
    fact that the confidence intervals of \(\bar{r}_{s, realized}\) and
    \(\bar{r}_{s, random}\) did overlap and by the fact that there was strong
    and significant correlation between ranks and associated frequency for
    realized steps.}
    \label{PermeabilityResults}
  \end{center}
\end{figure}

Results from the k-fold cross-validation suggested that our prediction was
significant and robust, as highlighted by the fact that the 95\%-CIs intervals
of \(\bar{r}_{s, realized}\) and \(\bar{r}_{s, random}\) did not overlap
(\Cref{PermeabilityResults}b). Likewise, the significant correlation between
ranks and corresponding frequencies for realized steps suggested a good fit
between predictions and observations (\Cref{PermeabilityResults}b).

\subsection{Permeability Surface}
Our prediction of landscape permeability revealed substantial differences across
regions in the study area (\Cref{PermeabilityMap}). Comparisons of median
permeability values (\Cref{PermeabilityComp}) showed that permeability inside
the KAZA-TFCA was more than two times as high as permeability outside it.
Permeability varied by country, with a five-fold permeability difference among
them. Angola and Botswana were characterized by comparably highly permeable
landscapes, Zimbabwe and Zambia were relatively impermeable, and Namibia ranged
in between the two extremes (\Cref{PermeabilityComp}). Visual inspection of our
covariate layers indicated that high permeability in Angola and Botswana was
mainly related to a combination of low human influence, low tree cover, high
shrubs/grassland cover, and a close distance to water. Although swamps,
wetlands, and permanent water themselves provided little permeability, their
surroundings acted as strong attractants to dispersers. The low permeability
that characterized Zambia and Zimbabwe, on the other hand, was mainly caused by
substantial human influence. Albeit the KAZA-TFCA covered most permeability
hot-spots, several highly permeable regions remained uncovered by its borders.
Across all countries, protected areas provided roughly double the permeability
of unprotected landscapes (\Cref{PermeabilityComp}).

\begin{figure}[hbtp]
  \begin{center}
    \begin{tikzpicture}
        \node[anchor=south west,inner sep=0] (image) at (0,0,0) {
          \begin{minipage}{\textwidth}
            \includegraphics[width = \textwidth]{99_PermeabilityMap.pdf}
          \end{minipage}
        };
        \begin{scope}[x={(image.south east)},y={(image.north west)}]
        \end{scope}
    \end{tikzpicture}
    \caption{Predicted permeability surface for the extent of the KAZA-TFCA.
    Permeability was predicted by calculating selection scores \(w(x) =
    exp(\beta_1 x_1 + \beta_2 x_2 + ... + \beta_n x_n)\) for each raster cell
    based on the raster cell's underlying covariates (\(x_i\)) and estimated
    selection strength (\(\beta_i\)). Areas that dispersers find easy to
    traverse are depicted in bright colors. Bold white lines delineate the
    borders of the KAZA-TFCA, whereas dashed white lines show country borders.}
    \label{PermeabilityMap}
  \end{center}
\end{figure}

\begin{table}[h]
  \begin{center}
  \caption{Comparison of median permeability (interquantile range in brackets)
  across countries, separated into areas within and outside the KAZA-TFCA, as
  well as within and outside formally protected areas. High values indicate high
  permeability, whereas low values correspond to low permeability.}
  \label{PermeabilityComp}
  \begin{tabular}{llllll}
    \toprule
    \multicolumn{1}{c}{} & \multicolumn{2}{c}{KAZA-TFCA} &
    \multicolumn{2}{c}{Protection Status} & \multicolumn{1}{c}{} \\
    \cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
    Country & Inside & Outside & Protected & Pastoral & Overall\\
    \midrule
    Angola & 0.35 (0.41) & 0.12 (0.32) & 0.35 (0.41) & 0.12 (0.32) & 0.19 (0.38)\\
    Botswana & 0.24 (0.30) & 0.14 (0.16) & 0.27 (0.35) & 0.14 (0.18) & 0.18 (0.25)\\
    Namibia & 0.20 (0.30) & 0.12 (0.17) & 0.22 (0.30) & 0.10 (0.14) & 0.14 (0.24)\\
    Zambia & 0.05 (0.09) & 0.02 (0.05) & 0.04 (0.09) & 0.03 (0.05) & 0.03 (0.06)\\
    Zimbabwe & 0.06 (0.16) & 0.05 (0.04) & 0.07 (0.17) & 0.04 (0.04) & 0.05 (0.06)\\
    \hline
    Overall & 0.15 (0.29) & 0.06 (0.14) & 0.14 (0.30) & 0.06 (0.14) & 0.08 (0.21)\\
    \bottomrule
  \end{tabular}
  \end{center}
\end{table}

\subsection{Least-Cost Paths \& Least-Cost Corridors}
Our least-cost analysis revealed three major movement corridors of which all
were well-contained within the KAZA-TFCA boundaries (\Cref{LeastCost}). One
major corridor ran SE-NW and connected the Okavango-Linyanti ecosystem in
Botswana with Luengue-Luiana NP in Angola. A second corridor ran W-E between
Chobe NP in Botswana and Zimbabwe's Hwange NP. A third major corridor ran NE-SW,
completely across unprotected areas, and connected Kafue NP in Zambia with more
central regions of the KAZA-TFCA. Several minor corridors branched off from
these three major corridors; these included a southward connection between the
Okavango-Linyanti and the Central Kalahari Game Reserve, a southwesterly
corridor connecting Luengue-Luiana NP with Namibia's Khaudum NP, and a
northeasterly extension of the Hwange corridor into Zimbabwe's Matusadona NP.
According to our predictions, the landscapes in the Okavango-Linyanti region
were the highest frequented dispersal routes within the KAZA-TFCA
(\Cref{LeastCost}b). Our model did not detect any significant direct corridors
between Zimbabwe and Zambia or Zambia and Angola, and only a very limited W-E
direct connection between the Okavango region and Namibia's Khaudum NP. Except
for the corridor into the Central Kalahari National Park, our model did not
detect any significant connectivity outside the boundaries of the KAZA-TFCA.
Furthermore, we found little to no direct connectivity between peripheral
points; that is, most paths and corridors connecting two adjacent peripheral
points ran through more central regions before heading towards their destination
at the periphery (\Cref{LeastCost}).

\begin{figure}[hbtp]
  \begin{center}
    \begin{minipage}{0.95\textwidth}
      \includegraphics[width = 0.95\textwidth]{99_LeastCostPaths.pdf}
      \includegraphics[width = 0.95\textwidth]{99_LeastCostCorrs.pdf}
    \end{minipage}
    \caption{(a) Source points (black semicircles along the map border) and
    corresponding least-cost paths between them. Continuous thin black lines
    indicate the borders of the KAZA-TFCA, whereas dashed black lines delineate
    country-borders. (b) Least-cost corridors between the same source points as
    illustrated in subfigure (a). For ease of spatial reference, we also labeled
    some national parks (NPs, in dark-grey).}
    \label{LeastCost}
  \end{center}
\end{figure}

\section{Discussion}
We used GPS relocation data collected on dispersing African wild dogs to
investigate whether their main movement corridors are contained within the
boundaries of the world's largest transboundary conservation area, namely the
KAZA-TFCA. Our analysis suggests that the KAZA-TFCA indeed encompasses all major
corridors of African wild dogs, demonstrating the potential value of such an
initiative. We thus exemplified how pertinent dispersal data of a highly mobile
species can be used to assess the adequacy of already existing or planned
protected areas. Our approach is neither limited to the African wild dog, nor to
our study area and thus applicable to any study system. All covariates used
throughout this study are readily available on a global scale and many of them
are likely to be important determinants of movement behavior, landscape
permeability, and connectivity for other species \citep{Zeller.2012,
Thurfjell.2014}. Interestingly, our predicted network of least cost-paths and
corridors for African wild dogs shows surprising similarities to corridors of
dispersing lions inhabiting the same ecosystem \citep{Elliot.2014,
Cushman.2018}. This not only reinforces confidence in our own predictions but
also suggests potential synergies for the conservation of these two, and
possibly more, species. Expanding our analytical framework to additional species
will likely yield important insights on the consistency of inter-specific
movement corridors, thus highlighting areas that are exceptionally valuable for
the conservation of several species (e.g. \citealp{Brennan.2020}).

Our results emphasize that human influences constitute some of the main barriers
to connectivity among wild dog populations. This conforms to findings on
dispersing wild dogs from eastern Africa \citep{Masenga.2016, Oneill.2020} but
conflicts with findings from South Africa by \cite{DaviesMostert.2012}, who
reported a high willingness of dispersers to cross human-dominated landscapes.
Such differences may arise from the fact that our model infers preferences by
comparing \textit{used} and \textit{available} habitats, whereas
\cite{DaviesMostert.2012} only recorded net dispersal distances, thereby
precluding such an analysis. Thus, we believe that differences to
\cite{DaviesMostert.2012} may be explained by the unavailability of alternative
routes through natural landscapes, which may have forced dispersers in South
Africa to cross human dominated landscapes despite a strong aversion to do so.
In this regard, our representation of dispersal corridors and the resulting
connectivity appear conservative, as dispersers may be able to make the best out
of a bad situation and cross landscapes characterized by considerably
unfavorable conditions \citep{Palomares.2000, Elliot.2014}. Nevertheless,
successful conservation of this species relies on policymakers' and local
authorities' willingness and ability to provide and conserve natural areas that
remain free from anthropogenic pressures. This is not only paramount in light of
increasing connectivity and facilitating dispersal, but also in terms of
reducing human-caused mortality during dispersal. In fact, previous studies have
shown that human-caused mortality represents a major threat to wild dogs'
ability to disperse \citep{Woodroffe.2019, Cozzi.2020}.

Besides human influence, we identified water as additional obstacle to
dispersal. This corroborates earlier studies showing that water bodies are
almost impenetrable to resident packs \citep{Abrahms.2017} and only infrequently
crossed by dispersing individuals \citep{Cozzi.2020}. An accurate and dynamic
representation of water is thus imperative and particularly relevant in seasonal
or flood-pulsing ecosystems such as the Okavango Delta.

Although dispersers avoided moving through water, they selected locations in its
vicinity. This behavior may be caused by the occurence of prey close to water
\citep{Bonyongo.2005}. For the same reason, however, competitors such as lions,
spotted hyenas, and resident wild dogs may also use areas close to water
\citep{Valeix.2010}, thereby occasionally forcing dispersing wild dogs to switch
behavior and move into prey-poorer areas away from water \citep{Creel.2002,
Mills.1997}. This may explain the large confidence intervals for the
corresponding \(\beta\)-estimate of \textit{distance to water}. Given the
influence that resident conspecifics, competitors, and prey can have on
dispersers \citep{Cozzi.2018, Armansin.2019} future studies should strive to
collect and incorporate intra- and interspecific relationships into analyses of
landscape connectivity.

Overall, our findings on habitat selection during dispersal coincide with
findings from dipsersing wild dogs in Kenya \citep{Oneill.2020} and Tanzania
\citep{Masenga.2016}, suggesting that there are strong commonalities between
dispersers from these very different ecosystems. Thus, despite wild dogs'
ability to cope with diverse habitats and adapt to changing conditions
\citep{Woodroffe.2011}, the fundamental factors included in our study appear to
influence wild dogs from other ecosystems alike. Nevertheless, expanding our
analysis to dispersers emigrating from other source populations would invaluably
contribute to our understanding of dispersal.

Locally, we identified the Okavango-Linyanti region as a potential dispersal hub
through which dispersing wild dogs gain access to more peripheral regions of the
KAZA-TFCA. It appears that the absence of human activities, the central position
within the KAZA-TFCA, and the presence of relatively impermeable water bodies
(e.g. Okavango Delta, Linyanti Swamp) funnel dispersal movements, resulting in a
highly frequented corridor. Furthermore, the lack of permeable areas between
peripheral source points often resulted in corridors stretching longer Euclidean
distances, in an arc-shaped route via a stretch of suitable habitat through the
Okavango-Linyanty ecosystem. This is an expected outcome in case structural and
functional connectivity coincide, i.e. when dispersers follow suitable habitats
to disperse \citep{Fattebert.2015, Hauenstein.2019}. The key role of the
Okavango-Linyanti region for overall connectivity within the KAZA-TFCA thus
calls for actions to secure its protection status in the future. While the
region is currently a Wildlife Management Area, it has neither the status of a
National Park nor that of a Game Reserve. A similar case of non-formally
protected but key dispersal landscape is represented by the area south of Kafue
NP in Zambia, for which a disruption of its main and narrow dispersal corridor
would result in considerable isolation of its subpopulations. We also revealed a
potential southwards corridor between the Okavango-Linyanti ecosystem and the
Central Kalahari National Park. \cite{Elliot.2014} and \cite{Cushman.2018}
identified a similar corridor for dispersing lions, suggesting that upholding
and protecting a link between those ecosystems is pivotal. Some areas through
which the corridor runs are neither part of the KAZA-TFCA nor profit from any
form of protection status. Human presence and activities along the national road
that longitudinally traverses this corridor may limit functional connectivity
\citep{Cozzi.2020}.

Our approach of identifying movement corridors based on pre-defined start and
end points implicitly assumes that individuals know the end point of their
dispersal journey and that they have almost complete knowledge of associated
movement costs \citep{Panzacchi.2016}. Since dispersers often move into unknown
territory, this may not necessarily be the case \citep{Abrahms.2017,
Cozzi.2020}. However, specification of pre-defined end points might not be
necessary, as the parametrized iSSF model can be used as mechanistic movement
model to simulate dispersal from known source points, yet without restricting
the domain of potential end points \citep{Signer.2017}. Consequently, movement
corridors would emerge more naturally as the result of a myriad of simulated
dispersal events \citep{Allen.2016, Zeller.2020}.

Besides estimating corridors, individual based simulations may be used to
generate permeability surfaces \citep{Avgar.2016, Signer.2017}. Such simulation
based surfaces have been shown to reduce the risk of overestimating permeability
(\(w(x)\)) and conesquently connectivity, particularly in areas that lie far
from suitable habitats \citep{Signer.2017}. While a simulation-based approach is
conceptually straightforward, computational requirements for such a large
spatial extent as the KAZA-TFCA are very high, making the use of this approach
challenging. We therefore urge future studies to optimize the simulation of
movement from iSSFs to capture a more mechanistic model of dispersal.

Our work shows how dispersal data of a highly mobile species can be used to
identify movement corridors and to assess the adequacy of protected areas. In
our case, the predicted movement corridors of African wild dogs were well
contained within the boundaries of the world's largest transboundary
conservation area, namely the KAZA-TFCA, suggesting that it will significantly
contribute to the long-term viability of this species. Moreover, our
connectivity network allowed revealing potential dispersal hubs through which
dispersers gain access to more remote regions of the study area. Finally, our
investigations showed that human influence constitutes one of the main barriers
to dispersal and substantially reduces landscape connectivity. Successful
conservation of wide-ranging species, such as exemplified by the African wild
dog, will therefore be contingent on the willingness of local authorities,
policymakers, and land managers to preserve areas that remain free from human
strains. Ultimately, our work contributes to the growing field of connectivity
studies and provides and easily expandable framework for assessing the adequacy
of already-existing or planned protected areas.

\section{Authors' Contributions}
D.D.H., D.M.B., A.O. and G.C. conceived the study and designed methodology;
D.M.B., G.C., and J.W.M. collected the data; D.D.H. and D.M.B. analysed the
data; G.C. and A.O. assisted with modelling; D.D.H., D.M.B., and G.C. wrote the
first draft of the manuscript and all authors contributed to the drafts at
several stages and gave final approval for publication.

\section{Data Availability}
GPS movement data of dispersing coalitions will be made available on dryad at
the time of publication. R-code showcasing the main steps for sections 2.5 to
2.6 can be downloaded from GitHub
(\url{https://github.com/DavidDHofmann/LeastCostAnalysis}).

\section{Acknowledgements}
We thank the Ministry of Environment and Tourism of Botswana for granting
permission to conduct this research. We thank C. Botes, I. Clavadetscher, and G.
Camenisch for assisting with wild dog immobilizations. We also thank B. Abrahms
for sharing her data of three dispersing wild dogs. Furthermore, we are indebted
to Prof. J. Fieberg, who consulted all statistical aspects of this work and P.
Wolski, from the Okavango Research Institute, who assisted us in generating
dynamic water maps. We also thank the anonymous reviewers for their valuable
comments and suggestions. This study was funded by Albert-Heim Stiftung, Basler
Stiftung für Biologische Forschung, Claraz Foundation, Idea Wild, Jacot
Foundation, National Geographic Society, Parrotia Stiftung, Stiftung Temperatio,
Wilderness Wildlife Trust Foundation, Forschungkredit der Universität Zürich,
and a Swiss National Science Foundation Grant (31003A\_182286) to A. Ozgul.

\newpage
\begingroup
\singlespacing
\bibliography{Literatur}
\endgroup

\end{document}