### 2021-04-22 Chapter_1/04_Manuscript/99_MovementModelAICs.tex
Worked on manuscript for 1 chap

% latex table generated in R 3.6.3 by xtable 1.8-4 package
% Thu Apr 22 09:09:15 2021
\begin{tabular}{lrrrrr}
  \toprule
Covariates & AIC & DeltaAIC & Weight & LogLik & ID \\ 
  \midrule
cos(ta) + log(sl) + Base Model + sl:LA + WA:sl + log\_sl:cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) & 89392.88 & 0.00 & 0.15 & -44670.44 &   1 \\ 
  cos(ta) + log(sl) + Base Model + sl:LA + WA:sl + log\_sl:cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + SH:log\_sl & 89393.92 & 1.04 & 0.09 & -44669.96 &   2 \\ 
  cos(ta) + log(sl) + Base Model + sl:LA + WA:sl + log\_sl:cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + DTW:log\_sl & 89394.13 & 1.25 & 0.08 & -44670.06 &   3 \\ 
  cos(ta) + log(sl) + Base Model + sl:LA + WA:sl + log\_sl:cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + WO:log\_sl & 89394.25 & 1.37 & 0.08 & -44670.13 &   4 \\ 
  cos(ta) + log(sl) + Base Model + sl:LA + WA:sl + log\_sl:cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl & 89394.36 & 1.48 & 0.07 & -44672.18 &   5 \\ 
  cos(ta) + log(sl) + Base Model + sl:LA + WA:sl + log\_sl:cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + log\_sl:LA & 89394.44 & 1.56 & 0.07 & -44670.22 &   6 \\ 
  cos(ta) + log(sl) + Base Model + sl:LA + WA:sl + log\_sl:cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + HI:sl & 89394.56 & 1.68 & 0.07 & -44670.28 &   7 \\ 
  cos(ta) + log(sl) + Base Model + sl:LA + WA:sl + log\_sl:cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + WA:log\_sl & 89394.57 & 1.69 & 0.07 & -44670.29 &   8 \\ 
  cos(ta) + log(sl) + Base Model + sl:LA + WA:sl + log\_sl:cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + WO:cos(ta) & 89394.59 & 1.71 & 0.07 & -44670.30 &   9 \\ 
  cos(ta) + log(sl) + Base Model + sl:LA + WA:sl + log\_sl:cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + WA:cos(ta) & 89394.63 & 1.75 & 0.06 & -44670.31 &  10 \\ 
  cos(ta) + log(sl) + Base Model + sl:LA + WA:sl + log\_sl:cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + sl:cos(ta) & 89394.68 & 1.80 & 0.06 & -44672.34 &  11 \\ 
  cos(ta) + log(sl) + Base Model + sl:LA + WA:sl + log\_sl:cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + HI:log\_sl & 89394.69 & 1.81 & 0.06 & -44670.35 &  12 \\ 
  cos(ta) + log(sl) + Base Model + sl:LA + WA:sl + log\_sl:cos(ta) + DTW:cos(ta) + WO:sl + HI:cos(ta) + SH:sl + DTW:sl + sl:cos(ta) + SH:cos(ta) & 89394.84 & 1.96 & 0.06 & -44670.42 &  13 \\ 
   \bottomrule
\end{tabular}