### 2022-02-07 Chapter_1/04_Manuscript/Manuscript.tex
updates

\documentclass[abstract=on,10pt,a4paper,bibliography=totocnumbered]{article}
\usepackage[paper=a4paper,left=35mm,right=35mm,top=25mm,bottom=30mm]{geometry}
\usepackage[doublespacing]{setspace}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[round]{natbib}
\usepackage{amsmath}
\usepackage{colortbl}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{gensymb}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage[hidelinks]{hyperref}
\usepackage[nameinlink]{cleveref}
% \usepackage{lineno}
\usepackage{multirow}
\usepackage{arydshln}
\usepackage[flushleft]{threeparttable}
\usepackage[nomarkers, nolists]{endfloat}
\usepackage[colorinlistoftodos]{todonotes}

%------------------------------------------------------------------------------
%	Some Styling
%------------------------------------------------------------------------------
% Creating some TikZ styles
\tikzset{
  nonterminal/.style = {rectangle
    , minimum size = 6mm
    , very thick
    , draw = black!
  }
}

% Changing the style of captions in figures etc.
\captionsetup{labelfont=bf, format=plain, font=small}

% Change how equations are referenced
\renewcommand{\theequation}{Equation \arabic{equation}}%

%------------------------------------------------------------------------------
%	Titlepage: Header
%------------------------------------------------------------------------------
\title{A Thee-Step Approach for Assessing Landscape Connectivity via Simulated
Dispersal: African Wild Dog Case Study}

% List of Authors
\author{
  David D. Hofmann\textsuperscript{1,2,\S} \and
  Gabriele Cozzi\textsuperscript{1,2} \and
  John W. McNutt\textsuperscript{2} \and
  Arpat Ozgul\textsuperscript{1} \and
  Dominik M. Behr\textsuperscript{1,2}
}

% Reduce spacing between authors
\makeatletter
\def\and{%
  \end{tabular}%
  \hskip -0.5em \@plus.17fil\relax
  \begin{tabular}[t]{c}}
\makeatother

% Current Date
% \date{\today}

% And here the masterpiece begins
\begin{document}

% Change page numbering
\pagenumbering{gobble}

% Required to be able to cite
\bibliographystyle{apalike}

% Create Titlepage
\maketitle

%------------------------------------------------------------------------------
%	Titlepage: Additional Info
%------------------------------------------------------------------------------
\begin{flushleft}

\vspace{0.5cm}

\textsuperscript{1} Department of Evolutionary Biology and Environmental
Studies, University of Zurich, Winterthurerstarsse 190, 8057 Zurich,
Switzerland.

\textsuperscript{2} Botswana Predator Conservation Program, Private Bag 13,
Maun, Botswana.

\textsuperscript{\S} Corresponding author (david.hofmann2@uzh.ch)

\vspace{4cm}

\textbf{Running Title:} Simulating Wild Dog Dispersal Trajectories to Assess
Landscape Connectivity

\vspace{0.5cm}

\textbf{Keywords:} dispersal, simulation, movement, integrated step-selection
function, Kavango-Zambezi Transfrontier Conservation Area, landscape
connectivity, Lycaon pictus

\end{flushleft}

%------------------------------------------------------------------------------
%	Abstract
%------------------------------------------------------------------------------
% \newpage
% \begin{abstract}
% %------------------------------------------------------------------------------
% % Short Version
% %------------------------------------------------------------------------------
% \textbf{Short Version }
% The ability to disperse is contingent on a sufficient degree of landscape
% connectivity, which is why the identification and preservation of movement
% corridors that promote connectivity has become a task of extraordinary
% importance. Currently, ecologists rely on least-cost analysis or circuit theory
% to investigate connectivity, although both methods make several assumptions that
% are hardly met in reality. To address these issues, simulations from
% individual-based movement models have been proposed, yet a unified framework to
% simulate dispersal and quantify connectivity is lacking.
%
% Here, we propose a simple three-step approach that combines several
% already-existing methods to assess connectivity using simulated dispersal
% trajectories. In step one, we use integrated step-selection functions to
% parametrize a mechanistic movement model rendering dispersal behavior. In step
% two, we apply the parametrized model as an individual-based movement model to
% simulate dispersal trajectories. In step three, we combine simulated
% trajectories into three complementary connectivity maps, each focusing on a
% different aspect of landscape connectivity.
%
% We showcase the application of the proposed approach using empirical data of
% dispersing African wild dogs (\textit{Lycaon pictus}) and assess landscape
% connectivity within the world's largest transboundary conservation area in
% Southern Africa. We thereby shed light into dispersing wild dogs' habitat and
% movement preferences, while also uncovering crucial dispersal corridors. With
% this analysis, we demonstrate that simulations from integrated step-selection
% functions offer a simple, yet powerful alternative to traditional connectivity
% modeling techniques.
% \end{abstract}

\newpage
\begin{abstract}
%------------------------------------------------------------------------------
%  LONG VERSION
%------------------------------------------------------------------------------
Dispersal of individuals contributes to long-term population persistence, yet
requires a sufficient degree of landscape connectivity. To date, connectivity
has mainly been investigated using least-cost analysis and circuit theory, two
methods that make assumptions that are hardly applicable to dispersal.
Least-cost analysis assumes that animals move towards a known endpoint and are
knowledgeable about the most favorable route. Circuit theory presumes a complete
random walk and fails to incorporate directional persistence. While these
assumptions can be relaxed by simulating dispersal movements explicitly across
the landscape, a unified approach for such simulations is lacking.

Here, we present a simple three-step approach to simulate dispersal movements
and to assess connectivity using empirical GPS movement data and a set of
habitat covariates. In step one, integrated step-selection functions are used to
fit a mechanistic movement model describing habitat and movement preferences of
dispersing individuals. In step two, the parameterized model is used to simulate
dispersal trajectories across the study area. In step three, a set of three
complementary connectivity maps is derived: a heatmap that highlights frequently
traversed areas, a betweenness map that pinpoints dispersal corridors, and a map
of inter-patch connectivity that indicates the presence and intensity of
functional links between habitat patches. As a case study, we applied the
approach to GPS data collected on dispersing individuals of the endangered
African wild dog (\textit{Lycaon pictus}) inhabiting the Kavango-Zambezi
Transfrontier Conservation Area (KAZA-TFCA) in southern Africa.

Besides providing insights into the movement and habitat preferences of
dispersing individuals, our movement model also rendered how movement behavior
depended on habitat conditions. The model thereby substantially outperformed a
model that omitted such interactions. While we simulated a total of 80,000
dispersal trajectories, metrics of connectivity only marginally changed beyond
10,500 simulations, suggesting that fewer simulations would have sufficed. The
three connectivity maps derived from simulations revealed several dispersal
hotspots and corridors across the extent of our study area. Each map emphasized
a different aspect of landscape connectivity, thus highlighting their
complementary nature. Overall, our case study demonstrates that a
simulation-based approach utilizing step-selection functions offers a simple yet
powerful alternative to traditional connectivity modeling techniques. Our
approach not only makes fewer unrealistic assumptions about dispersal but also
permits a more mechanistic understanding of dispersal movements and landscape
connectivity. It is thus useful for a variety of applications in ecological,
evolutionary, and conservation research.

\end{abstract}

%------------------------------------------------------------------------------
%	Main Text
%------------------------------------------------------------------------------
\newpage

\onehalfspacing
\tableofcontents
\doublespacing

% Change page numbering
\newpage
\pagenumbering{arabic}

% Create linenumbers
% \linenumbers

\section{Introduction}

% Importance of Dispersal & Connectivity
\subsection{Importance of Connectivity \& Connectivity Models}
Dispersal of individuals is a vital process that allows species to maintain
genetic diversity \citep{Perrin.1999, Perrin.2000, Frankham.2002, Leigh.2012,
Baguette.2013}, rescue non-viable populations \citep{Brown.1977}, and colonize
unoccupied habitats \citep{Hanski.1999b, MacArthur.2001}. However, the ability
to disperse depends on a sufficient degree of landscape connectivity
\citep{Fahrig.2003, Clobert.2012}, making the identification and protection of
dispersal corridors that promote connectivity a task of fundamental importance
\citep{Nathan.2008, Doerr.2011, Rudnick.2012}. Identifying dispersal corridors
not only necessitates a comprehensive understanding of the factors that limit
dispersal, but also an appropriate model to estimate connectivity
\citep{Baguette.2013, Vasudev.2015, Hofmann.2021}. To date, the most commonly
used models are least-cost path analysis (LCPA; \citealp{Adriaensen.2003}) and
circuit theory (CT; \citealp{McRae.2006, McRae.2008}). Unfortunately, both
models rest on assumptions that appear unsuitable for dispersers, thus calling
for the development of alternative approaches. One promising alternative is to
assess landscape connectivity via simulated dispersal trajectories generated
from individual-based movement models (IBMMs, \citealp{Diniz.2019}). However,
the large number of subjective modeling decisions required by IBMMs renders
among-system comparisons difficult.

% Issues with Both Methods
\subsection{Issues with Traditional Connectivity Models}
Traditional connectivity models make assumptions that are rarely met for
dispersers. LCPA, for instance, assumes that individuals move towards a
preconceived endpoint and choose a cost-minimizing route accordingly
\citep{Sawyer.2011, Abrahms.2017}. While this assumption may be justifiable for
migrating animals, it is unlikely to hold for dispersers, as dispersers
typically move across unfamiliar territory towards an unknown endpoint
\citep{Koen.2014, Cozzi.2020}. CT, on the contrary, posits that animals move
according to a random walk, entailing that autocorrelation between subsequent
movements cannot be rendered \citep{Diniz.2019}. For dispersers, however,
autocorrelated movements are regularly observed \citep{Cozzi.2020,
Hofmann.2021}, meaning that dispersal trajectories are usually strongly
directional. Moreover, because both models require static permeability or
resistance surfaces as input, they are unable to reflect the temporal dimension
of dispersal, such that statements about the expected duration for moving
between habitat patches are impossible \citep{Martensen.2017, Diniz.2019}.

% What about IBMMs?
\subsection{What about IBMMs?}
The shortcomings inherent to LCPA and CT can be overcome by simulating dispersal
using IBMMs and by converting simulated trajectories into meaningful measures of
connectivity \citep{Diniz.2019}. In contrast to LCPA and CT, IBMMs allow to
explicitly simulate how individuals move across and interact with the
encountered landscape \citep{Kanagaraj.2013, Clark.2015, Allen.2016,
Hauenstein.2019, Zeller.2020}, as well as to render potential interactions
between movement behavior and habitat conditions \citep{Avgar.2016}. This
strictly shifts the focus from a structural to a more functional view on
connectivity \citep{Tischendorf.2000, Kanagaraj.2013, Hauenstein.2019}.
Furthermore, IBMMs generate movement sequentially, i.e. they generate a series
of steps, so that the temporal dimension of dispersal becomes explicit and
allows modeling autocorrelation between consecutive steps \citep{Diniz.2019}.
Finally, simulations from IBMMs do not enforce movement or connections towards
preconceived endpoints, thereby preventing biases arising from misplaced
endpoints. Despite these advantages, a unifying framework to simulate dispersal
and assess connectivity using IBMMs is lacking. Considering the large number of
subjective decisions entailed by IBMMs, an approach that streamlines and unifies
the application of dispersal simulations to assess connectivity will, however,
be critical to safeguard interspecfic comparability and reproducability among
studies.

% Proposed Solution: Three-Step approach
\subsection{Proposed Solution: Three-Step Approach}
Here, we propose and exemplify a simple three-step approach for simulating
dispersal and assessing landscape connectivity (\Cref{GraphicalAbstract}). In
step one, GPS movement data of dispersing individuals is combined with habitat
covariates to fit a mechanistic movement model via integrated step-selection
functions (ISSFs, \citealp{Avgar.2016}). ISSFs allow inference on the study
species' habitat kernel (i.e. habitat preferences), its movement kernel (i.e.
movement preferences/capabilities), as well as potential interactions between
the two \citep{Avgar.2016, Fieberg.2021}. In step two, the parametrized movement
model is used to simulate individual dispersal trajectories starting from
randomly placed source points. Comparable simulations have already been applied
to estimate steady-state utilization distributions of resident individuals
\citep{Potts.2013, Signer.2017} and to model landscape connectivity, yet
disregarding interdependencies between habitat and movement kernels
\citep{Clark.2015, Zeller.2020}. Finally, in step three, the simulated
trajectories are converted into three complementary connectivity maps; (i) a
heatmap revealing areas that are frequently traversed by dispersers (e.g.
\citealp{Hauenstein.2019, Zeller.2020}), (ii) a betweenness-map delineating
dispersal corridors and bottlenecks (e.g. \citealp{BastilleRousseau.2018}),
(iii) and a map of inter-patch connectivity, depicting the presence and
intensity of functional links between habitat patches, as well as the average
dispersal duration required for the realization of those connections (e.g.
\citealp{Gustafson.1996, Kanagaraj.2013}).

% Here, we use GPS data of 16 dispersing wild dogs originating from a
% free-ranging population in northern Botswana to parametrize a mechanistic
% movement model, which we then employed to simulate 80,000 dispersal
% trajectories across the landscape of the world's largest transboundary
% conservation area, the Kavango-Zambezi Transfrontier Conservation Area
% (KAZA-TFCA).

% Introduction of the study species
\subsection{Case Study}
We showcase the application of the proposed approach (\Cref{GraphicalAbstract})
using GPS movement data collected on dispersing individuals of the endangered
African wild dog (\textit{Lycaon pictus}). The African wild dog is a highly
mobile species whose population persistence heavily relies on the availability
of large, natural or semi-natural landscapes and a sufficient degree of
connectivity among remaining subpopulations. Once common throughout sub-Saharan
Africa, this species has disappeared from much of its historic range, largely
due to human persecution, habitat fragmentation, and disease outbreaks
\citep{Woodroffe.2012}. Wild dogs typically disperse in single-sex coalitions
\citep{McNutt.1996, Behr.2020} and are capable of dispersing several hundred
kilometers \citep{DaviesMostert.2012, Masenga.2016, Cozzi.2020}. Although
previous studies have investigated connectivity for this species using LCPA
\citep{Hofmann.2021} or CT \citep{Brennan.2020}, a more comprehensive and
mechanistic understanding of dispersal and connectivity is missing (but see
\citealp{Creel.2020}). Nevertheless, with about 6,000 free-ranging wild dogs
remaining in fragmented subpopulations \citep{Woodroffe.2012}, reliable
information on dispersal behavior and landscape connectivity is essential for
the conservation of this endangered carnivore. We anticipated that a
connectivity assessment based on our three-step approach would overcome several
of the above highlighted conceptual shortcomings of traditional connectivity
models, while providing a more detailed view on movement behavior during
dispersal.

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width = \textwidth]{99_GraphicalAbstract2.pdf}
    \caption{Flowchart of the simulation-based connectivity analysis as proposed
    in this article. First, GPS data and habitat covariates must be
    collected. The combined data is then analyzed in an integrated step
    selection model, which enables the parametrization of the focal species'
    habitat and movement kernels and results in a mechanistic movement model.
    The parametrized model is then treated as an individual-based movement model
    and used to simulate dispersal trajectories. Ultimately, simulated
    trajectories serve to produce a set of maps that are pertinent to landscape
    connectivity. This includes a heatmap, indicating the traversal frequency
    across each spatial unit of the study area, a betweenness map, highlighting
    movement corridors and bottlenecks, and, finally, an inter-patch
    connectivity map, where the frequency of connections and their average
    duration can be depicted.}
    \label{GraphicalAbstract}
  \end{center}
\end{figure}

\section{Methods}
\subsection{Case Study}
\subsubsection{GPS Data}
We applied the three step approach mentioned above to GPS movement data from 16
dispersing African wild dog coalitions (7 female and 9 male coalitions). This
data has been collected between 2011 and 2019 from a free-ranging wild dog
population in northern Botswana (details on the data collection can be found in
\cite{Cozzi.2020} and \cite{Hofmann.2021}). During dispersal, GPS collars
recorded a fix every 4 hours and regularly transmitted data over the Iridium
satellite system. To ensure regular time intervals between GPS fixes, we removed
any fixes that were not successfully obtained at the desired 4-hour schedule
(allowing for a tolerance of \( \pm \) 15 minutes). To prepare the data for
step-selection analysis, we converted the fixes (n = 4'169) into steps, where
each step represented the straight-line movement between two consecutive GPS
fixes \citep{Turchin.1998}.

% To delineate periods of dispersal, we determined the exact timing of emigration
% from the natal pack and settlement in a new territory using direct field
% observations and through visual inspection of the net squared displacement (NSD)
% metric. The NSD metric measures the squared Euclidean distance of a GPS
% relocation to a reference point \citep{Borger.2012}, which we set to the center
% of the territory of each dispersers' natal pack. Thus, dispersal was deemed to
% have started once an individual left its natal territory and ended once the NSD
% metric remained constant, indicating settlement. Because behavior during
% dispersal is more pertinent to landscape connectivity than behavior during
% residence \citep{Elliot.2014, Abrahms.2017}, we only considered data collected
% during dispersal for our analysis.

\subsubsection{Study Area}
Our simulation of dispersal trajectories and assessment of connectivity spanned
across the entire Kavango-Zambezi Transfrontier Conservation Area (KAZA-TFCA,
\Cref{StudyArea}a and b). The KAZA-TFCA is the world's largest transboundary
conservation area and comprises parts of Angola, Botswana, Namibia, Zimbabwe,
and Zambia, thus hosting a rich diversity of landscapes, ranging from savannah
to grassland and from dry to moist woodland habitats. In its center lies the
Okavango Delta, a dominant hydro-geographical feature and the world's largest
flood-pulsing inland delta. Large portions of the KAZA-TFCA are formally
protected in the form of national parks (NPs) or other protected areas, yet a
considerable portion of the landscape is still human-dominated (e.g. roads,
agricultural sites, and settlements). We used NPs as patches to calculate
inter-patch connectivity (cf. Section 2.4.3). The decision to only consider NPs
as patches was purely out of simplicity and does not imply that dispersal
between other protected areas does not occur.

\begin{figure}[h]
  \begin{center}
    \begin{tikzpicture}
        \node[anchor=south west,inner sep=0] (image) at (0,0,0) {
        \includegraphics[width=\textwidth]{99_StudyArea.png}
        };
        \begin{scope}[x={(image.south east)},y={(image.north west)}]
            % % next four lines will help you to locate the point needed by forming a grid.
            % % comment these four lines in the final picture.
            % \draw[help lines,xstep=.1,ystep=.1] (0,0) grid (1,1);
            % \draw[help lines,xstep=.05,ystep=.05] (0,0) grid (1,1);
            % \foreach \x in {0,1,...,9} { \node [anchor=north] at (\x/10,0) {0.\x}; }
            % \foreach \y in {0,1,...,9} { \node [anchor=east] at (0,\y/10) {0.\y};}
            % % upto here
            \draw[densely dotted, blue] (0.169, 0.222) -- (0.364, 0.955);
            \draw[densely dotted, blue] (0.169, 0.157) -- (0.364, 0.074);
        \end{scope}
    \end{tikzpicture}
    \caption{Illustration of the study area located in southern Africa. (a) The
    study area was confined by a bounding box spanning the entire KAZA-TFCA
    which comprises parts of Angola, Namibia, Botswana, Zimbabwe, and Zambia.
    (b) The KAZA-TFCA currently represents the world's largest terrestrial
    transfrontier conservation area, covering a total area of 520'000
    km\textsuperscript{2}. Its main purpose is to re-establish connectivity
    between already-existing NPs (dark green) and other protected areas (light
    green).}
    \label{StudyArea}
  \end{center}
\end{figure}

\subsubsection{Covariates}
We represented the physical landscape in our study area by the habitat
covariates \textsf{water-cover, distance-to-water, woodland-cover,
shrub/grassland-cover, and human-influence}. To render the seasonal dynamics in
water-cover of major water bodies for the extent of the Okavango Delta, we
applied an algorithm that enabled us to obtain weekly updated raster-layers for
\textsf{water-cover} and \textsf{distance-to-water} from MODIS satellite imagery
\citep{Wolski.2017, Hofmann.2021}. This algorithm is now implemented in the {\tt
floodmapr} package (available on GitHub;
\url{https://github.com/DavidDHofmann/floodmapr}). To ensure a consistent
resolution across habitat covariates, we coarsened or interpolated all layers to
a resolution of 250 m x 250 m. A detailed description of how we prepared each
habitat covariate is provided in \cite{Hofmann.2021}.

Besides habitat covariates, we computed movement metrics that we used as
movement covariates in the ISSF models \citep{Avgar.2016, Fieberg.2021}.
Specifically, we computed for each step the step length (\textsf{sl}), its
natural logarithm (\textsf{log(sl)}), and the cosine of the relative turning
angle (\textsf{cos(ta)}), which is a measure of directionality
\citep{Turchin.1998}. Because wild dog activity is low during the hot midday
hours \citep{Cozzi.2012}, we additionally created the variable
\textsf{LowActivity}, indicating whether a step was realized during periods of
low wild dog activity (09:00 to 17:00 local time) or high wild dog activity
(17:00 to 09:00 local time). We performed all data preparations, spatial
computations, and statistical analysis in {\tt R}, version 3.6.6 \citep{R.2020}.
Some helper functions were written in {\tt C++} and imported into {\tt R} using
the {\tt Rcpp} package \citep{Eddelbuettel.2011, Eddelbuettel.2013}.

\subsection{Step 1 - Movement Model}
We used ISSFs \citep{Avgar.2016} to parametrize a mechanistic movement model
starting from GPS movement data collected on dispersing individuals. More
specifically, we paired each realized (i.e. observed) step with a set of 24
randomly generated alternative steps. A realized step plus its 24 random steps
formed a 25-step-stratum and received a unique identifier. As suggested by
\cite{Avgar.2016}, we generated random steps by sampling random turning angles
from a uniform distribution (\(-\pi, +\pi\)) (which is equivalent to a von Mises
distribution with location and concentration parameters \(\mu = \kappa = 0\))
and step lengths from a gamma distribution that was fitted to realized steps
(scale \(\theta\) = 6'308 and shape \(k\) = 0.37).

Along each realized and random step, we extracted values from underlying habitat
covariate layers using the {\tt velox} package \citep{Hunziker.2021} and we
computed averages of each covariate along the steps. For each step, we further
calculated the movement metrics \textsf{sl}, \textsf{log(sl)}, and
\textsf{cos(ta)}. To facilitate model convergence, we standardized all
continuous covariates to a mean of zero and a standard deviation of one.
Correlations among covariates were low (\(|r| < 0.6\); \citealp{Latham.2011}),
so we retained all of them for modeling.

To contrast realized steps (scored 1) and random steps (scored 0), we assumed
that animals assigned a selection score \(w(x)\) to each step (\ref{EQ1};
\citealp{Fortin.2005}), where \(w(x)\) depended on the step's associated
covariates (\(x_1, x_2, ..., x_n\)) and on the animal's preferences (i.e.
relative selection strengths; \citealp{Avgar.2017}) towards these covariates
(\(\beta_1, \beta_2, ..., \beta_n\)):

\begin{equation}
\label{EQ1}
  w(x) = exp(\beta_1 x_1 + \beta_2 x_2 + ... + \beta_n x_n)
\end{equation}

\noindent The probability of a step \(i\) being realized was then contingent on
the step's selection score, as well as on the selection scores of all other step
in the same stratum:

\begin{equation}
\label{EQ2}
  P(Y_{i} = 1 | Y_{1} + Y_{2} + ... + Y_{i} = 1) =
  \frac{w(x_{i})}{w(x_{1}) + w(x_{2}) + ... + w(x_{i})}
\end{equation}

\noindent To estimate preferences (i.e. the \(\beta\)-coefficients), we used
mixed effects conditional logistic regression analysis \citep{Muff.2020},
implemented through the r-package {\tt glmmTMB} \citep{Brooks.2017}. The method
introduced by \cite{Muff.2020} allows to model random slopes, yet requires to
fix the variance of the stratum specific intercept to a large value. We
therefore fixed the variance of the stratum specific intercept to an arbitrary
high value of \(10 ^ 6\) and used disperser identity to model random slopes for
all covariates.

The structure of our movement model was based on a habitat selection model that
was previously developed for dispersing wild dogs (hereafter referred to as
\textit{base model}, \citealp{Hofmann.2021}). In the base model, no interactions
among habitat covariates and movement covariates were considered, so we here
expanded the model and allowed for interactions between all movement covariates
and habitat covariates, thus reflecting that movement behavior could depend on
habitat conditions (details in Appendix A1). To determine the most parsimonious
movement model among model candidates, we ran stepwise forward model selection
based on Akaike's Information Criterion (AIC, \citealp{Burnham.2002}). We
validated the predictive power of the most parsimonious model using k-fold
cross-validation for case-control studies as described in \cite{Fortin.2009}.
This validation proves a significant prediction in case the Spearman rank
correlation coefficient of predicted step-ranks and associated frequencies under
the movement model is significantly greater than under the assumption of random
preferences (details in Appendix A2).

\subsection{Step 2 - Dispersal Simulation}
We used the most parsimonious movement model to simulate 80'000 individual
dispersal trajectories across the study area. The simulation of a dispersal
trajectory resembled an ``inverted'' ISSF and was set up as follows. (1) We
defined a source point and assumed a random initial orientation of the simulated
animal. (2) Starting from the source point, we generated 25 random steps by
sampling turning angles from a uniform distribution (\(-\pi, +\pi\)) and step
lengths from our fitted gamma distribution. (3) Along each random step, we
extracted and averaged values from the different habitat covariate layers and
computed the movement metrics \textsf{sl}, \textsf{log(sl)}, and
\textsf{cos(ta)}. To ensure compatible scales with the fitted movement model, we
standardized covariate values using means and standard deviations from the
empirical data. (4) We applied the parametrized movement model to predict the
selection score \(w(x)\) for each step using \ref{EQ1} and we converted
predicted scores into probabilities using \ref{EQ2}. (5) We randomly sampled one
of the generated random steps based on assigned probabilities and determined the
animal's new position. We repeated steps (2) to (5) until 2,000 steps were
realized.

As source points for the simulations, we distributed 50,000 points at random
locations inside protected areas that were larger than the average home range of
resident wild dog packs (i.e. \(>\) 700 km\textsuperscript{2};
\citealp{Pomilia.2015}). We placed another 30,000 points randomly inside the
buffer zone, mimicking potential immigration into the study area (Figure S1).

To mitigate edge effects and to deal with random steps leaving the study area,
we followed \cite{Koen.2010} and artificially expanded all covariate layers by a
100 km wide buffer zone. Within the buffer zone, we randomized covariate values
by resampling values from the original covariate layers. Through this buffer
zone, simulated dispersers were able to leave and re-enter the main study area.
In cases where random steps crossed the outer border of this buffer zone, we
resampled steps until they fully lied within the buffer zone, essentially
forcing individuals to remain within the expanded study area.

To ensure reliable connectivity estimates, we determined the number of simulated
dispersal trajectories required for connectivity to reach a ``steady state''.
For this purpose, we distributed 1,000 rectangular ``checkpoints'', each with an
arbitrary extent of 5 km x 5 km, at random coordinates within the study area
(excluding the buffer). We then determined the relative frequency at which each
checkpoint was traversed by simulated dispersal trajectories (hereafter referred
to as relative traversal frequency) as we gradually increased the number of
simulated trajectories from 1 to 50,000. To assess variability in the relative
traversal frequency, we repeatedly subsampled 100 times from all 50'000
trajectories and computed the mean traversal frequency across replicates, as
well as its 95\% prediction-interval for each checkpoint. We considered
connectivity to have reached a steady state once the width of the
prediction-interval dropped below a value of 0.01 for all checkpoints.

\subsection{Step 3 - Connectivity Maps}
\subsubsection{Heatmap}
To identify dispersal hotspots within the study area, we created a heatmap
indicating the absolute frequency at which different areas were traversed by
simulated dispersal trajectories (e.g. \citealp{Peer.2008, Hauenstein.2019,
Zeller.2020}). Specifically, we rasterized all simulated trajectories onto a
raster with 1 km x 1 km resolution and tallied resulting layers into a single
map. This procedure ensured that every trajectory was only counted once, even if
it traversed the same raster-cell multiple times, thus reducing potential biases
caused by individuals that were surrounded by unfavorable habitat and ``moved in
circles''. To achieve high performance rasterization, we used the R-package {\tt
terra} \citep{Hijmans.2021b}.

\subsubsection{Betweenness Map}
To pinpoint movement corridors and bottlenecks, we converted simulated
trajectories into a network and calculated betweenness scores for all
raster-cells in the study area \citep{BastilleRousseau.2018}. Betweenness is a
pertinent metric for connectivity as it measures how often a specific
network-node (in our case a raster-cell) lies on a shortest path between any
other pair of nodes \citep{BastilleRousseau.2018}. To convert simulated
trajectories into a network, we followed \cite{BastilleRousseau.2018} and
overlaid the study area (including the buffer) with a raster containing 5 km x 5
km raster-cells, where the center of each raster-cell served as node in the
final network. To identify edges (i.e. connections) between the nodes, we used
the simulated trajectories and determined all transitions occurring from one
cell to another, as well as the frequency at which those transitions occurred.
This resulted in an edge-list that we translated into a weighted network using
the r-package {\tt igraph} \citep{Gabor.2006}. The final weight of each edge was
determined by the frequency of transitions, yet because {\tt igraph} handles
edge weights (\(\omega\)) as costs, we inverted the traversal-frequency through
each raster-cell by applying \(\omega = \frac{mean(Traversal
Frequency)}{Traversal Frequency_i}\). Consequently, regularly used edges
received small weights (i.e. low costs) and vice versa. Finally, we used the
weighted network to calculate betweenness scores for all network nodes.

\subsubsection{Inter-Patch Connectivity Map}
To examine the presence and intensity of functional links (i.e. connections)
between patches within the study area, we calculated inter-patch connectivity
(e.g. \citealp{Gustafson.1996}, \citealp{Kanagaraj.2013}). For this, we computed
the relative frequency at which dispersers originating from one patch
successfully moved into another patch. We considered movements between patches
as successful if an individual's dispersal trajectory intersected with the
target patch at least once. For each trajectory we also recorded the number of
steps required to reach the first intersection with the respective patch,
allowing us to compute the average dispersal durations from one patch to
another. In summary, we determined \textit{if} and \textit{how often} dispersers
moved between certain patches, as well as \textit{how long} individuals had to
move to make these connections.

\section{Results}
\subsection{Movement Model}
The most parsimonious movement model consisted of movement covariates, habitat
covariates and several of their interactions, suggesting that movement behavior
during dispersal depended on habitat conditions (\Cref{MovementModel}a, Table S1
and Table S2). Although multiple models received an AIC weight \(>\) 0 (Table
S1), we only considered results from the most parsimonious model for simplicity.
This decision only marginally influenced subsequent steps as all models with
positive AIC weights retained similar covariates (Table S1). The k-fold
cross-validation showed that the final model substantially outperformed a random
guess and provided reliable predictions (i.e. confidence intervals of
\(\bar{r}_{s, realized}\) and \(\bar{r}_{s, random}\) did not overlap).
Moreover, the model correctly assigned high selection scores to realized steps
(\Cref{MovementModel}b), indicating a good fit between predictions and
observations. Compared to the base model (\(\bar{r}_{s, realized} = -0.55;
95\%-CI = [-0.57, -0.52]\); \citealp{Hofmann.2021}), the inclusion of several
interactions between movement and habitat covariates significantly improved
model performance (\(\bar{r}_{s, realized} = -0.65; 95\%-CI = [-0.67, -0.64]\)).

Plots that aid with the interpretation of the most parsimonious movement model
are provided in Figure S2 and suggest that, under average conditions, dispersing
wild dogs avoided moving through water, woodlands, and areas dominated by
humans, but preferred moving across shrublands or grasslands
(\Cref{MovementModel}a). Dispersers realized shorter steps (indicating slower
movements) in areas covered by water or woodland, while realizing larger steps
in areas dominated by shrubs or grass (\Cref{MovementModel}a). We found a
particularly strong effect for the variable \textsf{LowActivity}, suggesting
that dispersing wild dogs moved faster during twilight and at night (i.e.
between 17:00 and 09:00 o'clock) in comparison to the rest of the day
(\Cref{MovementModel}a). Although dispersers revealed a preference for
directional movements (i.e. low turning angles), especially when moving quickly,
they did less so in proximity to humans or water, resulting in more tortuous
movements in such areas (\Cref{MovementModel}a).

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{99_MovementModel}
    \caption{(a) Most parsimonious movement model for dispersing wild dogs. The
    model comprises a habitat kernel, a movement kernel, as well as their
    interactions. The horizontal line segments delineate the 90\%, 95\%, and
    99\% confidence-intervals for the respective \(\beta\)-coefficients.
    Significance codes: * \(p < 0.10\), ** \(p < 0.05\), *** \(p < 0.01\). (b)
    Results from the k-fold cross validation procedure. The upper plot shows
    rank frequencies of realized steps according to model predictions with known
    preferences, whereas the lower plot shows rank frequencies of realized steps
    when assuming random preferences. The blue ribbon shows the prediction
    interval around a loess smoothing regression that we fitted to ease the
    interpretation of the plots. The significant correlation between rank and
    associated frequency in (b1) highlights that the most parsimonious model
    successfully outperforms a random guess (b2) and assigns comparably high
    selection scores to realized steps.}
    \label{MovementModel}
  \end{center}
\end{figure}

\subsection{Dispersal Simulation}
Dispersal simulations based on the most parsimonious movement model proved
useful for assessing landscape connectivity. Of the 50,000 simulated dispersal
trajectories with starting point within the main study area, only 4.5\% reached
a map boundary, suggesting minimal biases due to boundary effects. Moreover, our
examination of the relative traversal frequency across all checkpoints showed
that connectivity reached a steady state after 10,500 simulated dispersal
trajectories (Figure S3). Although variability in relative traversal frequency
kept decreasing as we increased the number of simulated dispersers, the marginal
benefit of simulating additional trajectories diminished quickly (Figure S3).

\subsection{Heatmap}
The heatmap (\Cref{Heatmap}), which resulted from the sum of all simulated
dispersal trajectories, allowed us to pinpoint areas that were intensely used by
simulated dispersers. This enabled us to compare areas inside and outside the
KAZA-TFCA borders with respect to the frequency at which they were visited. For
instance, we could deduct that areas inside the KAZA-TFCA were frequently
traversed by dispersers (median traversal frequency inside KAZA-TFCA = 166, IQR
= 274, Figure S6a), whereas areas beyond the KAZA-TFCA boundary were
comparatively rarely visited (median traversal frequency outside KAZA-TFCA = 61,
IQR = 133, Figure S6a). Most notably, the region in northern Botswana south of
the Linyanti swamp appeared to serve as highly frequented dispersal hotspot
(median traversal frequency = 987, IQR = 558). Aside from revealing movement
hotspots, the heatmap also provided information on areas that act as movement
barriers. The presence of extensive water bodies, such as, for example, the
Okavango Delta, the Makgadikgadi Pan, and the Linyanti swamp, substantially
restricted dispersal movements and limited realized connectivity inside the
KAZA-TFCA. Although the heatmap improved our understanding of the frequency at
which areas were traversed by simulated dispersers, it seemed impractical to
pinpoint dispersal corridors.

\begin{figure}
  \includegraphics[width=\textwidth]{99_Heatmap.png}
  \caption{Heatmap showing traversal frequencies of 80'000 simulated dispersers
  moving 2'000 steps across the KAZA-TFCA. Simulations were based on an
  integrated step-selection model that we fitted to the movement data of
  dispersing African wild dogs. To generate the heatmap, we rasterized and
  tallied all simulated trajectories. Consequently, the map highlights areas
  that are frequently traversed by virtual dispersers. For spatial reference we
  plotted a few selected NPs (dark gray). Additional heatmaps showing the
  traversal frequency when individuals move fewer than 2'000 steps are provided
  in Figure S4.}
  \label{Heatmap}
\end{figure}

\subsection{Betweenness}
The betweenness map (\Cref{Betweenness}) revealed distinct dispersal corridors
that run across the study area. In comparison to the heatmap, the betweenness
map thus facilitated the identification of narrower and more linear routes that
are used to disperse between distinct regions. Again, northern Botswana emerged
as a wild dog dispersal corridor that connected more remote regions in the study
area. Towards east, the extension of this corridor ran through Chobe NP into
Hwange NP. From there, a further extension connected to Matusadona NP in
Zimbabwe. Northwest of the Linyanti ecosystem, a major corridor expanded into
Angola, where it split and finally traversed over a long stretch of unprotected
area into Zambia's Kafue NP. Several additional corridors with lower betweenness
scores emerged, yet most of them ran within the KAZA-TFCA boundaries (median
betweenness inside KAZA-TFCA = 6.947 \(\times\) 10\textsuperscript{6}, IQR =
54.311 \(\times\) 10\textsuperscript{6}, Figure S6b). In general, there were few
corridors that directly linked the peripheral regions of the KAZA-TFCA and
passed through unprotected areas outside the KAZA-TFCA (mean betweenness outside
KAZA-TFCA = 2.685 \(\times\) 10\textsuperscript{6}, IQR = 9.891 \(\times\)
10\textsuperscript{6}, Figure S6b).

%  9 2000  Betweenness Inside KAZA-TFCA  6946817.   54311358.
% 10 2000  Betweenness Outside KAZA-TFCA 2685174.    9890504.

\begin{figure}
  \includegraphics[width=\textwidth]{99_Betweenness.png}
  \caption{Map of betweenness scores, highlighting distinct dispersal corridors
  and potential bottlenecks across the extent of the KAZA-TFCA. Betweenness
  measures the number of shortest paths traversing through each node
  (raster-cell). Hence, a high betweenness score indicates that the respective
  area is exceptionally important for connecting different regions in the study
  area. The metric is therefore useful to pinpoint discrete movement corridors
  \citep{BastilleRousseau.2018}. Note that we square-rooted betweenness scores
  to improve visibility of corridors with comparably low scores. Additional
  betweenness maps showing betweenness scores when individuals move fewer than
  2'000 steps are provided in Figure S4.}
  \label{Betweenness}
\end{figure}

\subsection{Inter-Patch Connectivity}
The inter-patch connectivity map showed that the relative frequency at which
simulated dispersal trajectories moved from one NP to another varied, as did the
average dispersal duration required to make these connections
(\Cref{InterpatchConnectivity}). For some NPs, we also detected imbalances
between the number of ingoing and outgoing links, hinting at possible
source-sink dynamics. From Chobe NP, for instance, 510 individuals reached
Moremi NP, yet the opposite route was only realized by 340 individuals. However,
relative to the number of simulated individuals, these numbers imply fractions
of 50\% and 68\%, respectively. Overall, inter-patch connectivity between NPs in
Angola, Namibia, Botswana, and Zimbabwe appeared to be high; between 54\% and
87\% of individuals originating from a NP in these countries successfully moved
into some other NP (Figure S7a). Conversely, only 19\% of the dispersers leaving
from a NP in Zambia managed to find their way into some other NP (Figure S7b).
Prior to reaching another NP, individuals from Angola, Namibia, Botswana,
Zimbabwe, and Zambia had to move for an average of 630, 640, 940, 1045, and 890
steps, respectively. Furthermore, it appeared that the dispersal corridor
between Angola's NPs and the Kafue NP in Zambia identified in
\Cref{InterpatchConnectivity} is only rarely realized.

\begin{figure}
  \includegraphics[width=\textwidth]{99_InterpatchConnectivity.png}
  \caption{Map of inter-patch connectivity in relation to dispersal duration,
  highlighting connections between NPs (dark green). Yellow bubbles represent
  the center of the different NPs and are sized in relation to the number of
  simulated dispersers originating from each park. Black dots represent NPs that
  were smaller than 700 km\textsuperscript{2} and therefore did not serve as
  source areas. Arrows between NPs illustrate between which NPs the simulated
  dispersers successfully moved and the color of each arrow shows the average
  number of steps (4-hourly movements) that were necessary to realize those
  connections. Additionally, the line thickness indicates the relative number of
  dispersers originating from a NP that realized those connections. Note that a
  similar network view could be adopted to investigate connectivity between
  other protected areas need not to be restricted to NPs.}
  \label{InterpatchConnectivity}
\end{figure}

\section{Discussion}

% Short Summary
\subsection{Short Summary}
Here, we proposed a three-step approach to assess landscape connectivity via
simulated dispersal trajectories, and demonstrated its application using
empirical data from a free-ranging population of African wild dogs. In step one,
we used ISSFs to parametrize a fully mechanistic movement model describing how
individuals move through the landscape. In step two, we employed the movement
model to simulate dispersal trajectories across the landscape. In step three, we
translated the simulated trajectories into three complementary connectivity
maps, each emphasizing a different aspect of landscape connectivity (e.g.
frequently traversed areas, critical dispersal corridors and bottlenecks, and
the presence and intensity of functional links between suitable patches). Such
simulations from ISSFs overcome several conceptual shortcomings inherent to
traditional connectivity modeling techniques, such as LCPA and CT, and enable a
more detailed view on landscape connectivity.

% Movement Model
\subsection{Movement Model}
Our results on habitat preferences showed that dispersers avoid areas dominated
by humans and covered by water, but select for regions with open grassland in
the vicinity to water bodies. This largely complied with previous studies that
investigated habitat selection by dispersing wild dogs
\citep{DaviesMostert.2012, Masenga.2016, Woodroffe.2019, Oneill.2020,
Hofmann.2021}. However, by also accounting for movement preferences, we were
able to model several additional complexities common to dispersal. For instance,
our model attributed a strong effect to the binary variable
\textsf{LowActivity}, demonstrating the necessity to consider different activity
period of the focal species. By including an interaction between turning angle
and step length we could furthermore accommodate that dispersers exhibit step
lengths that are correlated with turning angles, meaning turning angles are
smaller when individuals move fast. Although similar autocorrelations could be
incorporated by sampling step lengths and turning angles from copula probability
distributions \citep{Hodel.2021a, Hodel.2021b}, the ISSF framework allowed us to
conveniently include correlations in the movement model. We only considered
first order autocorrelation, i.e. correlation between two consecutive steps,
although higher order autocorrelation is conceivable and may be desirable to
model \citep{Dray.2010, McClintock.2012}. This will, however, require vast
amounts of GPS data that are not interrupted by missing fixes; something that is
rarely achieved in reality \citep{Graves.2006}. The power and flexibility of
ISSFs to model additive effects between habitat and movement covariates
\citep{Avgar.2016, Signer.2017} furthermore allowed us to formally capture that
dispersing wild dogs move slower and more tortuous in areas covered by water,
something that surely was to be expected. Overall, the inclusion of interactions
between habitat and movement covariates in our movement model lead to a
significant improvement in predictive performance compared to an earlier model
that omitted such interactions \citep{Hofmann.2021}.

\subsection{Maps}
Each of the three connectivity maps derived from simulated dispersal
trajectories highlighted a different aspect of landscape connectivity. The
heatmap was most suitable for pinpointing frequently traversed areas and showed
that an exceptionally large number of dispersers moved through the regions of
the Moremi NP and the Chobe NP in northern Botswana. \cite{Hofmann.2021}
previously identified the same area as potential dispersal hotspot using LCPA,
however, following their analysis it was not clear whether this was the
consequence of the central location of the region and connections being enforced
between predefined start and endpoints. Contrary to LCPA, a simulation-based
approach as presented here does not require predefined endpoints, as endpoints
emerge naturally from the simulated dispersal trajectories. Not having to
predefine endpoints is especially useful for dispersal studies, as known
endpoints are usually an unrealistic assumption \citep{Elliot.2014,
Abrahms.2017, Cozzi.2020}. Simulations also enable the detection of potential
routes that do not lead into suitable habitats \citep{Dwernychuk.1972,
VanDerMeer.2014} but into areas with a high susceptibility for human wildlife
conflicts \citep{Cushman.2018}.

In contrast to the heatmap, the betweenness map emphasized relatively narrow and
linear movement routes and thus facilitated the identification of discrete
movement corridors. While in some cases both the heatmap and the betweenness map
attributed a high importance to the same areas (e.g. northern Botswana), little
consensus was found for other regions. For instance, while the stretch of
unprotected land between Luengue-Luiana NP in Angola and the Kafue NP in Zambia
was characterized by a high betweenness-score, it only received a low score on
the heatmap. The reason for this differences can be found in the differential
way in which the maps view connectivity. While the heatmap assigns high
connectivity to areas that are frequently traversed, it does not distinguish
between areas that truly bring individuals into different parts of the study
area and regions that lead into ecological traps. The converse is true on the
betweenness map, which strictly highlights regions that promote movement into
other areas of the landscape and thus might promote gene-flow. Because neither
of the two maps provides insights into functional links between distinct habitat
patches or how connections depend on the dispersal duration, we also produced a
map of inter-patch connectivity. The map depicted the frequency at which
simulated individuals moved between NPs as well as the average duration (in
steps) required to realize them. Calculating dispersal durations was possible
because dispersal trajectories were simulated spatially and temporally
explicitly, something that is currently unfeasible with LCPA or CT. An explicit
representation of time enables answerings questions such as: ``\textit{How long
will it take a disperser to move from A to B?}'' or ``\textit{Is it possible for
a disperser to move from A to B within X days?}''. Moreover, it yields
opportunities to incorporate seasonality and to investigate whether dispersal
corridors exist seasonally or all-year round (\textit{dynamic connectivity};
\citealp{Zeller.2020}). With LCPA or CT, seasonality can currently only be
incorporated by repeatedly running a connectivity analysis using an array of
seasonal permeability surfaces (e.g. \citealp{Benz.2016, Osipova.2019}). In
contrast, simulations from ISSFs allow the environment to change ``as the
dispersers move'', so that simulated trajectories can dynamically respond to
seasonal fluctuations in the environment.

Overall, the contrasts highlight the complementary nature of the presented
connectivity maps and emphasize the value of consulting multiple metrics when
assessing connectivity.

\subsection{Disadvantages of ISSF Simulations}
Despite the many benefits and great flexibility offered by simulations from
ISSFs, one must also be aware of the associated limitations. For example, while
our approach of simulating dispersal proved usedful to assess landscape
connectivity, it was computationally costly. Simulating 80,000 dispersal
trajectories for 2'000 steps across the KAZA-TFCA required five days of
computation on a regular desktop machine (AMD Ryzen 7 2700X processor with 8 x
3.6 GHz and 16 logical cores, 64 GB of RAM). The long simulation time was
primarily caused by the massive extent of the study area considered (ca. 1.8 Mio
km\textsuperscript{2}) and the large number of trajectories simulated. Most
connectivity studies focus on smaller study areas (e.g. \citealp{Kanagaraj.2013,
Clark.2015, McClure.2016, Abrahms.2017, Zeller.2020}) and will therefore require
fewer simulations and achieve faster simulation times (given the same spatial
resolution). We also believe that fewer simulated trajectories will often
suffice, as the relative traversal frequency by simulated trajectories through
randomly placed checkpoints across our study area converged already after 10,500
runs. The number of required simulations to achieve reliable estimates of
connectivity will, however, vary depending on the structure of the landscape and
the dispersal capabilities of the focal species \citep{Gustafson.1996}. For
species that disperse short distances through homogeneous environments, few
simulations may suffice to gauge connectivity, whereas for species that disperse
over long distances through heterogeneous habitats, a large number of
simulations will be necessary to sufficiently explore the spectrum of possible
routes.

Aside from the computational requirements, simulations further entail several
non-trivial but important modeling decisions, on four of which we would like to
further elaborate: (1) the number of simulated individuals, (2) the location of
source points, (3) the simulated dispersal duration, and (4) behavior at map
boundaries.

(1) When simulating dispersal trajectories, the modeler needs to decide on the
number of simulated individuals. A higher number is always desirable, as each
additional trajectory provides information about landscape connectivity.
However, each additional simulation entails computational costs, so a trade-off
needs to be managed. Here, we followed \cite{Signer.2017} who suggested to
simulate additional individuals only until the metrics of interest converge
towards a steady state. The exact number of required individuals might, however,
vary depending on the target metric and the anticipated connectivity map. More
sophisticated target metrics tailored to each of the presented connectivity
maps need to be developed in the future.

(2) To initiate dispersers, a modeler needs to provide a set of source points at
which the virtual disperser will be released. We placed source points within
protected areas large enough to sustain viable wild dog populations, implicitly
assuming wild dogs primarily survive in large, formally protected areas
\citep{Woodroffe.1999, DaviesMostert.2012, Woodroffe.2012, VanDerMeer.2014}. We
lacked precise knowledge about the presence and abundance of wild dogs in the
different protected areas, so we distributed source randomly points within them.
In cases where such data is available, source points could be distributed
accordingly, reflecting that the number of simulated trajectories does not
necessarily scales with the size of the source area. Alternatively, source
points could be distributed homogeneously and only later be weighted when
computing the heatmap, betweenness map, or inter-patch connectivity map. In any
case, the challenge of selecting meaningful source points is not unique to
individual-based simulations but also applies to LCPA or CT.

(3) The use of ISSFs to simulate dispersers requires deciding on the number of
simulated steps (i.e. dispersal durations). If sufficient dispersal data of the
focal species has been collected, dispersal durations can be sampled from
observed dispersal events. Due to the low number of observed dispersal events,
we opted against this solution and instead simulated all individuals for 2,000
steps, which is at the upper end of observed dispersal durations
\citep{DaviesMostert.2012, Masenga.2016, Cozzi.2020, Hofmann.2021}. This
approach also had the advantage that it allowed us to subsample simulated
trajectories to shorter durations after their simulation and thereby to
investigate the sensitivity of our results with respect to exact dispersal
durations (Figures S4 and S5).

(4) Unless simulated dispersal trajectories are strongly drawn towards a point
of attraction (e.g. \citealp{Signer.2017}), some trajectories will inevitably
approach a map boundary and one or more of the generated random steps will leave
the study area. One solution to deal with such cases would be to simply
terminate the simulation of the affected trajectory, implicitly assuming that
the respective animal left the study area. However, this approach might produce
ambiguous results in cases where individuals are released near map borders,
since already a single random step leaving the study area will break the
simulation, thus resulting in biased connectivity estimates along map borders.
Rather than halting the simulation, we created a buffer zone \citep{Koen.2010}
and resampled random steps until they fully lied within the study area. This
proved to be an effective solution to overcome problems with boundary effects.

\subsection{Conclusion}
In summary, we proposed and applied a simple three-step approach that relies on
ISSF-analysis and enables the simulation of dispersal trajectories and the
assessment of landscape connectivity. The proposed approach overcomes several of
the conceptual shortcomings inherent to LCPA and CT, such as the assumption of
known endpoints, and provides a highly flexible tool for investigating
connectivity. With this work, we hope to have sparked interest in the
application, optimization, or creation of methods to investigate dispersal
movements and connectivity via individual-based simulations, while at the same
time stressing some of the non-trivial modeling decisions involved. We also hope
to provide a useful framework that helps unifying and streamlining the
application of individual-based simulation for assessing landscape connectivity.

\section{Authors' Contributions}
D.D.H., D.M.B., A.O. and G.C. conceived the study and designed methodology;
D.M.B., G.C., and J.W.M. collected the data; D.D.H. and D.M.B. analysed the
data; G.C. and A.O. assisted with modeling; D.D.H., D.M.B., and G.C. wrote the
first draft of the manuscript and all authors contributed to the drafts at
several stages and gave final approval for publication.

\section{Data Availability}
GPS movement data of dispersing wild dogs is available on dryad
\citep{Hofmann.2021b}. Access to R-scripts that exemplify the application of the
proposed approach using simulated data are provided through Github
(\url{https://github.com/DavidDHofmann/DispersalSimulation}). In addition, all
codes required to reproduce the African wild dog case study will be made
available through an online repository at the time of publication.

\section{Acknowledgements}
We thank the Ministry of Environment and Tourism of Botswana for granting
permission to conduct this research. We thank C. Botes, I. Clavadetscher, and G.
Camenisch for assisting with wild dog immobilizations. We also thank B. Abrahms
for sharing her data of three dispersing wild dogs. Furthermore, we would like
to thank Johannes Signer for assisting with the simulation algorithm. This study
was funded by Albert-Heim Foundation, Basler Stiftung für Biologische Forschung,
Claraz Foundation, Idea Wild, Jacot Foundation, National Geographic Society,
Parrotia Stiftung, Stiftung Temperatio, Wilderness Wildlife Trust Foundation,
Forschungkredit der Universität Zürich, and a Swiss National Science Foundation
Grant (31003A\_182286) to A. Ozgul.

\newpage
\begingroup
\singlespacing
\bibliography{Literatur}
\endgroup

\end{document}