### 2022-09-13 Chapter_8/02_R-Scripts/02_Visualization/Interpatch.Rmd
Major update. Adjusted generation of source areas.

---
title: "Interpatch"
author: "David Hofmann"
date: "2022-09-06"
output: pdf_document
---
This is just a test document to prepare the table of inter-patch connectivity.

```{r, echo = F, warning = F, message = F}
# Clear R's brain
rm(list = ls())

# Load required packages
library(tidyverse)    # To wrangle data
library(kableExtra)   # To generate a nice table

# Load results on interpatch connectivity
dat <- read_rds("/home/david/ownCloud/University/15. PhD/Chapter_8/03_Data/03_Results/BootstrappedInterpatchConnectivity.rds")
dat <- subset(dat, FloodLevel != "Mean")

################################################################################
#### Frequency
################################################################################
# Reshape the data
dat_arr_freq <- dat %>%
  select(-c(StepNumber, StepNumberSE)) %>%
  mutate(Combined = paste(round(dat$Freq), "$\\pm$", sprintf("%.2f", round(dat$FreqSE, 2)))) %>%
  select(-c(Freq, FreqSE)) %>%
  pivot_wider(names_from = SourceArea, values_from = Combined) %>%
  arrange(CurrentArea, desc(FloodLevel)) %>%
  select(To = CurrentArea, everything()) %>%
  select(-FloodLevel)

# Make a nice table
table <- kbl(dat_arr_freq, booktabs = T) %>%
  # add_header_above(c("", "1", "2", "3", "4", "5", "6", "7", "8", "9")) %>%
  # add_header_above(c("", "Main" = 5, "Buffer" = 4)) %>%
  add_header_above(c("", "From" = 9)) %>%
  column_spec(1, border_right = T) %>%
  collapse_rows(1, latex_hline = c("custom"), custom_latex_hline = 1)

```