### 2022-12-22 Chapter_8/04_Manuscript/Manuscript.tex
upd

\documentclass[abstract=on,10pt,a4paper,bibliography=totocnumbered]{article}
\usepackage[paper=a4paper,left=35mm,right=35mm,top=25mm,bottom=30mm]{geometry}
\usepackage[doublespacing]{setspace}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[round]{natbib}
\usepackage{amsmath}
\usepackage{colortbl}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{gensymb}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage[hidelinks]{hyperref}
\usepackage[nameinlink]{cleveref}
% \usepackage{lineno}
\usepackage{multirow}
\usepackage{arydshln}
\usepackage[flushleft]{threeparttable}
\usepackage[nomarkers, nolists]{endfloat}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{scalerel}
\usepackage{tikz}
\usetikzlibrary{svg.path}

%------------------------------------------------------------------------------
%	Some Styling
%------------------------------------------------------------------------------
% Creating some TikZ styles
\tikzset{
  nonterminal/.style = {rectangle
    , minimum size = 6mm
    , very thick
    , draw = black!
  }
}

% Changing the style of captions in figures etc.
\captionsetup{labelfont=bf, format=plain, font=small}

% Change how equations are referenced
\renewcommand{\theequation}{Equation \arabic{equation}}%

% To be able to put an ORCID
\definecolor{orcidlogocol}{HTML}{A6CE39}
\tikzset{
  orcidlogo/.pic={
    \fill[orcidlogocol] svg{M256,128c0,70.7-57.3,128-128,128C57.3,256,0,198.7,0,128C0,57.3,57.3,0,128,0C198.7,0,256,57.3,256,128z};
    \fill[white] svg{M86.3,186.2H70.9V79.1h15.4v48.4V186.2z}
                 svg{M108.9,79.1h41.6c39.6,0,57,28.3,57,53.6c0,27.5-21.5,53.6-56.8,53.6h-41.8V79.1z M124.3,172.4h24.5c34.9,0,42.9-26.5,42.9-39.7c0-21.5-13.7-39.7-43.7-39.7h-23.7V172.4z}
                 svg{M88.7,56.8c0,5.5-4.5,10.1-10.1,10.1c-5.6,0-10.1-4.6-10.1-10.1c0-5.6,4.5-10.1,10.1-10.1C84.2,46.7,88.7,51.3,88.7,56.8z};
  }
}
\newcommand\orcid[1]{\href{https://orcid.org/#1}{\mbox{\scalerel*{
\begin{tikzpicture}[yscale=-1,transform shape]
  \pic{orcidlogo};
\end{tikzpicture}
}{|}}}}

%------------------------------------------------------------------------------
%	Titlepage: Header
%------------------------------------------------------------------------------
\title{Flooding of the Okavango Delta influences Connectivity for Dispersing
African Wild Dogs}

% List of Authors
\author{
  David D. Hofmann\textsuperscript{1,2,\S} \orcid{0000-0003-3477-4365} \and
  Dominik M. Behr\textsuperscript{1,2} \orcid{0000-0001-7378-8538} \and
  John W. McNutt\textsuperscript{2} \and
  Arpat Ozgul\textsuperscript{1} \orcid{0000-0001-7477-2642} \and
  Gabriele Cozzi\textsuperscript{1,2} \orcid{0000-0002-1744-1940}
}

% Reduce spacing between authors
\makeatletter
\def\and{%
  \end{tabular}%
  \hskip -0.5em \@plus.17fil\relax
  \begin{tabular}[t]{c}}
\makeatother

% Current Date
% \date{\today}

% And here the masterpiece begins
\begin{document}

% Change page numbering
\pagenumbering{gobble}

% Required to be able to cite
\bibliographystyle{apalike}

% Create Titlepage
\maketitle

%------------------------------------------------------------------------------
%	Titlepage: Additional Info
%------------------------------------------------------------------------------
\begin{flushleft}

\vspace{0.5cm}

\textsuperscript{1} Department of Evolutionary Biology and Environmental
Studies, University of Zurich, Winterthurerstarsse 190, 8057 Zurich,
Switzerland.

\textsuperscript{2} Botswana Predator Conservation Program, Private Bag 13,
Maun, Botswana.

\textsuperscript{\S} Corresponding author: david.hofmann2@uzh.ch

\vspace{4cm}

\textbf{Running Title:} Seasonal Flooding of the Okavango Delta and its
Consequences for African Wild Dog Dispersal and Connectivity

\vspace{0.5cm}

\textbf{Keywords:} movement, connectivity, seasonality, dispersal, conservation,
permeability

\end{flushleft}

%------------------------------------------------------------------------------
%	Abstract
%------------------------------------------------------------------------------
\newpage
\begin{abstract}
Climate change is expected to profoundly impact the life history of wild-living
animal populations. While the impact of climate change on the demographics of
local subpopulations has been studied repeatedly, little is known about the
consequences of environmental change on dispersal and connectivity.

We capitalize on a ``natural experimental setup'', the flood-pulse driven
environmental change across the Okavango Delta in northern Botswana, to
investigate the impact of changing environmental conditions on dispersal
patterns and connectivity of the endangered African wild dog (\textit{Lycaon
pictus}). For this, we simulate dispersal trajectories across the landscape of
the Okavango Delta under two extreme environmental scenarios; one assuming a
maximum flood extent, one assuming a minimum flood extent.

During maximum flood, we observe a reduction in connectivity and an increase in
dispersal durations between distinct habitat patches. Notably, dispersal into
the center of the Okavango delta is reduced by 80\% during maximum flooding. At
minimum flooding, conversely, the delta reveals vital dispersal corridors and
increases chances of successful dispersal into neighboring habitat patches by xx
percent. Concomitantly, dispersal durations to move between patches is reduced,
on average by xx days,at low flood levels.

Climate change is expected to critically impact the flood-pulsing cycles of the
Okavango delta and may therefore substantially alter landscape connectivity in
this area. Our analysis suggests that anticipating such alterations in
environmental conditions is key when implementing conservation strategies.
Besides a better understanding of the conservation needs for the African wild
dog, our study also presents a first step towards incorporating environmental
changes due to seasonality or climate change in dispersal and connectivity
analyses and demonstrates the usefulness of individual-based dispersal
simulations as a pertinent conservation tool to study impacts of environmental
change on dispersal success.

\end{abstract}

%------------------------------------------------------------------------------
%	Main Text
%------------------------------------------------------------------------------
\newpage

\onehalfspacing
\tableofcontents
\doublespacing

% Change page numbering
\newpage
\pagenumbering{arabic}

% Create linenumbers
% \linenumbers

\section{Introduction}
Climate change is expected to profoundly impact ecosystems worldwide with
far-reaching consequences for species living therein \citep{Ozgul.2010,
Radchuk.2019, IPCC.2022}. By altering environmental conditions, climate change
affects animal behavior \citep{Fuller.2016}, resource availability
\citep{Durant.2007}, population dynamics \citep{Paniw.2021}, and the
distribution of wild living animal populations \citep{Thomas.2004,
Thuiller.2006}. One potential life-history pathway through which species may
mediate the effects of environmental change is dispersal \citep{Anderson.2012}.
By bringing individuals away from their natal location \citep{Clobert.2012},
dispersal equips species with a means to adapt to climate change by tracking
favorable habitat conditions \citep{Raia.2012} and to shift into a different
part of their fundamental niches \citep{Kokko.2022}. Dispersal furthermore
provides resilience against changing conditions by ensuring genetic diversity
\citep{Frankham.2002, Leigh.2012, Baguette.2013}, facilitating the colonization
empty habitats \citep{Gustafson.1996, Hanski.1999b, MacArthur.2001}, and
promoting the reinforcement of weakened and small subpopulations
\citep{Brown.1977}. As such, it is little surpsing that improving connectivity
is viewed as the number one strategy to combat the adverse effects of climate
change \citep{Heller.2009}.

While dispersal offers a means to offset the demographic consequences of climate
change \citep{Kokko.2006, Hodgson.2009, Travis.2013}, it itself is a function of
climatic and environmental conditions (e.g. \citep{Behr.2020}). The link between
dispersal and climaticor environmental can either be indirect, for example if
the propensity of individuals to disperse depends on environmental conditions,
or direct, when the biophysical environment through which dispersers move
affects dispersal prospects \citep{Travis.2013}. As the latter case indicates,
dispersal is also inextricably linked to the concept landscape connectivity
\citep{Baguette.2013}, which is understood as the degree by which the landscape
facilitates or impedes movements \citep{Taylor.1993}. Only if a sufficient
degree of landscape connectivity is maintained, successful dispersal of
individuals can be guaranteed \citep{Fahrig.2003}. Unfortunately, our
understanding how changing environmental conditions influence connectivity and
dispersal is limited \citep{Kokko.2006, Travis.2013}. Because collecting
information on dispersing animals is difficult and because it is difficult to
assess how changing environmental conditions affect connectivity...

The continued degradation and destruction habitats across the globe impedes
dispersal, promotes inbreeding and genetic drift and increases the
likelihood of local extirpation due to stochastic events \citep{Sawyer.2011}.
Combined, these factors limit species ability to adapt to changing environmental
conditions \citep{Fahrig.2003}.

\cite{Ashrafzadeh.2019} found that cimate change is likely to substantially
reduce connectivity for mountain newts in Iran. Similarly, \cite{Luo.2021}
predicted significant reductions in connectivity for the giant spiny frog in
South-East Asia.

However, failing to take into account potential changes in climatic and
environmental conditions imperils our ability to design cost-effective
conservation strategies.

Here, we used naturally occuring seasonal flood-cycles to approximate the likely
changes in connectivity that will arise under amplified climatic conditions due
to climate change. While a better understanding of the ODs flooding regime under global change is of major interest, reliable predictions remain challenging due to the complex interplay between climate,
anthropogenic water use, animal behavior, and topography of the region.

Climate change
could lead to the disruption of existing dispersal corridors and the emergence
of new dispersal barriers but may also provide possibilities to access
previously unsuitable dispersal routes. Designing and implementing effective
conservation strategies that aim to improve landscape connectivity, such as
habitat restoration or the creation of wildlife corridors, requires better
understanding of such dynamics.

Albeit the impacts of climate change on connectivity cannot be directly
estimated, one can study how changing environmental conditions due to
seasonality influence patterns of dispersal and connectivity.

Habitat degradation and fragmentation lead to isolated subpopulations that have
a high associated risk of extinction due to stochastic events and generally
decreased genetic variability \citep{Melbourne.2008}.

More recently, individual-based movement models (IBMMs) have been uprising and
are recommended for their ability to incorporate several crucial realities of
animal behavior, such as, for example, autocorrelation, that are by omitted by
necessity using resistance-based approaches \citep{Hofmann.2023,
Unnithan.2022a}. IBMMs allow much greater flexibility and are thus considered
powerful tools for examining connectivity under different landscape
configurations \citep{Littlefield.2019}.

Botswana is among the countries most heavily influenced by climate change .

For multispecies approaches to connectivity within the KAZA-TFCA, see
\cite{Brennan.2020} and \cite{Lines.2021}.

Talk about connectivity models
Talk about functional connectivity
Talk about habitat fragmentation and destruction

Osipova

Unfortunately, little is known about the specific responses of different species
to climate change, especially for regions most vulnerable to altered conditions
\citep{Paniw.2021}.

Only if there

The challenges associated with studying dispersing animals have thus far
impaired the collection of data of dispersers at the appropriate temporal and
spatial scale \citep{Graves.2014, Vasudev.2015} and weakened our ability to
project dispersal prospects under changing environmental conditions into the
future. Nevertheless, a more comprehensive understanding of dispersal under
changing environmental conditions is critical for designing effective
conservation strategies in the light of climate change. By predicting and
modeling dispersal patterns under different scenarios, we can better anticipate
the impacts of climate change on species and design effective conservation
strategies to protect them. This is especially important for endangered species
that are already at the verge of extinction.

The Okavango delta in Southern Africa poses a unique opportunity to study the
impacts of environmental change on species dispersal ability and connectivity in a
large scale natural experiment setup. According to projections by the IPCC,
Botswana and its surroundings are among the most vulnerable to climate change.
While global temperatures are expected to increase by xx degrees until the end
of the 21st century, For example, \cite{Engelbrecht.2015} predicts an increase
between 4 and 6\degree C for regions in southern Africa. One species that
heavily relies on dispersal for long-term persistence and that inhabits such an
arid ecosystem in Southern Africa is the African wild dog (\textit{Lycaon
pictus}). The African wild dog is a lightly built canid lives in cohesive packs
comprising up to 20 individuals. Wild dogs are social breeders and the majority
of reproduction is monopolized by the dominant alpha couple who are supported by
subordinates in taking care of their pups. The species exhibits strong
inbreeding avoidance, hence juvenile individuals born into the pack disperse in
same-sex sibling coalitions after reaching sexual maturity \citep{McNutt.1996,
Behr.2020}. Notably, the African wild dog is the sole extant representative of
its genus and thus considered as high-priority species for conservation efforts
\citep{Leigh.2012}. However, fragmentation and destruction of their historic
habitat may have decreased dispersal success for this species and thus
jeopardizes connectivity \citep{Leigh.2012}. Since the impacts of climate change
are not uniformly distributed across the planet, but particularly pronounced at
high latitudes and arid ecosystems, the species may be seen as most vulnerable
to such ecosystem shifts.

Talk about Zeller und Osipova

In a previous study, \citep{Hofmann.2022} have investigated landscape
connectivity in the same ecosystem, yet assuming a static landscape.

Here, we capitalize on a previously parameterized dispersal model and use
dispersal simulations to investigate dispersal success and connectivity patterns
for African wild dogs under two extreme scenarios; one assuming maximum flooding
of the Okavango delta and one assuming minimum flooding of the Okavango delta.
Given that water poses a barrier to dispersing individuals, we anticipated that
dispersal prospects and connectivity during maximum flood are low. During
minimum flood, in contrast, we expect to reveal the presence of several
dispersal corridors that may be used to freely and quickly move between
remaining habitat patches.

\section{Materials and Methods}
We conducted all analyses using the programming language \textsf{R}. Any spatial
data manipulation was completed using the \textsf{terra} and \textsf{spatstat}
packages. Several helper functions for the simulation algorithm were written in
\textsf{C++} and imported to R using the \textsf{Rcpp} package. Figures were
generated using \textsf{ggplot2}, \textsf{igraph}, and \textsf{ggnetwork}. All
R-scripts used to conduct our analyses are provided through an online
repository.

\subsection{Study Area}
The study area for this analysis was focused on the Okavango delta (OD) and its
surroundings in Southern Africa, comprising parts of Angola, Namibia, Botswana,
Zimbabwe, and Zambia (\Cref{StudyArea}). The OD is the world's largest inland
delta and the main driver of seasonal environmental change in the region. While
our primary focus lied on the immediate surroundings of the Okavango Delta, we
considered a large rectangular extent stretching from 20\degree 30' E to
26\degree E. (totaling to an area of 300'000 km\textsuperscript{2}) to
accommodate for the long distance dispersal events commonly observed in African
wild dogs (e.g. \citealp{Davies-Mostert.2012, Masenga.2016, Cozzi.2020}). The
flood-pulsing rhythm of the OD is mainly dictated by precipitation in the
catchment areas of the Angolian highlands, where rainwater is collected and
channeled into the OD through the Okavango River. Although precipitation in
Angola peaks between December and March, water only slowly descends through the
Okavango river and its distributaries, reaching the distal ends of the delta in
July or August, where the water percolates at the Thamalakane and Kunyere
Faults. At minimum extent, the flood covers an area of 3'600
km\textsuperscript{2}, during maximum flood more than 9'000
km\textsuperscript{2}. Vegetation in the study area is dominated by mopane
forest, mixed acacia woodland, and grassland. Human influence is low and mainly
concentrated around small villages at the western periphery of the delta as well
as the city of Maun at the south-eastern tip of the OD. Large portions of land
are dedicated national parks, game reserves or forest reserves. The study area
is also part of the world's largest transboundary conservation initiative, the
Kavango-Zambezi Transfrontier Conservation Area, which aims to restore
connectivity between protected areas in Southern Africa.

\begin{figure}
  \begin{center}
  \includegraphics[width = \textwidth]{99_StudyArea.png}
  \caption{Study area across which we simulated dispersal. Simulated dispersers
  were released at random locations within the orange source areas distributed
  across the delta. Emigration zones (purple) served as checkpoints and enabled
  us to identify if and where simulated dispersers left the close surroundings
  of the Okavango delta. These zones were generated using a set of cutlines
  originating from the center of the delta and roughly cutting an elliptical
  buffer zone into sections of equal size.}
  \label{StudyArea}
  \end{center}
\end{figure}

\subsection{Spatial Habitat Layers}
We represented the physical landscape through which dispersers could move by a
set of spatially referenced habitat layers, each resolved at 250m x 250m. The
set of layers included water-cover, distance-to-water, tree-cover,
shrub/grassland-cover, and a human influence layer depicting anthropogenic
influences through villages, roads, and agriculture. A detailed description of
the different habitat layers is provided in \Cref{Hofmann.2021}. Importantly,
the water-cover and derived distance-to water layers were generated using MODIS
Terra MCD43A4 satellite imagery that was classified using a ``floodmapping''
algorithm developed by \citep{Wolski.2017} and available through the R-package
\textsf{floodmapr}. The algorithm allowed us to generate almost weekly updated
``floodmaps'', thus providing detailed information about the flood-extent at any
given point in time. In total, we generated 700 floodmaps between the years 2000
and 2019. Based on these maps, we generated a minimal and maximum flood
scenario. To create the minimum flood scenario, we averaged the 50 floodmaps
with smallest flood extent and generated a binary image using areas that were
inundated in at least 50\% of the maps. Similarly, we created an average image
for high flood using the 50 most flooded maps. The final maps are depicted in
\Cref{FloodExtent}.

\begin{figure}
  \begin{center}
  \includegraphics[width = \textwidth]{99_FloodExtent.png}
  \caption{Flood extent in the two scenarios considered. In the left panel, the
  flood is at an extremely low level, stretching across 3'558
  km\textsuperscript{2}, whereas in the right panel the flood is at an extremely
  high level and covers 9'397 km\textsuperscript{2}. The two maps were generated
  using 700 remote sensed MODIS MCD43A4 satellite images spanning the years 2000
  to 2019.}
  \label{FloodExtent}
  \end{center}
\end{figure}

\subsection{Dispersal Model}
Our dispersal model was based on a previously parametrized and validated
integrated step-selection function (iSSF, \citealp{Avgar.2016}) fitted to GPS
data of dispersing AWDs \citep{Hofmann.2021}. In step selection functions (SSFs,
\citealp{Fortin.2005}), observed GPS locations are converted into steps (the
straight-line traveled between two GPS recordings \citep{Turchin.1998a}) and
compared to a set of \textit{random} steps in a (mixed effects) conditional
logistic regression framework \citep{Fortin.2005, Thurfjell.2014, Muff.2020,
Fieberg.2021}. The model presented in \citep{Hofmann.2021} used dispersal data
of 16 dispersing African wild dogs from a free-ranging wild dog population in
northern Botswana. GPS data during dispersal was collected at 4-hourly intervals
and translated into steps of similar duration. Observed steps were then paired
step with 24 random steps that were generated using a uniform distribution for
turning angles (\(-\pi, +\pi\)) and step lengths from a gamma distribution
fitted to observed steps (scale \(\theta\) = 6'308 and shape \(k\) = 0.37). It
was then assumed that animals assigned to each observed and random step a
selection score of the form \citep{Fortin.2005}:

\begin{equation}
\label{EQ1}
  w(x) = exp(\beta_1 x_1 + \beta_2 x_2 + ... + \beta_n x_n)
\end{equation}

Where (\(x_1, x_2, ..., x_n\)) represent the covariate values along each of the
steps and the (\(\beta_1, \beta_2, ..., \beta_n\)) are the animal's relative
selection strengths \citealp{Avgar.2017} towards these covariates. The benefit
of \textit{integrated} SSFs over regular SSFs is that they provide a means to
render two complementary ``kernels''. A movement kernel that describes general
movement behavior of dispersing AWDs and a habitat kernel that describes
preferences of AWDs with regards to environmental conditions
\citep{Fieberg.2021}. iSSFs also allow interactions among the two kernels and
are thus suitable to render that movement behavior may change depending on
habitat conditions. A brief summary of the fitted dispersal model is provided in
Appendix xx.

\subsection{Source Areas and Emigration Zones}
We simulated dispersing AWDs originating from six distinct source areas located
in the vicinity of the Okavango delta (\Cref{StudyArea}). For areas one to six
we selected locations at the delta's periphery that remained dry in both
scenarios. At those locations, we generated circular buffers with a radius of 20
km. For source area six, we isolated a polygon covering Chief's Island, a
peninsula located at the OD's center. Besides source areas, we also generated
``zones of emigration'' that we used as checkpoints to determine if and where
simulated individuals left the delta's vicinity (\Cref{StudyArea}). We generated
emigration zones by first overlaying the OD with an elliptic buffer zone that we
dissected using a set of cutlines that originated from the ODs center and spread
according to cardinal points (\Cref{StudyArea}).

\subsection{Dispersal Simulation}
For each source area we simulated 1'000 individuals, once assuming a minimum
flood, once assuming a maximum flood. This resulted in the simulation of 6'000
individuals for each environmental scenario, hence 12'000 individuals in total.
The simulation algorithm was based on the algorithm described in
\Cref{Hofmann.2022} and works as follows. A random location within the source
area is chosen as a starting point. Originating from the starting point, a set
of 25 random steps is generated by sampling step lengths from a gamma
distribution fitted to observed steps (what is a step?) (shape = , scale = ) and
turning angles from a uniform distibution ($-\pi, +\pi$). Along each random step
the underlying spatial covariates are extracted and relevant movement metrics
are computed (i.e. log(sl), cos(ta), ta). The parametrized dispersal model is
used to predict the probability of each step for being chosen given the steps
covariate values. One of the steps is sampled based on assigned probabilities
and the location of the animal is updated. The procedure is then repated until
the desired number of steps is realized. Here, we simulated each individual for
2'000 steps, which correspons to the longest dispersal duration recorded in our
data. The original model was trained using 4-hourly steps, thus a simulated step
also resembled the movement conducted within four hours. Trajectories resulting
from such a simulation can be understood as correlated random walks that take
into account both habitat and movement preferences of dispersing individuals.

% We used a previously parametrized dispersal model to simulate dispersal
% trajectories of African wild dogs across the study area. To simulate dispersing
% wild dogs, we employed the simulation framework presented in
% \cite{Hofmann.2022}. In this framework, a movement model that renders habitat
% and movement preferences of dispersing individuals is parametrized using
% step-selection functions. Once parametrized, the model can be employed to
% simulate virtual dispersers moving across the landscape.
%
% We released virtual dispersers at random locations within the four distinct
% source areas depicted in \Cref{StudyArea}. From each source area we simulated xx
% individuals moving across the landscape at minimum flood and another xx
% individuals at maximum flood. For comparison, we also ran the simulations for a
% medium flood extent, yet the results for this will be presented in the appendix.

\subsection{Dispersal Prospects and Connectivity}
Based on simulated dispersal trajectories in the two scenarios we quantified
dispersal sucess and connectivity using three complementary connectivity metrics
as outlined in \cite{Hofmann.2022}. First, we generated heatmaps depicting the
frequency at which different areas in the landscape were visited by simulated
dispersers. Such heatmaps serve to detect dispersal hotspots and areas of
intense use. However, they are less suitable for detecting pinchpoints and
bottlenecks that are critical in linking distinct patches. Hence, we also
computed spatially explicit betweenness scores which are useful to highlight
exactly such pinchpoints \citep{Bastille-Rousseau.2018, Bastille-Rousseau.2021}.
To compute betweenness, we overlayed the study area with a regular grid with 2.5
km x 2.5 km grid cells and determined how often simulated individuals
transitioned from one grid-cell to another. Note that in case the same
individual repeatedly realized the same cell-transition (e.g. repeatedly moved
between A-B-A-B...), we only counted a single transition to avoid emphasis on
regions where individuals moved in circles. With on the so generated
transitions, we generated a network using the centers of all grid-cells as nodes
and cell-transitions as weighted connections between the nodes. Based on this
network we computed weighted betweenness scores using the R-package
\textsf{igraph}. As a final connectivity metric and metric of dispersal success,
we calculated the number of successful dispersal events between the different
source areas as well as towards the emigration zones. We coin this type of
connectivity ``inter-patch connectivity'' as it relates to the movement between
distinct patches. Dispersal between two areas was said to be successful whenever
a trajectory leaving one area intersected with the target area. To gauge the
dispersal duration needed to move between patches (be consistent with
``patches'', ``source-areas'' etc.), we also recorded the minimum number of
steps that individuals moved before arriving at the respective patch.

\section{Results}
\subsection{General}
See \Cref{Emigration}
\begin{figure}
  \begin{center}
  \includegraphics[width = \textwidth]{99_Emigration.png}
  \caption{Caption}
  \label{Emigration}
  \end{center}
\end{figure}

\subsection{Heatmaps}
Heatmaps produced from simulated dispersal trajectories reveal that the OD acts
as a major dispersal barrier during periods of high flood, but reveals viable
dispersal corridors during periods of low flood (\Cref{Heatmaps}). During
minimum flood, the area north-west of Maun appears to serve as vital dispersal
habitat. The same area is entirely avoided during maximum flood. Besides
striking differences in connectivity for the close vicinity of the delta, the
remainder of the study area shows only marginal differences in connectivity
between the two scenarios. For instance, in both scenarios the area south of the
Linyanti swamp appears as frequently visited dispersal habitat. Additional
heatmaps highlighting differences in connectivity for each source area
separately are provided in Appendix SX.

\begin{figure}
  \begin{center}
  \includegraphics[width = \textwidth]{99_Heatmaps.png}
  \caption{Heatmaps depicting the number of simulated dispersal trajectories
  traversing each grid-cell in the study area. The left panel shows results for
  the minimum flood scenario, whereas the right panel shows results for the
  maximum flood scenario. Source areas (numbered 1-6) from which dispresers were
  released, and emigration zones (numbered 7-14) are shaded in dark gray.}
  \label{Heatmaps}
  \end{center}
\end{figure}

\subsection{Betweenness}
The betweenness maps reveal a similar pattern in that connectivity through the
OD is only pronounced during periods of low floods and vanishes entirely during
maximum flood (\Cref{Betweenness}). A set of four dispersal corridors meets on
the central peninsula (source area 5, \Cref{Betweenness}) at minimum flood but
the same corridors are absent when the flood reaches a maximum extent. Instead,
a narrow corridor runs north west of Maun, connecting source areas one and two.
Again, the remainder of the study area is only marginally affected by flooding
patterns. Additional betweenness maps highlighting differences in connectivity
for each source area separately are provided in Appendix SX.

\begin{figure}
  \begin{center}
  \includegraphics[width = \textwidth]{99_Betweenness.png}
  \caption{Betweenness scores highlighting potential dispersal corridors for
  both the minimum and maximum flood scenario. Areas with high betweenness
  scores (bright yellow) are used by many simulated individuals to move into
  adjacent regions and can thus be understood as critical pinch-points. Source
  areas (numbered 1-6) from which dispresers were released, and emigration zones
  (numbered 7-14) are shaded in light gray.}
  \label{Betweenness}
  \end{center}
\end{figure}

\subsection{Inter-Patch Connectivity}
Our analysis of interpatch connectivity demonstrates notable differences in
dispersal prospects and duration depending on the extent of the flood
(\Cref{IPCTable}). While $\input{99_NumberReachingOthersMinFlood.tex}$ simulated
dispersers reach another source area during minimum extent, only
$\input{99_NumberReachingOthersMaxFlood.tex}$ do so during maximum extent. The
differences are particularly pronounced for individuals dispersing from or into
the source area located at the OD's center (\Cref{Interpatch}). While the area
is reached by $\input{99_NumberReachingArea6MinFlood.tex}$ simulated individuals
during minimum flood, only $\input{99_NumberReachingArea6MaxFlood.tex}$
dispersers arrive there during maximum flood. Furthermore, the dispersal
duration into source area six from any other source area increases from
$\input{99_StepsToReachingArea6MinFlood.tex}$ to
$\input{99_StepsToReachingArea6MaxFlood.tex}$. Across all simulations, the
average dispersal duration before reaching another source area increases from xx
to xx from the minimum to the maximum flood scenario. Nevertheless, connectivity
into some areas increases during maximum flooding. Emigration increased slightly
from xx to xx

For instance, while dispersers .. Additional maps highlighting differences in
inter-patch connectivty for each source area separately are provided in Appendix
SX.

\begin{figure}
  \begin{center}
  \includegraphics[width = \textwidth]{99_IPCTable.png}
  \caption{Dispersal frequency (a) and duration (b) (in steps) between source
  areas and emigration zones during minimum and maximum flood.}
  \label{IPCTable}
  \end{center}
\end{figure}

\begin{figure}
  \begin{center}
  \includegraphics[width = \textwidth]{99_Interpatch.png}
  \caption{Inter-patch connectivity into and from source area number six. This
  patch experienced the most drastic reduction in connectivity in result to a
  high flood level. During minimum flood there is ample connectivity from and
  into source areas 1-3. During maximum flood, connectivity is lower and mainly
  limited to source area 2.}
  \label{Interpatch}
  \end{center}
\end{figure}

% \input{IPC_Frequency.tex}
% \input{IPC_Duration.tex}
% \input{IPC.tex}

\section{Discussion}
According to our simulations, the propensity to move between the eastern and
western part of the delta is much lower during maximum extent. This is mainly
due to the flood-waters and the city of Maun acting as dispersal barriers.
During maximum extent, the floodwaters of the delta close a gap between the
delta and Maun that otherwise would serve as dispersal corridor. Anecdotal
evidence supports this hypothesis, for the only dispersing individuals recorded
to move from the eastern to the western part of the delta moved at times of low
flood. In line with this, it appears that a large flood extent pushes dispersing
individuals to move closer to human inhabited areas such as the village of Maun.

Predicting how climate change will impact the dispersal ability of AWDs is
challenging for multiple reasons. First of all, predicting the flooding patterns
of the OD under climate change is merely impossible due to the complex feedbacks
between surface-temperature, soil conditions, precipitation patterns and the
associated changes in vegetation. Second, the delta is not only prone to changes
in environmental conditions, but also to changes in anthropogenic use of the
inflowing water. Finally, it is unclear how AWDs, in fact, how any
species, will cope with environmental change due to global warming. Even though
some studies predict that AWD populations are likely to decline under increasing
temperatures, these studies fail to account for the behavioral plasticity of
their focal species. AWDs respiratory system, for instance, has evolved as a
perfect adaptation to high temperatures and AWDs may, in fact, profit from a
comparative advantage (cite an economist) over their competitors and prey under
rising temperatures. Although the theory of comparative advantages is a
fundamental concept in economics, it has yet to find its way into ecological
studies.

\cite{Murray-Hudson.2006} predicted that increased temperatures, additional human
abstractions, and reduced river flows might lead to a ``Delta dying'' and that
the impact of climate will be much more pronounced than the impact of
anthropogenic water use.

Although local rainfalls in Botswana are expected to increase in terms of
intensity, simulations show that the length of the rainy season will decline,
more than offsetting the incline in precipitation \cite{Akinyemi.2019}.

Although dispersal is generally believed to be beneficial for the long term viability of species, it can also lead to the net-loss of individuals in areas where edge effects and anthropogenic mortality are high \citep{Leigh.2012}.

We studied a population of African wild dogs that resides in a natural
environment with little human influence. This is only representative for a small
share of the extant wild dog populations, as most individuals reside in areas
that are prone to substantial edge effects. For these populations, the benefit
of dispersal is disputed, as dispersal under high anthropogenic mortality may
lead to a net-loss of individuals \citep{Leigh.2012}.

Dispersal is an important, if not the most important, driver of landscape
connectivity and therefore of major interest to conservation authorities. It has
also been demonstrated that dispersal may lead to population declines in areas
where anthropogenic mortality is high and dispersal prospects low
\citep{Leigh.2012}.

Efforts aiming at restoring or improving landscape connectivity should consider
and take into account changes in landscape characteristics due to climate
change.

We assessed the implications of environmental change on the dispersal prospects,
yet we did not consider how changing conditions alter dispersal propensity.

According to our simulations, dispersers are able to cover larger distances
during periods of low flood. This finding is little surprising, considering that
inundated areas act as dispersal barriers and force dispersers to detour and
circumvent water-covered areas. However, it still leads to an interesting
hypothesis. Previous studies have shown that the euclidean dispersal distance of
female coalitions is larger than that of male coalitions. This has led to ...
However, demographic analyses have also revealed that female offspring tend to
emigrate from their pack at younger ages and earlier in the year, when
floodwaters are still at a relatively low level (Behr). It is thus conceivable
that the sex-differences in reported dispersal distances is mainly a consequence
of environmental conditions during dispersal, rather than owed to physiological
differences between sexes.

While our analysis marks an important step into incorporating environmental
change into studies of connectivity, there are several critical additions that
should be considered in the future. We studied dispersal and connectivity under
two different environmental scenarios, yet our movement model assumed that
dispersers had identical habitat and movement preferences in both scenarios. In
reality, however, it can be expected that movement and habitat kernels of
dispersers differ depending on the season considered (examples).

An additional complication arises when species movement is not solely driven by
environmental conditions, but also affected by itra- and inter-specific factors.
For instance, ... has shown that dispersers...
Rendering such conditions alone is challenging, yet rendering the conditions
under changing environmental conditions is merely impossible.

Climate change will not only change species distribution but also community
composition \citep{Thuiller.2006}.

To address such differences, researchers could model
habitat and movement preferences using season-dependent models, or,
alternatively, by combining hidden markov models with step-selection functions.
(cite papers that fieberg sent)

For the discussion: Through changing environmental conditions, climate change
will also impact the distribution of various species with far-reaching
consequences on inter- and intra-specific competition (Abrahms). Due to a lack
of data, we omitted any inclusion of social factors in our dispersal model,
albeit they are known to be important determinants of dispersal (Armansin).
Studying how climate change influences species distribution and their
interactions, and ultimately how this alters dispersal, will be challenging but
necessary to more realistically simulate dispersal.

Only recently, it has been discovered that AWDs communicate using shared marking
sites. The role of such marking sites for dispersing coalitions remains to be
investigated, yet it is is likely that, akin to resident packs, use SMS as
navigation waypoint and demarcation lines. Chemical analyses suggest that the
compounds used for communication are highly volatile and may not persist in
extreme climate conditions. In result, dispersers may lose their ability to
effectively navigate across the landscape and locate potential mates with whom
to settle. This would reduce pack-formation prospects and undermine...

Validating predictions from individual-based dispersal models is challenging and
requires additional dispersal data, which is inherently scarce anyways. Sciticen
science may help to fill this gap by augmenting observed GPS data with
occasional sightings of uncollared dispersing coalitions. This is especially
ciritical for species that disperse across borders and beyond confined study
areas. The African carnivore wildbook offers ...

Although the OD is arguably the main driver of seasonal change across the
studied ecosystem, there are several other factors that undergoe seasonal
change, including vegetation and the abundance and distribution of prey or
predators.

The OD is an important driver of species distribution and it has been found that
an expanding flood limits available habitat, thus leading to more inter-specific
competition, particularly between AWDs and lions.

Although validating the predictions resulting from our dispersal model is
challenging, the simulations enable us to generate several interesting
hypotheses...

\subsection{Conclusion}
Our dispersal simulations across two extreme climatic scenarios reveal striking
differences in dispersal prospects and landscape connectivity for dispersing
AWDs. This implies that (1) climatic variation, be it due to seasonality or
climate change, must be included in anlayses dealing with dispersal and (2) that
projected climate change is likely to have profound impacts on landscape
connectivity.

\section{Authors' Contributions}
D.D.H., D.M.B., A.O. and G.C. conceived the study and designed methodology;
D.M.B., G.C., and J.W.M. collected the data; D.D.H. and D.M.B. analysed the
data; G.C. and A.O. assisted with modeling; D.D.H., D.M.B., and G.C. wrote the
first draft of the manuscript and all authors contributed to the drafts at
several stages and gave final approval for publication.

\section{Data Availability}
GPS movement data of dispersing wild dogs is available on dryad
\citep{Hofmann.2021b}. Access to R-scripts that exemplify the application of the
proposed approach using simulated data are provided through Github
(\url{https://github.com/DavidDHofmann/DispersalSimulation}). In addition, all
codes required to reproduce the African wild dog case study will be made
available through an online repository at the time of publication.

\section{Acknowledgements}
We thank the Ministry of Environment and Tourism of Botswana for granting
permission to conduct this research. We thank C. Botes, I. Clavadetscher, and G.
Camenisch for assisting with wild dog immobilizations. We also thank B. Abrahms
for sharing her data of three dispersing wild dogs. Furthermore, we would like
to thank Johannes Signer for assisting with the simulation algorithm. This study
was funded by Albert-Heim Foundation, Basler Stiftung für Biologische Forschung,
Claraz Foundation, Idea Wild, Jacot Foundation, National Geographic Society,
Parrotia Stiftung, Stiftung Temperatio, Wilderness Wildlife Trust Foundation,
Forschungkredit der Universität Zürich, and a Swiss National Science Foundation
Grant (31003A\_182286) to A. Ozgul.

\newpage
\begingroup
\singlespacing
\bibliography{Literature}
\endgroup

\end{document}