### 2023-03-14 Chapter_8/04_Manuscript/Manuscript_2023-01-19.tex

\documentclass[abstract=on,10pt,a4paper,bibliography=totocnumbered]{article}
\usepackage[paper=a4paper,left=35mm,right=35mm,top=25mm,bottom=30mm]{geometry}
\usepackage[doublespacing]{setspace}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[round]{natbib}
\usepackage{amsmath}
\usepackage{colortbl}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{gensymb}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage[hidelinks]{hyperref}
\usepackage[nameinlink]{cleveref}
% \usepackage{lineno}
\usepackage{multirow}
\usepackage{arydshln}
\usepackage[flushleft]{threeparttable}
\usepackage[nomarkers, nolists]{endfloat}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{scalerel}
\usepackage{tikz}
\usetikzlibrary{svg.path}

%------------------------------------------------------------------------------
%	Some Styling
%------------------------------------------------------------------------------
% Creating some TikZ styles
\tikzset{
  nonterminal/.style = {rectangle
    , minimum size = 6mm
    , very thick
    , draw = black!
  }
}

% Changing the style of captions in figures etc.
\captionsetup{labelfont=bf, format=plain, font=small}

% Change how equations are referenced
\renewcommand{\theequation}{Equation \arabic{equation}}%

% To be able to put an ORCID
\definecolor{orcidlogocol}{HTML}{A6CE39}
\tikzset{
  orcidlogo/.pic={
    \fill[orcidlogocol] svg{M256,128c0,70.7-57.3,128-128,128C57.3,256,0,198.7,0,128C0,57.3,57.3,0,128,0C198.7,0,256,57.3,256,128z};
    \fill[white] svg{M86.3,186.2H70.9V79.1h15.4v48.4V186.2z}
                 svg{M108.9,79.1h41.6c39.6,0,57,28.3,57,53.6c0,27.5-21.5,53.6-56.8,53.6h-41.8V79.1z M124.3,172.4h24.5c34.9,0,42.9-26.5,42.9-39.7c0-21.5-13.7-39.7-43.7-39.7h-23.7V172.4z}
                 svg{M88.7,56.8c0,5.5-4.5,10.1-10.1,10.1c-5.6,0-10.1-4.6-10.1-10.1c0-5.6,4.5-10.1,10.1-10.1C84.2,46.7,88.7,51.3,88.7,56.8z};
  }
}
\newcommand\orcid[1]{\href{https://orcid.org/#1}{\mbox{\scalerel*{
\begin{tikzpicture}[yscale=-1,transform shape]
  \pic{orcidlogo};
\end{tikzpicture}
}{|}}}}

%------------------------------------------------------------------------------
%	Titlepage: Header
%------------------------------------------------------------------------------
% \title{Flooding of the Okavango Delta influences Connectivity for Dispersing
% African Wild Dogs}
\title{African Wild Dog Dispersal and Connectivity under Climate Change -
Lessons Learned from Seasonal Flood Extremes}

% List of Authors
\author{
  David D. Hofmann\textsuperscript{1,2,\S} \orcid{0000-0003-3477-4365} \and
  Dominik M. Behr\textsuperscript{1,2} \orcid{0000-0001-7378-8538} \and
  John W. McNutt\textsuperscript{2} \and
  Arpat Ozgul\textsuperscript{1} \orcid{0000-0001-7477-2642} \and
  Gabriele Cozzi\textsuperscript{1,2} \orcid{0000-0002-1744-1940}
}

% Reduce spacing between authors
\makeatletter
\def\and{%
  \end{tabular}%
  \hskip -0.5em \@plus.17fil\relax
  \begin{tabular}[t]{c}}
\makeatother

% Current Date
% \date{\today}

% And here the masterpiece begins
\begin{document}

% Change page numbering
\pagenumbering{gobble}

% Required to be able to cite
\bibliographystyle{apalike}

% Create Titlepage
\maketitle

%------------------------------------------------------------------------------
%	Titlepage: Additional Info
%------------------------------------------------------------------------------
\begin{flushleft}

\vspace{0.5cm}

\textsuperscript{1} Department of Evolutionary Biology and Environmental
Studies, University of Zurich, Winterthurerstarsse 190, 8057 Zurich,
Switzerland.

\textsuperscript{2} Botswana Predator Conservation Program, Private Bag 13,
Maun, Botswana.

\textsuperscript{\S} Corresponding author: david.hofmann2@uzh.ch

\vspace{4cm}

\textbf{Running Title:} Seasonal Flooding of the Okavango Delta and its
Consequences for African Wild Dog Dispersal and Connectivity

\textbf{Running Title:} Seasonal Flooding of the Okavango Delta and its
Consequences for African Wild Dog Connectivity under Climate Change

\vspace{0.5cm}

\textbf{Keywords:} movement, connectivity, climate change, seasonality,
dispersal, conservation, individual-based

\end{flushleft}

%------------------------------------------------------------------------------
%	Abstract
%------------------------------------------------------------------------------
\newpage
\begin{abstract}
Climate change is expected to profoundly impact the life history of wild-living
animal populations. While the impacts of climate change on the demographics of
local subpopulations have been studied repeatedly, little is known about the
consequences of environmental change on dispersal and connectivity.

We capitalize on a ``natural experimental setup'', the flood-pulse driven change
in surface-water across the Okavango Delta in northern Botswana, to investigate
the impact of changing environmental conditions on dispersal patterns and
connectivity of the endangered African wild dog (\textit{Lycaon pictus}). For
this, we simulate dispersal trajectories across the Okavango Delta under two
extreme environmental scenarios; one assuming a maximum flood, one assuming a
minimum flood. These scenarios serve to approximate environmental conditions
under amplified climatic variability, as it is expected under continued climate
change.

During maximum flood, the Okavango Delta poses an important dispersal barrier
that reduces dispersal prospects between distinct areas and increases dispersal
durations. At maximum flood, dispersal into the central habitats of the Okavango
Delta is reduced by $\input{99_NumberReachingArea6Percentage.tex}$\% and
dispersal durations increased by
$\input{99_StepsToReachingArea6Percentage.tex}$\%. Across the entire study area,
the differences are less dramatic, with
$\input{99_NumberReachingOthersPercentage.tex}$\% lower dispersal success and
$\input{99_StepsToReachingOthersPercentage.tex}$\% longer dispersal durations
during maximum flood. We also found that a maximum flood shifted dispersal
bottlenecks and areas with a high potential for human-wildlife conflict closer
towards densely populated villages. We provide a rich suite of maps to visualize
the differences between minimum and maximum flood.

Whilst the exact impacts of climate change on the flooding regime of the
Okavango Delta remain unknown, our results suggest that connectivity will vastly
differ depending on future flood conditions. Considering such differences will
be key to designing future proof conservation strategies, particularly in light
of ongoing climate change. Our results also have direct applicability for the
conservation of the endangered African wild dog as we highlight critical
dispersal corridors and human-wildlife conflict zones in two distinct scenarios.

\end{abstract}

%------------------------------------------------------------------------------
%	Main Text
%------------------------------------------------------------------------------
\newpage

\onehalfspacing
\tableofcontents
\doublespacing

% Change page numbering
\newpage
\pagenumbering{arabic}

% Create linenumbers
% \linenumbers

\section{Introduction}
\subsection{Climate Change and Dispersal}
Climate change is expected to profoundly impact ecosystems across the globe with
far-reaching consequences for the species living therein \citep{Ozgul.2010,
Radchuk.2019, IPCC.2022}. By altering environmental conditions, climate change
affects animal behavior \citep{Fuller.2016}, resource availability
\citep{Durant.2007}, population dynamics \citep{Paniw.2021}, and the
distribution of wild living animal populations \citep{Thomas.2004,
Thuiller.2006}. An important life-history pathway through which species may
mediate the negative consequences of climate change is dispersal
\citep{Anderson.2012}, i.e. the movement of individuals away from their natal
location to the site of first reproduction \citep{Clobert.2012}. Through
dispersal, species may adapt to climate change by tracking favorable habitat
conditions \citep{Raia.2012} and by shifting into a different region of their
fundamental niche \citep{Kokko.2006}. Dispersal also facilitates the
colonization empty habitats \citep{Gustafson.1996, Hanski.1999a,
MacArthur.2001}, promotes the reinforcement of weakened and small subpopulations
\citep{Brown.1977}, and safeguards genetic diversity \citep{Frankham.2002,
Leigh.2012, Baguette.2013}, thus providing additional resilience against
changing environmental conditions \citep{Kokko.2006, Fahrig.2003}.

\subsection{Connectivity}
While dispersal offers a means to offset the negative demographic consequences
of climate change \citep{Kokko.2006, Hodgson.2009, Travis.2013}, it itself is a
function of climatic and environmental conditions (e.g. \citealp{Elliot.2014,
Behr.2020}). The link between dispersal and the environment can either be
indirect, for example if the propensity of individuals to disperse depends on
environmental conditions, or direct, when the biophysical environment through
which dispersers move affects dispersal prospects \citep{Travis.2013}. The
latter highlights that dispersal is also inextricably linked to the concept
landscape connectivity \citep{Baguette.2013}, which is understood as the degree
to which the landscape facilitates or impedes movements \citep{Taylor.1993}. A
sufficient degree of landscape connectivity is thus a critical prerequisite for
successful dispersal \citep{Fahrig.2003}, yet the continued degradation and
destruction habitats worldwide continues to imperil the dispersal prospects of
many species \citep{Melbourne.2008, Sawyer.2011}. Conservation strategies that
aim at facilitating dispersal by improving landscape connectivity are therefore
often viewed as pinnacle of all conservation strategies \citep{Heller.2009}.
Despite this, our understanding of dispersal and its implications for
connectivity is limited, especially in light of changing environmental
conditions.

\subsection{Modeling Connectivity}
To study dispersal and connectivity, various modeling techniques have emerged
(see e.g. \citealp{Etherington.2016} and \citealp{Diniz.2019} for overviews).
Initially, the techniques were limited to examining structural aspects of
connectivity by focusing on the composition and configuration of habitat
patches, while ignoring species' responses to the landscape matrix
\citep{Tischendorf.2000, Doerr.2011}. With the increasing availability of
telemetry data and methods to study species' habitat and movement preferences
\citep{Boyce.2002, Fortin.2009, Cushman.2010, Avgar.2016}, preferably during
dispersal \citep{Elliot.2014}, however, the focus has shifted towards a more
functional view on connectivity, which also takes into account how species
interact with their surroundings \citep{Tischendorf.2000, Doerr.2011}.
Currently, the most prominent \textit{functional} connectivity models are based
on least-cost path analysis (LCPA, \citealp{Adriaensen.2003}) and circuit theory
(CT, \citealp{McRae.2008}), two graph-based methods that estimate conductance of
the landscape by means of a resistance (or inversely permeability) surface
\citep{Zeller.2012}. Such a surface is meant to reflect the ease or willingness
at which the focal species traverses a specific area and is generated by
consolidating multiple habitat layers into a single layer of resistance
\citep{Zeller.2012}. Since both LCPA and CT approaches make assumptions that are
rarely met by dispersing individuals, individual-based movement models (IBMMs),
in which dispersal movements are simulated explicitly, have also gained some
momentum \citep{Kanagaraj.2013, Allen.2016, Hauenstein.2019, Diniz.2019,
Zeller.2020, UnnithanKumar.2022, UnnithanKumar.2022a, Hofmann.2023}. IBMMs
provide great modeling flexibility and are thus considered powerful tools for
examining connectivity under different landscape configurations
\citep{Littlefield.2019, UnnithanKumar.2022a}. However, most connectivity
studies focus on a snapshot in time and fail to account for changing
environmental conditions, such as those akin to climate change. Moreover, the
challenges associated with studying dispersing animals further impairs the
collection of data during dispersal at the appropriate temporal and spatial
scale \citep{Graves.2014, Vasudev.2015} and weakens our ability to project
dispersal prospects under changing environmental conditions into the future.

\subsection{Climate Change and Seasonality}
Predicting the impacts of climate change on dispersal and connectivity is
non-trivial and typically requires spatial information about future climatic or
environmental conditions over the area of interest \citep{Littlefield.2019}.
This information can then be used in various ways. \cite{Ashrafzadeh.2019}, for
example, combined climatic predictions until 2070 with a species-distribution
model for mountain newts (\textit{Neurergus kaiseri}) in Iran to demonstrate a
decrease in connectivity due to increased habitat fragmentation. Similarly,
\cite{Luo.2021} mapped the future distribution of the giant spiny frog
(\textit{Quasipaa spinosa}) under different representative climate pathways and
reported a reduction in connectivity for the species across South-East Asia. In
these studies, the focus lies on the impacts of climate change on species
distribution and subsequent changes in connectivity due to the configuration of
habitat patches, yet less on the habitat matrix and its implications for
dispersal. For martens, (\textit{Martes americana}), \cite{Wasserman.2012}
developed several resistance layers emerging under different climate scenarios
and find that already low warmings will result in increased isolation of
remaining subpopulations. While not primarily focused on climate change, another
body of literature captures environmental variability by generating resistance
surfaces for different scenarios. \cite{Mui.2017}, for instance, developed
seasonal resistance maps for Blanding's turtle \textit{Emydoidea blandingii}
showing that connectivity was substantially lower in late summer compared to
spring. Similarly, \cite{Osipova.2019} studied connectivity for African
elephants (\textit{Loxodonta africana}) during wet and dry season and found that
ignoring seasonality resulted in an underestimation of connectivity during the
wet season and an overestimation during the dry season. For the same species,
\cite{Kaszta.2021} provide monthly updated connectivity maps revealing that
connectivity varies strongly across a typical year. Finally, \cite{Zeller.2020}
use dynamic resistance surfaces showing differences in connectivity for black
bears \textit{Ursus americano}. Altogether, the studies exemplify that
connectivity should not be regarded as static in time, but dynamic across and
within years.

In many cases, anticipating environmental conditions under climate change is not
viable as relevant data is not available or entails major uncertainty
\citep{Collins.2012}. This is particularly true for complex ecosystems with
intricate feedback loops and in cases where one is interested in landscape
characteristics, rather than climatic conditions. In general, it is accepted
that aside from increasing temperatures, climate change will also amplify the
frequency and magnitude of extreme events, such as severe droughts, heavy
precipitation, floods, and storms \citep{Stott.2016, Ummenhofer.2017,
IPCC.2022}. Thus, instead of attempting to study the impacts of climate change
directly, one may capitalize on naturally occurring fluctuations of the
environment to gauge the likely consequences of shifting the system towards what
is currently considered an extreme.

\subsection{Okavango Delta}
The Okavango Delta (OD) in Southern Africa poses a unique opportunity to study
the impacts of extreme environmental change on species dispersal ability and
connectivity in a large scale natural experiment setup. The OD is the world's
largest inland delta and characterized by substantial seasonal differences in
surface-water cover. Throughout the course of a year, the area covered by the
OD's floodwaters can fluctuate between 3'000 and 10'000 km\textsuperscript{2}
with striking variability within and between years \citep{Gumbricht.2004,
Wolski.2017}. Importantly, the region is among the most vulnerable to climate
change, as a temperature increase of 4 to 6\degree C above pre-industrial levels
is expected within the 21\textsuperscript{st} century \citep{Engelbrecht.2015,
Akinyemi.2019}, which is far beyond the global average \citep{IPCC.2022}. A
keystone predator in this ecosystem and an umbrella species for conservation
efforts is the African wild dog (AWD, \textit{Lycaon pictus}). While the species
was once widespread across across entire Sub-Saharan Africa, it has disappeared
from a vast majority of its historic range, mainly due to human persecution,
deadly diseases, and continued destruction and degradation of its habitats
\citep{Woodroffe.2012}. AWDs are characterized by an unsurpassed dispersal
ability, as young individuals that leave their natal pack can cover several
hundred kilometers within a few days \citep{Davies-Mostert.2012, Masenga.2016,
Cozzi.2020}. Dispersal typically happens in dispersal coalitions of same-sex
siblings  \citep{McNutt.1996}. Previous research has shown that the ODs
floodwaters pose a major dispersal barrier, yet the analysis was centered around
an average flooding scenario \citep{Hofmann.2021}.

\subsection{What We Did}
Here, we utilize a previously parameterized and validated dispersal model as
IBMM to simulate dispersal and assess dispersal success and connectivity
patterns for African wild dogs under two extreme scenarios: one assuming maximum
flooding of the Okavango delta and one assuming minimum flooding of the Okavango
delta. The IBMM was trained using GPS data collected during dispersal and
frequently updated environmental data, thus providing a high degree of realism
\citep{Hofmann.2021}. Given that dispersers avoid crossing through water (albeit
we do occasionally observe it in the field), we anticipated that dispersal
prospects and connectivity during maximum flood would be low. Moreover,  when
the flood extent of the OD is at its maximum, the water extends almost into the
densely populated village of Maun. Since both the flood and humans are avoided
by dispersing wild dogs, we anticipated that a fully flooded Delta would result
in a total halt of movement between the Western and Eastern side of the OD.

\section{Materials and Methods}
We conducted all data preparation and analyses using the programming language
\textsf{R} \citep{RCoreTeam.2022}. For any spatial data manipulation, we used
the packages \textsf{terra} \citep{Hijmans.2022} and \textsf{spatstat}
\citep{Baddeley.2015}. Several helper functions for the dispersal simulation
algorithm were written in \textsf{C++} and imported to R using the \textsf{Rcpp}
package \citep{Eddelbuettel.2011}. Network analysis was achieved in
\textsf{igraph} \citep{Csardi.2006} and figures were generated using
\textsf{ggplot2} \citep{Wickham.2016} and \textsf{ggnetwork}
\citep{Briatte.2021}. All R-scripts required to replicate our analyses are
provided through an online repository.

\subsection{Study Area}
The study area for this analysis was focused on the Okavango delta (OD) and its
surroundings in Southern Africa, comprising parts of Angola, Namibia, Botswana,
Zimbabwe, and Zambia (\Cref{StudyArea}). While our primary focus lied on the
immediate surroundings of the Okavango Delta, we considered an extended
rectangular extent stretching from 20\degree 30' E to 26\degree E. This
encompasses a total of 300'000 km\textsuperscript{2}) and comprises all long
distance dispersal events previously recorded in this area \citep{Cozzi.2020,
Hofmann.2021}. The annual flood-pulsing rhythm of the OD is mainly dictated by
precipitation in the Angolan highlands, which serve as catchment areas from
which water is further channeled into the Okavango River and transported into
the OD \citep{Wolski.2017}. Although precipitation reaches its maximum between
December and March, the collected water only slowly descends through the
Okavango River and its distributaries, reaching the distal ends of the delta
only towards July or August. In result, peak flooding is out of sync with local
precipitation, such that the flood usually arrives in the OD during the peak dry
season. Once the water reaches the OD's distal ends, it percolates at the
Thamalakane and Kunyere Faults, two natural faultlines at which the waterflow is
hindered. During minimum extent, the flood covers an area of only 3'600
km\textsuperscript{2}, whereas during maximum flood more than 9'000
km\textsuperscript{2} are flooded. Vegetation in this study area is dominated by
mopane forest, mixed acacia woodland, and grassland. Human influence is low and
mainly concentrated around small villages at the western periphery of the OD as
well as the city of Maun at the south-eastern tip of the OD. Large portions of
land are dedicated national parks, game reserves or forest reserves. The study
area is also part of the world's largest transboundary conservation initiative,
the Kavango-Zambezi Transfrontier Conservation Area. Previous studies have
attributed a high potential of this initiative for improving connectivity for
various species \citep{Brennan.2020, Lines.2021, Hofmann.2021}.

\begin{figure}
  \begin{center}
  \includegraphics[width = \textwidth]{99_StudyArea.png}
  \caption{(a) Location of the study area in Southern Africa across which we
  simulated dispersing African wild dogs. To mitigate edge effects, our study
  area comprised a buffer zone (red polygon) within which we randomized
  covariate values of the habitat layers. (b) The study area was centered on the
  Okavango Delta and encompassed its immediate surroundings. We initiated
  simulated dispersers at random locations within one of the six source areas
  (orange polygons) that we distributed across the delta. Emigration zones
  (purple polygons) served as checkpoints to identify if and where simulated
  dispersers left the close surroundings of the Okavango delta. These zones were
  generated using a set of cutlines (black dotted lines) originating from the
  center of the delta that dissected an elliptical buffer surrounding the delta
  into sections of equal size and in accordance with the cardinal directions.}
  \label{StudyArea}
  \end{center}
\end{figure}

\subsection{Spatial Habitat Layers}
We represented the physical landscape through which dispersers could move by a
set of spatially referenced habitat layers, each with a resolution of 250 m. The
set of layers included water-cover, distance-to-water, tree-cover,
shrub/grassland-cover, and a human influence layer depicting anthropogenic
influences through villages, roads, and agriculture. A detailed description of
the different habitat layers is provided in \citep{Hofmann.2021}. Importantly,
the water-cover and derived distance-to water layers were generated using MODIS
Terra MCD43A4 satellite imagery that was classified into binary water-cover maps
using a ``floodmapping'' algorithm developed by \citep{Wolski.2017}. This
allowed us to generate almost weekly updated ``floodmaps'', thus providing
detailed information about the flood-extent at any given point in time. In
total, we generated 700 floodmaps between the years 2000 and 2019, based on
which we generated a minimal and maximum flood scenario. To create the minimum
(maximum) flood scenario, we averaged the 50 floodmaps with the smallest
(highest) flood extent and generated a binary layer by masking all pixels that
were inundated in less than 50\% of the maps. The resulting maps are presented
in \Cref{FloodExtent}. Ultimately, we combined the habitat layers into two
stacks, one representing the minimum flood scenario, one representing the
maximum flood scenario. To mitigate edge effects during the dispersal
simulation, we followed \cite{Koen.2010} and expanded the spatial extent of the
stacked layers by 20\% and randomized habitat values within the so created
buffer zone (red rectangle in \Cref{StudyArea} a).

\begin{figure}
  \begin{center}
  \includegraphics[width = \textwidth]{99_FloodExtent.png}
  \caption{Flood extent in the minimum (a) and maximum (b) flood scenarios. In
  the minimum flood scenario (a), water stretched across 3'558
  km\textsuperscript{2}, whereas during maximum flood (b) it covered 9'397
  km\textsuperscript{2}. The two maps were generated using 700 remote sensed
  MODIS MCD43A4 satellite images spanning the years 2000 to 2019.}
  \label{FloodExtent}
  \end{center}
\end{figure}

\subsection{Source Areas and Emigration Zones}
We simulated dispersing AWDs originating from six distinct source areas located
in the vicinity of the OD (\Cref{StudyArea}). We placed source areas in regions
that remained dry during both the minimum and maximum flooding scenario and are
known to host viable wild living wild dog populations. While source areas one to
five were located across the delta's periphery, source area six laid in the OD's
center. The selection of distinct source areas served to facilitate the
identification and quantification of the number of successful dispersal events
between different regions of the OD. Besides source areas, we also generated
``emigration zones'' that we used as checkpoints to determine if and where
simulated individuals left the delta's immediate surroundings
(\Cref{StudyArea}). We generated these zones by first overlaying the OD with an
elliptic that we dissected into roughly equally sized polygons in accordance
with cardinal points (\Cref{StudyArea}).

\subsection{Dispersal Simulation}
We used a previously parameterized and validated dispersal model to simulate
dispersal of AWDs. The dispersal model was trained using GPS data of 16 wild dog
coalitions dispersing across northern Botswana \citep{Hofmann.2023} which was
fed into an integrated step-selection function (iSSF, \citealp{Avgar.2016}). In
iSSFs, consecutive GPS locations are converted into steps (the straight-line
traveled between two GPS recordings \citep{Turchin.1998a}) and compared to a set
of \textit{random} steps in a conditional logistic regression framework
\citep{Fortin.2005, Thurfjell.2014, Muff.2020, Fieberg.2021}. The model
presented in \cite{Hofmann.2023} comprised of a movement kernel, describing how
dispersers move across the landscape in the absence of habitat selection, a
habitat kernel, indicating preferred or avoided habitat features, and
interactions among the two, i.e. how movement behavior changes depending on
habitat conditions. According to this model, the main characteristics of AWD
dispersal movements are avoidance of water, avoidance of areas influenced by
humans, and a preference for directional and fast movements. The model
parameters are provided in Appendix SX and explained in \citealp{Hofmann.2023}.

Originating from each of the six source areas, we simulated 2'000 individuals
dispersing for a total of 2'000 steps. 1'000 individuals were simulated assuming
a minimum flood, the remaining 1'000 assuming a maximum flood. This resulted in
the simulation of a total of 12'000 individuals. The simulation procedure was
based on the algorithm described in \Cref{Hofmann.2023} and works as follows. A
random location within the source area is defined as starting point. Originating
from the starting point, a set of 25 random steps is generated by sampling step
lengths from a gamma distribution fitted to observed steps (shape = 0.37, scale
= 6'316) and turning angles from a uniform distribution ($-\pi, +\pi$). Along
each random step the underlying spatial covariates are extracted, and relevant
movement metrics are computed. \(\beta-\)estimates from the fitted model are
used to predict the probability of each step for being chosen, given the steps
associated covariates. Among the 25 proposed steps, one is chosen at random
based on assigned probabilities. The location of the animal is updated, and the
procedure is repeated until the desired number of steps is realized. Here, we
simulated each individual for 2'000 steps, corresponding to a dispersal duration
of 400 days and the longest dispersal duration recorded in this study area
\citep{Cozzi.2020, Hofmann.2021}. The simulated trajectories can be understood
as correlated random walks.

\subsection{Derived Metrics}
Based on simulated dispersal trajectories we quantified connectivity and
identified areas of elevated potential for human wildlife conflict. Our
assessment of connectivity was based on the three complementary connectivity
metrics for IBMMs discussed in \cite{Hofmann.2023}. The set of metrics comprised
of \textit{heatmaps}, depicting areas of intense use, \textit{betweenness maps},
highlighting dispersal corridors and bottlenecks and \textit{maps of inter-patch
connectivity}, visualizing dispersal success, and duration into distinct habitat
patches. We generated heatmaps by superimposing the study area with a grid with
a spatial resolution of 1 km and quantifying the frequency of simulated
trajectories traversing each grid cell. To compute spatially mapped betweenness
scores, we overlaid the study area with a grid that had a resolution of 2.5 km
and determined the frequency at which simulated individuals transitioned from
one grid-cell to another. A cell-transition was said to occur whenever a
simulated step crossed from one grid-cell across or into another. In case the
same individual repeatedly realized the same cell-transition, we only counted a
single transition to avoid emphasis on regions where individuals moved in
circles. This resulted in a weighted edge-list that we used to compute weighted
betweenness scores for each grid-cell, i.e. the importance of the respective
grid-cell in facilitating movement into adjacent areas
\citep{Bastille-Rousseau.2018, Bastille-Rousseau.2021}. Betweenness was computed
using the \textsf{igraph} R-package \citep{Csardi.2006}. Because the
computations associated with calculating betweenness scores are computationally
more demanding, we deemed the grid size of 2.5 km a sensible compromise between
efficiency and resolution. As a final connectivity metric, we computed the
number of successful dispersal events between each of the six distinct source
areas. We coin this type of connectivity ``inter-patch connectivity''. Dispersal
between two areas was said to be successful whenever the trajectory of an
individual leaving one area intersected with the polygon of another area. We
also estimated the number of individuals that left the OD's vicinity and moved
into an emigration zone. To quantify the dispersal durations required to move
between source areas or into emigration zones, we recorded the minimum number of
steps that individuals moved before arriving at the respective destination.
Besides connectivity, we also identified zones with a high potential for human
wildlife conflict. For this, we isolated all simulated locations where simulated
individuals moved within 500 meters of the nearest human-influenced grid-cell.
Based on the so isolated coordinates we generated density maps. To highlight
differences between derived metrics during maximum and minimum flooding, we
computed difference maps for the heatmap, betweenness map, and human wildlife
conflict maps.

\section{Results}
Figures depicting the derived connectivity and human-wildlife conflict maps are
provided in \Cref{Metrics}. Difference maps to visualize the differences between
minimum and maximum flood are given in \Cref{Difference}. Source-specific maps
are presented in the Appendix. As the heatmaps in \Cref{Metrics}a reveal, the OD
acts as major dispersal barrier during maximum flood yet reveals vital
dispersal habitat during minimum flood. Differences between maximum and minimum
flood (\Cref{Difference}a) are particularly pronounced for the region between
source areas 1 and 2, where few dispersers occur during times of maximum flood.
In fact, because the floodwaters of the OD reach almost into Maun, the OD
creates a line of separation between its eastern and western sections. The
separation is further amplified as the city of Maun is avoided by dispersers in
both scenarios. Similar patterns are observed on the betweenness maps
(\Cref{Metrics}b), where several pinch-points and bottlenecks linking source area
6 to the surrounding source areas exist during minimum flood. During maximum
flood, however, these links vanish and instead a single corridor at the
south-eastern tip of the OD emerges (\Cref{Difference}b). Despite its apparent
importance in linking the eastern and western sections of the delta, it is
evident from (\Cref{Metrics}a) that this corridor is only rarely used,
especially during the maximum flood scenario. As for the potential for human
wildlife conflict, two clusters emerge (\Cref{Metrics}c). The first cluster lies
at the inflow of the Okavango Delta between source areas 4 and 5 and is most
pronounced during minimum flood (\Cref{Difference}c). Another, albeit visually
less distinct, cluster covers the area at the distal end of the OD, stretching
from lake Ngami to Maun. This area appears particularly relevant at maximum
flood (\Cref{Difference}c). Our analysis of inter-patch connectivity further
demonstrates notable differences in dispersal prospects and dispersal durations
depending on the extent of the flood (\Cref{Metrics}d and \Cref{IPCTable}).
While $\input{99_NumberReachingOthersMinFlood.tex}$ simulated dispersers reach
another source area during minimum extent, only
$\input{99_NumberReachingOthersMaxFlood.tex}$ do so during maximum extent, thus
indicating an overall decrease in dispersal success of
$\input{99_NumberReachingOthersPercentage.tex}$\% during maximum flood.
Concomitantly, the average minimum dispersal durations increases by
$\input{99_StepsToReachingOthersPercentage.tex}$\%, i.e. from
$\input{99_NumberReachingOthersMinFlood}$ steps to
$\input{99_NumberReachingOthersMaxFlood}$ steps during maximum flood. These
differences are particularly pronounced for individuals dispersing into source
area 6. While the area is reached by
$\input{99_NumberReachingArea6MinFlood.tex}$ simulated individuals during
minimum flood, only $\input{99_NumberReachingArea6MaxFlood.tex}$, i.e.
$\input{99_NumberReachingArea6Percentage.tex}$\% less, arrive there during
maximum flood. Furthermore, the dispersal duration into source area six from any
other source area increases by $\input{99_StepsToReachingArea6Percentage}$\%
from $\input{99_StepsToReachingArea6MinFlood.tex}$ steps to
$\input{99_StepsToReachingArea6MaxFlood.tex}$ steps.

\begin{figure}
  \begin{center}
  \includegraphics[width = 0.8\textwidth]{99_Metrics.png}
  \caption{(a) Heatmaps, (b) betweenness maps, (c) maps of human wildlife
  conflict, and (d) maps of inter-patch connectivity, derived from simulated
  dispersal events. Left panels were derived from the minimum flood scenario,
  right panels from the maximum flood scenario. Source areas from which
  dispersers were released are numbered 1-6.  The color scale for betweenness
  scores in (b) was square-rooted to improve visibility of corridors with lower
  values. Note that for clarity in (d) we only present links between adjacent
  source areas. Additional, source-specific maps for each of the four metrics
  are provided in the appendix.)}
  \label{Metrics}
  \end{center}
\end{figure}

\begin{figure}
  \begin{center}
  \includegraphics[width = \textwidth]{99_Difference.png}
  \caption{Difference maps generated from the (a) heatmaps, (b) betweenness
  maps, and (c) maps of human wildlife conflict in \Cref{Metrics}. The maps
  depict the difference between maximum and minimum flood for each metric (i.e.
  $Metric_{flood:max} - Metric_{flood:min}$). Orange regions indicate that the
  respective metric was higher during minimum flood, blue regions that the
  metric was higher during maximum flood.}
  \label{Difference}
  \end{center}
\end{figure}

\begin{table}
  \caption{(a) Dispersal frequency and (b) duration (in steps) between source
  areas (labeled 1 to 6) and emigration zones (labeled 7 to 14) during minimum
  and maximum flood.}
  \label{IPCTable}
  \includegraphics[width = \textwidth]{99_IPCTable.png}
\end{table}

\section{Discussion}
\subsection{Brief Summary}
According to our simulations, the propensity to move between the eastern and
western part of the delta is much lower during maximum extent. This is mainly
due to the flood-waters and the city of Maun acting as dispersal barriers.
During maximum extent, the floodwaters of the delta close a gap between the
delta and Maun that otherwise would serve as dispersal corridor. Anecdotal
evidence supports this hypothesis, for the only dispersing individuals recorded
to move from the eastern to the western part of the delta moved at times of low
flood. In line with this, it appears that a large flood extent pushes dispersing
individuals to move closer to human inhabited areas such as the village of Maun.

Predicting how climate change will impact the dispersal ability of AWDs is
challenging for multiple reasons. First of all, predicting the flooding patterns
of the OD under climate change is non-trivial due to complex feedback-loops
between surface-temperature, soil conditions, precipitation patterns, and the
associated changes in vegetation.

Second, the delta is not only prone to changes
in environmental conditions, but also to changes in anthropogenic use of the
inflowing water. Finally, it is unclear how AWDs, in fact, how any
species, will cope with environmental change due to global warming. Even though
some studies predict that AWD populations are likely to decline under increasing
temperatures, these studies fail to account for the behavioral plasticity of
their focal species. AWDs respiratory system, for instance, has evolved as a
perfect adaptation to high temperatures and AWDs may, in fact, profit from a
comparative advantage (cite an economist) over their competitors and prey under
rising temperatures. Although the theory of comparative advantages is a
fundamental concept in economics, it has yet to find its way into ecological
studies.

\cite{Murray-Hudson.2006} predicted that increased temperatures, additional human
abstractions, and reduced river flows might lead to a ``Delta dying'' and that
the impact of climate will be much more pronounced than the impact of
anthropogenic water use. However, the paper also found that predicted conditions
vary from much dryer to much wetter depending on the employed climate model.

While precipitation in the OD's catchment areas are expected to increase, it
remains unclear wheather the effect will be offset by increased
evapotranspiration. \cite{Wolski.2018} predict changes that can range from much
``wetter'' to much  ``drier'' conditions depending on the employed climate
model. In conclusion, there is too much uncertainty to make reliable
predictions, meaning that conservation practicioners will need to retain
flexibility in their planning.

Climate models consistently predict an above global average increase in
temperatures for Southern Africa.

There is large uncertainty regarding the future anthropogenic use and
abstractions from the Okavango river \citep{Hughes.2011}.

It has been suggested that climate change will intensivy human wildlife conflict
due to increased competition for scarce resources \cite{Abrahms.2021}. Instead
of \textit{increased} human wildlife conflict, our results suggest that human
wildlife conflict will shift. Given that these areas have had little exposure to
such conflicts in the past, inexperience and lack of knowledge may explain the
resulting apparent increases in conflict situations.

Besides the directional change in climate, it is agreed that there will be an
increased frequence in extreme climate events.

Along a similar vein, the shrinking of the sea-ice has resulted in a higher
incidence of problem polar bears in Canada over the past decades
\citep{Towns.2009}. Similarly, it has been found that the annual range shifts of
polar bears, in result to seasonally melting sea-ice, results in increased
frequency of HWC.

Step-selection functions (or path selection functions) provide better estimates
of connectivity than other resource selection approaches \citep{Zeller.2016}.

Information on habitat selection or connectivity are suitable for predicting
areas with an elevated potential for human wildlife conflict
\citep{Buchholtz.2020}.

Although local rainfalls in Botswana are expected to increase in terms of
intensity, simulations show that the length of the rainy season will decline,
more than offsetting the incline in precipitation \cite{Akinyemi.2019}.

Although dispersal is generally believed to be beneficial for the long term
viability of species, it can also lead to the net-loss of individuals in areas
where edge effects and anthropogenic mortality are high \citep{Leigh.2012}.

We studied a population of African wild dogs that resides in a natural
environment with little human influence. This is only representative for a small
share of the extant wild dog populations, as most individuals reside in areas
that are prone to substantial edge effects. For these populations, the benefit
of dispersal is disputed, as dispersal under high anthropogenic mortality may
lead to a net-loss of individuals \citep{Leigh.2012}.

Dispersal is an important, if not the most important, driver of landscape
connectivity and therefore of major interest to conservation authorities. It has
also been demonstrated that dispersal may lead to population declines in areas
where anthropogenic mortality is high and dispersal prospects low
\citep{Leigh.2012}.

Efforts aiming at restoring or improving landscape connectivity should consider
and take into account changes in landscape characteristics due to climate
change.

Maybe cite the \cite{Gusset.2009} paper.

We assessed the implications of environmental change on the dispersal prospects,
yet we did not consider how changing conditions alter dispersal propensity.

We studied the indirect (through the landscape), but not the direct (propensity
to disperse) conesquences of climate change on dispersal.

According to our simulations, dispersers are able to cover larger distances
during periods of low flood. This finding is little surprising, considering that
inundated areas act as dispersal barriers and force dispersers to detour and
circumvent water-covered areas. However, it still leads to an interesting
hypothesis. Previous studies have shown that the euclidean dispersal distance of
female coalitions is larger than that of male coalitions. This has led to ...
However, demographic analyses have also revealed that female offspring tend to
emigrate from their pack at younger ages and earlier in the year, when
floodwaters are still at a relatively low level (Behr). It is thus conceivable
that the sex-differences in reported dispersal distances is mainly a consequence
of environmental conditions during dispersal, rather than owed to physiological
differences between sexes.

Unsurprisingly, \cite{Michalski.2006} show that increased proximity between
wildlife and humans significantly increases the likelihood of human-wildlife
conflict.

\subsection{Effect of Climate Change on the Okavango Delta}
\cite{Akinyemi.2019} predicts fewer rain days in Botswana (may want to look at
predictions for Angola though)

While a better understanding of the ODs flooding regime under global change is
of major interest, the specific impacts of climate change on the OD are not
fully understood. This is owed to the complex interplay between climate,
anthropogenic water use, animal behavior, and topography of the region. Talk a
bit about anthropogenic water use (e.g. Namibia is building a hydropower station
in the Zambezi Region) Talk a bit about droughts, etc. Rising sea-level ->
Intrusion of salt water Simulations by \cite{Wolski.2006} reveal that upstream
abstractions have a smaller infuence on flooding patterns of the OD than climate
change. In combination, however, Simulated upstream abstractions and climate
change resulted in significant ``delta-dying'', i.e....

\subsection{Species Distribution}
Climate change and anthropogenic use will result in trophic ripple effects
\citep{Wolski.2006}. Climate change affects species distribution through habitat
suitability, which determines structural connectivity which ultimately
determines functional connectivity. How climate change will affect the
distribution of African wild dogs is not known. Given that the species has
occured across the entire Sub-Saharan continent suggests that this species is
highly adaptable and limited mainly by the presence of humans, yet less by the
habitat itself. In fact, wild dogs have been found on mount Kilimandjaro.

\subsection{Social Landscape}
The OD is an important driver of species distribution and it has been found that
an expanding flood limits available habitat, thus leading to more inter-specific
competition, particularly between AWDs and lions (Robynne).
Although the OD is arguably the main driver of seasonal change across the
studied ecosystem, there are several other factors that undergoe seasonal
change, including vegetation and the abundance and distribution of prey or
predators.An additional complication arises when species movement is not solely driven by
environmental conditions, but also affected by itra- and inter-specific factors.
For instance, ... has shown that dispersers...
Rendering such conditions alone is challenging, yet rendering the conditions
under changing environmental conditions is merely impossible.
As highlighted in the papers presented in the intro, climate change change
environmental conditions, and in return species distribution, which will, may
ultimately culminate in alterations of community composition
\citep{Thuiller.2006}. Previous studies have shown that intra-specific factors
are imortant determinants of dispersal movements \citep{Cozzi.2018} and it can
be expected that inter-specific factors are similarly important
\citep{Armansin.2019}. Through changing environmental conditions, climate change
will also impact the distribution of various species with far-reaching
consequences on inter- and intra-specific competition (Abrahms). Due to a lack
of data, we omitted any inclusion of social factors in our dispersal model,
albeit they are known to be important determinants of dispersal (Armansin).
Studying how climate change influences species distribution and their
interactions, and ultimately how this alters dispersal, will be challenging but
necessary to more realistically simulate dispersal.

\subsection{Seasonality in Dispersal Models}
While our analysis marks an important step into incorporating environmental
change into studies of connectivity, there are several critical additions that
should be considered in the future. We studied dispersal and connectivity under
two different environmental scenarios, yet our movement model assumed that
dispersers had identical habitat and movement preferences in both scenarios. In
reality, however, it can be expected that movement and habitat kernels of
dispersers differ depending on the season considered (examples).
Can incporporate seasonality by keeping environmental conditions constant, but
rendering changes in species habitat preferences. Or, keep species habitat
preferences constant and vary the environemnt. Or both. To address such
differences, researchers could model habitat and movement preferences using
season-dependent models, or, alternatively, by combining hidden markov models
with step-selection functions. (cite papers that fieberg sent)

Connectivity analyses that omit dynamism, be it due to seasonality or climate
change, not only fail to capture an important aspect of reality, but also forego
the ability to gain an understanding of connectivity under ``non-average''
conditions, such as those expected under climate change. Reqires more data on
dispersers -> Difficult to get -> Link to citizen science

\subsection{Citizen Science}
Validating predictions from individual-based dispersal models is challenging and
requires additional dispersal data, which is inherently scarce anyways. Sciticen
science may help to fill this gap by augmenting observed GPS data with
occasional sightings of uncollared dispersing coalitions. This is especially
ciritical for species that disperse across borders and beyond confined study
areas. The African carnivore wildbook offers ...

\subsection{Conclusion}
Our dispersal simulations across two extreme climatic scenarios reveal striking
differences in dispersal prospects and landscape connectivity for dispersing
AWDs. This implies that (1) climatic variation, be it due to seasonality or
climate change, must be included in anlayses dealing with dispersal and (2) that
projected climate change is likely to have profound impacts on landscape
connectivity.

Nevertheless, a more comprehensive understanding of dispersal under
changing environmental conditions is critical for designing effective
conservation strategies in the light of climate change. By predicting and
modeling dispersal patterns under different scenarios, we can better anticipate
the impacts of climate change on species and design effective conservation
strategies to protect them. This is especially important for endangered species
that are already at the verge of extinction.

\section{Authors' Contributions}
D.D.H., G.C., D.M.B., A.O. and conceived the study and designed methodology;
D.D.H., G.C., D.M.B., and J.W.M. collected the data; D.D.H. analysed the data;
G.C., D.M.B., and A.O. assisted with modeling; D.D.H., G.C., and D.M.B.wrote the
first draft of the manuscript and all authors contributed to the drafts at
several stages and gave final approval for publication.

\section{Data Availability}
Access to R-scripts to replicate our anlaysis will be provided through an online
repository at the time of publication.

\section{Acknowledgements}
We thank the Ministry of Environment and Tourism of Botswana for granting
permission to conduct this research. We thank C. Botes, I. Clavadetscher, and G.
Camenisch for assisting with wild dog immobilizations. We also thank B. Abrahms
for sharing her data of three dispersing wild dogs. This study was funded by
Albert-Heim Foundation, Basler Stiftung für Biologische Forschung, Claraz
Foundation, Idea Wild, Jacot Foundation, National Geographic Society, Parrotia
Stiftung, Stiftung Temperatio, Wilderness Wildlife Trust Foundation,
Forschungkredit der Universität Zürich, and a Swiss National Science Foundation
Grant (31003A\_182286) to A. Ozgul.

\newpage
\begingroup
\singlespacing
\bibliography{Literature}
\endgroup

\end{document}