### 2024-07-01 Chapter_4/02_R-Scripts/03_AppliedExamples/Example_01.Rmd

---
title : 'Example 1: Dynamic + Model Approach with Simulated Data'
author: "David D. Hofmann, Gabriele Cozzi, John Fieberg"
date  : "2024-01-30"
output:
  html_document:
    toc      : true
    toc_depth: 2
---
```{r setup, echo = FALSE}

knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

```

## Purpose
This example analysis serves to illustrate the application of the _dynamic +
model_ approach with simulated data. We're first going to simulate a trajectory
with known movement kernel and habitat-selection parameters (we call this the
truth). We will then introduce data-missingness by randomly removing some of the
simulated data-points. Finally, we will apply the _dynamic + model_ approach and
try to obtain the true simulation parameters.

## Disclaimer
In this script, we'll utilize a set of custom functions that are stored in the
`Functions.R` file. They are used to simulate covariates and movement, compute
step-metrics, generate random-steps, and fit step distributions dynamically.
Most of the functions can be replaced by much more generalized functions from
the `amt` R-package. However, `amt`'s `random_steps()` function doesn't cope
well with sampling random steps from distributions fit to different
step-durations. We'll therefore use of our custom code. If you'd like to apply
it to your own data, inspect the functions closels and verify that they suit
your purposes.

## Preamble
Let's start by preparing our R-Session and loading all necessary packages. We
also load the custom functions stored in the `Functions.R` file.

```{r, warning = F, message = F}

# Set notation standards
options(scipen = 999)

# Clear R's brain
rm(list = ls())

# Load required packages
library(tidyverse)     # For data wrangling
library(sf)            # For plotting spatial features
library(lubridate)     # To handle dates and times
library(raster)        # For spatial data manipulation
library(amt)           # To fit step-length and turning-angle distributions
library(survival)      # To fit conditional logistic regression model
library(ggh4x)         # For nested ggplot facets

# Load custom functions
source("02_R-Scripts/Functions.R")

```

## Simulation
We start by simulating a set of raster layers that represent the landscape on
which we want to simulate movement. For this, we'll use the functions
`simDistance()`, `simElevation()`, and `simForest()`, which allow simulating
spatially auto-correlated layers. You may also use functions from other
packages, such as the `NLMR` or `gstat` packages, to generate such layers.

```{r}

# Set a seed for reproducability
seed <- 123456
set.seed(seed)

# Simulate landscape covariates (n x n)
n      <- 300
dist   <- simDistance(n, x = n / 2, y = n / 2)
elev   <- simElevation(n, autocorr_range = 10, seed = seed)
forest <- simForest(n, autocorr_range = 10, prop = 0.5, seed = seed + 1)
cov <- stack(dist, elev, forest)
names(cov) <- c("dist", "elev", "forest")

# Visualize the simulated covariates
cov %>%
  as.data.frame(xy = T) %>%
  pivot_longer(dist:forest, names_to = "Covariate", values_to = "Value") %>%
  ggplot(aes(x = x, y = y, fill = Value)) +
    geom_raster() +
    scale_fill_viridis_c() +
    scale_color_datetime(low = "red", high = "green", name = "Timestamp") +
    facet_wrap(~ Covariate) +
    coord_equal() +
    theme_minimal() +
    theme(strip.background = element_rect(fill = "gray95", color = "white"))

```

On this landscape, we now simulate a trajectory using the `move()` function. The
function requires quite a few inputs, such as the starting location (`xy`), the
covariate layers (`covars`), as well as a model formula (`formula`) and
associated betas (`prefs`) that determine the habitat-selection function.
Furthermore, we need to provide the step-length and turning-angle distributions
from which random steps are sampled (`sl_dist` and `ta_dist`). The `move`
function assumes a gamma distribution (with shape and scale parameters) for
step-lengths and a von Mises distribution (with a concentration parameter kappa)
for turning-angles. Finally, there are some simulation parameters that we need
to define, including the extent on which simulated animals are allowed to move
(`ext`), the number of steps simulated (`n_steps`), the number of random-steps
proposed at each relocation (`n_rsteps`), whether the simulation should stop if
one of the random steps leaves the study area (`stop`) and if a
simulation-progressbar should be printed (`messages`).

```{r, message = F}

# Simulate a trajectory across this landscape
sim <- move(
    xy       = cbind(n / 2, n / 2)
  , covars   = cov
  , formula  = ~ dist + elev + forest
  , prefs    = c(-20, 1, -1)
  , sl_dist  = list(name = "gamma", params = list(shape = 3, scale = 1))
  , ta_dist  = list(name = "vonmises", params = list(kappa = 0.5, mu = 0))
  , ext      = as(extent(c(0, n, 0, n)), "SpatialPolygons")
  , n_steps  = 1000
  , n_rsteps = 25
  , stop     = F
  , messages = F
)

# Give the individual a random name
sim$id <- "Hermes"

# Add timestamps
sim$timestamp <- ymd_hms("2023-01-01 09:00:00") + hours(1) * sim$step_number

# Check it
head(sim, n = 10)

```

The `sim` data-frame represents the full "TRUE" movement trajectory of the
simulated animal. Note that we simulated that fixes were collected exactly on
the hour. This is usually not the case in real applications, but simplifies the
subsequent code a bit. In any case, it's straight forward to round timestamps of
real data to the nearest hour or minute to avoid issues with slight deviations
from the anticipated schedule. To mimic what real data would look like, we
should also remove the step-metrics (`absta`, `ta`, `sl`), as they are usually
not observed in reality and later have to be inferred from observed coordinates.

```{r}

# Remove columns that are not observed in reality
sim[, c("absta", "ta", "sl", "step_number")] <- NULL
head(sim, n = 10)

# Visualize track
ggplot(sim, aes(x = x, y = y, col = timestamp)) +
  geom_path() +
  scale_color_datetime(low = "red", high = "green", name = "Timestamp") +
  coord_equal() +
  theme_minimal()

```

For later, we should keep track of the true simulation parameters. This will
allow us to compare our model estimates from the iSSF model to the true
parameters.

```{r}

# Keep track of true simulation parameters
truth <- data.frame(
    Term  = c("dist", "elev", "forest", "shape", "scale", "kappa")
  , Truth = c(-20, 1, -1, 3, 1, 0.5)
)

```

Now we rarify the full trajectory and create an "incomplete" dataset where some
locations are missing.

```{r}

# Rarify track by randomly removing 25% of the fixes. You'll see that the number
# of fixes drops from 1000 to 750. We can then give each step an id (unique
# across individuals if there are many) and number (not unique across
# individuals if there are many)
sim_rarified <- sim %>%
  rarifyData(missingness = 0.25) %>%
  mutate(
      step_id     = 1:n()
    , step_number = 1:n()
  )

# Compare number of fixes
cbind(Original = nrow(sim), Rarified = nrow(sim_rarified))

```

The rarified trajectory is our point of departure and represents what we would
observe in reality. Let's again visualize everything.

```{r}

# Visualize covariates and the rarified track
cov_df <- cov %>%
  as.data.frame(xy = T) %>%
  pivot_longer(dist:forest, names_to = "Covariate", values_to = "Value")
ggplot() +
  geom_raster(data = cov_df, aes(x = x, y = y, fill = Value)) +
  geom_path(data = sim_rarified, aes(x = x, y = y, col = timestamp)) +
  scale_fill_viridis_c() +
  scale_color_datetime(low = "red", high = "green", name = "Timestamp") +
  facet_wrap(~ Covariate) +
  coord_equal() +
  theme_minimal() +
  theme(strip.background = element_rect(fill = "gray95", color = "white"))

```

## Fitting Tentative Distributions
Because the dataset is incomplete, the time between consecutive datapoints is
not always equal. Let's figure out the regular (i.e. most frequent)
step-duration.

```{r}

# Assess the regular step-duration of our data
samp <- sim_rarified %>%
  computeDurations() %>%
  count(duration)
samp

```

As expected (since we simulated so), the regular step-duration is 1 hour.
However, we also observe that in a few cases the duration between two fixes can
be as long as `r max(samp$duration)` hours. In regular iSSF, we would only
consider bursts within which step-durations are somewhat comparable. Here,
however, we are tolerant towards unequal step-durations, as we'll later account
for differing step-durations in the model. Consequently, we would like to
accommodate for differing durations by fitting tentative distributions for
different step-durations. Since there are barely any fixes that are separated >
3 hours apart, having a forgiveness larger than 3 won't add much. Let's
therefore fix our forgiveness to 3 and fit the associated step-length and
turning angle distributions using the `fitDists()` function. Depending on the
missingness, some step-durations may not occur very often, making it difficult
to parametrize the associated gamma and von Mises distributions. We can
alleviate this problem by resampling the observed data to different
step-durations (option `resample = T`) or by increasing missingness even more
(option `rarify = T`). Here, we chose to use the resampling option.

```{r}

# Fit step-length and turning angle distributions to different durations
dists <- fitDists(sim_rarified
  , durations = 1:3 # Return fitted distributions for durations 1-3
  , regular   = 1   # The regular step-duration is 1
  , dynamic   = T   # Fit distributions dynamically
  , resample  = T   # Resample track to the different durations
  , rarify    = F   # Don't increase missingness
)

# Show the fitted values
dists

```

You can see that we now have different gamma and von Mises parameters for steps
from different durations. We can later use the so fitted distributions to
generate random steps matching different step-durations.

## Generating Random Steps
Before generating random steps, we need to split our data into bursts within
which the step-duration does not exceed 3.

```{r}

# Compute bursts within which the step duration does not exceed our forgiveness
sim_rarified_bursted <- computeBursts(sim_rarified, max_duration = 3)
head(sim_rarified_bursted)

```

In addition, we need to compute step-metrics (step-lengths and turning-angles)
per burst.

```{r}

# Compute step metrics within bursts
sim_rarified_metrics <- computeMetrics(sim_rarified_bursted)
head(sim_rarified_metrics)

```

Finally, we can use the `computeSSF()` function to generate 25 random steps for
each valid step (a valid step has a step-length and a turning-angle associated).
Note that we cna use the dynamically fitted distributions `dists` to match
sampled random steps with the step-durations of observed steps by picking the
`dynamic+model` approach. The function `computeSSF()` will also automatically
drop any invalid steps or steps with a duration that occurs less than 5 times in
the dataset. This is to prevent issues when including the step-duration in the
iSSF model.

```{r}

# Generate random steps
sim_rarified_ssf <- computeSSF(sim_rarified_metrics
  , n_rsteps = 25
  , dists    = dists
  , approach = "dynamic+model"
)
head(sim_rarified_ssf)

```

Let's visualize some of the observed and random steps to ensure everything is
working as expected.

```{r}

# Verify that steps of different durations follow different sl- and
# ta-distributions
sim_rarified_ssf %>%
  dplyr::select(case, duration, sl, relta) %>%
  pivot_longer(sl:relta, names_to = "Metric", values_to = "Value") %>%
  mutate(Type = ifelse(case == 1, "Observed", "Random")) %>%
  ggplot(aes(x = Value, col = as.factor(duration))) +
    geom_density() +
    facet_nested_wrap(~ Type + Metric, scale = "free") +
    scale_color_viridis_d(name = "Step-Duration") +
    theme_minimal() +
    theme(strip.background = element_rect(fill = "gray95", color = "white"))

# Visualize some observed and random steps
sim_rarified_ssf %>%
  mutate(Type = ifelse(case == 1, "Observed", "Random")) %>%
  subset(step_number %in% 1:1000) %>%
  ggplot(aes(x = x, y = y, xend = x_to, yend = y_to, col = as.factor(duration))) +
    geom_segment(linewidth = 0.1) +
    scale_color_viridis_d(begin = 0.3, name = "Duration") +
    theme_minimal() +
    coord_equal() +
    facet_wrap(~ Type) +
    theme(strip.background = element_rect(fill = "gray95", color = "white"))

```

## Computing Covariates
We still need to compute all covariates that we want include our iSSF model.
This includes spatial covariates at the end-point of each step, as well as
step-metrics. We can use the function `computeCovars` for this.

```{r}

# Compute covariates for all steps (the function will extract spatial covariates
# at the end of each step and also compute log_sl and cos_ta)
sim_rarified_covars <- computeCovars(sim_rarified_ssf
  , covars    = cov
  , extract   = "end"
)
head(sim_rarified_covars)

```

## Running the Model

We are now ready to estimate parameters using conditional logistic regression.
This can be achieved using the `clogit()` function from the `survival` package.
Alternatively, one can use the wrapper `fit_clogit()` from the `amt` package.
Importantly, we will include interactions among the step-descriptors $sl$,
$log(sl)$, and $cos(ta)$ and the step-duration $duration$.

```{r}

# Ensure the step-duration enters as factor
sim_rarified_covars$duration <- as.factor(sim_rarified_covars$duration)

# Fit the model
mod <- clogit(data = sim_rarified_covars, formula = case ~
  + dist            # Habitat-selection
  + elev            # Habitat-selection
  + forest          # Habitat-selection
  + sl              # Movement Kernel
  + log_sl          # Movement Kernel
  + cos_ta          # Movement Kernel
  + sl:duration     # Movement x Step-Duration
  + log_sl:duration # Movement x Step-Duration
  + cos_ta:duration # Movement x Step-Duration
  + strata(step_id)
)
summary(mod)

```

This looks good. Now we extract the model coefficients from the fitted model and
use them to update our "tentative" distribution parameters for the gamma and von
Mises distribution. Since the true data was simulated with $\Delta t = 1$, we
will only update parameters for the case when $\Delta t = 1$.

```{r}

# Compute confidence intervals
ci <- confint(mod, level = 0.95)

# Put relevant results together
coefs <- summary(mod)$coefficients
coefs <- data.frame(
    Coefficient = rownames(coefs)
  , Estimate    = coefs[, "coef"]
  , SE          = coefs[, "se(coef)"]
  , Z           = coefs[, "z"]
  , LCI         = ci[, 1]
  , UCI         = ci[, 2]
)

# We can use this dataframe to update our tentative distribution parameters.
corrected <- correctDists(list(
    Dists = dists
  , Coefs = coefs
))
corrected$Dists$updated

```

Finally, we put all results together into a `data.frame` and compare them to the
truth.

```{r}

# Now put everything together into a dataframe
mov_estimates <- corrected$Dists
hab_estimates <- corrected$Coefs

# Compile the updated distribution parameters (for step duration of one)
mov_estimates <- tibble(
    shape = as.numeric(mov_estimates$updated$sl$shape)
  , scale = as.numeric(mov_estimates$updated$sl$scale)
  , kappa = as.numeric(mov_estimates$updated$ta$kappa)
  ) %>%
  pivot_longer(shape:kappa, names_to = "Term", values_to = "Estimate") %>%
  mutate(Kernel = "Movement")

# Compile the habitat selection coefficients
hab_estimates <- hab_estimates %>%
  dplyr::select(Term = Coefficient, Estimate) %>%
  subset(Term %in% c("forest", "dist", "elev")) %>%
  mutate(Kernel = "Habitat")

# Put them together
estimates <- rbind(hab_estimates, mov_estimates, make.row.names = F)
estimates

# Compare to truth
comparison <- left_join(truth, estimates, by = "Term")
comparison

```

Wonderful. It looks like we were able to approximate the true simulation
parameters quite well.

```{r}
sessionInfo()
```